---
title: "Figure Drug distinct responses and SASP"
author: "Joanna Yeung"
date: '2025-01-27'
output: html_document
---

```{r}
# data wrangling
library(preprocessCore)
library(readr)
library(readxl)
library(stringr)
library(dplyr)

# plotting
library(gridExtra)
library(ggplot2)
library(pheatmap)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(scales)
library(ggrepel)

# genomics data wrangling and analysis
library(rtracklayer)
library(Biostrings)
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(DESeq2)
library(apeglm)
library(ChIPseeker)
library(clusterProfiler)
library(msigdbr)
library(MotifDb)
library(motifmatchr)

# genome annotation databases
library(EnsDb.Hsapiens.v79)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(BSgenome.Hsapiens.UCSC.hg38)
library(org.Hs.eg.db)
```

# make function to specify code chunk to run from other Rmd file to avoid copying and pasting large lines of code. 
```{r}
# arguments: 
# Rmd_file=file path & name of Rmd file with code chunk
# chunk_name= name of code chunk to run from Rmd file. 
run_code_chunk <- function(Rmd_file, chunk_name){ 
  # Read the Rmd file
lines <- readLines(Rmd_file)

# Identify the start and end of the specific chunk
chunk_start <- grep(paste0("^```\\{r\\s*", chunk_name), lines)  # Find chunk start
chunk_end <- grep("^```$", lines)  # Find all chunk end lines

# Find the first `chunk_end` after `chunk_start`
chunk_end <- chunk_end[chunk_end > chunk_start][1]

  if (is.na(chunk_end)) {
    stop(paste("End of chunk", chunk_name, "not found. Check the Rmd file for missing delimiters."))
  }

# specify lines inside chunk, removing lines that make up the start and end of the chunk. 
chunk_to_run <- lines[(chunk_start[1]+1):(chunk_end[1]-1)]

# Evaluate the extracted code
  eval(parse(text = paste(chunk_to_run, collapse = "\n")), envir = .GlobalEnv)
}
```

# annotation colors for pheatmap
```{r}
hm_cols_RNA <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("Cycling"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD"), 
                Biorep=c("1"="orange", "2"="light blue", "3"="pink"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "DoxovsPalbo"="#CCEBC5"), 
                Clusters=c("2"="#66C2A5","5"="#A6D854","7"="#8DA0CB", "1"="#FC8D62", "4"="#E78AC3","6"="#FFD92F","3"="#E5C494"), 
                SASP = c("Yes"="dark green", "No"="white"),
                Cell_Cycle = c("Yes"="dark blue", "No"="white"), 
                clusters_cutree8=brewer.pal(8, "Set2"), 
                p53_All=c("Yes"="#00B0F0", "No"="white"),
                NFKB_All=c("Yes"="red", "No"="white"))
names(hm_cols_RNA$clusters_cutree8) <- as.factor(1:8)


hm_cols_ATAC <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("None"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "PvsD"="#CCEBC5"),
                Biorep=c("1"="orange", "2"="light blue", "3"="pink", "4"="light green"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "PvsD"="#CCEBC5"),
                Peak_anno=c("Distal"="#B15A28", "Gene Body"="#7F54A2", "Promoter"="#A7CEE2"), 
                Doxo=c("Yes"="#B3C543", "No"="white"),
                Palbo=c("Yes"="#E0ADCD", "No"="white")
                )
```

# download msigdb databases, to be used for GO enrichment analysis. 
```{r}
# msigdb GO Biological Processes database
msigdbr_df_GO_BP <- as.data.frame(msigdbr(species = "Homo sapiens", category="C5", subcategory ="GO:BP"))
msigdbr_t2g_GO_BP = msigdbr_df_GO_BP %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_GO_BP$gs_name <- sub("GOBP_", "", msigdbr_t2g_GO_BP$gs_name)
msigdbr_t2g_GO_BP$gs_name <- gsub("_", " ", msigdbr_t2g_GO_BP$gs_name)

# msigdb GO Molecular Function Processes database
msigdbr_df_GO_MF <- as.data.frame(msigdbr(species = "Homo sapiens", category="C5", subcategory ="GO:MF"))
msigdbr_t2g_GO_MF = msigdbr_df_GO_MF %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_GO_MF$gs_name <- sub("GOMF_", "", msigdbr_t2g_GO_MF$gs_name)
msigdbr_t2g_GO_MF$gs_name <- gsub("_", " ", msigdbr_t2g_GO_MF$gs_name)

# msigdb Hallmark database
msigdbr_df_H <- as.data.frame(msigdbr(species = "Homo sapiens", category="H", subcategory =""))
msigdbr_t2g_H = msigdbr_df_H %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_H$gs_name <- sub("HALLMARK_", "", msigdbr_t2g_H$gs_name)
msigdbr_t2g_H$gs_name <- gsub("_", " ", msigdbr_t2g_H$gs_name)

# msigdb KEGG pathway database
msigdbr_df_KEGG <- as.data.frame(msigdbr(species = "Homo sapiens", category="C2", subcategory ="CP:KEGG"))
msigdbr_t2g_KEGG = msigdbr_df_KEGG %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_KEGG$gs_name <- sub("KEGG_", "", msigdbr_t2g_KEGG$gs_name)
msigdbr_t2g_KEGG$gs_name <- gsub("_", " ", msigdbr_t2g_KEGG$gs_name)

# msigdb Reactome database
msigdbr_df_REACT <- as.data.frame(msigdbr(species = "Homo sapiens", category="C2", subcategory="CP:REACTOME"))
msigdbr_t2g_REACT = msigdbr_df_REACT %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_REACT$gs_name <- sub("REACTOME_", "", msigdbr_t2g_REACT$gs_name)
msigdbr_t2g_REACT$gs_name <- gsub("_", " ", msigdbr_t2g_REACT$gs_name)
```


#### universe (background gene set) for GO enrichment analysis is all genes expressed
```{r}
universe <- genes_anno_type[match(gsub('\\.[0-9]*$', '', row.names(acrossGroups)), genes_anno_type$ensembl_id), ]
```

# GO enrichment analysis on k-means clusters from sigMat (significantly changing genes across all conditions, based on Likelihood ratio test)

## *function to get GO enriched terms*:
## arguments:
### gene= vector of genes to test for GO enrichment (gene ensembl format)
### universe=background gene set 
### Database=database containing groups of genes to GO annotation term e.g. msigdb GO:BP
```{r getGOenrichment function}
getGOenrichment <-function(gene,universe,Database){
  #gene_entrezID <-  AnnotationDbi::select(EnsDb.Hsapiens.v79, gene[!is.na(gene)], "ENTREZID", "SYMBOL") #convert gene IDs from SYMBOL format to ENTREZ ID format (was using v79, new code is using v107)
  gene_entrezID <- genes_anno_type[match(gene, genes_anno_type$gene_id), ]
  gene_go <- enricher(gene_entrezID$entrezid, TERM2GENE=Database, universe=universe) #enrichment analysis with hypergeometric test
  gene_go <- setReadable(gene_go, OrgDb = org.Hs.eg.db, keyType="ENTREZID") #convert enrichment results from ENTREZ ID format to SYMBOL for easier interpretation
return(gene_go)
} 
```

# assign significant genes to relevant k-means clusters & drug + cell cycle arrest categories. 
```{r genes_in_clusters}
# load in workspace with DESeq results and rlog counts matrix of genes. 
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Global_transcriptome_and_chromatin_accessibility_is_dynamic_in_both_types_of_therapy_induced_senescence/workspace.RData")
genes_in_clusters <- lapply(levels(clusterDF$Clusters), function(x){clusterDF[clusterDF$Clusters %in% x, ]$gene_id})
cluster_names <- c("CellCycle1", "CellCycle2", "Palbo1", "Palbo2", "Doxo1", "Doxo2", "Shared") 
names(genes_in_clusters) <- cluster_names
```

# RNA-seq: GO enrichment on significant genes
```{r RNA-seq GO enrichment analysis}
# GO enrichment analysis with msigdb GO:BP database:
GOBP_enrichment_clusters <- list()
# GO enrichment analysis with msigdb REACT database:
GO_enrichment_clusters_KEGG <- list()
# GO enrichment analysis with msigdb Hallmark database:
GO_enrichment_clusters_HallMark <- list()
# GO enrichment analysis with msigdb REACT database:
GO_enrichment_clusters_REACT <- list()
# GO enrichment analysis with msigdb GO:MF database:
GO_enrichment_clusters_MF <- list()

for(i in 1:length(genes_in_clusters)){
  
GOBP_enrichment_clusters[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_GO_BP) # do GO enrichment analysis with background gene set set as all gene expressed, database used is msigdb GO:BP

GO_enrichment_clusters_KEGG[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_KEGG)

GO_enrichment_clusters_HallMark[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_H)

GO_enrichment_clusters_REACT[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_REACT)

GO_enrichment_clusters_MF[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_GO_MF)
}
```

### RNA-seq: format GO enrichment terms from each cluster into matrix of -log10(padj values)
```{r RNAseq GO enrichment results}
# convert each clusters' GO enrichment terms into dataframes
GO_enrichment_clusters_KEGG_df <- list()
 GO_enrichment_clusters_REACT_df <- list()
 GO_enrichment_clusters_HallMark_df <- list()
 GOBP_enrichment_clusters_df <- list()
GO_enrichment_clusters_MF_df <- list()
for(i in 1:length(GO_enrichment_clusters_REACT)){
  GO_enrichment_clusters_KEGG_df[[i]] <- as.data.frame(GO_enrichment_clusters_KEGG[[i]])
   GO_enrichment_clusters_REACT_df[[i]] <- as.data.frame(GO_enrichment_clusters_REACT[[i]])
 GO_enrichment_clusters_HallMark_df[[i]] <- as.data.frame(GO_enrichment_clusters_HallMark[[i]])
 GOBP_enrichment_clusters_df[[i]] <- as.data.frame(GOBP_enrichment_clusters[[i]])
  GO_enrichment_clusters_MF_df[[i]] <- as.data.frame(GO_enrichment_clusters_MF[[i]])
}

 # indicate names of each list as the cluster number
 names(GO_enrichment_clusters_KEGG_df) <- cluster_names 
 names(GO_enrichment_clusters_REACT_df) <- cluster_names 
 names(GO_enrichment_clusters_HallMark_df) <- cluster_names 
 names(GOBP_enrichment_clusters_df) <- cluster_names 
names(GO_enrichment_clusters_MF_df) <- cluster_names 
 
 # remove any cluster of genes that have 0 GO enrichment terms
GO_enrichment_clusters_KEGG_df <-  GO_enrichment_clusters_KEGG_df[sapply( GO_enrichment_clusters_KEGG_df, nrow)>0]
GO_enrichment_clusters_REACT_df <-  GO_enrichment_clusters_REACT_df[sapply( GO_enrichment_clusters_REACT_df, nrow)>0]
GO_enrichment_clusters_HallMark_df <-  GO_enrichment_clusters_HallMark_df[sapply( GO_enrichment_clusters_HallMark_df, nrow)>0]
GOBP_enrichment_clusters_df <-  GOBP_enrichment_clusters_df[sapply( GOBP_enrichment_clusters_df, nrow)>0]
GO_enrichment_clusters_MF_df <-  GO_enrichment_clusters_MF_df[sapply( GO_enrichment_clusters_MF_df, nrow)>0]
 
 # indicate which cluster and database each dataframe comes from 
 for(i in 1:length(GO_enrichment_clusters_KEGG_df)){
  GO_enrichment_clusters_KEGG_df[[i]]$Cluster <- names(GO_enrichment_clusters_KEGG_df)[i]
  GO_enrichment_clusters_KEGG_df[[i]]$Database <- "KEGG"
 }
  for(i in 1:length(GO_enrichment_clusters_REACT_df)){
   GO_enrichment_clusters_REACT_df[[i]]$Cluster <-  names(GO_enrichment_clusters_REACT_df)[i]
   GO_enrichment_clusters_REACT_df[[i]]$Database <- "REACTOME"
  }
  for(i in 1:length(GO_enrichment_clusters_HallMark_df)){
 GO_enrichment_clusters_HallMark_df[[i]]$Cluster <- names(GO_enrichment_clusters_HallMark_df)[i]
 GO_enrichment_clusters_HallMark_df[[i]]$Database <- "Hallmark"
  }
 
  for(i in 1:length(GOBP_enrichment_clusters_df)){
 GOBP_enrichment_clusters_df[[i]]$Cluster <- names(GOBP_enrichment_clusters_df)[i]
  GOBP_enrichment_clusters_df[[i]]$Database <- "GO:BP"
  }

  for(i in 1:length(GO_enrichment_clusters_MF_df)){
 GO_enrichment_clusters_MF_df[[i]]$Cluster <- names(GO_enrichment_clusters_MF_df)[i]
 GO_enrichment_clusters_MF_df[[i]]$Database <- "GO:MF"
  }

# combine GO enrichment results from different databases into 1 dataframe for each cluster of genes. 
GO_results_list <- vector()
GO_results_list$CellCycle1 <- rbind(GO_enrichment_clusters_KEGG_df$CellCycle1,
                                    GO_enrichment_clusters_HallMark_df$CellCycle1,
                                    GO_enrichment_clusters_REACT_df$CellCycle1,
                                    GOBP_enrichment_clusters_df$CellCycle1,
                                    GO_enrichment_clusters_MF_df$CellCycle1)

GO_results_list$CellCycle2 <- rbind(GO_enrichment_clusters_KEGG_df$CellCycle2,
                                    GO_enrichment_clusters_HallMark_df$CellCycle2,
                                    GO_enrichment_clusters_REACT_df$CellCycle2,
                                    GOBP_enrichment_clusters_df$CellCycle2,
                                    GO_enrichment_clusters_MF_df$CellCycle2)

GO_results_list$Palbo1 <- rbind(GO_enrichment_clusters_KEGG_df$Palbo1,     
                                GO_enrichment_clusters_HallMark_df$Palbo1 , 
                                GO_enrichment_clusters_REACT_df$Palbo1, 
                                GOBP_enrichment_clusters_df$Palbo1, 
                                GO_enrichment_clusters_MF_df$Palbo1)

GO_results_list$Palbo2 <- rbind(GO_enrichment_clusters_KEGG_df$Palbo2,
                                GO_enrichment_clusters_HallMark_df$Palbo2,
                                GO_enrichment_clusters_REACT_df$Palbo2,
                                GOBP_enrichment_clusters_df$Palbo2, 
                                GO_enrichment_clusters_MF_df$Palbo2)

GO_results_list$Doxo1 <- rbind(GO_enrichment_clusters_KEGG_df$Doxo1,
                               GO_enrichment_clusters_HallMark_df$Doxo1,
                               GO_enrichment_clusters_REACT_df$Doxo1,
                               GOBP_enrichment_clusters_df$Doxo1, 
                               GO_enrichment_clusters_MF_df$Doxo1)

GO_results_list$Doxo2 <- rbind(GO_enrichment_clusters_KEGG_df$Doxo2,
                               GO_enrichment_clusters_HallMark_df$Doxo2,
                               GO_enrichment_clusters_REACT_df$Doxo2,
                               GOBP_enrichment_clusters_df$Doxo2, 
                               GO_enrichment_clusters_MF_df$Doxo2)

GO_results_list$Shared <- rbind(GO_enrichment_clusters_KEGG_df$Shared,
                                GO_enrichment_clusters_HallMark_df$Shared,
                                GO_enrichment_clusters_REACT_df$Shared,
                                GOBP_enrichment_clusters_df$Shared, 
                                GO_enrichment_clusters_MF_df$Shared)

# order GO enrichment results from different databases by padj values
for(i in 1:length(GO_results_list)){
  GO_results_list[[i]] <- GO_results_list[[i]][order(GO_results_list[[i]]$p.adjust), ]
}

# combine dataframes of each cluster of gene's GO enrichment results into 1 master dataframe
GO_results_df <- do.call(rbind, GO_results_list)

# create matrix of -log10pvalues of the top 50 enriched pathways for each cluster

GO_results_list_mat <- matrix(nrow=length(unique(unlist(lapply(GO_results_list, function(GO_results_list){GO_results_list$ID[1:50]})))), ncol=length(GO_results_list)) # make empty matrix, number of columns are the clusters of genes and number of total top GO terms of all clusters

row.names(GO_results_list_mat) <- unique(unlist(lapply(GO_results_list, function(GO_results_list){GO_results_list$ID[1:50]}))) # row names are top 12 GO enrichment terms for each cluster combined
colnames(GO_results_list_mat) <- cluster_names  # column names are clusters

# pathways that are not enriched in the cluster are listed as NA so just change the value to 0 so that it can be plotted in heatmap
for(i in 1:ncol(GO_results_list_mat)){
GO_results_list_mat[ ,i] <- -log10(GO_results_list[[i]][match(row.names(GO_results_list_mat), GO_results_list[[i]]$ID), ]$p.adjust)
GO_results_list_mat[ ,i] <- ifelse(GO_results_list_mat[ ,i] %in% NA, 0, GO_results_list_mat[ ,i])
}
GO_results_list_mat <- GO_results_list_mat[!is.na(row.names(GO_results_list_mat)), ] # remove padj values that are NA 

# specify GO results matrix annotation colors: based on ggsci:pal_aaas palette. 
GO_anno_colors <-  list(database=c("GO:BP"="#3B4992FF", "Hallmark"="#EE0000FF", "KEGG"="#008B45FF", "REACTOME" ="#631879FF", "GO:MF"="#808180FF"), 
                        Cluster=c("CellCycle1"="#66C2A5", "CellCycle2"="#A6D854", "Palbo1"="#8DA0CB", "Palbo2"= "#FC8D62","Doxo1"="#E78AC3","Doxo2"="#FFD92F", "Shared"="#E5C494"))
names(GO_anno_colors$Cluster) <- cluster_names 

# annotation dataframe for rows of heatmap to indicate which database GO term came from 
GO_anno <- data.frame(row.names=row.names(GO_results_list_mat), database=GO_results_df[match(row.names(GO_results_list_mat), GO_results_df$ID), ]$Database)
# annotation dataframe for columns of heatmap to color by Cluster colors. 
Cluster_anno <- data.frame(Cluster=cluster_names , row.names = colnames(GO_results_list_mat))
```

### MEME-AME analysis on DESeq_Peaks pairwise comparison 
```{bash}
cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq

# all up peaks- background peak set set to corresponding down peaks of the same DESeq comparison. 
sbatch -p risc,hpc --array=1-16 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoDown/inputfiles_annoDown_DESeq_peaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoUp/inputfiles_annoUp_DESeq_peaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_output/DESeq/annoUp,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoUp/folderlist_annoUp_DESeq_peaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm

# all down peaks- background peak set set to corresponding up peaks of the same DESeq comparison. 
sbatch -p risc,hpc --array=1-16 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoUp/inputfiles_annoUp_DESeq_peaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoDown/inputfiles_annoDown_DESeq_peaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_output/DESeq/annoDown,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoDown/folderlist_annoDown_DESeq_peaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm
```

# RNA-seq: GSEA Analysis
```{r RNAseq GSEA analysis}
ranked_res_forGSEA <- list()
for(i in 1:length(LFCResultsall_RNA)){
  
  rank <- LFCResultsall_RNA[[i]]
  
  # Assign a default p-value for genes with NA padj (maximum p-value)
  rank$pvalue[is.na(rank$padj)] <- 1 
  
  # Create a ranked list based on log2FoldChange and -log10(p-value)
  rank$rank_metric <- sign(rank$log2FoldChange) * -log10(rank$pvalue)
  
  # Sort by the ranking metric
  rank <- rank[order(rank$rank_metric, decreasing = TRUE), ]
  
  # Store the ranked values
  ranked_res_forGSEA[[i]] <- rank$rank_metric
  names(ranked_res_forGSEA[[i]]) <- row.names(rank)

  # Convert gene IDs to Entrez IDs if available
  entrez_ids <- genes_anno_type[match(names(ranked_res_forGSEA[[i]]), genes_anno_type$ensembl_id), ]$entrezid
  names(ranked_res_forGSEA[[i]]) <- ifelse(is.na(entrez_ids), names(ranked_res_forGSEA[[i]]), entrez_ids)
}

names(ranked_res_forGSEA) <- names(LFCResultsall_RNA)

# run GSEA analysis using: 
GSEA_Results_H <- list() # msigdb Hallmark database
GSEA_Results_KEGG <- list() # msigdb KEGG database
GSEA_Results_REACT <- list() # msigdb Reactome database
GSEA_Results_MF <- list() # msigdb GO:MF database
for(i in 1:length(ranked_res_forGSEA)){
 GSEA_Results_H[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_H) 
 GSEA_Results_KEGG[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_KEGG) 
 GSEA_Results_REACT[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_REACT) 
 GSEA_Results_MF[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_GO_MF) 
}

names(GSEA_Results_H) <- names(LFCResultsall_RNA)
names(GSEA_Results_KEGG) <- names(LFCResultsall_RNA)
names(GSEA_Results_REACT) <- names(LFCResultsall_RNA)
names(GSEA_Results_MF) <- names(LFCResultsall_RNA)
```

# RNA-seq: generate input files for IPA pathways. 
```{r}
# clusters of genes 
for(i in 1:length(genes_in_clusters)){
  
  IPA_input <- as.data.frame(genes_in_clusters[[i]])
  colnames(IPA_input) <- "ID"
  write.table(IPA_input, file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/input/", cluster_names[i], ".csv"), row.names=F, sep=",")
}
# DESeq Wald Test results for pairwise comparisons. 
for(i in 1:length(LFCResultsall_RNA)){
  write.table(as.data.frame(LFCResultsall_RNA[[i]]), file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files/DESeq_results_input/", names(LFCResultsall_RNA)[i], ".csv"), sep=",")
}

# log2FC values of each sig changing gene clusters. 
for(i in 1:length(drug_cat_list)){
table <- sig_LFCMat[row.names(sig_LFCMat) %in% row.names(drug_cat_list[[i]]), ]
    write.table(table, file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/input/sigGenes_log2FC/", cluster_names[i], "_log2FC.csv"), sep=",", row.names = T)
}
```

# RNA-seq: import IPA upstream regulators & pathways
```{r import IPA results}
IPA_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/output"
IPA_upstream_regs <- list()
IPA_pathways <- list()
# read in information for enriched IPA pathways per hierarchical cluster (clustering based on scaled rlog normalized counts)
for(i in 1:length(cluster_names)){
IPA_pathways[[i]]<- read_tsv(paste0(IPA_dir, "/", cluster_names[i], "_IPA_pathways.txt"), skip=2)
# read in information for enriched upstream regulators 
IPA_upstream_regs[[i]] <- read.table(paste0(IPA_dir, "/", cluster_names[i], "_IPA_upstream_regs.txt"), sep="\t", skip=2, header=T)
IPA_upstream_regs[[i]] <- IPA_upstream_regs[[i]][!grepl("mir|miR|Mir", IPA_upstream_regs[[i]]$Upstream.Regulator), ] # remove upstream regulators that are miRNAs (uncertain if they will be informative & there is a lot of them)
}

names(IPA_upstream_regs) <- cluster_names
names(IPA_pathways) <- cluster_names

# filter out upstream regulators that are not expressed in this cell line. 

#extract ids of all expressed genes
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL

# keep only upstream regulators that are part of expressed genes or have lower case letters (these are upstream regulators that could be miRNAs or protein complexes)
#IPA_upstream_regs <- lapply(IPA_upstream_regs, function(IPA_upstream_regs){IPA_upstream_regs[c(grep('[:lower:]', IPA_upstream_regs$Upstream.Regulator), which(IPA_upstream_regs$Upstream.Regulator %in% expressedGenes)), ]})
```

# RNA-seq: plotting padj values of top IPA predicted upstream regulators of all clusters
```{r IPA results matrix}
create_ipa_regs_matrix <- function(IPA_upstream_regs, cluster_names, filter_keyword = "complex", top_n = 25) {
  # Extract unique upstream regulators based on the filter keyword
  unique_upstream_regs <- unique(unlist(lapply(IPA_upstream_regs, function(x) {
    x$Upstream.Regulator[grepl(filter_keyword, x$Molecule.Type)][1:top_n]
  })))

  # Create a matrix to store -log10 p-values
  IPA_complex_regs_mat <- matrix(
    nrow = length(unique_upstream_regs),
    ncol = length(IPA_upstream_regs)
  )
  row.names(IPA_complex_regs_mat) <- unique_upstream_regs # Set row names to upstream regulators
  colnames(IPA_complex_regs_mat) <- cluster_names        # Set column names to cluster names

  # Fill in the matrix with -log10 p-values
  for (i in seq_len(ncol(IPA_complex_regs_mat))) {
    # Match and extract p-values for each cluster
    IPA_complex_regs_mat[, i] <- -log10(IPA_upstream_regs[[i]][
      match(row.names(IPA_complex_regs_mat), IPA_upstream_regs[[i]]$Upstream.Regulator),
      "p.value.of.overlap"
    ])

    # Replace NA values with 0 for heatmap compatibility
    IPA_complex_regs_mat[, i][is.na(IPA_complex_regs_mat[, i])] <- 0
  }

  return(IPA_complex_regs_mat)
}

IPA_group_regs_mat <- create_ipa_regs_matrix(
  IPA_upstream_regs = IPA_upstream_regs,
  cluster_names = cluster_names,
  filter_keyword = "group",
  top_n = 25
)

IPA_TF_regs_mat <- create_ipa_regs_matrix(
  IPA_upstream_regs = IPA_upstream_regs,
  cluster_names = cluster_names,
  filter_keyword = "transcription",
  top_n = 25
)

# plot heatmap of -log10 pvalue of overlap of enriched complex regulators
pheatmap(IPA_group_regs_mat, cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 10,length.out = 100))

pheatmap(IPA_TF_regs_mat, cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 10,length.out = 100))
```

# RNA-seq: import comparison analysis of IPA upstream regulators and pathways for sig changing genes
```{r import_IPA_comparison_results}
setwd("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/output/sigGenes_log2FC")
IPA_Compare_upstream_regs <- list()
IPA_Compare_pathways <- list()
# read in information for enriched IPA pathways per hierarchical cluster (clustering based on scaled rlog normalized counts)

upstream_regs_files <- dir(pattern="_IPA_upstream_regs.txt")
pathways_files <- dir(pattern="_IPA_pathways.txt")
for(i in 1:length(pathways_files)){
# read in information for enriched pathways
IPA_Compare_pathways[[i]]<- read_tsv(pathways_files[i], skip=2)
# read in information for enriched upstream regulators 
IPA_Compare_upstream_regs[[i]] <- read_tsv(upstream_regs_files[i], skip=2)

  IPA_Compare_pathways[[i]] <- IPA_Compare_pathways[[i]] %>%
  mutate(across(where(is.character), ~ replace(.x, .x == "N/A", "0"))) %>%
  mutate(across(-`Canonical Pathways`, as.numeric)) # Convert all except the first column to numeric
  
  IPA_Compare_upstream_regs[[i]] <- IPA_Compare_upstream_regs[[i]] %>%
  mutate(across(where(is.character), ~ replace(.x, .x == "N/A", "0"))) %>%
    mutate(across(-`Upstream Regulators`, as.numeric)) # Convert all except the first column to numeric
}

names(IPA_Compare_upstream_regs) <- gsub("_.*", "", pathways_files)
names(IPA_Compare_pathways) <- gsub("_.*", "", pathways_files)

# convert IPA results into matrix for plotting. 
IPA_Compare_pathways_mat <- lapply(IPA_Compare_pathways, function(x){
  mat <- as.matrix(x[ ,-1])
  row.names(mat) <- x$`Canonical Pathways`
 remove <- which(rowSums(mat != 0) == 0) # Find rows where the entire row (all columns) is exactly 0 to remove upstream regulators that have no activation score across all columns (DESeq comparisons)
    return(mat[-remove, ])
})

IPA_Compare_upstream_regs_mat <- lapply(IPA_Compare_upstream_regs, function(x){
  mat <- as.matrix(x[ ,-1])
  row.names(mat) <- x$`Upstream Regulators`
  remove <- which(rowSums(mat != 0) == 0) # Find rows where the entire row (all columns) is exactly 0 to remove upstream regulators that have no activation score across all columns (DESeq comparisons)
  return(mat[-remove, ])
})
```

# RNA-seq: make matrix with log2FC values of significant genes
```{r log2FC sig genes matrix}
# get log2FC values of significantly changing genes over time
sig_LFCMat <- matrix(ncol=length(LFCResultsall_RNA), nrow=nrow(clusterDF))
colnames(sig_LFCMat) <- names(LFCResultsall_RNA)
row.names(sig_LFCMat) <-row.names(LFCResultsall_RNA[[10]][row.names(LFCResultsall_RNA[[10]]) %in% sigGenes, ])
for(i in 1:ncol(sig_LFCMat)){
sig_LFCMat[ ,i] <- LFCResultsall_RNA[[i]][row.names(LFCResultsall_RNA[[i]]) %in% sigGenes, ]$log2FoldChange
}

row.names(sig_LFCMat) <-  make.unique(ifelse(genes_anno_type[match(row.names(sig_LFCMat), genes_anno_type$ensembl_id), ]$gene_id %in% NA, row.names(sig_LFCMat), genes_anno_type[match(row.names(sig_LFCMat), genes_anno_type$ensembl_id), ]$gene_id))

sig_LFCMat <- sig_LFCMat[match(row.names(ordered_sigMat), row.names(sig_LFCMat)), ] # order based on k-means & hierarchical ordering of rlog counts. 

# get log2FC values of all expressed genes over time 
LFCMat <- matrix(ncol=length(LFCResultsall_RNA), nrow=nrow(LFCResultsall_RNA[[1]]))
colnames(LFCMat) <- names(LFCResultsall_RNA)
row.names(LFCMat) <-row.names(LFCResultsall_RNA[[1]])
for(i in 1:ncol(LFCMat)){
LFCMat[ ,i] <- LFCResultsall_RNA[[i]]$log2FoldChange
}

row.names(LFCMat) <-  ifelse(genes_anno_type[match(row.names(LFCMat), genes_anno_type$ensembl_id), ]$gene_id %in% NA, row.names(LFCMat), genes_anno_type[match(row.names(LFCMat), genes_anno_type$ensembl_id), ]$gene_id)


anno_sigLFC <- data.frame(
  row.names = colnames(sig_LFCMat),
  log2FC = c(rep("vs Cyc", 4), rep("vs day3", 3), rep("vs Cyc", 5), rep("vs day3", 4), rep("DoxovsPalbo",4)),
  Treatment = c(rep("Doxo", 7), rep("Palbo", 9), rep("DoxovsPalbo",4)),
  Timepoint = c("3", "10", "14", "21", "10", "14", "21", "3", "10", "14", "21", "28", "10", "14", "21", "28","3", "10", "14", "21")
)
```

# ATAC-seq: create input fasta files for MEME-AME analysis on all up or down peaks. 
```{r}
directory <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq"
outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input"


# specify file paths for input to MEME-AME analysis for up peaks
folderlist <- c("day10PalbovsDoxoUpPeaks_vsCycling", 
                "day10PalbovsDoxoUpPeaks_vsday3", 
                "day14PalbovsDoxoUpPeaks_vsCycling", 
                "day14PalbovsDoxoUpPeaks_vsday3",
                "day21PalbovsDoxoUpPeaks_vsCycling", 
                "day21PalbovsDoxoUpPeaks_vsday3",
                "day28Palbovsday21DoxoUpPeaks_vsCycling", 
                "day28Palbovsday21DoxoUpPeaks_vsday3",
                "day3PalbovsDoxoUpPeaks_vsCycling")

writeLines(dir(paste0(directory, "/annoUp"), full.names=T, pattern="Palbovs"), con=paste0(outdir, "/inputfiles_annoUp_DESeq_Palbopeaks.txt")) # directory paths to fasta files of Up peaks

writeLines(c(dir(paste0(directory, "/annoUp"), full.names=T, pattern="Doxovs")[1:6], dir(paste0(directory, "/annoUp"), full.names=T, pattern="Doxovs")[5:7]), con=paste0(outdir, "/inputfiles_annoUp_DESeq_Doxopeaks.txt")) # directory paths to fasta files of Up peaks 
# since Doxo doesn't have day 28 timepoint, repeat day 21 DESeq comparison files for comparing to day28 Palbo. 

writeLines(folderlist, con=paste0(outdir, "/folderlist_annoUp_DESeq_PalbovsDoxoPeaks.txt")) # write folder names
```

### ATAC-seq: MEME-AME analysis on DESeq_Peaks pairwise comparison between up Palbo and up Doxo peaks to find differences in enriched motifs between upregulated drug distinct peaks. 
```{bash}
cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input

# what is enriched in Palbo up peaks that is different from Doxo up peaks? set Doxo up peaks as background. 
sbatch -p risc,hpc --array=1-9 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/inputfiles_annoUp_DESeq_Doxopeaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/inputfiles_annoUp_DESeq_Palbopeaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_output/annoUp_PalbovsDoxo,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/folderlist_annoUp_DESeq_PalbovsDoxoPeaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm

# what is enriched in Doxo up peaks that is different from Palbo up peaks? set Palbo up peaks as background.  
sbatch -p risc,hpc --array=1-9 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/inputfiles_annoUp_DESeq_Palbopeaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/inputfiles_annoUp_DESeq_Doxopeaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_output/annoUp_DoxovsPalbo,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/folderlist_annoUp_DESeq_PalbovsDoxoPeaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm
```

# ATAC-seq: import MEME-AME results into dataframe: MEME-AME was done to find differential motifs under up peaks between Doxo & Palbo at different timepoints relative to cycling or day 3. 
```{r import_differential_MEMEAME_results}
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL
expressedGenes[grep("TP53$|TP63$|TP73$", expressedGenes)] <- c("P63", "P73", "P53")
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Functions/process_motif_data.RData")

# differential motifs: Doxo vs Palbo
annoUp_DvsP_AME <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_output/annoUp_DoxovsPalbo", expressed_genes = expressedGenes)

# Add column to specify which RNA cluster
annoUp_DvsP_AME$Comparison <- gsub('\\.[0-9]*$', "",  row.names(annoUp_DvsP_AME)) 
annoUp_DvsP_AME$Comparison <- factor(annoUp_DvsP_AME$Comparison, levels=unique(annoUp_DvsP_AME$Comparison)[c(9,1,3,5,7,2,4,6,8)])
# Add column to specify whether up peaks are relative to cycling or day 3 controls. 
annoUp_DvsP_AME$Group <- ifelse(grepl("vsCycling", annoUp_DvsP_AME$Comparison), "vsCycling", "vsday3")

# differential motifs: Palbo vs Doxo
annoUp_PvsD_AME <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_output/annoUp_PalbovsDoxo", expressed_genes = expressedGenes)

# Add column to specify which RNA cluster
annoUp_PvsD_AME$Comparison <- gsub('\\.[0-9]*$', "",  row.names(annoUp_PvsD_AME)) 
annoUp_PvsD_AME$Comparison <- factor(annoUp_PvsD_AME$Comparison, unique(annoUp_DvsP_AME$Comparison)[c(9,1,3,5,7,2,4,6,8)])

# Add column to specify whether up peaks are relative to cycling or day 3 controls. 
annoUp_PvsD_AME$Group <- ifelse(grepl("vsCycling", annoUp_PvsD_AME$Comparison), "vsCycling", "vsday3")
```

-------------------------------------------------------------------------------------------------------------------------
# *FIGURES*

# RNA-seq: plot heatmap of drug distinct genes, clustered based on log2FC from day 3 treatment controls & Cycling controls
```{r}
sig_LFCMat_drugdistinct <- sig_LFCMat[row.names(sig_LFCMat) %in% clusterDF[clusterDF$Clusters %in% c("7", "1", "4", "6"), ]$gene_id & !row.names(sig_LFCMat) %in% c("Y_RNA", "Metazoa_SRP"), ]

pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"

pdf(paste0(pdf_dir, "/RNA_log2FC_drug_distinct_sig_LFCmat.pdf"), width=4.35, height=5)
pheatmap(sig_LFCMat_drugdistinct, show_rownames = F, show_colnames=F, cluster_cols = F, cluster_rows=F, colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100), breaks = seq(-4.5, 4.5,length.out = 100), annotation_row = clusterDF[ ,c("SASP", "Clusters")], annotation_col = anno_sigLFC, annotation_colors = hm_cols_RNA, border_color = NA, main="drug distinct upreg genes")
dev.off()
```

# RNA-seq: plot master heatmap of -log10 pvalues of GO enrichment results for all clusters
```{r}
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"

pdf(paste0(pdf_dir, "/GO_results_list_mat.pdf"), height=20, width=15)
pheatmap(GO_results_list_mat, cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100),breaks = seq(0, 10,length.out = 100), annotation_row = GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, clustering_distance_rows = "correlation", border_color = "black")
dev.off()
```

# RNA-seq: plot padj values of GO enrichment results for top 5 drug distinct up genes
```{r}
# Still showing the pvalues of the other clusters
GO_results_mat_clusters <- list()
for(i in 1:length(GO_results_list)){

GO_results_mat_clusters[[i]] <- GO_results_list_mat[row.names(GO_results_list_mat) %in% GO_results_list[[i]]$ID[1:25], ]
GO_results_mat_clusters[[i]] <- GO_results_mat_clusters[[i]][order(GO_results_mat_clusters[[i]][ ,i], decreasing=T), ]
}

# plot padj values for top 5 GO terms for each cluster individually. Still showing the pvalues of the other clusters as columns but only the top GO terms from the cluster of interest. 
GO_results_mat_clusters <- list()
for(i in 1:length(GO_results_list)){

GO_results_mat_clusters[[i]] <- GO_results_list_mat[row.names(GO_results_list_mat) %in% GO_results_list[[i]]$ID[1:10], ]
GO_results_mat_clusters[[i]] <- GO_results_mat_clusters[[i]][order(GO_results_mat_clusters[[i]][ ,i], decreasing=T), ]
}

pdf(paste0(pdf_dir, "/GO_results_Palbo1_clusters.pdf"), width=8, height=2)
pheatmap(GO_results_mat_clusters[[3]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 20,length.out = 100), cluster_rows = F, main=cluster_names[3], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()

pdf(paste0(pdf_dir, "/GO_results_Palbo2_clusters.pdf"), width=5.5, height=2)
pheatmap(GO_results_mat_clusters[[4]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 10,length.out = 100), cluster_rows = F, main=cluster_names[4], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()

pdf(paste0(pdf_dir, "/GO_results_Doxo1_clusters.pdf"), width=9, height=2)
pheatmap(GO_results_mat_clusters[[5]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 5,length.out = 100), cluster_rows = F, main=cluster_names[5], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()

pdf(paste0(pdf_dir, "/GO_results_Doxo2_clusters.pdf"), width=6.5, height=2)
pheatmap(GO_results_mat_clusters[[6]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 8,length.out = 100), cluster_rows = F, main=cluster_names[6], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()
```

# create function that gets GSEA enrichmentScores that trend opposite between drugs to figure out which enriched terms to plot runningScore plots with. 
```{r get_opposing_GSEA_enrichment}
get_opposing_GSEA_enrichment <- function(Results_enriched=GSEA_Results_H$`Palbo14 vs Cycling0`, Results_opposite=GSEA_Results_H$`Doxo14 vs Cycling0`, names=c("Palbo", "Doxo"), database="Hallmark") {
  # Extract positively enriched terms in drug114 vs Cycling0
  GSEApos <- Results_enriched[
    Results_enriched$enrichmentScore > 0, 
    c("ID", "enrichmentScore", "p.adjust")
  ]
  
 # Find terms that are positively enriched in drug1 but negatively enriched in drug2
GSEAopp <- Results_opposite[
  Results_opposite$ID %in% GSEApos$ID & 
    Results_opposite$enrichmentScore < 0, 
  c("ID", "enrichmentScore", "p.adjust")
]

# Initialize an empty dataframe for opp_df
opp_df <- data.frame()

# Only proceed if GSEAopp has rows
if (nrow(GSEAopp) > 0) {
  opp_df <- data.frame(
    ID = GSEAopp$ID,
    Resultsopp_enrichmentScore = GSEAopp$enrichmentScore,
    Resultsenr_enrichmentScore = GSEApos[match(GSEAopp$ID, GSEApos$ID), ]$enrichmentScore,
    database = database
  )
  
  # Sort by drug1 enrichment score in descending order
  opp_df <- opp_df[order(opp_df$Resultsenr_enrichmentScore, decreasing = TRUE), ]
}
  # Find enriched terms in drug1 that are not enriched in drug2
  enriched_GSEA <- Results_enriched[
    !Results_enriched$ID %in% Results_opposite$ID, 
    c("ID", "enrichmentScore", "p.adjust")
  ]
  
  # Sort by drug1 enrichment score in descending order
  enriched_GSEA <- enriched_GSEA[order(enriched_GSEA$enrichmentScore, decreasing = TRUE), ]
  
  # Create dataframe for terms enriched in drug1 but not found in drug2
  notinopp_df <- data.frame(
    ID = enriched_GSEA$ID,
    Resultsopp_enrichmentScore = 0,
    Resultsenr_enrichmentScore = enriched_GSEA$enrichmentScore,
   database= database
  )
  
  # Merge results
  positiveenr_Results1 <- rbind(opp_df, notinopp_df)
  colnames(positiveenr_Results1) <- c("ID", paste0(names[2], "enrichmentScore"), paste0(names[1], "enrichmentScore"), "database")
  return(positiveenr_Results1 )
}
```

# RNA-seq: find GSEA enrichment Scores that are opposing between drug treatments
```{r}
# find positive enriched terms in day 14 Palbo vs Cycling
Palbo_enriched_GSEA <- rbind(get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_H$`Palbo14 vs Cycling0`, Results_opposite=GSEA_Results_H$`Doxo14 vs Cycling0`, names=c("Palbo", "Doxo"), database="Hallmark"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_KEGG$`Palbo14 vs Cycling0`, Results_opposite=GSEA_Results_KEGG$`Doxo14 vs Cycling0`, names=c("Palbo", "Doxo"), database="KEGG"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_REACT$`Palbo14 vs Cycling0`, Results_opposite=GSEA_Results_REACT$`Doxo14 vs Cycling0`, names=c("Palbo", "Doxo"), database="REACTOME"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_MF$`Palbo14 vs Cycling0`, Results_opposite=GSEA_Results_MF$`Doxo14 vs Cycling0`, names=c("Palbo", "Doxo"), database="GO:MF")
                            )
# find positive enriched terms in day 14 Doxo vs Cycling
Doxo_enriched_GSEA <- rbind(get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_H$`Doxo14 vs Cycling0`, Results_opposite=GSEA_Results_H$`Palbo14 vs Cycling0`, names=c("Doxo", "Palbo"), database="Hallmark"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_KEGG$`Doxo14 vs Cycling0`, Results_opposite=GSEA_Results_KEGG$`Palbo14 vs Cycling0`, names=c("Doxo", "Palbo"), database="KEGG"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_REACT$`Doxo14 vs Cycling0`, Results_opposite=GSEA_Results_REACT$`Palbo14 vs Cycling0`, names=c("Doxo", "Palbo"), database="REACTOME"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_MF$`Doxo14 vs Cycling0`, Results_opposite=GSEA_Results_MF$`Palbo14 vs Cycling0`, names=c("Doxo", "Palbo"), database="GO:MF")
                            )
# Palbo: order by highest to lowest enrichment scores and corresponding Doxo enrichment Score is negative
Palbo_enriched_GSEA <- Palbo_enriched_GSEA[Palbo_enriched_GSEA$DoxoenrichmentScore < 0, ]
Palbo_enriched_GSEA <- Palbo_enriched_GSEA[order(Palbo_enriched_GSEA$PalboenrichmentScore, decreasing=T), ]

# Doxo: order by highest to lowest enrichment scores and corresponding Palbo enrichment Score is negative
Doxo_enriched_GSEA <- Doxo_enriched_GSEA[order(Doxo_enriched_GSEA$DoxoenrichmentScore, decreasing=T) & Doxo_enriched_GSEA$PalboenrichmentScore < 0, ]
```

# RNA-seq: save GSEA running score enriched terms trending in opposite directions between drugs
```{r}
library(enrichplot)
library(patchwork)

# Define the gene sets and corresponding titles
genesetID <- c("ANGIOGENESIS", "EPITHELIAL MESENCHYMAL TRANSITION")
# Get padj value of GSEA results to be plotted
GSEA_padj <- data.frame(ID=GSEA_Results_H$`Palbo14 vs Cycling0`@result[GSEA_Results_H$`Palbo14 vs Cycling0`@result$ID %in% genesetID, ]$ID, 
                        palbo_p.adjust= GSEA_Results_H$`Palbo14 vs Cycling0`@result[GSEA_Results_H$`Palbo14 vs Cycling0`@result$ID %in% genesetID, ]$p.adjust, 
                        doxo_p.adjust= GSEA_Results_H$`Doxo14 vs Cycling0`@result[GSEA_Results_H$`Doxo14 vs Cycling0`@result$ID %in% genesetID, ]$p.adjust)

# Open PDF for saving plots
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
pdf(paste0(pdf_dir, "/GSEA_day14Doxo_day14Palbo_vs_Cyc_Hallmark.pdf"), width=3, height=5)
# Loop through each gene set
for (i in 1:length(genesetID)) {
  
  # Generate GSEA plots for Doxo and Palbo
  p_doxo <- gseaplot(GSEA_Results_H$`Doxo14 vs Cycling0`, geneSetID = genesetID[i], by="runningScore", 
                      title = paste("day 14 Doxo vs Cycling:", genesetID[i]), 
                      color.line="#B3C543", color="#B3C543", color.vline ="#B3C543") +
            theme_classic() + ylim(-0.75, 0.85)
  
  p_palbo <- gseaplot(GSEA_Results_H$`Palbo14 vs Cycling0`, geneSetID = genesetID[i], by="runningScore", 
                      title = paste("day 14 Palbo vs Cycling:", genesetID[i]), 
                      color.line="#E0ADCD", color="#E0ADCD", color.vline ="#E0ADCD") +
            theme_classic() + ylim(-0.75, 0.85)
  
  # Plot the figures
  print(plot_grid(p_doxo, p_palbo, ncol=1))
}

# Close PDF
dev.off()

# Define the gene sets and corresponding titles
genesetID <- c("CROSSLINKING OF COLLAGEN FIBRILS", "ECM PROTEOGLYCANS", "INTEGRIN CELL SURFACE INTERACTIONS", "MEIOTIC RECOMBINATION", "DISEASES OF PROGRAMMED CELL DEATH", "DEPOSITION OF NEW CENPA CONTAINING NUCLEOSOMES AT THE CENTROMERE", "OXIDATIVE STRESS INDUCED SENESCENCE")
titles <- genesetID
# Get padj value of GSEA results to be plotted
GSEA_padj <- data.frame(ID=GSEA_Results_REACT$`Palbo14 vs Cycling0`@result[GSEA_Results_REACT$`Palbo14 vs Cycling0`@result$ID %in% genesetID, ]$ID, 
                        palbo_p.adjust= GSEA_Results_REACT$`Palbo14 vs Cycling0`@result[GSEA_Results_REACT$`Palbo14 vs Cycling0`@result$ID %in% genesetID, ]$p.adjust, 
                        doxo_p.adjust= GSEA_Results_REACT$`Doxo14 vs Cycling0`@result[GSEA_Results_REACT$`Doxo14 vs Cycling0`@result$ID %in% genesetID, ]$p.adjust)

# Open PDF for saving plots
pdf(paste0(pdf_dir, "/GSEA_day14Doxo_day14Palbo_vs_Cyc_REACTOME.pdf"), width=3, height=5)
# Loop through each gene set
for (i in 1:length(genesetID)) {
  
  # Generate GSEA plots for Doxo and Palbo
  p_doxo <- gseaplot(GSEA_Results_REACT$`Doxo14 vs Cycling0`, geneSetID = genesetID[i], by="runningScore", 
                      title = paste("day 14 Doxo vs Cycling:", genesetID[i]), 
                      color.line="#B3C543", color="#B3C543", color.vline ="#B3C543") +
            theme_classic() + ylim(-0.8, 0.85)
  
  p_palbo <- gseaplot(GSEA_Results_REACT$`Palbo14 vs Cycling0`, geneSetID = genesetID[i], by="runningScore", 
                      title = paste("day 14 Palbo vs Cycling:", genesetID[i]), 
                      color.line="#E0ADCD", color="#E0ADCD", color.vline ="#E0ADCD") +
            theme_classic() + ylim(-0.8, 0.85)
  
  # Plot the figures
  print(plot_grid(p_doxo, p_palbo, ncol=1))
}

# Close PDF
dev.off()
```


# RNA-seq: plot padj values for top 5 IPA upstream regs for each cluster individually. 
```{r}
# Still showing the pvalues of the other clusters as columns but only the top IPA upstream regs from the cluster of interest. 
IPA_TF_regs_clusters <- list()
IPA_group_regs_clusters <- list()
for(i in 1:length(IPA_upstream_regs)){
  #transcription regulators
IPA_TF_regs_clusters[[i]] <-IPA_TF_regs_mat[row.names(IPA_TF_regs_mat) %in% IPA_upstream_regs[[i]]$Upstream.Regulator[grepl("transcription", IPA_upstream_regs[[i]]$Molecule.Type)][1:5], ]
IPA_TF_regs_clusters[[i]] <- IPA_TF_regs_clusters[[i]][order(IPA_TF_regs_clusters[[i]][ ,i], decreasing=T), ]
  # complex regulators
IPA_group_regs_clusters[[i]] <-IPA_group_regs_mat[row.names(IPA_group_regs_mat) %in% IPA_upstream_regs[[i]]$Upstream.Regulator[grepl("group", IPA_upstream_regs[[i]]$Molecule.Type)][1:5], ]
IPA_group_regs_clusters[[i]] <- IPA_group_regs_clusters[[i]][order(IPA_group_regs_clusters[[i]][ ,i], decreasing=T), ] # sort by decreasing padj value
}

# make heatmap of top 5 IPA predicted upstream regulators for drug distinct upregulated genes
pdf(file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure Drug distinct responses and SASP/RNA_IPA_TF_regs_drug_distinct_up_genes.pdf", width=3, height=2)
pheatmap(IPA_TF_regs_clusters[[6]], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 100,length.out = 100), cluster_rows = F, main="Shared down genes: predicted TF regulators", border_color = "black",  annotation_col=Cluster_anno, annotation_colors = GO_anno_colors)
dev.off()

pdf(file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure Drug distinct responses and SASP/RNA_IPA_group_regs_SharedDown_genes.pdf", width=3.8, height=2)
pheatmap(IPA_group_regs_clusters[[1]], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 100,length.out = 100), cluster_rows = F, main="Shared down genes: predicted group regulators", border_color = "black", annotation_col=Cluster_anno, annotation_colors = GO_anno_colors)
dev.off()
```

# ATAC-seq, merged RNA clusters: plot motif enrichment results of all clusters as heatmap of p-adj values for each type ("PalboDownvsCyc", "PalboDownvsday3",  "DoxoDownvsCyc","DoxoDownvsday3", "PalboUpvsCyc", "PalboUpvsday3", "DoxoUpvsCyc", "DoxoUpvsday3") of sig changing peaks
```{r}
# load in required functions 
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Functions/create_AME_peaks_matrix.RData")

# get top enriched motifs for Doxo vs Palbo MEME-AME analysis results
AME_padj_mat_DvsP <- create_AME_peaks_matrix(annoUp_DvsP_AME)

filt_AME_padj_mat_DvsP_vsCyc <- AME_padj_mat_DvsP[rowSums(AME_padj_mat_DvsP[ , 1:5] > 0) >= 5, ] # filter by motifs that are enriched in all cycling comparisons. 
filt_AME_padj_mat_DvsP_vsCyc <- filt_AME_padj_mat_DvsP_vsCyc[order(rowSums(filt_AME_padj_mat_DvsP_vsCyc), decreasing=T), ] # order by overall enrichment across comparisons vs cycling. 
filt_AME_padj_mat_DvsP_vsday3 <- AME_padj_mat_DvsP[rowSums(AME_padj_mat_DvsP[ , 6:9] > 0) >= 4, ] # filter by motifs that are enriched in all day3 comparisons. 
filt_AME_padj_mat_DvsP_vsday3 <- filt_AME_padj_mat_DvsP_vsday3[order(rowSums(filt_AME_padj_mat_DvsP_vsday3), decreasing=T), ] # order by overall enrichment across comparisons vs day3. 

# make list of filtered matrices of -log10(padj) to loop through
filt_matrices <- c(
  rep(list(filt_AME_padj_mat_DvsP_vsCyc), 5),
  rep(list(filt_AME_padj_mat_DvsP_vsday3), 4)
)
comparisons=levels(annoUp_DvsP_AME$Comparison)
AME_padj_dots_DvsP <- vector("list", length(comparisons))

#  filter for overall top enriched motifs from dataframe of MEME-AME results for input to make as dotplot. 
for (i in seq_along(comparisons)) {
  AME_padj_dots_DvsP[[i]] <- annoUp_DvsP_AME[annoUp_DvsP_AME$Comparison %in% comparisons[i], ]
  
  max_number <- min(15, nrow(filt_matrices[[i]]))
  
  AME_padj_dots_DvsP[[i]] <- AME_padj_dots_DvsP[[i]][
    AME_padj_dots_DvsP[[i]]$motif_ID %in% row.names(filt_matrices[[i]][1:max_number, ]), 
  ]
  
  AME_padj_dots_DvsP[[i]]$motif_ID <- factor(
    AME_padj_dots_DvsP[[i]]$motif_ID, 
    levels = rev(unique(AME_padj_dots_DvsP[[i]]$motif_ID))
  )
}

AME_padj_dots_DvsP <- do.call(rbind, AME_padj_dots_DvsP)


# get top enriched motifs for Doxo vs Palbo MEME-AME analysis results
AME_padj_mat_PvsD <- create_AME_peaks_matrix(annoUp_PvsD_AME)

filt_AME_padj_mat_PvsD_vsCyc <- AME_padj_mat_PvsD[rowSums(AME_padj_mat_PvsD[, 1:5] > 20) >= 3, ] # filter for motifs that are enriched in at least 4/5 comparisons to cycling
filt_AME_padj_mat_PvsD_vsCyc <- filt_AME_padj_mat_PvsD_vsCyc[order(rowSums(filt_AME_padj_mat_PvsD_vsCyc), decreasing=T), ] # order by overall enrichment across comparisons vs cycling. 
filt_AME_padj_mat_PvsD_vsday3 <- AME_padj_mat_PvsD[rowSums(AME_padj_mat_PvsD[, 6:9] > 20) >= 2, ] # filter for motifs that are enriched in at least 3/4 comparisons to day3
filt_AME_padj_mat_PvsD_vsday3<- filt_AME_padj_mat_PvsD_vsday3[order(rowSums(filt_AME_padj_mat_PvsD_vsday3), decreasing=T), ] # order by overall enrichment across comparisons vs cycling. 

# since the enrichment patterns are different for Palbo and not exclusive to day 21 timepoint looking very different from the rest maybe its best to try to go back to filtering by overall enrichment across both Cycling and day3 instead of treating them separately. 

filt_matrices <- c(
  rep(list(filt_AME_padj_mat_PvsD_vsCyc), 5),
  rep(list(filt_AME_padj_mat_PvsD_vsday3), 4)
)
comparisons=levels(annoUp_DvsP_AME$Comparison)
AME_padj_dots_PvsD <- vector("list", length(comparisons))

#  filter for overall top enriched motifs from dataframe of MEME-AME results for input to make as dotplot. 
for (i in seq_along(comparisons)) {
  AME_padj_dots_PvsD[[i]] <- annoUp_PvsD_AME[annoUp_PvsD_AME$Comparison %in% comparisons[i], ]
  
  max_number <- min(15, nrow(filt_matrices[[i]]))
  
  AME_padj_dots_PvsD[[i]] <- AME_padj_dots_PvsD[[i]][
    AME_padj_dots_PvsD[[i]]$motif_ID %in% row.names(filt_matrices[[i]][1:max_number, ]), 
  ]
  
  AME_padj_dots_PvsD[[i]]$motif_ID <- factor(
    AME_padj_dots_PvsD[[i]]$motif_ID, 
    levels = rev(unique(AME_padj_dots_PvsD[[i]]$motif_ID))
  )
}

AME_padj_dots_PvsD <- do.call(rbind, AME_padj_dots_PvsD)

filt_AME_padj_mat_PvsD <- AME_padj_mat_PvsD[order(rowSums(AME_padj_mat_PvsD), decreasing=T), ] # filter for motifs that are enriched in at least 4/5 comparisons to cycling

# create function for plotting motif enrichment results as dotplot
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=50, size_limit=4.5 ){
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient(name="-log10(padj)", 
                        low="light yellow", high="dark orange",
                        limits=c(0, fill_limit), 
                        oob=scales::squish) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=c(0, size_limit), 
                          range=c(1, 10)) +  # Adjust dot sizes
    ggtitle(Title) +
    theme_classic()+
    theme(
      text = element_text(size=8), 
      axis.text.x = element_text(size=8, angle=90), 
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    ) +
    facet_wrap(~Group, nrow=1, scales="free") +
    labs(x="Comparison", y="motif ID")
}


pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
pdf(paste0(pdf_dir, "/ATAC_PalbovsDoxo_top_differential_motifs_UpPeaks.dotplots.pdf"), width=5, height=10)
create_AME_dotplots(AME_padj_dots_DvsP, Title="Doxo vs Palbo Up peaks", fill_limit = 200, size_limit = 35) 
create_AME_dotplots(AME_padj_dots_PvsD, Title="Palbo vs Doxo Up peaks", fill_limit = 100, size_limit = 25) 
dev.off()


# plot all differentially enriched motifs as heatmap to refer to. 
pdf(paste0(pdf_dir, "/ATAC_PalbovsDoxo_all_differential_motifs_UpPeaks.heatmap.pdf"), width=5, height=20)
pheatmap(AME_padj_mat_DvsP, cluster_cols = F, cluster_rows=T,  colorRampPalette(brewer.pal(n = 7, name ="Reds"))(100), breaks = seq(0, 200 ,length.out = 100), main="-log10(padj)")
pheatmap(AME_padj_mat_PvsD, cluster_cols = F, cluster_rows=T,  colorRampPalette(brewer.pal(n = 7, name ="Reds"))(100), breaks = seq(0, 100 ,length.out = 100), main="-log10(padj)")
dev.off()
```

# RNA-seq: plotting padj values of top IPA predicted upstream regulators of drug distinct gene clusters. 
```{r}
filter_upstream_regulators <- function(mat, threshold_neg1 = -0.5, threshold_pos1 = 0.5,  threshold_neg2 = -0.5, threshold_pos2 = 0.5) {
  # Identify upstream regulators that are differentially regulated between Doxo & Palbo
  filt_negDoxo_posPalbo <- which(rowSums(mat[, 1:5]) <= threshold_neg1 & rowSums(mat[, 6:10]) >= threshold_pos1)
  filt_posDoxo_negPalbo <- which(rowSums(mat[, 1:5]) >= threshold_pos2 & rowSums(mat[, 6:10]) <= threshold_neg2)
  
  # Combine filtered indices
  filt_mat <- mat[c(filt_negDoxo_posPalbo, filt_posDoxo_negPalbo), , drop = FALSE]
  
  # Remove random drugs/chemicals (upstream regulators that contain at least 2 uppercase letters)
  filt_mat <- filt_mat[grep("[A-Z].*[A-Z]", rownames(filt_mat)), , drop = FALSE]
  
  return(filt_mat)
}

# plot most differentially activated upstream regulators of Palbo1 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Palbo1, threshold_neg1 = -8.5, threshold_pos1 = 10,  threshold_neg2 = -6, threshold_pos2 = 8) # set differential thresholds for Palbo vs Doxo. 
filt_mat <- filt_mat[!row.names(filt_mat) %in% c("SB203580", "LY294002", "SB-431542"), ]  # remove upstream regulators that are drugs/chemicals

length(row.names(filt_mat)[row.names(filt_mat) %in% expressedGenes])/length(row.names(filt_mat)) # double check that the upstream regulators are expressed genes
# only"TGF beta" dropped out which is fine because its a family, not a specific gene. 

pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
# plot heatmap of -log10 pvalue of overlap of enriched complex regulators
pdf(paste0(pdf_dir, "/RNA_IPA_Palbo1_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,6:10]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-3, 3,length.out = 100), main="Activation scores")
dev.off() # order by top to bottom sum of scores. 

# plot most differentially activated upstream regulators of Palbo2 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Palbo2, threshold_neg1 = 1.2, threshold_pos1 = 6.2, threshold_neg2 = -0.5, threshold_pos2 =0.5)

length(row.names(filt_mat)[row.names(filt_mat) %in% expressedGenes])/length(row.names(filt_mat)) # double check that the upstream regulators are expressed genes

# finding genes that did not appear in expressed genes using their ENSG identifier to double check it really isn't being expressed. 
# normCounts[grepl("ENSG00000112116", row.names(normCounts)), ]

filt_mat <- filt_mat[grep("IL5|IL17F|VH032|IOX2", rownames(filt_mat), invert = T), , drop = FALSE]
nrow(filt_mat) # filter out chemicals and non-expressed genes 

# only"TGF beta" dropped out which is fine because its a family, not a specific gene. 

pdf(paste0(pdf_dir, "/RNA_IPA_Palbo2_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,6:10]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")
dev.off()

# plot most differentially activated upstream regulators of Doxo1 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Doxo1, threshold_neg1 = -4.2, threshold_pos1 = 4.2, threshold_neg2 = -4.2, threshold_pos2 =4.2)

length(row.names(filt_mat)[row.names(filt_mat) %in% expressedGenes])/length(row.names(filt_mat)) # double check that the upstream regulators are expressed genes
row.names(filt_mat)[!row.names(filt_mat) %in% expressedGenes]
filt_mat <- filt_mat[grep("GW9662|PG3-Oc", rownames(filt_mat), invert = T), , drop = FALSE]  # filter out chemicals and non-expressed genes. AURK and p38 MAPK are encoded by AURKA, MAPK14 respectively 

pdf(paste0(pdf_dir, "/RNA_IPA_Doxo1_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,1:5]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")
dev.off()

# plot most differentially activated upstream regulators of Doxo2 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Doxo2, threshold_neg1 = -6.5, threshold_pos1 = 0,  threshold_neg2 = -2, threshold_pos2 = 0.5)

length(row.names(filt_mat)[row.names(filt_mat) %in% expressedGenes])/length(row.names(filt_mat)) # double check that the upstream regulators are expressed genes
row.names(filt_mat)[!row.names(filt_mat) %in% expressedGenes]
normCounts[grep("ENSG00000182968|ENSG00000134595|ENSG00000222028", row.names(normCounts)), ]
filt_mat <- filt_mat[grep("GW9662|PG3-Oc|PSMB11|SOX1|SOX3", rownames(filt_mat), invert = T), , drop = FALSE]  # filter out chemicals and non-expressed genes. AURK and p38 MAPK are encoded by AURKA, MAPK14 respectively 

pdf(paste0(pdf_dir, "/RNA_IPA_Doxo2_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,1:5]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")
dev.off()

```

# plot differential upstream regs based on thresholding of the largest differences between overall sum of activation scores between Doxo & Palbo instead of thresholding. 
```{r}
mat <- IPA_Compare_upstream_regs_mat$Palbo2
diff_regs <- names(rev(sort(rowSums(mat[ ,6:10]) - rowSums(mat[ ,1:5]))))[1:20]
diff_regs <- diff_regs[!diff_regs %in% c("arsenic trioxide", "cadmium chloride", "VH032", "IOX2")]

#pdf(paste0(pdf_dir, "/RNA_IPA_Palbo2_diff_regs_bydifference.pdf"), width=3, height=3)
pheatmap(mat[row.names(mat) %in% diff_regs[1:12], ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")

mat <- IPA_Compare_upstream_regs_mat$Palbo1
diff_regs <- names(rev(sort(rowSums(mat[ ,6:10]) - rowSums(mat[ ,1:5]))))[1:20]
diff_regs <- diff_regs[!diff_regs %in% c("arsenic trioxide", "cadmium chloride", "VH032", "IOX2")]
```


# run code necessary for matching top activated/inhibited transcriptional regulators with ATAC-seq motif enrichment results from up peaks across time. 
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="create_AME_peaks_matrix")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="create_AME_dotplots")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="annoUpPeaks_AME_df")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="annoDownPeaks_AME_df")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="UpPeaks_AME_mat")

row.names(Doxo_UpPeaks_AME_mat)[grep("P53|P63|P73", row.names(Doxo_UpPeaks_AME_mat))] <- c("TP53", "TP73", "TP63")
```

# RNA-seq + ATAC-seq: instead of differentially activated upstream regs between drug treatments, plot top activated & top inhibited upstream Transcriptional regulators 
```{r}
# read in database matching upstream regulators to molecule type and target molecules. 
IPA_TFs_db <- list() 
db_files <- dir("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/output/sigGenes_log2FC/databases", pattern="transcription_regulators_database.txt", full.names = T)
for(i in 1:length(db_files)){
IPA_TFs_db[[i]] <- read_delim(db_files[i], 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE, skip = 2)
}
names(IPA_TFs_db) <- gsub("_.*", "", dir("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/output/sigGenes_log2FC/databases", pattern="transcription_regulators_database.txt"))

# function to filter for top upstream regulators and plot activation scores as heatmap
plot_upstream_reg_heatmaps <- function(mat, cluster_name, tf_db, zscore_cols, atac_peaks, pdf_dir) {
  # Ensure the output directory exists
  if (!dir.exists(pdf_dir)) {
    dir.create(pdf_dir, recursive = TRUE)
  }

  # Order the matrix by highest to lowest overall activation z-scores
  ordered_mat <- mat[order(rowSums(zscore_cols), decreasing = TRUE), ]

  # Filter by transcriptional regulators
  filt_mat <- ordered_mat[row.names(ordered_mat) %in% tf_db, ]

  # Define filtering conditions
  filter_conditions <- list(
    list(
      name = paste0("RNA_IPA_",cluster_name, "_top_act_inh_upstream_regs"),
      data = filt_mat
    ),
    list(
      name = paste0("RNA_IPA_" ,cluster_name, "_top_act_inh_filt_by_ATAC_upstream_regs"),
      data = filt_mat[row.names(filt_mat) %in% row.names(atac_peaks), ]
    )
  )

  # Loop through each filtering condition and generate heatmaps
  for (condition in filter_conditions) {
    filt_data <- condition$data

    # Get top activated & inhibited upstream transcriptional regulators
    top_regs <- c(row.names(head(filt_data)), row.names(tail(filt_data)))

    # Define the output PDF filename
    pdf_filename <- paste0(pdf_dir, "/", condition$name, ".pdf")

    # Generate heatmap and save to PDF
    pdf(pdf_filename, width = 3, height = 3)
    pheatmap(
      filt_data[row.names(filt_data) %in% top_regs, ],
      cluster_cols = FALSE,
      cluster_rows = FALSE,
      colorRampPalette(rev(brewer.pal(n = 7, name = "PuOr")))(100),
      breaks = seq(-4, 4, length.out = 100),
      main = "Activation scores"
    )
    dev.off()
  }
}
```

# RNA-seq + ATAC-seq: generate plots of top activated and inhibited upstream transcriptional regulators for each cluster of sig changing upregulated genes. 
```{r}
plot_upstream_reg_heatmaps(
  mat = IPA_Compare_upstream_regs_mat$Doxo1, 
  cluster_name = "Doxo1", 
  tf_db = IPA_TFs_db$Doxo1$`Upstream Regulator`, 
  zscore_cols = IPA_Compare_upstream_regs_mat$Doxo1[, 1:5], 
  atac_peaks = Doxo_UpPeaks_AME_mat, 
  pdf_dir = "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
)

plot_upstream_reg_heatmaps(
  mat = IPA_Compare_upstream_regs_mat$Doxo2, 
  cluster_name = "Doxo2", 
  tf_db = IPA_TFs_db$Doxo2$`Upstream Regulator`, 
  zscore_cols = IPA_Compare_upstream_regs_mat$Doxo2[, 1:5], 
  atac_peaks = Doxo_UpPeaks_AME_mat, 
  pdf_dir = "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
)

plot_upstream_reg_heatmaps(
  mat = IPA_Compare_upstream_regs_mat$Palbo1, 
  cluster_name = "Palbo1", 
  tf_db = IPA_TFs_db$Palbo1$`Upstream Regulator`, 
  zscore_cols = IPA_Compare_upstream_regs_mat$Palbo1[, 6:10], 
  atac_peaks = Palbo_UpPeaks_AME_mat, 
  pdf_dir = "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
)

plot_upstream_reg_heatmaps(
  mat = IPA_Compare_upstream_regs_mat$Palbo2, 
  cluster_name = "Palbo2", 
  tf_db = IPA_TFs_db$Palbo2$`Upstream Regulator`, 
  zscore_cols = IPA_Compare_upstream_regs_mat$Palbo2[, 6:10], 
  atac_peaks = Palbo_UpPeaks_AME_mat, 
  pdf_dir = "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
)

plot_upstream_reg_heatmaps(
  mat = IPA_Compare_upstream_regs_mat$Shared, 
  cluster_name = "Shared", 
  tf_db = IPA_TFs_db$Shared$`Upstream Regulator`, 
  zscore_cols = IPA_Compare_upstream_regs_mat$Shared[, 1:10], 
  atac_peaks = Palbo_UpPeaks_AME_mat, 
  pdf_dir = "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
)
```


# plot ESR1 expression for Justin (not related to this paper)
```{r}
normCounts <- counts(RNAddsall, normalized=T)
normCounts_ESR1 <- log2(normCounts[grep("ENSG00000091831", row.names(normCounts)), ]+1)

NormCounts_df_ESR1 <- data.frame(Counts=normCounts_ESR1, metaDataall)
ggplot(NormCounts_df_ESR1, aes(x=as.numeric(Timepoint), y=Counts, color=Treatment, shape=Biorep))+geom_point(size=3)+scale_color_manual(values=hm_cols_RNA$Treatment)+theme_minimal()+ylab("log2 Normalized Counts")+xlab("Length of Treatment (days)")+ggtitle("ESR1")

violinplot(log2(normCounts+1), las = 2, col = "lightblue", main = "Expression Levels", 
        ylab = "log2 normalized counts")

```

