---
title: "Figure Drug distinct responses and SASP"
author: "Joanna Yeung"
date: '2025-01-27'
output: html_document
---

```{r}
# data wrangling
library(preprocessCore)
library(readr)
library(readxl)
library(stringr)
# plotting
library(gridExtra)
library(ggplot2)
library(pheatmap)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(scales)
library(ggrepel)

# genomics data wrangling and analysis
library(rtracklayer)
library(Biostrings)
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(DESeq2)
library(apeglm)
library(ChIPseeker)
library(clusterProfiler)
library(msigdbr)
library(MotifDb)
library(motifmatchr)

# genome annotation databases
library(EnsDb.Hsapiens.v79)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(BSgenome.Hsapiens.UCSC.hg38)
library(org.Hs.eg.db)
```

# annotation colors for pheatmap
```{r}
hm_cols_RNA <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("Cycling"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "DoxovsPalbo"="#CCEBC5"), 
                Biorep=c("1"="orange", "2"="light blue", "3"="pink"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "DoxovsPalbo"="#CCEBC5"), 
                Clusters=c("2"="#66C2A5","5"="#A6D854","1"="#FC8D62", "7"="#8DA0CB","4"="#E78AC3","6"="#FFD92F","3"="#E5C494"), 
                SASP = c("Yes"="dark green", "No"="white"),
                Cell_Cycle = c("Yes"="dark blue", "No"="white"), 
                clusters_cutree8=brewer.pal(8, "Set2"), 
                p53_All=c("Yes"="#00B0F0", "No"="white"),
                NFKB_All=c("Yes"="red", "No"="white"))
names(hm_cols_RNA$clusters_cutree8) <- as.factor(1:8)


hm_cols_ATAC <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("None"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "PvsD"="#CCEBC5"),
                Biorep=c("1"="orange", "2"="light blue", "3"="pink", "4"="light green"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "PvsD"="#CCEBC5"),
                Peak_anno=c("Distal"="#B15A28", "Gene Body"="#7F54A2", "Promoter"="#A7CEE2"), 
                Doxo=c("Yes"="#B3C543", "No"="white"),
                Palbo=c("Yes"="#E0ADCD", "No"="white")
                )
```

# load in workspace with DESeq results and rlog counts matrix of genes. 
```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure Global transcriptome and chromatin accessibility is dynamic in both types of therapy induced senescence/workspace.RData")
```

#### universe (background gene set) for GO enrichment analysis is all genes expressed
```{r}
universe <- genes_anno_type[match(gsub('\\.[0-9]*$', '', row.names(acrossGroups)), genes_anno_type$ensembl_id), ]
```

# GO enrichment analysis on k-means clusters from sigMat (significantly changing genes across all conditions, based on Likelihood ratio test)

## *function to get GO enriched terms*:
## arguments:
### gene= vector of genes to test for GO enrichment (gene ensembl format)
### universe=background gene set 
### Database=database containing groups of genes to GO annotation term e.g. msigdb GO:BP
```{r getGOenrichment function}
getGOenrichment <-function(gene,universe,Database){
  #gene_entrezID <-  AnnotationDbi::select(EnsDb.Hsapiens.v79, gene[!is.na(gene)], "ENTREZID", "SYMBOL") #convert gene IDs from SYMBOL format to ENTREZ ID format (was using v79, new code is using v107)
  gene_entrezID <- genes_anno_type[match(gene, genes_anno_type$gene_id), ]
  gene_go <- enricher(gene_entrezID$entrezid, TERM2GENE=Database, universe=universe) #enrichment analysis with hypergeometric test
  gene_go <- setReadable(gene_go, OrgDb = org.Hs.eg.db, keyType="ENTREZID") #convert enrichment results from ENTREZ ID format to SYMBOL for easier interpretation
return(gene_go)
} 
```

# assign significant genes to relevant k-means clusters & drug + cell cycle arrest categories. 
```{r genes_in_clusters}
genes_in_clusters <- lapply(levels(clusterDF$Clusters), function(x){clusterDF[clusterDF$Clusters %in% x, ]$gene_id})
cluster_names <- c("CellCycle1", "CellCycle2", "Palbo1", "Palbo2", "Doxo1", "Doxo2", "Shared") 
names(genes_in_clusters) <- cluster_names
```

# RNA-seq: GO enrichment on significant genes
```{r RNA-seq GO enrichment analysis}
# GO enrichment analysis with msigdb GO:BP database:
GOBP_enrichment_clusters <- list()
# GO enrichment analysis with msigdb REACT database:
GO_enrichment_clusters_KEGG <- list()
# GO enrichment analysis with msigdb Hallmark database:
GO_enrichment_clusters_HallMark <- list()
# GO enrichment analysis with msigdb REACT database:
GO_enrichment_clusters_REACT <- list()
# GO enrichment analysis with msigdb GO:MF database:
GO_enrichment_clusters_MF <- list()

for(i in 1:length(genes_in_clusters)){
  
GOBP_enrichment_clusters[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_GO_BP) # do GO enrichment analysis with background gene set set as all gene expressed, database used is msigdb GO:BP

GO_enrichment_clusters_KEGG[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_KEGG)

GO_enrichment_clusters_HallMark[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_H)

GO_enrichment_clusters_REACT[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_REACT)

GO_enrichment_clusters_MF[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_GO_MF)
}
```

### RNA-seq: format GO enrichment terms from each cluster into matrix of -log10(padj values)
```{r RNAseq GO enrichment results}
# convert each clusters' GO enrichment terms into dataframes
GO_enrichment_clusters_KEGG_df <- list()
 GO_enrichment_clusters_REACT_df <- list()
 GO_enrichment_clusters_HallMark_df <- list()
 GOBP_enrichment_clusters_df <- list()
GO_enrichment_clusters_MF_df <- list()
for(i in 1:length(GO_enrichment_clusters_REACT)){
  GO_enrichment_clusters_KEGG_df[[i]] <- as.data.frame(GO_enrichment_clusters_KEGG[[i]])
   GO_enrichment_clusters_REACT_df[[i]] <- as.data.frame(GO_enrichment_clusters_REACT[[i]])
 GO_enrichment_clusters_HallMark_df[[i]] <- as.data.frame(GO_enrichment_clusters_HallMark[[i]])
 GOBP_enrichment_clusters_df[[i]] <- as.data.frame(GOBP_enrichment_clusters[[i]])
  GO_enrichment_clusters_MF_df[[i]] <- as.data.frame(GO_enrichment_clusters_MF[[i]])
}

 # indicate names of each list as the cluster number
 names(GO_enrichment_clusters_KEGG_df) <- cluster_names 
 names(GO_enrichment_clusters_REACT_df) <- cluster_names 
 names(GO_enrichment_clusters_HallMark_df) <- cluster_names 
 names(GOBP_enrichment_clusters_df) <- cluster_names 
names(GO_enrichment_clusters_MF_df) <- cluster_names 
 
 # remove any cluster of genes that have 0 GO enrichment terms
GO_enrichment_clusters_KEGG_df <-  GO_enrichment_clusters_KEGG_df[sapply( GO_enrichment_clusters_KEGG_df, nrow)>0]
GO_enrichment_clusters_REACT_df <-  GO_enrichment_clusters_REACT_df[sapply( GO_enrichment_clusters_REACT_df, nrow)>0]
GO_enrichment_clusters_HallMark_df <-  GO_enrichment_clusters_HallMark_df[sapply( GO_enrichment_clusters_HallMark_df, nrow)>0]
GOBP_enrichment_clusters_df <-  GOBP_enrichment_clusters_df[sapply( GOBP_enrichment_clusters_df, nrow)>0]
GO_enrichment_clusters_MF_df <-  GO_enrichment_clusters_MF_df[sapply( GO_enrichment_clusters_MF_df, nrow)>0]
 
 # indicate which cluster and database each dataframe comes from 
 for(i in 1:length(GO_enrichment_clusters_KEGG_df)){
  GO_enrichment_clusters_KEGG_df[[i]]$Cluster <- names(GO_enrichment_clusters_KEGG_df)[i]
  GO_enrichment_clusters_KEGG_df[[i]]$Database <- "KEGG"
 }
  for(i in 1:length(GO_enrichment_clusters_REACT_df)){
   GO_enrichment_clusters_REACT_df[[i]]$Cluster <-  names(GO_enrichment_clusters_REACT_df)[i]
   GO_enrichment_clusters_REACT_df[[i]]$Database <- "REACTOME"
  }
  for(i in 1:length(GO_enrichment_clusters_HallMark_df)){
 GO_enrichment_clusters_HallMark_df[[i]]$Cluster <- names(GO_enrichment_clusters_HallMark_df)[i]
 GO_enrichment_clusters_HallMark_df[[i]]$Database <- "Hallmark"
  }
 
  for(i in 1:length(GOBP_enrichment_clusters_df)){
 GOBP_enrichment_clusters_df[[i]]$Cluster <- names(GOBP_enrichment_clusters_df)[i]
  GOBP_enrichment_clusters_df[[i]]$Database <- "GO:BP"
  }

  for(i in 1:length(GO_enrichment_clusters_MF_df)){
 GO_enrichment_clusters_MF_df[[i]]$Cluster <- names(GO_enrichment_clusters_MF_df)[i]
 GO_enrichment_clusters_MF_df[[i]]$Database <- "GO:MF"
  }

# combine GO enrichment results from different databases into 1 dataframe for each cluster of genes. 
GO_results_list <- vector()
GO_results_list$CellCycle1 <- rbind(GO_enrichment_clusters_KEGG_df$CellCycle1,
                                    GO_enrichment_clusters_HallMark_df$CellCycle1,
                                    GO_enrichment_clusters_REACT_df$CellCycle1,
                                    GOBP_enrichment_clusters_df$CellCycle1,
                                    GO_enrichment_clusters_MF_df$CellCycle1)

GO_results_list$CellCycle2 <- rbind(GO_enrichment_clusters_KEGG_df$CellCycle2,
                                    GO_enrichment_clusters_HallMark_df$CellCycle2,
                                    GO_enrichment_clusters_REACT_df$CellCycle2,
                                    GOBP_enrichment_clusters_df$CellCycle2,
                                    GO_enrichment_clusters_MF_df$CellCycle2)

GO_results_list$Palbo1 <- rbind(GO_enrichment_clusters_KEGG_df$Palbo1,     
                                GO_enrichment_clusters_HallMark_df$Palbo1 , 
                                GO_enrichment_clusters_REACT_df$Palbo1, 
                                GOBP_enrichment_clusters_df$Palbo1, 
                                GO_enrichment_clusters_MF_df$Palbo1)

GO_results_list$Palbo2 <- rbind(GO_enrichment_clusters_KEGG_df$Palbo2,
                                GO_enrichment_clusters_HallMark_df$Palbo2,
                                GO_enrichment_clusters_REACT_df$Palbo2,
                                GOBP_enrichment_clusters_df$Palbo2, 
                                GO_enrichment_clusters_MF_df$Palbo2)

GO_results_list$Doxo1 <- rbind(GO_enrichment_clusters_KEGG_df$Doxo1,
                               GO_enrichment_clusters_HallMark_df$Doxo1,
                               GO_enrichment_clusters_REACT_df$Doxo1,
                               GOBP_enrichment_clusters_df$Doxo1, 
                               GO_enrichment_clusters_MF_df$Doxo1)

GO_results_list$Doxo2 <- rbind(GO_enrichment_clusters_KEGG_df$Doxo2,
                               GO_enrichment_clusters_HallMark_df$Doxo2,
                               GO_enrichment_clusters_REACT_df$Doxo2,
                               GOBP_enrichment_clusters_df$Doxo2, 
                               GO_enrichment_clusters_MF_df$Doxo2)

GO_results_list$Shared <- rbind(GO_enrichment_clusters_KEGG_df$Shared,
                                GO_enrichment_clusters_HallMark_df$Shared,
                                GO_enrichment_clusters_REACT_df$Shared,
                                GOBP_enrichment_clusters_df$Shared, 
                                GO_enrichment_clusters_MF_df$Shared)

# order GO enrichment results from different databases by padj values
for(i in 1:length(GO_results_list)){
  GO_results_list[[i]] <- GO_results_list[[i]][order(GO_results_list[[i]]$p.adjust), ]
}

# combine dataframes of each cluster of gene's GO enrichment results into 1 master dataframe
GO_results_df <- do.call(rbind, GO_results_list)

# create matrix of -log10pvalues of the top 15 enriched pathways for each cluster

GO_results_list_mat <- matrix(nrow=length(unique(unlist(lapply(GO_results_list, function(GO_results_list){GO_results_list$ID[1:50]})))), ncol=length(GO_results_list)) # make empty matrix, number of columns are the clusters of genes and number of total top GO terms of all clusters

row.names(GO_results_list_mat) <- unique(unlist(lapply(GO_results_list, function(GO_results_list){GO_results_list$ID[1:50]}))) # row names are top 12 GO enrichment terms for each cluster combined
colnames(GO_results_list_mat) <- cluster_names  # column names are clusters

# pathways that are not enriched in the cluster are listed as NA so just change the value to 0 so that it can be plotted in heatmap
for(i in 1:ncol(GO_results_list_mat)){
GO_results_list_mat[ ,i] <- -log10(GO_results_list[[i]][match(row.names(GO_results_list_mat), GO_results_list[[i]]$ID), ]$p.adjust)
GO_results_list_mat[ ,i] <- ifelse(GO_results_list_mat[ ,i] %in% NA, 0, GO_results_list_mat[ ,i])
}
GO_results_list_mat <- GO_results_list_mat[!is.na(row.names(GO_results_list_mat)), ] # remove padj values that are NA 

# specify GO results matrix annotation colors: based on ggsci:pal_aaas palette. 
GO_anno_colors <-  list(database=c("GO:BP"="#3B4992FF", "Hallmark"="#EE0000FF", "KEGG"="#008B45FF", "REACTOME" ="#631879FF", "GO:MF"="#808180FF"), 
                        Cluster=c("CellCycle1"="#66C2A5", "CellCycle2"="#A6D854", "Palbo1"="#FC8D62", "Palbo2"="#8DA0CB","Doxo1"="#E78AC3","Doxo2"="#FFD92F", "Shared"="#E5C494"))
names(GO_anno_colors$Cluster) <- cluster_names 

# annotation dataframe for rows of heatmap to indicate which database GO term came from 
GO_anno <- data.frame(row.names=row.names(GO_results_list_mat), database=GO_results_df[match(row.names(GO_results_list_mat), GO_results_df$ID), ]$Database)
# annotation dataframe for columns of heatmap to color by Cluster colors. 
Cluster_anno <- data.frame(Cluster=cluster_names , row.names = colnames(GO_results_list_mat))
```

### MEME-AME analysis on DESeq_Peaks pairwise comparison 
```{bash}
cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq

# all up peaks- background peak set set to corresponding down peaks of the same DESeq comparison. 
sbatch -p risc,hpc --array=1-16 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoDown/inputfiles_annoDown_DESeq_peaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoUp/inputfiles_annoUp_DESeq_peaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_output/DESeq/annoUp,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoUp/folderlist_annoUp_DESeq_peaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm

# all down peaks- background peak set set to corresponding up peaks of the same DESeq comparison. 
sbatch -p risc,hpc --array=1-16 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoUp/inputfiles_annoUp_DESeq_peaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoDown/inputfiles_annoDown_DESeq_peaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_output/DESeq/annoDown,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoDown/folderlist_annoDown_DESeq_peaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm
```

# RNA-seq: GSEA Analysis
```{r RNAseq GSEA analysis}
ranked_res_forGSEA <- list()
for(i in 1:length(LFCResultsall_RNA)){
# Filter for significant genes (optional)
  
rank <- LFCResultsall_RNA[[i]][!is.na(LFCResultsall_RNA[[i]]$padj), ]

# Create a ranked list based on a metric (e.g., log2 fold change or -log10(p-value))
# Here, we rank by the sign of log2FoldChange * -log10(p-value)
rank$rank_metric <- sign(rank$log2FoldChange) * -log10(rank$pvalue)

rank <- rank[order(rank$rank_metric, decreasing = TRUE), ]

# Sort the LFCResultsall_RNA results by the ranking metric
ranked_res_forGSEA[[i]] <- rank$rank_metric
names(ranked_res_forGSEA[[i]]) <- row.names(rank)

names(ranked_res_forGSEA[[i]]) <- ifelse(genes_anno_type[match(names(ranked_res_forGSEA[[i]]), genes_anno_type$ensembl_id), ]$entrezid %in% NA, names(ranked_res_forGSEA[[i]]), genes_anno_type[match(names(ranked_res_forGSEA[[i]]), genes_anno_type$ensembl_id), ]$entrezid)
} # convert to entrez id for input into GSEA function. 
names(ranked_res_forGSEA) <- names(LFCResultsall_RNA)

# run GSEA analysis using: 
GSEA_Results_H <- list() # msigdb Hallmark database
GSEA_Results_KEGG <- list() # msigdb KEGG database
GSEA_Results_REACT <- list() # msigdb Reactome database
GSEA_plots_H <- list()
GSEA_plots_KEGG <- list()
GSEA_plots_REACT <- list()
for(i in 1:length(ranked_res_forGSEA)){
 GSEA_Results_H[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_H) 
 GSEA_Results_KEGG[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_KEGG) 
 GSEA_Results_REACT[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_REACT) 
}

names(GSEA_Results_H) <- names(LFCResultsall_RNA)
names(GSEA_Results_KEGG) <- names(LFCResultsall_RNA)
names(GSEA_Results_REACT) <- names(LFCResultsall_RNA)

# generate running score plots for GSEA results
# GSEA Hallmark
for(i in 1:length(GSEA_Results_H)){
  maxplots <- length(GSEA_Results_H[[i]]$ID)
GSEA_plots_H[[i]] <- list()
GSEA_plots_KEGG[[i]] <- list()
GSEA_plots_REACT[[i]] <- list()
  for(j in 1:maxplots){
  GSEA_plots_H[[i]][[j]] <- gseaplot(GSEA_Results_H[[i]], geneSetID = j, by = "runningScore", title = paste0(names(GSEA_Results_H)[i], ": " , GSEA_Results_H[[i]]$ID[j]) )
  }
  names(GSEA_plots_H[[i]]) <- GSEA_Results_H[[i]]$ID[1:maxplots]
}

# GSEA KEGG
for (i in 1:length(GSEA_Results_KEGG)) {
  # Check if the current object is not empty
  if (!is.null(GSEA_Results_KEGG[[i]]) && length(GSEA_Results_KEGG[[i]]$ID) > 0) {
    maxplots <- length(GSEA_Results_KEGG[[i]]$ID)
    GSEA_plots_KEGG[[i]] <- list()  # Initialize a list to store plots for this object
    
    for (j in 1:maxplots) {
      GSEA_plots_KEGG[[i]][[j]] <- gseaplot(
        GSEA_Results_KEGG[[i]], 
        geneSetID = j, 
        by = "runningScore", 
        title = GSEA_Results_KEGG[[i]]$ID[j]
      )
    }
    # Assign names to the plots
    names(GSEA_plots_KEGG[[i]]) <- GSEA_Results_KEGG[[i]]$ID[1:maxplots]
  } else {
    # If empty, set the current slot in GSEA_plots_KEGG to NULL
    GSEA_plots_KEGG[[i]] <- NULL
  }
}

# GSEA Reactome
for (i in 1:length(GSEA_Results_REACT)) {
  # Check if the current object is not empty
  if (!is.null(GSEA_Results_REACT[[i]]) && length(GSEA_Results_REACT[[i]]$ID) > 0) {
    maxplots <- length(GSEA_Results_REACT[[i]]$ID)
    GSEA_plots_REACT[[i]] <- list()  # Initialize a list to store plots for this object
    
    for (j in 1:maxplots) {
      GSEA_plots_REACT[[i]][[j]] <- gseaplot(
        GSEA_Results_REACT[[i]], 
        geneSetID = j, 
        by = "runningScore", 
        title = GSEA_Results_REACT[[i]]$ID[j]
      )
    }
    # Assign names to the plots
    names(GSEA_plots_REACT[[i]]) <- GSEA_Results_REACT[[i]]$ID[1:maxplots]
  } else {
    # If empty, set the current slot in GSEA_plots_REACT to NULL
    GSEA_plots_REACT[[i]] <- NULL
  }
}

# assign names to plots based on which DESeq comparison GSEA was done on. 
names(GSEA_plots_H) <- names(LFCResultsall_RNA)
names(GSEA_plots_KEGG) <- names(which(unlist(lapply(GSEA_Results_KEGG, function(x){length(x$ID)})) >0))
names(GSEA_plots_REACT) <- names(which(unlist(lapply(GSEA_Results_REACT, function(x){length(x$ID)})) >0))
```

# RNA-seq: generate input files for IPA pathways. 
```{r}
# clusters of genes 
for(i in 1:length(genes_in_clusters)){
  
  IPA_input <- as.data.frame(genes_in_clusters[[i]])
  colnames(IPA_input) <- "ID"
  write.table(IPA_input, file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/input/", cluster_names[i], ".csv"), row.names=F, sep=",")
}
# DESeq Wald Test results for pairwise comparisons. 
for(i in 1:length(LFCResultsall_RNA)){
  write.table(as.data.frame(LFCResultsall_RNA[[i]]), file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files/DESeq_results_input/", names(LFCResultsall_RNA)[i], ".csv"), sep=",")
}

# log2FC values of each sig changing gene clusters. 
for(i in 1:length(drug_cat_list)){
table <- sig_LFCMat[row.names(sig_LFCMat) %in% row.names(drug_cat_list[[i]]), ]
    write.table(table, file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/input/sigGenes_log2FC/", cluster_names[i], "_log2FC.csv"), sep=",", row.names = T)
}
```

# RNA-seq: import IPA upstream regulators & pathways
```{r import IPA results}
setwd("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/output")
IPA_upstream_regs <- list()
IPA_pathways <- list()
# read in information for enriched IPA pathways per hierarchical cluster (clustering based on scaled rlog normalized counts)
for(i in 1:length(cluster_names)){
IPA_pathways[[i]]<- read_tsv(paste0(cluster_names[i], "_IPA_pathways.txt"), skip=2)
# read in information for enriched upstream regulators 
IPA_upstream_regs[[i]] <- read.table(paste0(cluster_names[i], "_IPA_upstream_regs.txt"), sep="\t", skip=2, header=T)
IPA_upstream_regs[[i]] <- IPA_upstream_regs[[i]][!grepl("mir|miR|Mir", IPA_upstream_regs[[i]]$Upstream.Regulator), ] # remove upstream regulators that are miRNAs (uncertain if they will be informative & there is a lot of them)
}

names(IPA_upstream_regs) <- cluster_names
names(IPA_pathways) <- cluster_names

# filter out upstream regulators that are not expressed in this cell line. 

#extract ids of all expressed genes
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL

# keep only upstream regulators that are part of expressed genes or have lower case letters (these are upstream regulators that could be miRNAs or protein complexes)
#IPA_upstream_regs <- lapply(IPA_upstream_regs, function(IPA_upstream_regs){IPA_upstream_regs[c(grep('[:lower:]', IPA_upstream_regs$Upstream.Regulator), which(IPA_upstream_regs$Upstream.Regulator %in% expressedGenes)), ]})

```

# RNA-seq: plotting padj values of top IPA predicted upstream regulators of all clusters
```{r IPA results matrix}
create_ipa_regs_matrix <- function(IPA_upstream_regs, cluster_names, filter_keyword = "complex", top_n = 25) {
  # Extract unique upstream regulators based on the filter keyword
  unique_upstream_regs <- unique(unlist(lapply(IPA_upstream_regs, function(x) {
    x$Upstream.Regulator[grepl(filter_keyword, x$Molecule.Type)][1:top_n]
  })))

  # Create a matrix to store -log10 p-values
  IPA_complex_regs_mat <- matrix(
    nrow = length(unique_upstream_regs),
    ncol = length(IPA_upstream_regs)
  )
  row.names(IPA_complex_regs_mat) <- unique_upstream_regs # Set row names to upstream regulators
  colnames(IPA_complex_regs_mat) <- cluster_names        # Set column names to cluster names

  # Fill in the matrix with -log10 p-values
  for (i in seq_len(ncol(IPA_complex_regs_mat))) {
    # Match and extract p-values for each cluster
    IPA_complex_regs_mat[, i] <- -log10(IPA_upstream_regs[[i]][
      match(row.names(IPA_complex_regs_mat), IPA_upstream_regs[[i]]$Upstream.Regulator),
      "p.value.of.overlap"
    ])

    # Replace NA values with 0 for heatmap compatibility
    IPA_complex_regs_mat[, i][is.na(IPA_complex_regs_mat[, i])] <- 0
  }

  return(IPA_complex_regs_mat)
}

IPA_group_regs_mat <- create_ipa_regs_matrix(
  IPA_upstream_regs = IPA_upstream_regs,
  cluster_names = cluster_names,
  filter_keyword = "group",
  top_n = 25
)

IPA_TF_regs_mat <- create_ipa_regs_matrix(
  IPA_upstream_regs = IPA_upstream_regs,
  cluster_names = cluster_names,
  filter_keyword = "transcription",
  top_n = 25
)

# plot heatmap of -log10 pvalue of overlap of enriched complex regulators
pheatmap(IPA_group_regs_mat, cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 10,length.out = 100))

pheatmap(IPA_TF_regs_mat, cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 10,length.out = 100))
```

# RNA-seq: import comparison analysis of IPA upstream regulators and pathways for sig changing genes
```{r import IPA results}
setwd("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/output/sigGenes_log2FC")
IPA_Compare_upstream_regs <- list()
IPA_Compare_pathways <- list()
# read in information for enriched IPA pathways per hierarchical cluster (clustering based on scaled rlog normalized counts)

upstream_regs_files <- dir(pattern="_IPA_upstream_regs.txt")
pathways_files <- dir(pattern="_IPA_pathways.txt")
for(i in 1:length(pathways_files)){
# read in information for enriched pathways
IPA_Compare_pathways[[i]]<- read_tsv(pathways_files[i], skip=2)
# read in information for enriched upstream regulators 
IPA_Compare_upstream_regs[[i]] <- read_tsv(upstream_regs_files[i], skip=2)

  IPA_Compare_pathways[[i]] <- IPA_Compare_pathways[[i]] %>%
  mutate(across(where(is.character), ~ replace(.x, .x == "N/A", "0"))) %>%
  mutate(across(-`Canonical Pathways`, as.numeric)) # Convert all except the first column to numeric
  
  IPA_Compare_upstream_regs[[i]] <- IPA_Compare_upstream_regs[[i]] %>%
  mutate(across(where(is.character), ~ replace(.x, .x == "N/A", "0"))) %>%
    mutate(across(-`Upstream Regulators`, as.numeric)) # Convert all except the first column to numeric
}

names(IPA_Compare_upstream_regs) <- gsub("_.*", "", pathways_files)
names(IPA_Compare_pathways) <- gsub("_.*", "", pathways_files)

# convert IPA results into matrix for plotting. 
IPA_Compare_pathways_mat <- lapply(IPA_Compare_pathways, function(x){
  mat <- as.matrix(x[ ,-1])
  row.names(mat) <- x$`Canonical Pathways`
 remove <- which(rowSums(mat != 0) == 0) # Find rows where the entire row (all columns) is exactly 0 to remove upstream regulators that have no activation score across all columns (DESeq comparisons)
    return(mat[-remove, ])
})

IPA_Compare_upstream_regs_mat <- lapply(IPA_Compare_upstream_regs, function(x){
  mat <- as.matrix(x[ ,-1])
  row.names(mat) <- x$`Upstream Regulators`
  remove <- which(rowSums(mat != 0) == 0) # Find rows where the entire row (all columns) is exactly 0 to remove upstream regulators that have no activation score across all columns (DESeq comparisons)
  return(mat[-remove, ])
})
```

# RNA-seq: make matrix with log2FC values of significant genes
```{r log2FC sig genes matrix}
# get log2FC values of significantly changing genes over time
sig_LFCMat <- matrix(ncol=length(LFCResultsall_RNA), nrow=nrow(clusterDF))
colnames(sig_LFCMat) <- names(LFCResultsall_RNA)
row.names(sig_LFCMat) <-row.names(LFCResultsall_RNA[[10]][row.names(LFCResultsall_RNA[[10]]) %in% sigGenes, ])
for(i in 1:ncol(sig_LFCMat)){
sig_LFCMat[ ,i] <- LFCResultsall_RNA[[i]][row.names(LFCResultsall_RNA[[i]]) %in% sigGenes, ]$log2FoldChange
}

row.names(sig_LFCMat) <-  make.unique(ifelse(genes_anno_type[match(row.names(sig_LFCMat), genes_anno_type$ensembl_id), ]$gene_id %in% NA, row.names(sig_LFCMat), genes_anno_type[match(row.names(sig_LFCMat), genes_anno_type$ensembl_id), ]$gene_id))

sig_LFCMat <- sig_LFCMat[match(row.names(ordered_sigMat), row.names(sig_LFCMat)), ] # order based on k-means & hierarchical ordering of rlog counts. 

# get log2FC values of all expressed genes over time 
LFCMat <- matrix(ncol=length(LFCResultsall_RNA), nrow=nrow(LFCResultsall_RNA[[1]]))
colnames(LFCMat) <- names(LFCResultsall_RNA)
row.names(LFCMat) <-row.names(LFCResultsall_RNA[[1]])
for(i in 1:ncol(LFCMat)){
LFCMat[ ,i] <- LFCResultsall_RNA[[i]]$log2FoldChange
}

row.names(LFCMat) <-  ifelse(genes_anno_type[match(row.names(LFCMat), genes_anno_type$ensembl_id), ]$gene_id %in% NA, row.names(LFCMat), genes_anno_type[match(row.names(LFCMat), genes_anno_type$ensembl_id), ]$gene_id)
```

# ATAC-seq: create input fasta files for MEME-AME analysis on all up or down peaks. 
```{r}
directory <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq"
outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input"


# specify file paths for input to MEME-AME analysis for up peaks
folderlist <- c("day10PalbovsDoxoUpPeaks_vsCycling", 
                "day10PalbovsDoxoUpPeaks_vsday3", 
                "day14PalbovsDoxoUpPeaks_vsCycling", 
                "day14PalbovsDoxoUpPeaks_vsday3",
                "day21PalbovsDoxoUpPeaks_vsCycling", 
                "day21PalbovsDoxoUpPeaks_vsday3",
                "day28Palbovsday21DoxoUpPeaks_vsCycling", 
                "day28Palbovsday21DoxoUpPeaks_vsday3",
                "day3PalbovsDoxoUpPeaks_vsCycling")

writeLines(dir(paste0(directory, "/annoUp"), full.names=T, pattern="Palbovs"), con=paste0(outdir, "/inputfiles_annoUp_DESeq_Palbopeaks.txt")) # directory paths to fasta files of Up peaks

writeLines(c(dir(paste0(directory, "/annoUp"), full.names=T, pattern="Doxovs")[1:6], dir(paste0(directory, "/annoUp"), full.names=T, pattern="Doxovs")[5:7]), con=paste0(outdir, "/inputfiles_annoUp_DESeq_Doxopeaks.txt")) # directory paths to fasta files of Up peaks 
# since Doxo doesn't have day 28 timepoint, repeat day 21 DESeq comparison files for comparing to day28 Palbo. 

writeLines(folderlist, con=paste0(outdir, "/folderlist_annoUp_DESeq_PalbovsDoxoPeaks.txt")) # write folder names
```

### ATAC-seq: MEME-AME analysis on DESeq_Peaks pairwise comparison between up Palbo and up Doxo peaks to find differences in enriched motifs between upregulated drug distinct peaks. 
```{bash}
cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input

# what is enriched in Palbo up peaks that is different from Doxo up peaks? set Doxo up peaks as background. 
sbatch -p risc,hpc --array=1-9 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/inputfiles_annoUp_DESeq_Doxopeaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/inputfiles_annoUp_DESeq_Palbopeaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_output/annoUp_PalbovsDoxo,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/folderlist_annoUp_DESeq_PalbovsDoxoPeaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm

# what is enriched in Doxo up peaks that is different from Palbo up peaks? set Palbo up peaks as background.  
sbatch -p risc,hpc --array=1-9 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/inputfiles_annoUp_DESeq_Palbopeaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/inputfiles_annoUp_DESeq_Doxopeaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_output/annoUp_DoxovsPalbo,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_input/folderlist_annoUp_DESeq_PalbovsDoxoPeaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm
```

# ATAC-seq: import MEME-AME results into dataframe: MEME-AME was done to find differential motifs under up peaks between Doxo & Palbo at different timepoints relative to cycling or day 3. 
```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/RNAseq/post_kallisto_analysis/workspaces/expressedGenes.RData")
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Functions/process_motif_data.RData")

# differential motifs: Doxo vs Palbo
annoUp_DvsP_AME <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_output/annoUp_DoxovsPalbo", expressed_genes = expressedGenes)

# Add column to specify which RNA cluster
annoUp_DvsP_AME$Comparison <- gsub('\\.[0-9]*$', "",  row.names(annoUp_DvsP_AME)) 
annoUp_DvsP_AME$Comparison <- factor(annoUp_DvsP_AME$Comparison, levels=unique(annoUp_DvsP_AME$Comparison)[c(9,1,3,5,7,2,4,6,8)])
# Add column to specify whether up peaks are relative to cycling or day 3 controls. 
annoUp_DvsP_AME$Group <- ifelse(grepl("vsCycling", annoUp_DvsP_AME$Comparison), "vsCycling", "vsday3")

# differential motifs: Palbo vs Doxo
annoUp_PvsD_AME <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/motif_analysis_output/annoUp_PalbovsDoxo", expressed_genes = expressedGenes)

# Add column to specify which RNA cluster
annoUp_PvsD_AME$Comparison <- gsub('\\.[0-9]*$', "",  row.names(annoUp_PvsD_AME)) 
annoUp_PvsD_AME$Comparison <- factor(annoUp_PvsD_AME$Comparison, unique(annoUp_DvsP_AME$Comparison)[c(9,1,3,5,7,2,4,6,8)])

# Add column to specify whether up peaks are relative to cycling or day 3 controls. 
annoUp_PvsD_AME$Group <- ifelse(grepl("vsCycling", annoUp_PvsD_AME$Comparison), "vsCycling", "vsday3")
```

-------------------------------------------------------------------------------------------------------------------------
# *FIGURES*

# RNA-seq: plot master heatmap of -log10 pvalues of GO enrichment results for all clusters
```{r}
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
pdf(paste0(pdf_dir, "/GO_results_list_mat.pdf"), height=20, width=15)
pheatmap(GO_results_list_mat, cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100),breaks = seq(0, 10,length.out = 100), annotation_row = GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, clustering_distance_rows = "correlation", border_color = "black")
dev.off()
```

# RNA-seq: plot padj values of GO enrichment results for top 5 drug distinct up genes
```{r}
# Still showing the pvalues of the other clusters
GO_results_mat_clusters <- list()
for(i in 1:length(GO_results_list)){

GO_results_mat_clusters[[i]] <- GO_results_list_mat[row.names(GO_results_list_mat) %in% GO_results_list[[i]]$ID[1:25], ]
GO_results_mat_clusters[[i]] <- GO_results_mat_clusters[[i]][order(GO_results_mat_clusters[[i]][ ,i], decreasing=T), ]
}

# plot padj values for top 5 GO terms for each cluster individually. Still showing the pvalues of the other clusters as columns but only the top GO terms from the cluster of interest. 
GO_results_mat_clusters <- list()
for(i in 1:length(GO_results_list)){

GO_results_mat_clusters[[i]] <- GO_results_list_mat[row.names(GO_results_list_mat) %in% GO_results_list[[i]]$ID[1:10], ]
GO_results_mat_clusters[[i]] <- GO_results_mat_clusters[[i]][order(GO_results_mat_clusters[[i]][ ,i], decreasing=T), ]
}

pdf(paste0(pdf_dir, "/GO_results_Palbo1_clusters.pdf"), width=8, height=2)
pheatmap(GO_results_mat_clusters[[3]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 20,length.out = 100), cluster_rows = F, main=cluster_names[3], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()

pdf(paste0(pdf_dir, "/GO_results_Palbo2_clusters.pdf"), width=5.5, height=2)
pheatmap(GO_results_mat_clusters[[4]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 10,length.out = 100), cluster_rows = F, main=cluster_names[4], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()

pdf(paste0(pdf_dir, "/GO_results_Doxo1_clusters.pdf"), width=9, height=2)
pheatmap(GO_results_mat_clusters[[5]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 5,length.out = 100), cluster_rows = F, main=cluster_names[5], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()

pdf(paste0(pdf_dir, "/GO_results_Doxo2_clusters.pdf"), width=6.5, height=2)
pheatmap(GO_results_mat_clusters[[6]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 8,length.out = 100), cluster_rows = F, main=cluster_names[6], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()
```

# RNA-seq: plot padj values for top 5 IPA upstream regs for each cluster individually. 
```{r}
# Still showing the pvalues of the other clusters as columns but only the top IPA upstream regs from the cluster of interest. 
IPA_TF_regs_clusters <- list()
IPA_group_regs_clusters <- list()
for(i in 1:length(IPA_upstream_regs)){
  #transcription regulators
IPA_TF_regs_clusters[[i]] <-IPA_TF_regs_mat[row.names(IPA_TF_regs_mat) %in% IPA_upstream_regs[[i]]$Upstream.Regulator[grepl("transcription", IPA_upstream_regs[[i]]$Molecule.Type)][1:5], ]
IPA_TF_regs_clusters[[i]] <- IPA_TF_regs_clusters[[i]][order(IPA_TF_regs_clusters[[i]][ ,i], decreasing=T), ]
  # complex regulators
IPA_group_regs_clusters[[i]] <-IPA_group_regs_mat[row.names(IPA_group_regs_mat) %in% IPA_upstream_regs[[i]]$Upstream.Regulator[grepl("group", IPA_upstream_regs[[i]]$Molecule.Type)][1:5], ]
IPA_group_regs_clusters[[i]] <- IPA_group_regs_clusters[[i]][order(IPA_group_regs_clusters[[i]][ ,i], decreasing=T), ] # sort by decreasing padj value
}

# make heatmap of top 5 IPA predicted upstream regulators for drug distinct upregulated genes
pdf(file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure Drug distinct responses and SASP/RNA_IPA_TF_regs_drug_distinct_up_genes.pdf", width=3, height=2)
pheatmap(IPA_TF_regs_clusters[[6]], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 100,length.out = 100), cluster_rows = F, main="Shared down genes: predicted TF regulators", border_color = "black",  annotation_col=Cluster_anno, annotation_colors = GO_anno_colors)
dev.off()

pdf(file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure Drug distinct responses and SASP/RNA_IPA_group_regs_SharedDown_genes.pdf", width=3.8, height=2)
pheatmap(IPA_group_regs_clusters[[1]], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 100,length.out = 100), cluster_rows = F, main="Shared down genes: predicted group regulators", border_color = "black", annotation_col=Cluster_anno, annotation_colors = GO_anno_colors)
dev.off()
```

# plot log2FC of E2F target genes at day 14 Palbo & day 14 Doxo relative to Cycling. 
```{r}

log2FC_E2F <- rbind(
          data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Palbo3 vs Cycling0"], 
           Condition="day 3 Palbo", 
           Comparison="vs Cycling"),
  
          data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Palbo14 vs Cycling0"], 
           Condition="day 14 Palbo", 
           Comparison="vs Cycling" ), 
           
           data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Palbo21 vs Cycling0"], 
           Condition="day 21 Palbo", 
           Comparison="vs Cycling" ), 
           
           data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Palbo28 vs Cycling0"], 
           Condition="day 28 Palbo", 
           Comparison="vs Cycling" ),
           
           data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Doxo3 vs Cycling0"], 
           Condition="day 3 Doxo", 
           Comparison="vs Cycling"),
      
      data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Doxo14 vs Cycling0"], 
           Condition="day 14 Doxo", 
           Comparison="vs Cycling"), 
      
          data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Doxo21 vs Cycling0"], 
           Condition="day 21 Doxo", 
           Comparison="vs Cycling"), 
      
      data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Palbo14 vs Palbo3"], 
           Condition="day 14 Palbo",
           Comparison="vs day 3"), 
      
       data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Palbo21 vs Palbo3"], 
           Condition="day 21 Palbo",
           Comparison="vs day 3" ), 
           
           data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Palbo28 vs Palbo3"], 
           Condition="day 28 Palbo",
           Comparison="vs day 3" ),
      
      data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Doxo14 vs Doxo3"], 
           Condition="day 14 Doxo",
           Comparison="vs day 3"),
      
      data.frame(genes= E2F_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% E2F_genes, "Doxo21 vs Doxo3"], 
           Condition="day 21 Doxo",
           Comparison="vs day 3")
)
# function to create violin plot of log2FC values of genes
create_violin_plot <- function(data=log2FC_E2F, plot_title="RNA: log2FC of p53 target genes") {
  library(ggplot2)
  library(ggbeeswarm)
  
  ggplot(
    data,
    aes(x = Condition, y = log2FC, fill = Condition)
  ) +
    geom_violin() +
    geom_quasirandom(
      color = "black", alpha = 0.2,
      method = "pseudorandom", # Aligns jitter points to violin shape
      groupOnX = TRUE,
      size = 0.5
    ) +
    stat_summary(
      fun = "median",
      geom = "crossbar",
      color = "white",
      width = 0.5
    ) +
    theme_classic() +
    theme(
      axis.text.x = element_text(angle = 90, size = 8),
      axis.text.y = element_text(size = 8),
      axis.title.y = element_text(size = 10)
    ) +
    facet_wrap(~Comparison, scales = "free_x") +
    scale_fill_manual(
      values = c(
        "day 3 Doxo" = "#B3C543",
        "day 14 Doxo" = "#B3C543",
        "day 21 Doxo" = "#B3C543",
        "day 3 Palbo" = "#E0ADCD",
        "day 14 Palbo" = "#E0ADCD",
        "day 21 Palbo" = "#E0ADCD",
        "day 28 Palbo" = "#E0ADCD"
      )
    ) +
    ylab("log2FC") +
    xlab("") +
    ggtitle(plot_title)
}

pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure Drug distinct responses and SASP/RNA_log2FC_E2F_targets.pdf", width=7, height=3)
library(ggbeeswarm)
create_violin_plot(
  data = log2FC_E2F,
  plot_title = "RNA: log2FC of E2F Target genes"
)
dev.off()

pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure Drug distinct responses and SASP/RNA_log2FC_day14_E2F_targets.pdf", width=4, height=3)
library(ggbeeswarm)
create_violin_plot(
  data = log2FC_E2F[log2FC_E2F$Condition %in% c("day 14 Palbo", "day 14 Doxo"), ],
  plot_title = "RNA: log2FC of E2F Target genes"
)
dev.off()
```

# ATAC-seq, merged RNA clusters: plot motif enrichment results of all clusters as heatmap of p-adj values for each type ("PalboDownvsCyc", "PalboDownvsday3",  "DoxoDownvsCyc","DoxoDownvsday3", "PalboUpvsCyc", "PalboUpvsday3", "DoxoUpvsCyc", "DoxoUpvsday3") of sig changing peaks
```{r}
# load in required functions 
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Functions/create_AME_peaks_matrix.RData")

# get top enriched motifs for Doxo vs Palbo MEME-AME analysis results
AME_padj_mat_DvsP <- list()
comparisons <- levels(annoUp_DvsP_AME$Comparison)

for(i in 1:length(comparisons)){
AME_padj_mat_DvsP[[i]] <- create_AME_peaks_matrix(annoUp_DvsP_AME[annoUp_DvsP_AME$Comparison %in% comparisons[i], ])

AME_padj_mat_DvsP[[i]] <- AME_padj_mat_DvsP[[i]][order(rowSums(AME_padj_mat_DvsP[[i]]), decreasing=T), ] # sort motif IDs by overall significance across clusters. 
}

names(AME_padj_mat_DvsP) <- comparisons

AME_padj_dots_DvsP <- list()
for(i in 1:length(comparisons)){
AME_padj_dots_DvsP[[i]] <- annoUp_DvsP_AME[annoUp_DvsP_AME$Comparison %in% comparisons[i], ]

max_number <- ifelse(nrow(AME_padj_mat_DvsP[[i]]) >=7, 7, nrow(AME_padj_mat_DvsP[[i]]))
AME_padj_dots_DvsP[[i]] <- AME_padj_dots_DvsP[[i]][AME_padj_dots_DvsP[[i]]$motif_ID %in% row.names(AME_padj_mat_DvsP[[i]][1:max_number, ]), ] # show only top 15 most significant padj values motif IDs
AME_padj_dots_DvsP[[i]]$motif_ID <- factor(AME_padj_dots_DvsP[[i]]$motif_ID, levels=rev(unique(AME_padj_dots_DvsP[[i]]$motif_ID)))
}

AME_padj_dots_DvsP <- do.call(rbind, AME_padj_dots_DvsP)


# get top enriched motifs for Doxo vs Palbo MEME-AME analysis results
AME_padj_mat_PvsD <- list()
comparisons <- levels(annoUp_PvsD_AME$Comparison)

for(i in 1:length(comparisons)){
AME_padj_mat_PvsD[[i]] <- create_AME_peaks_matrix(annoUp_PvsD_AME[annoUp_PvsD_AME$Comparison %in% comparisons[i], ])

AME_padj_mat_PvsD[[i]] <- AME_padj_mat_PvsD[[i]][order(rowSums(AME_padj_mat_PvsD[[i]]), decreasing=T), ] # sort motif IDs by overall significance across clusters. 
}

names(AME_padj_mat_PvsD) <- comparisons

AME_padj_dots_PvsD <- list()
for(i in 1:length(comparisons)){
AME_padj_dots_PvsD[[i]] <- annoUp_PvsD_AME[annoUp_PvsD_AME$Comparison %in% comparisons[i], ]

max_number <- ifelse(nrow(AME_padj_mat_PvsD[[i]]) >=7, 7, nrow(AME_padj_mat_PvsD[[i]]))
AME_padj_dots_PvsD[[i]] <- AME_padj_dots_PvsD[[i]][AME_padj_dots_PvsD[[i]]$motif_ID %in% row.names(AME_padj_mat_PvsD[[i]][1:max_number, ]), ] # show only top 15 most significant padj values motif IDs
AME_padj_dots_PvsD[[i]]$motif_ID <- factor(AME_padj_dots_PvsD[[i]]$motif_ID, levels=rev(unique(AME_padj_dots_PvsD[[i]]$motif_ID)))
}

AME_padj_dots_PvsD <- do.call(rbind, AME_padj_dots_PvsD)

# create function for plotting motif enrichment results as dotplot
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=50, size_limit=4.5 ){
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient(name="-log10(padj)", 
                        low="light yellow", high="dark orange",
                        limits=c(0, fill_limit), 
                        oob=scales::squish) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=c(0, size_limit), 
                          range=c(1, 10)) +  # Adjust dot sizes
    ggtitle(Title) +
    theme_classic()+
    theme(
      text = element_text(size=8), 
      axis.text.x = element_text(size=8, angle=90), 
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    ) +
    facet_wrap(~Group, nrow=1, scales="free") +
    labs(x="Comparison", y="motif ID")
}


pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
pdf(paste0(pdf_dir, "/ATAC_PalbovsDoxo_top_differential_motifs_UpPeaks.dotplots.pdf"), width=5, height=7)
create_AME_dotplots(AME_padj_dots_DvsP, Title="Doxo vs Palbo Up peaks", fill_limit = 200, size_limit = 35) 
create_AME_dotplots(AME_padj_dots_PvsD, Title="Palbo vs Doxo Up peaks", fill_limit = 100, size_limit = 25) 
dev.off()
```

# RNA-seq: plotting padj values of top IPA predicted upstream regulators of drug distinct gene clusters. 
```{r}
filter_upstream_regulators <- function(mat, threshold_neg1 = -0.5, threshold_pos1 = 0.5,  threshold_neg2 = -0.5, threshold_pos2 = 0.5) {
  # Identify upstream regulators that are differentially regulated between Doxo & Palbo
  filt_negDoxo_posPalbo <- which(rowSums(mat[, 1:5]) <= threshold_neg1 & rowSums(mat[, 6:10]) >= threshold_pos1)
  filt_posDoxo_negPalbo <- which(rowSums(mat[, 1:5]) >= threshold_pos2 & rowSums(mat[, 6:10]) <= threshold_neg2)
  
  # Combine filtered indices
  filt_mat <- mat[c(filt_negDoxo_posPalbo, filt_posDoxo_negPalbo), , drop = FALSE]
  
  # Remove random drugs/chemicals (upstream regulators that contain at least 2 uppercase letters)
  filt_mat <- filt_mat[grep("[A-Z].*[A-Z]", rownames(filt_mat)), , drop = FALSE]
  
  return(filt_mat)
}

# plot most differentially activated upstream regulators of Palbo1 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Palbo1, threshold_neg1 = -8.5, threshold_pos1 = 10,  threshold_neg2 = -6, threshold_pos2 = 8) # set differential thresholds for Palbo vs Doxo. 

remove_id <- which(grepl("drug|chemical", IPA_upstream_regs$Palbo1[match(row.names(filt_mat), IPA_upstream_regs$Palbo1$Upstream.Regulator), ]$Molecule.Type)) # remove upstream regulators that are drugs/chemicals
filt_mat <- filt_mat[-remove_id, ]

pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP"
# plot heatmap of -log10 pvalue of overlap of enriched complex regulators
pdf(paste0(pdf_dir, "/RNA_IPA_Palbo1_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,6:10]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-3, 3,length.out = 100), main="Activation scores")
dev.off()

# plot most differentially activated upstream regulators of Palbo2 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Palbo2, threshold_neg1 = 1, threshold_pos1 = 7, threshold_neg2 = -0.5, threshold_pos2 =0.5)
filt_mat <- filt_mat[grep("-", rownames(filt_mat), invert = T), , drop = FALSE]

pdf(paste0(pdf_dir, "/RNA_IPA_Palbo2_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,6:10]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")
dev.off()

# plot most differentially activated upstream regulators of Doxo1 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Doxo1, threshold_neg1 = -4.2, threshold_pos1 = 4.2, threshold_neg2 = -4.2, threshold_pos2 =4.2)
filt_mat <- filt_mat[grep("-|GW9662", rownames(filt_mat), invert = T), , drop = FALSE] # remove anything with - or drug GW9662 from final mat to plot. 

pdf(paste0(pdf_dir, "/RNA_IPA_Doxo1_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,1:5]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")
dev.off()

# plot most differentially activated upstream regulators of Doxo2 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Doxo2, threshold_neg1 = -6.5, threshold_pos1 = 0,  threshold_neg2 = -2, threshold_pos2 = 0.5)

mat <- IPA_Compare_upstream_regs_mat$Doxo2
pdf(paste0(pdf_dir, "/RNA_IPA_Doxo2_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,1:5]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")
dev.off()
```