---
title: "Supplemental_Tables"
author: "Joanna Yeung"
date: '2025-03-22'
output: html_document
---
```{r}
# data wrangling
library(preprocessCore)
library(readr)
library(readxl)
library(stringr)
library(dplyr)
library(writexl)

# plotting
library(gridExtra)
library(ggplot2)
library(pheatmap)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(scales)
library(ggrepel)

# genomics data wrangling and analysis
library(rtracklayer)
library(Biostrings)
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(DESeq2)
library(apeglm)
library(ChIPseeker)
library(clusterProfiler)
library(msigdbr)
library(MotifDb)
library(motifmatchr)

# genome annotation databases
library(EnsDb.Hsapiens.v79)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(BSgenome.Hsapiens.UCSC.hg38)
library(org.Hs.eg.db)
```

```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Global_transcriptome_and_chromatin_accessibility_is_dynamic_in_both_types_of_therapy_induced_senescence/workspace.RData")
```

# make function to specify code chunk to run from other Rmd file to avoid copying and pasting large lines of code. 
```{r}
# arguments: 
# Rmd_file=file path & name of Rmd file with code chunk
# chunk_name= name of code chunk to run from Rmd file. 
run_code_chunk <- function(Rmd_file, chunk_name){ 
  # Read the Rmd file
lines <- readLines(Rmd_file)

# Identify the start and end of the specific chunk
chunk_start <- grep(paste0("^```\\{r\\s*", chunk_name), lines)  # Find chunk start
chunk_end <- grep("^```$", lines)  # Find all chunk end lines

# Find the first `chunk_end` after `chunk_start`
chunk_end <- chunk_end[chunk_end > chunk_start][1]

  if (is.na(chunk_end)) {
    stop(paste("End of chunk", chunk_name, "not found. Check the Rmd file for missing delimiters."))
  }

# specify lines inside chunk, removing lines that make up the start and end of the chunk. 
chunk_to_run <- lines[(chunk_start[1]+1):(chunk_end[1]-1)]

# Evaluate the extracted code
  eval(parse(text = paste(chunk_to_run, collapse = "\n")), envir = .GlobalEnv)
}
```

# RNA: make supplemental table of significantly changing genes. 
```{r}
IPA_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/output"
IPA_upstream_regs <- list()
IPA_pathways <- list()
# read in information for enriched IPA pathways per hierarchical cluster (clustering based on scaled rlog normalized counts)
for(i in 1:length(cluster_names)){
IPA_pathways[[i]]<- read_tsv(paste0(IPA_dir, "/", cluster_names[i], "_IPA_pathways.txt"), skip=2)
# read in information for enriched upstream regulators 
IPA_upstream_regs[[i]] <- read.table(paste0(IPA_dir, "/", cluster_names[i], "_IPA_upstream_regs.txt"), sep="\t", skip=2, header=T)
IPA_upstream_regs[[i]] <- IPA_upstream_regs[[i]][!grepl("mir|miR|Mir", IPA_upstream_regs[[i]]$Upstream.Regulator), ] # remove upstream regulators that are miRNAs (uncertain if they will be informative & there is a lot of them)
}

names(IPA_upstream_regs) <- cluster_names
names(IPA_pathways) <- cluster_names


run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="log2FC sig genes matrix")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling.Rmd", chunk_name="p53 NFKB databases")

# filter for E2F target genes from IPA & Hallmark E2F Targets pathway term. 
E2F_Targets=unique(c(unlist(lapply(IPA_upstream_regs, function(x){str_split(x[grep("E2F", x$Upstream.Regulator), ]$Target.Molecules.in.Dataset, ",")})), 
              msigdbr_df_H[msigdbr_df_H$gs_name %in% "HALLMARK_E2F_TARGETS", ]$gene_symbol
))

# add annotation to significant genes based on if they are E2F targets. 
clusterDF$E2F <- ifelse(clusterDF$gene_id %in% E2F_Targets, "Yes", "No") # annotate sig genes that are E2F targets

# add ENSG id
clusterDF$ENSG_id <- row.names(LFCResultsall_RNA[[1]][match(clusterDF$gene_id, LFCResultsall_RNA[[1]]$SYMBOL), ])
# add in drug category column 
clusterDF$RNA_Cluster_Description <- ifelse(clusterDF$Clusters %in% c("2", "5"), "Shared downregulated", ifelse(clusterDF$Clusters %in% c("7", "1"), "Palbo upregulated", ifelse(clusterDF$Clusters %in% c("4", "6"), "Doxo upregulated", "Shared upregulated")))

# reformat number assigned by kmeans clustering to 1:7 as indicated in the figures. 
kmeansCluster_to_RNACluster <- data.frame(Cluster=levels(clusterDF$Clusters), RNA_cluster=factor(1:7))
clusterDF$RNA_Cluster <- kmeansCluster_to_RNACluster[match(clusterDF$Clusters, kmeansCluster_to_RNACluster$Cluster), ]$RNA_cluster
```
```{r}
# change column names for annotation of significantly changing genes. 
clusterDF <- clusterDF[ ,c("Clusters", "RNA_Cluster", "RNA_Cluster_Description", "gene_id", "ENSG_id", "SASP", "Cell_Cycle", "NFKB_All", "p53_All", "E2F")]
colnames(clusterDF) <- c("Kmeans_assigned_Cluster_number", "RNA_Cluster", "RNA_Cluster_Description", "gene_symbol", "gene_id", "SASP", "Cell_Cycle", "NFKB_target", "p53_target", "E2F_target")

# change column names for counts & log2FC matrix of sig changing genes. 
# match order of sigMat
metaDataall <- metaDataall[match(colnames(ordered_sigMat), row.names(metaDataall)), ]

new_LFCnames <- gsub("(Doxo|Palbo)(\\d+) vs (Doxo|Palbo|Cycling)(\\d*)", 
                 "day\\2 \\1 vs day\\4 \\3", colnames(sig_LFCMat))

# Remove "day " before "Cycling" since it shouldn't have a day value
new_LFCnames <- gsub("day  Cycling", "Cycling", new_LFCnames)

# change column names for counts & log2FC matrix of sig changing genes. 
colnames(ordered_sigMat) <- paste0("Biorep", metaDataall$Biorep, "_", metaDataall$Treatment, "_day", metaDataall$Timepoint)
colnames(sig_LFCMat) <- new_LFCnames

# convert to dataframe
ordered_sigMat <- data.frame(gene_symbol=row.names(ordered_sigMat), ordered_sigMat)
sig_LFCMat <-  data.frame(gene_symbol=row.names(sig_LFCMat), sig_LFCMat)

# all genes counts matrix, normalized counts 
all_genes_counts_matrix <- counts(RNAddsall, normalize=T)
colnames(all_genes_counts_matrix) <-  paste0("Biorep", metaDataall$Biorep, "_", metaDataall$Treatment, "_day", metaDataall$Timepoint)
all_genes_counts_matrix <- all_genes_counts_matrix[ ,c(1,9,15,2,10,16,3,17,4,11,5,12,18,6,19,7,13,8,14)]
# export all genes counts matrix for GEO submission
write.csv(all_genes_counts_matrix, file="/lustre/fs4/risc_lab/store/jyeung/Senescence_Liposarcoma_GEO_submission/RNAseq/Normalized_by_DESeq2_RNA_counts_matrix.csv")
```
```{r}
Table_num <- "1"
write_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Tables"
# Write to an Excel file with each table as a separate sheet
write_xlsx(list("Annotation" = clusterDF, "rlog_normalized_counts" = ordered_sigMat, "log2FoldChange"=sig_LFCMat), paste0(write_dir, "/Supplemental_Table", Table_num, "_RNA_significantly_changing_genes.xlsx"))
```

# Supplemental Table of vst normalized counts under ATAC-seq peaks
```{r}
ATAC_sigPeaks_Table <- function(matrix, sigLFCResults_ATAC=sigLFCResults_ATACUp, DoxoGR=mergedDiffGR$allUp_Doxo, PalboGR=mergedDiffGR$allUp_Palbo) {

# make dataframe to specify which drug treatment peak belongs to: Palbo, Doxo or both. 
drug_cat <- data.frame(
                       peak_id=row.names(matrix),
                       row.names=row.names(matrix), Palbo=ifelse(row.names(matrix) %in% PalboGR$name, "Yes", "No"), 
                       Doxo=ifelse(row.names(matrix) %in% DoxoGR$name, "Yes", "No"), 
                       Both=ifelse(row.names(matrix) %in% DoxoGR$name & row.names(matrix) %in% PalboGR$name, "Yes", "No") )

# order rows by Palbo, Doxo then both peaks
drug_cat_list <- list(Palbo=drug_cat[drug_cat$Palbo %in% "Yes" & drug_cat$Doxo %in% "No", ],
                      Shared=drug_cat[drug_cat$Both %in% "Yes", ],
                      Doxo=drug_cat[drug_cat$Palbo %in% "No" & drug_cat$Doxo %in% "Yes", ]
)
drug_cat <- do.call(rbind, drug_cat_list)

ordered_matrix <- matrix[match(drug_cat$peak_id, row.names(matrix)), ]

# Apply the function to create a list of data frames with dynamic column names
result_list <- lapply(names(sigLFCResults_ATAC), function(comp_name) {
  x <- sigLFCResults_ATAC[[comp_name]]
  df <- data.frame(
    peak_id = drug_cat$peak_id, 
    Comparison = ifelse(drug_cat$peak_id %in% row.names(x), "Yes", "No")
  )
  # Set the column name of 'Comparison' to the name of the current list element
   colnames(df)[2] <- comp_name
  return(df)
})

# Merge all data frames by the 'peak_id' column
merged_result <- do.call(cbind, result_list)
merged_result <- merged_result[ ,c(1, which(!colnames(merged_result) %in% "peak_id"))]
merged_anno <- cbind(drug_cat, merged_result)
merged_anno <- merged_anno[ ,c(1, which(!colnames(merged_anno) %in% "peak_id"))]
merged_anno <- cbind(merged_anno, as.data.frame(annomasterpeaks_pval0.2[match(merged_anno$peak_id, annomasterpeaks_pval0.2$name), ])[ ,c("annotation", "geneId", "distanceToTSS")])

colnames(ordered_matrix) <- gsub("None", "Cycling", paste0("Biorep", annoCol$Biorep, "_", annoCol$Treatment, "_day", annoCol$Timepoint))
# convert to dataframe
ordered_matrix <- data.frame(peak_id=row.names(ordered_matrix), ordered_matrix)

return(list(matrix=ordered_matrix, peak_anno=merged_anno))
}
```
```{r}
sigUp_ATACPeaksTable <- ATAC_sigPeaks_Table(matrix=mergedDiff_hm_rmOutlier$UpPalbo.UpDoxo, sigLFCResults_ATAC=sigLFCResults_ATACUp, DoxoGR=mergedDiffGR$allUp_Doxo, PalboGR=mergedDiffGR$allUp_Palbo)

sigDown_ATACPeaksTable <- ATAC_sigPeaks_Table(matrix=mergedDiff_hm_rmOutlier$DownPalbo.DownDoxo, sigLFCResults_ATAC=sigLFCResults_ATACDown, DoxoGR=mergedDiffGR$allDown_Doxo, PalboGR=mergedDiffGR$allDown_Palbo)

Table_num <- 2
# Write to an Excel file with each table as a separate sheet
write_xlsx(list("UpPeaks_Annotation" = sigUp_ATACPeaksTable$peak_anno, "vst_normalized_counts_UpPeaks" = sigUp_ATACPeaksTable$matrix, "DownPeaks_Annotation" = sigDown_ATACPeaksTable$peak_anno, "vst_normalized_counts_DownPeaks" = sigDown_ATACPeaksTable$matrix), paste0(write_dir, "/Supplemental_Table", Table_num, "_ATAC_significantly_changing_peaks.xlsx"))

# counts matrix for ATAC-seq peaks 
ATAC_normCounts_matrix <- counts(dds_filt, normalized=T)
ATAC_normCounts_matrix <- ATAC_normCounts_matrix[, c(1:5, 38, 40, 39, 41, 42, 28:32, 43:45, 6:7, 49:51, 11:15, 33:37, 46:48, 8:10, 52:54, 16:27)]
colnames(ATAC_normCounts_matrix) <- c("Cycling_Biorep1.1","Cycling_Biorep1.2","Cycling_Biorep2.1","Cycling_Biorep2.2","Cycling_Biorep2.3","Cycling_Biorep3.2","Cycling_Biorep3.3","Cycling_Biorep4.1","Cycling_Biorep4.2","Cycling_Biorep4.3","day3_Doxo_Biorep1.1","day3_Doxo_Biorep1.2","day3_Doxo_Biorep2.1","day3_Doxo_Biorep2.2","day3_Doxo_Biorep2.3","day10_Doxo_Biorep1.1","day10_Doxo_Biorep1.2","day10_Doxo_Biorep1.3","day14_Doxo_Biorep2.1","day14_Doxo_Biorep2.2","day14_Doxo_Biorep1.1","day14_Doxo_Biorep1.2","day14_Doxo_Biorep1.3","day21_Doxo_Biorep1.1","day21_Doxo_Biorep1.2","day21_Doxo_Biorep1.3","day21_Doxo_Biorep2.1","day21_Doxo_Biorep2.2","day3_Palbo_Biorep1.1","day3_Palbo_Biorep1.2","day3_Palbo_Biorep2.1","day3_Palbo_Biorep2.2","day3_Palbo_Biorep2.3","day10_Palbo_Biorep1.1","day10_Palbo_Biorep1.2","day10_Palbo_Biorep1.3","day14_Palbo_Biorep2.1","day14_Palbo_Biorep2.2","day14_Palbo_Biorep2.3","day14_Palbo_Biorep1.1","day14_Palbo_Biorep1.2","day14_Palbo_Biorep1.3","day21_Palbo_Biorep1.1","day21_Palbo_Biorep1.2","day21_Palbo_Biorep1.3","day21_Palbo_Biorep2.1","day21_Palbo_Biorep2.2","day21_Palbo_Biorep2.3","day28_Palbo_Biorep1.1","day28_Palbo_Biorep1.2","day28_Palbo_Biorep1.3","day28_Palbo_Biorep2.1","day28_Palbo_Biorep2.2","day28_Palbo_Biorep2.3")

# export all genes counts matrix for GEO submission
write.csv(ATAC_normCounts_matrix, file="/lustre/fs4/risc_lab/store/jyeung/Senescence_Liposarcoma_GEO_submission/ATACseq/Normalized_by_DESeq2_ATAC_counts_matrix.csv")
```

# Supplemental table of RNA databases of target genes
```{r}
SASP_db <- data.frame(SASP_genes = unique(SASP_all) )
CellCycle_db <- data.frame(CellCycle_genes=unique(c(msigdbr_df_H[msigdbr_df_H$gs_name %in% c("HALLMARK_E2F_TARGETS", "HALLMARK_G2M_CHECKPOINT"), ]$gene_symbol, msigdbr_df_REACT[msigdbr_df_REACT$gs_name %in% c("REACTOME_CELL_CYCLE", "REACTOME_CELL_CYCLE_CHECKPOINTS", "REACTOME_CELL_CYCLE_MITOTIC"), ]$gene_symbol, msigdbr_df_KEGG[msigdbr_df_KEGG$gs_name %in% "KEGG_CELL_CYCLE", ]$gene_symbol))
)
NFKB_Targets_db <- data.frame(NFKB_targets=unique(c(Gilmore_NFKB_Targets, NFKB_Hallmark, IPA_targets$NFKB))
)
p53_Targets_db <- data.frame(p53_targets=unique(c(Fischer_Oncogene_2017_p53_census_target_genes$`Gene Symbol`, p53_Hallmark))
)

E2F_Targets_db <- data.frame(E2F_targets=unique(E2F_Targets) )

Table_num <- 3
# Write to an Excel file with each table as a separate sheet
write_xlsx(list("SASP genes" = SASP_db, "Cell cycle genes" = CellCycle_db, "NFkB target genes" = NFKB_Targets_db, "p53 target genes" = p53_Targets_db, "E2F target genes"=E2F_Targets_db ), paste0(write_dir, "/Supplemental_Table", Table_num, "_RNA_Curated_databases_for_gene_annotation.xlsx"))
```

# Supplemental table of activation scores from IPA_Comparison_Analysis on upstream regulators from RNA log2FC comparisons.  
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name=" import_IPA_comparison_results")

IPA_Compare_upstream_regs_df <- lapply(IPA_Compare_upstream_regs_mat, function(x){
  new_LFCnames <- gsub("(Doxo|Palbo)(\\d+) vs (Doxo|Palbo|Cycling)(\\d*)", 
                 "day\\2 \\1 vs day\\4 \\3", colnames(x))

# Remove "day " before "Cycling" since it shouldn't have a day value
new_LFCnames <- gsub("day  Cycling", "Cycling", new_LFCnames)

colnames(x) <- new_LFCnames

df <- data.frame(upstream_regulator=row.names(x), x)
return(df)
})

names(IPA_Compare_upstream_regs_df) <- c("Shared down (RNA Cluster 1)", "Shared down (RNA Cluster 2)", "Doxo up (RNA Cluster 5)", "Doxo up (RNA Cluster 6)", "Palbo up (RNA Cluster 3)", "Palbo up (RNA Cluster 4)", "Shared up (RNA Cluster 7)" )

IPA_Compare_upstream_regs_df <- IPA_Compare_upstream_regs_df[c(1,2,5,6,3,4,7)]
Table_num <- 4
# Write to an Excel file with each table as a separate sheet
write_xlsx(IPA_Compare_upstream_regs_df, paste0(write_dir, "/Supplemental_Table", Table_num, "_RNA_Comparison_Analysis_upstream_regulators_activation_scores.xlsx"))
```

# Supplemental table of IPA upstream regulators & GO enrichment results from RNA-seq data (based on gene IDs input)
```{r }
universe <- genes_anno_type[match(gsub('\\.[0-9]*$', '', row.names(acrossGroups)), genes_anno_type$ensembl_id), ]

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="getGOenrichment function")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="genes_in_clusters")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="RNA-seq GO enrichment analysis")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name=" RNAseq GO enrichment results")

names(GO_results_list) <- c("Shared down (RNA Cluster 1)", "Shared down (RNA Cluster 2)", "Palbo up (RNA Cluster 3)", "Palbo up (RNA Cluster 4)", "Doxo up (RNA Cluster 5)", "Doxo up (RNA Cluster 6)", "Shared up (RNA Cluster 7)" )

GO_results_list <- lapply(seq_along(GO_results_list), function(i) {
  x <- GO_results_list[[i]]
  x$RNA_Cluster <- names(GO_results_list)[i]  # Assign the correct name
  x <- x[ , !colnames(x) %in% c("Description", "Cluster")]  # Avoid dropping to vector
  return(x)
})

names(IPA_upstream_regs) <- c("Shared down (RNA Cluster 1)", "Shared down (RNA Cluster 2)", "Palbo up (RNA Cluster 3)", "Palbo up (RNA Cluster 4)", "Doxo up (RNA Cluster 5)", "Doxo up (RNA Cluster 6)", "Shared up (RNA Cluster 7)" )

IPA_upstream_regs <- lapply(seq_along(IPA_upstream_regs), function(i) {
  x <- IPA_upstream_regs[[i]]
  x$RNA_Cluster <- names(IPA_upstream_regs)[i]  # Assign the correct name
  return(x)
})

GO_results_df <- do.call(rbind, GO_results_list)
IPA_upstream_regs_df <- do.call(rbind, IPA_upstream_regs)

Table_num <- 5
# Write to an Excel file with each table as a separate sheet
write_xlsx(list("GO_enrichment_results"=GO_results_df, "Predicted_Upstream_Regulators"=IPA_upstream_regs_df), paste0(write_dir, "/Supplemental_Table", Table_num, "_RNA_GO_enrichment_IPA_upstream_regulators_results.xlsx"))
```


# Supplemental table of ATAC motif enrichment results on up/down peaks at each timepoint. 
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="annoUpPeaks_AME_df")
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="annoDownPeaks_AME_df")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="create_AME_peaks_matrix")
 
Palbo_Doxo_AME_UpPeaks_Mat <- create_AME_peaks_matrix(annoUpPeaks_AME_df)

 set.seed(123)  # For reproducibility
  distance_matrix <- dist(Palbo_Doxo_AME_UpPeaks_Mat)
    hc <- hclust(distance_matrix, method = "complete")
    
    # Order the rows based on hierarchical clustering
    order <- hc$order
    
    # Reorder the original data matrix
    Palbo_Doxo_AME_UpPeaks_Mat <- Palbo_Doxo_AME_UpPeaks_Mat[order, ]
    
Palbo_Doxo_AME_DownPeaks_Mat <- create_AME_peaks_matrix(annoDownPeaks_AME_df)

 set.seed(123)  # For reproducibility
  distance_matrix <- dist(Palbo_Doxo_AME_DownPeaks_Mat)
    hc <- hclust(distance_matrix, method = "complete")
    
    # Order the rows based on hierarchical clustering
    order <- hc$order
    
    # Reorder the original data matrix
    Palbo_Doxo_AME_DownPeaks_Mat <- Palbo_Doxo_AME_DownPeaks_Mat[order, ]

    # rename columns to make more sense and remove unnecessary columns
annoUpPeaks_AME_df <- annoUpPeaks_AME_df[ ,!colnames(annoUpPeaks_AME_df) %in% c("rank", "motif_DB", "motif_alt_ID")]
annoDownPeaks_AME_df <- annoDownPeaks_AME_df[ ,!colnames(annoDownPeaks_AME_df) %in% c("rank", "motif_DB", "motif_alt_ID")]
colnames(annoUpPeaks_AME_df) <- c("motif_ID", "consensus", "p-value", "adj-pvalue", "E-value", "tests", "FASTA_max", "pos", "neg", "PWM_min", "True_Positive(TP)", "percent_TP", "False_Positive(FP)", "percent_FP", "Percent_TP - Percent_FP", "-log10(padj)value", "Comparison", "Group", "Timepoint")
colnames(annoDownPeaks_AME_df) <- colnames(annoUpPeaks_AME_df)

# convert matrices of -log10(padj) values to dataframe for writexl. 
Palbo_Doxo_AME_UpPeaks_Mat <- data.frame(motif_ID=row.names(Palbo_Doxo_AME_UpPeaks_Mat), Palbo_Doxo_AME_UpPeaks_Mat)
Palbo_Doxo_AME_DownPeaks_Mat <- data.frame(motif_ID=row.names(Palbo_Doxo_AME_DownPeaks_Mat), Palbo_Doxo_AME_DownPeaks_Mat)

Table_num <- 6
# Write to an Excel file with each table as a separate sheet
write_xlsx(list("MEME-AME Results Up Peaks"=annoUpPeaks_AME_df, "MEME-AME Results Down Peaks"=annoDownPeaks_AME_df, "-log10(padj)values Up Peaks"=Palbo_Doxo_AME_UpPeaks_Mat, "-log10(padj)values Down Peaks"=Palbo_Doxo_AME_DownPeaks_Mat), paste0(write_dir, "/Supplemental_Table", Table_num, "_ATAC_MEMEAME_Motif_enrichment_results_Dynamic_Peaks.xlsx"))
```

# Supplemental table of differentially enriched motifs between drugs in up ATAC-seq peaks
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="import_differential_MEMEAME_results")

annoUp_PvsD_AME_mat <- create_AME_peaks_matrix(annoUp_PvsD_AME)
annoUp_DvsP_AME_mat <- create_AME_peaks_matrix(annoUp_DvsP_AME)

annoUp_PvsD_AME <- annoUp_PvsD_AME[ ,!colnames(annoUp_PvsD_AME) %in% c("rank", "motif_DB", "motif_alt_ID", "Group")]
annoUp_DvsP_AME <- annoUp_DvsP_AME[ ,!colnames(annoUp_DvsP_AME) %in% c("rank", "motif_DB", "motif_alt_ID", "Group")]
colnames(annoUp_PvsD_AME) <- c("motif_ID", "consensus", "p-value", "adj-pvalue", "E-value", "tests", "FASTA_max", "pos", "neg", "PWM_min", "True_Positive(TP)", "percent_TP", "False_Positive(FP)", "percent_FP", "Percent_TP - Percent_FP", "-log10(padj)value", "Comparison")
colnames(annoUp_DvsP_AME) <- colnames(annoUp_PvsD_AME)

# rename column names (Comparisons) to something more readable. 
rename_Comparison_df <- data.frame(original_names=colnames(annoUp_PvsD_AME_mat), new_names=c("day3Palbo.vs.day3Doxo_UpPeaks.vs.Cycling", "day10Palbo.vs.day10Doxo_UpPeaks.vs.Cycling", "day14Palbo.vs.day14Doxo_UpPeaks.vs.Cycling", "day21Palbo.vs.day21Doxo_UpPeaks.vs.Cycling", "day28Palbo.vs.day21Doxo_UpPeaks.vs.Cycling", "day10Palbo.vs.day10Doxo_UpPeaks.vs.day3", "day14Palbo.vs.day14Doxo_UpPeaks.vs.day3", "day21Palbo.vs.day21Doxo_UpPeaks.vs.day3", "day28Palbo.vs.day21Doxo_UpPeaks.vs.day3")
                                   )
colnames(annoUp_PvsD_AME_mat) <- rename_Comparison_df$new_names
colnames(annoUp_DvsP_AME_mat) <- rename_Comparison_df$new_names
annoUp_PvsD_AME$Comparison <- rename_Comparison_df[match(annoUp_PvsD_AME$Comparison, rename_Comparison_df$original_names), ]$new_names
annoUp_DvsP_AME$Comparison <- rename_Comparison_df[match(annoUp_DvsP_AME$Comparison, rename_Comparison_df$original_names), ]$new_names

# convert matrices of -log10(padj) values to dataframe for writexl. 
annoUp_PvsD_AME_mat <- data.frame(motif_ID=row.names(annoUp_PvsD_AME_mat), annoUp_PvsD_AME_mat)
annoUp_DvsP_AME_mat <- data.frame(motif_ID=row.names(annoUp_DvsP_AME_mat ), annoUp_DvsP_AME_mat)

Table_num <- 7
# Write to an Excel file with each table as a separate sheet
write_xlsx(list("More enriched in Palbo"=annoUp_PvsD_AME, "More enriched in Doxo"=annoUp_DvsP_AME, "-log10(padj)-More enrichedPalbo"=Palbo_Doxo_AME_UpPeaks_Mat, "-log10(padj)-More enrichedDoxo"=Palbo_Doxo_AME_DownPeaks_Mat), paste0(write_dir, "/Supplemental_Table", Table_num, "_ATAC_MEMEAME_differential_enriched_motifs_between_drugs_Up_Peaks.xlsx"))
```
# Supplemental table of MEME-AME motifs enriched under nearby dynamic ATAC-seq peaks to RNA Cluster
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment.Rmd", chunk_name="import nearby peaks kmeans 7 RNA clusters")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment.Rmd", chunk_name="convert AME df to matrix")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment.Rmd", chunk_name="merged_nearby_enriched_motifs_mat") 

Cluster_anno_df <- data.frame(original_names=c("CellCycle1_genes", "CellCycle2_genes", "Palbo1_genes", "Palbo2_genes", "Doxo1_genes", "Doxo2_genes", "Shared_genes") , new_names=c("Shared down (RNA Cluster 1)", "Shared down (RNA Cluster 2)", "Palbo up (RNA Cluster 3)", "Palbo up (RNA Cluster 4)", "Doxo up (RNA Cluster 5)", "Doxo up (RNA Cluster 6)", "Shared up (RNA Cluster 7)" ))

Peaks_AME_Clusters$Cluster <- Cluster_anno_df[match(Peaks_AME_Clusters$Cluster, Cluster_anno_df$original_names), ]$new_names

Peaks_AME_Clusters <- Peaks_AME_Clusters[ ,!colnames(Peaks_AME_Clusters) %in% c("rank", "motif_DB", "motif_alt_ID")]
colnames(Peaks_AME_Clusters) <- c("motif_ID", "consensus", "p-value", "adj-pvalue", "E-value", "tests", "FASTA_max", "pos", "neg", "PWM_min", "True_Positive(TP)", "percent_TP", "False_Positive(FP)", "percent_FP", "Percent_TP - Percent_FP", "-log10(padj)value", "RNA_Cluster", "Comparison" , "Direction", "Group")

colnames(merged_Up_AME_mat) <- c(paste0(Cluster_anno_df$new_names, "_nearby_Up.Doxo.Peaks"), paste0(Cluster_anno_df$new_names, "_nearby_Up.Palbo.Peaks"))
colnames(merged_Down_AME_mat) <- colnames(merged_Up_AME_mat) 

# convert matrices of -log10(padj) values to dataframe for writexl. 
merged_Up_AME_mat <- data.frame(motif_ID=row.names(merged_Up_AME_mat), merged_Up_AME_mat)
merged_Down_AME_mat <- data.frame(motif_ID=row.names(merged_Down_AME_mat), merged_Down_AME_mat)

Table_num <- 8
# Write to an Excel file with each table as a separate sheet
write_xlsx(list("MEME-AME Results"=Peaks_AME_Clusters, "-log10(padj)-nearby.Up.Peaks"=merged_Up_AME_mat, "-log10(padj)-nearby.Down.Peaks" =merged_Down_AME_mat), paste0(write_dir, "/Supplemental_Table", Table_num, "_ATAC_MEMEAME__enriched_motifs_in_nearby_RNA_Clusters_dynamic_peaks.xlsx"))
```

# GEO to do list: 
# annotated master peak set
# MACS2 narrowPeak file per tech rep 
# merged bigwigs 
# merged bams



