---
title: "Supplemental_Tables"
author: "Joanna Yeung"
date: '2025-03-22'
output: html_document
---
```{r}
# data wrangling
library(preprocessCore)
library(readr)
library(readxl)
library(stringr)
library(dplyr)
library(writexl)

# plotting
library(gridExtra)
library(ggplot2)
library(pheatmap)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(scales)
library(ggrepel)

# genomics data wrangling and analysis
library(rtracklayer)
library(Biostrings)
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(DESeq2)
library(apeglm)
library(ChIPseeker)
library(clusterProfiler)
library(msigdbr)
library(MotifDb)
library(motifmatchr)

# genome annotation databases
library(EnsDb.Hsapiens.v79)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(BSgenome.Hsapiens.UCSC.hg38)
library(org.Hs.eg.db)
```

```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Global_transcriptome_and_chromatin_accessibility_is_dynamic_in_both_types_of_therapy_induced_senescence/workspace.RData")
```

# make function to specify code chunk to run from other Rmd file to avoid copying and pasting large lines of code. 
```{r}
# arguments: 
# Rmd_file=file path & name of Rmd file with code chunk
# chunk_name= name of code chunk to run from Rmd file. 
run_code_chunk <- function(Rmd_file, chunk_name){ 
  # Read the Rmd file
lines <- readLines(Rmd_file)

# Identify the start and end of the specific chunk
chunk_start <- grep(paste0("^```\\{r\\s*", chunk_name), lines)  # Find chunk start
chunk_end <- grep("^```$", lines)  # Find all chunk end lines

# Find the first `chunk_end` after `chunk_start`
chunk_end <- chunk_end[chunk_end > chunk_start][1]

  if (is.na(chunk_end)) {
    stop(paste("End of chunk", chunk_name, "not found. Check the Rmd file for missing delimiters."))
  }

# specify lines inside chunk, removing lines that make up the start and end of the chunk. 
chunk_to_run <- lines[(chunk_start[1]+1):(chunk_end[1]-1)]

# Evaluate the extracted code
  eval(parse(text = paste(chunk_to_run, collapse = "\n")), envir = .GlobalEnv)
}
```

# RNA Table S2: make supplemental table of log2FC & padj values of expressed genes. Include information on genes that were considered significantly changing. 
```{r}
IPA_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/output"
IPA_upstream_regs <- list()
IPA_pathways <- list()
# read in information for enriched IPA pathways per hierarchical cluster (clustering based on scaled rlog normalized counts)
for(i in 1:length(cluster_names)){
IPA_pathways[[i]]<- read_tsv(paste0(IPA_dir, "/", cluster_names[i], "_IPA_pathways.txt"), skip=2)
# read in information for enriched upstream regulators 
IPA_upstream_regs[[i]] <- read.table(paste0(IPA_dir, "/", cluster_names[i], "_IPA_upstream_regs.txt"), sep="\t", skip=2, header=T)
IPA_upstream_regs[[i]] <- IPA_upstream_regs[[i]][!grepl("mir|miR|Mir", IPA_upstream_regs[[i]]$Upstream.Regulator), ] # remove upstream regulators that are miRNAs (uncertain if they will be informative & there is a lot of them)
}

names(IPA_upstream_regs) <- cluster_names
names(IPA_pathways) <- cluster_names


run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="log2FC sig genes matrix")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling.Rmd", chunk_name="p53 NFKB databases")

# filter for E2F target genes from IPA & Hallmark E2F Targets pathway term. 
E2F_Targets=unique(c(unlist(lapply(IPA_upstream_regs, function(x){str_split(x[grep("E2F", x$Upstream.Regulator), ]$Target.Molecules.in.Dataset, ",")})), 
              msigdbr_df_H[msigdbr_df_H$gs_name %in% "HALLMARK_E2F_TARGETS", ]$gene_symbol
))

# add annotation to significant genes based on if they are E2F targets. 
clusterDF$E2F <- ifelse(clusterDF$gene_id %in% E2F_Targets, "Yes", "No") # annotate sig genes that are E2F targets

# add ENSG id
clusterDF$ENSG_id <- row.names(LFCResultsall_RNA[[1]][match(clusterDF$gene_id, LFCResultsall_RNA[[1]]$SYMBOL), ])
# add in drug category column 
clusterDF$RNA_Cluster_Description <- ifelse(clusterDF$Clusters %in% c("2", "5"), "Shared downregulated", ifelse(clusterDF$Clusters %in% c("7", "1"), "Palbo upregulated", ifelse(clusterDF$Clusters %in% c("4", "6"), "Doxo upregulated", "Shared upregulated")))

# reformat number assigned by kmeans clustering to 1:7 as indicated in the figures. 
kmeansCluster_to_RNACluster <- data.frame(Cluster=levels(clusterDF$Clusters), RNA_cluster=factor(1:7))
clusterDF$RNA_Cluster <- kmeansCluster_to_RNACluster[match(clusterDF$Clusters, kmeansCluster_to_RNACluster$Cluster), ]$RNA_cluster

# change column names for annotation of significantly changing genes. 
clusterDF <- clusterDF[ ,c("Clusters", "RNA_Cluster", "RNA_Cluster_Description", "gene_id", "ENSG_id", "SASP", "Cell_Cycle", "NFKB_All", "p53_All", "E2F")]
colnames(clusterDF) <- c("Kmeans_assigned_Cluster_number", "RNA_Cluster", "RNA_Cluster_Description", "gene_symbol", "gene_id", "SASP", "Cell_Cycle", "NFKB_target", "p53_target", "E2F_target")

# get log2FC values of all expressed genes over time 
LFCMat <- matrix(ncol=length(LFCResultsall_RNA), nrow=nrow(LFCResultsall_RNA[[1]]))
colnames(LFCMat) <- names(LFCResultsall_RNA)
row.names(LFCMat) <- row.names(LFCResultsall_RNA[[1]])
for(i in 1:ncol(LFCMat)){
LFCMat[ ,i] <- LFCResultsall_RNA[[i]]$log2FoldChange
}

# add in columns for padj values from DESeq LRT Test and gene symbol.
Table_S2 <- as.data.frame(LFCMat)
row.names(acrossGroups) <- gsub('\\.[0-9]*$', '', row.names(acrossGroups) )
Table_S2 <- cbind(Table_S2, acrossGroups[match(row.names(Table_S2), row.names(acrossGroups)), c("padj", "SYMBOL")])

# Supplemental table of RNA databases of target genes
target_genes_db <- list(SASP_genes=unique(SASP_all), 
                        CellCycle_genes=unique(c(msigdbr_df_H[msigdbr_df_H$gs_name %in% c("HALLMARK_E2F_TARGETS", "HALLMARK_G2M_CHECKPOINT"), ]$gene_symbol, msigdbr_df_REACT[msigdbr_df_REACT$gs_name %in% c("REACTOME_CELL_CYCLE", "REACTOME_CELL_CYCLE_CHECKPOINTS", "REACTOME_CELL_CYCLE_MITOTIC"), ]$gene_symbol, msigdbr_df_KEGG[msigdbr_df_KEGG$gs_name %in% "KEGG_CELL_CYCLE", ]$gene_symbol)),
                        NFKB_targets=unique(c(Gilmore_NFKB_Targets, NFKB_Hallmark, IPA_targets$NFKB)),
                        p53_targets=unique(c(Fischer_Oncogene_2017_p53_census_target_genes$`Gene Symbol`, p53_Hallmark)),
                        E2F_targets=unique(E2F_Targets)
)

# add in columns for which genes are target genes and SASP/CellCycle genes. 
for(i in 1:length(target_genes_db)) {
  Table_S2[[names(target_genes_db)[i]]] <- ifelse(Table_S2$SYMBOL %in% target_genes_db[[i]], "Yes", "No")
}

# add in columns containing info on significantly changing genes. 
Table_S2 <- cbind(Table_S2, clusterDF[match(row.names(Table_S2), clusterDF$gene_id), c("Kmeans_assigned_Cluster_number", "RNA_Cluster", "RNA_Cluster_Description")])
# add ensg id type as column
Table_S2$gene_id <- row.names(Table_S2)

# change column names for counts & log2FC matrix of sig changing genes. 
# match order of sigMat
metaDataall <- metaDataall[match(colnames(LFCMat), row.names(metaDataall)), ]

new_LFCnames <- gsub("(Doxo|Palbo)(\\d+) vs (Doxo|Palbo|Cycling)(\\d*)", 
                 "day\\2 \\1 vs day\\4 \\3", colnames(LFCMat))

# Remove "day " before "Cycling" since it shouldn't have a day value
new_LFCnames <- gsub("day0 Cycling", "Cycling", new_LFCnames)
new_LFCnames <- paste(new_LFCnames, "log2FC")

# rename columns
colnames(Table_S2) <- c(new_LFCnames, "padj (DESeq LRT)", "gene_symbol", colnames(Table_S2)[23:31])
# reorder columns
Table_S2 <- Table_S2[ ,c(31,22,1:21, 23:30)]
write_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Tables"
write_xlsx(TableS2, paste0(write_dir, "/Table", Table_num, "_RNA_information_on_expressed_genes.xlsx"))
```


# Supplemental table of IPA upstream regulators & GO enrichment results from RNA-seq data (based on gene IDs input)
```{r }
universe <- genes_anno_type[match(gsub('\\.[0-9]*$', '', row.names(acrossGroups)), genes_anno_type$ensembl_id), ]

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="getGOenrichment function")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="genes_in_clusters")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="RNA-seq GO enrichment analysis")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name=" RNAseq GO enrichment results")

names(GO_results_list) <- c("Shared down (RNA Cluster 1)", "Shared down (RNA Cluster 2)", "Palbo up (RNA Cluster 3)", "Palbo up (RNA Cluster 4)", "Doxo up (RNA Cluster 5)", "Doxo up (RNA Cluster 6)", "Shared up (RNA Cluster 7)" )

for(i in 1:length(IPA_upstream_regs)){
IPA_upstream_regs[[i]]$RNA_Cluster <- names(GO_results_list)[i]
}


GO_results_list <- lapply(seq_along(GO_results_list), function(i) {
  x <- GO_results_list[[i]]
  x$RNA_Cluster <- names(GO_results_list)[i]  # Assign the correct name
  x <- x[ , !colnames(x) %in% c("Description", "Cluster")]  # Avoid dropping to vector
  return(x)
})


GO_results_df <- do.call(rbind, GO_results_list)
IPA_upstream_regs_df <- do.call(rbind, IPA_upstream_regs)

Table_num <- "S3"
# Write to an Excel file with each table as a separate sheet (will change it after downloading so that both tables are side by side in 1 sheet)
write_xlsx(list("GO_enrichment_results"=GO_results_df, "Predicted_Upstream_Regulators"=IPA_upstream_regs_df), paste0(write_dir, "/Table", Table_num, "_RNA_GO_enrichment_IPA_upstream_regulators_results.xlsx"))
```

# Supplemental table of ATAC motif enrichment results on up/down peaks at each timepoint. 
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="annoUpPeaks_AME_df")
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="annoDownPeaks_AME_df")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks.Rmd", chunk_name="create_AME_peaks_matrix")
 
Palbo_Doxo_AME_UpPeaks_Mat <- create_AME_peaks_matrix(annoUpPeaks_AME_df)

 set.seed(123)  # For reproducibility
  distance_matrix <- dist(Palbo_Doxo_AME_UpPeaks_Mat)
    hc <- hclust(distance_matrix, method = "complete")
    
    # Order the rows based on hierarchical clustering
    order <- hc$order
    
    # Reorder the original data matrix
    Palbo_Doxo_AME_UpPeaks_Mat <- Palbo_Doxo_AME_UpPeaks_Mat[order, ]
    
Palbo_Doxo_AME_DownPeaks_Mat <- create_AME_peaks_matrix(annoDownPeaks_AME_df)

 set.seed(123)  # For reproducibility
  distance_matrix <- dist(Palbo_Doxo_AME_DownPeaks_Mat)
    hc <- hclust(distance_matrix, method = "complete")
    
    # Order the rows based on hierarchical clustering
    order <- hc$order
    
    # Reorder the original data matrix
    Palbo_Doxo_AME_DownPeaks_Mat <- Palbo_Doxo_AME_DownPeaks_Mat[order, ]

    # rename columns to make more sense and remove unnecessary columns
annoUpPeaks_AME_df <- annoUpPeaks_AME_df[ ,!colnames(annoUpPeaks_AME_df) %in% c("rank", "motif_DB", "motif_alt_ID")]
annoDownPeaks_AME_df <- annoDownPeaks_AME_df[ ,!colnames(annoDownPeaks_AME_df) %in% c("rank", "motif_DB", "motif_alt_ID")]
colnames(annoUpPeaks_AME_df) <- c("motif_ID", "consensus", "p-value", "adj-pvalue", "E-value", "tests", "FASTA_max", "pos", "neg", "PWM_min", "True_Positive(TP)", "percent_TP", "False_Positive(FP)", "percent_FP", "Percent_TP - Percent_FP", "-log10(padj)value", "Comparison", "Group", "Timepoint")
colnames(annoDownPeaks_AME_df) <- colnames(annoUpPeaks_AME_df)
# rename Group to be clear which direction peaks are. 
annoUpPeaks_AME_df$Group <- paste("Up", annoUpPeaks_AME_df$Group, "ATAC-seq peaks")
annoDownPeaks_AME_df$Group <- paste("Down", annoDownPeaks_AME_df$Group, "ATAC-seq peaks")

# convert matrices of -log10(padj) values to dataframe for writexl. 
Palbo_Doxo_AME_UpPeaks_Mat <- data.frame(motif_ID=row.names(Palbo_Doxo_AME_UpPeaks_Mat), Palbo_Doxo_AME_UpPeaks_Mat, Group="Up ATAC-seq peaks")
Palbo_Doxo_AME_DownPeaks_Mat <- data.frame(motif_ID=row.names(Palbo_Doxo_AME_DownPeaks_Mat), Palbo_Doxo_AME_DownPeaks_Mat, Group="Down ATAC-seq peaks")

# rbind dataframes so that everything is on 1 excel spreadsheet. 
annoPeaks_AME_df <- rbind(annoUpPeaks_AME_df, annoDownPeaks_AME_df)
AME_Peaks_Mat <- rbind(Palbo_Doxo_AME_UpPeaks_Mat, Palbo_Doxo_AME_DownPeaks_Mat)
Table_num <- "S4"
# Write to an Excel file with each table as a separate sheet (will merge the two spreadsheets into 1 manually after downloading)
write_xlsx(list("MEME-AME Results"=annoPeaks_AME_df, "-log10(padj)values MEME-AME"=AME_Peaks_Mat), paste0(write_dir, "/Table", Table_num, "_ATAC_MEMEAME_Motif_enrichment_results_Dynamic_Peaks.xlsx"))
```

#Supplemental Table: activation scores from IPA_Comparison_Analysis on upstream regulators from RNA log2FC comparisons.
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name=" import_IPA_comparison_results")

# Assign names to your list of matrices
names(IPA_Compare_upstream_regs_mat) <- c(
  "Shared down (RNA Cluster 1)", 
  "Shared down (RNA Cluster 2)", 
  "Doxo up (RNA Cluster 5)", 
  "Doxo up (RNA Cluster 6)", 
  "Palbo up (RNA Cluster 3)", 
  "Palbo up (RNA Cluster 4)", 
  "Shared up (RNA Cluster 7)"
)

# Process each matrix in the list and add cluster label
IPA_Compare_upstream_regs_df <- lapply(names(IPA_Compare_upstream_regs_mat), function(name) {
  x <- IPA_Compare_upstream_regs_mat[[name]]
  
  # Rename columns
  new_LFCnames <- gsub("(Doxo|Palbo)(\\d+) vs (Doxo|Palbo|Cycling)(\\d*)", 
                       "day\\2 \\1 vs day\\4 \\3", colnames(x))
  new_LFCnames <- gsub("day  Cycling", "Cycling", new_LFCnames)  # clean up extra space
  colnames(x) <- new_LFCnames

  # Convert to dataframe with upstream_regulator as a column
  df <- data.frame(upstream_regulator = rownames(x), x)
  
  # Add the RNA_Cluster label
  df[["RNA_Cluster"]] <- name
  return(df)
})


IPA_Compare_upstream_regs_df <- IPA_Compare_upstream_regs_df[c(1,2,5,6,3,4,7)]
IPA_Compare_upstream_regs_df <- do.call(rbind, IPA_Compare_upstream_regs_df)
Table_num <- "S5"
# Write to an Excel file with each table as a separate sheet
write_xlsx(IPA_Compare_upstream_regs_df, paste0(write_dir, "/Table", Table_num, "_RNA_Comparison_Analysis_upstream_regulators_activation_scores.xlsx"))
```
# Supplemental table of differentially enriched motifs between drugs in up ATAC-seq peaks
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="import_differential_MEMEAME_results")

annoUp_PvsD_AME_mat <- create_AME_peaks_matrix(annoUp_PvsD_AME)
annoUp_DvsP_AME_mat <- create_AME_peaks_matrix(annoUp_DvsP_AME)

annoUp_PvsD_AME <- annoUp_PvsD_AME[ ,!colnames(annoUp_PvsD_AME) %in% c("rank", "motif_DB", "motif_alt_ID", "Group")]
annoUp_DvsP_AME <- annoUp_DvsP_AME[ ,!colnames(annoUp_DvsP_AME) %in% c("rank", "motif_DB", "motif_alt_ID", "Group")]
colnames(annoUp_PvsD_AME) <- c("motif_ID", "consensus", "p-value", "adj-pvalue", "E-value", "tests", "FASTA_max", "pos", "neg", "PWM_min", "True_Positive(TP)", "percent_TP", "False_Positive(FP)", "percent_FP", "Percent_TP - Percent_FP", "-log10(padj)value", "Comparison")
colnames(annoUp_DvsP_AME) <- colnames(annoUp_PvsD_AME)

# rename column names (Comparisons) to something more readable. 
rename_Comparison_df <- data.frame(original_names=colnames(annoUp_PvsD_AME_mat),
                                   new_namesPvsD=c("day3Palbo.vs.day3Doxo_UpPeaks.vs.Cycling", "day10Palbo.vs.day10Doxo_UpPeaks.vs.Cycling", "day14Palbo.vs.day14Doxo_UpPeaks.vs.Cycling", "day21Palbo.vs.day21Doxo_UpPeaks.vs.Cycling", "day28Palbo.vs.day21Doxo_UpPeaks.vs.Cycling", "day10Palbo.vs.day10Doxo_UpPeaks.vs.day3", "day14Palbo.vs.day14Doxo_UpPeaks.vs.day3", "day21Palbo.vs.day21Doxo_UpPeaks.vs.day3", "day28Palbo.vs.day21Doxo_UpPeaks.vs.day3"), 
                                   new_namesDvsP=c("day3Doxo.vs.day3Palbo_UpPeaks.vs.Cycling", "day10Doxo.vs.day10Palbo_UpPeaks.vs.Cycling", "day14Doxo.vs.day14Palbo_UpPeaks.vs.Cycling", "day21Doxo.vs.day21Palbo_UpPeaks.vs.Cycling", "day21Doxo.vs.day28Palbo_UpPeaks.vs.Cycling", "day10Doxo.vs.day10Palbo_UpPeaks.vs.day3", "day14Doxo.vs.day14Palbo_UpPeaks.vs.day3", "day21Doxo.vs.day21Palbo_UpPeaks.vs.day3", "day21Doxo.vs.day28Palbo_UpPeaks.vs.day3")
)
colnames(annoUp_PvsD_AME_mat) <- rename_Comparison_df$new_namesPvsD
colnames(annoUp_DvsP_AME_mat) <- rename_Comparison_df$new_namesPvsD
annoUp_PvsD_AME$Comparison <- rename_Comparison_df[match(annoUp_PvsD_AME$Comparison, rename_Comparison_df$original_names), ]$new_namesPvsD
annoUp_DvsP_AME$Comparison <- rename_Comparison_df[match(annoUp_DvsP_AME$Comparison, rename_Comparison_df$original_names), ]$new_namesDvsP

# convert matrices of -log10(padj) values to dataframe for writexl. 
annoUp_PvsD_AME_mat <- data.frame(motif_ID=row.names(annoUp_PvsD_AME_mat), annoUp_PvsD_AME_mat, Group= "Motifs more enriched in Palbo-upregulated ATAC-seq peaks than Doxo-upregulated peaks")
annoUp_DvsP_AME_mat <- data.frame(motif_ID=row.names(annoUp_DvsP_AME_mat ), annoUp_DvsP_AME_mat, Group= "Motifs more enriched in Doxo-upregulated ATAC-seq peaks than Palbo-upregulated peaks")


annoUp_AME <- rbind(annoUp_PvsD_AME, annoUp_DvsP_AME)
annoUp_AME_mat <- rbind(annoUp_PvsD_AME_mat, annoUp_DvsP_AME_mat)
Table_num <- "S6"
# Write to an Excel file with each table as a separate sheet
write_xlsx(list("MEME-AME Results"=annoUp_AME , "-log10(padj)values MEME-AME"=annoUp_AME_mat), paste0(write_dir, "/Table", Table_num, "_ATAC_MEMEAME_differential_enriched_motifs_between_drugs_Up_Peaks.xlsx"))
```
# Supplemental table of MEME-AME motifs enriched under nearby dynamic ATAC-seq peaks to RNA Cluster
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment.Rmd", chunk_name="import nearby peaks kmeans 7 RNA clusters")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment.Rmd", chunk_name="convert AME df to matrix")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment.Rmd", chunk_name="merged_nearby_enriched_motifs_mat") 

Cluster_anno_df <- data.frame(original_names=c("CellCycle1_genes", "CellCycle2_genes", "Palbo1_genes", "Palbo2_genes", "Doxo1_genes", "Doxo2_genes", "Shared_genes") , new_names=c("Shared down (RNA Cluster 1)", "Shared down (RNA Cluster 2)", "Palbo up (RNA Cluster 3)", "Palbo up (RNA Cluster 4)", "Doxo up (RNA Cluster 5)", "Doxo up (RNA Cluster 6)", "Shared up (RNA Cluster 7)" ))

Peaks_AME_Clusters$Cluster <- Cluster_anno_df[match(Peaks_AME_Clusters$Cluster, Cluster_anno_df$original_names), ]$new_names

Peaks_AME_Clusters <- Peaks_AME_Clusters[ ,!colnames(Peaks_AME_Clusters) %in% c("rank", "motif_DB", "motif_alt_ID")]
colnames(Peaks_AME_Clusters) <- c("motif_ID", "consensus", "p-value", "adj-pvalue", "E-value", "tests", "FASTA_max", "pos", "neg", "PWM_min", "True_Positive(TP)", "percent_TP", "False_Positive(FP)", "percent_FP", "Percent_TP - Percent_FP", "-log10(padj)value", "RNA_Cluster", "Comparison" , "Direction", "Group")

colnames(merged_Up_AME_mat) <- c(paste0(Cluster_anno_df$new_names, "_nearby_Up.Doxo.Peaks"), paste0(Cluster_anno_df$new_names, "_nearby_Up.Palbo.Peaks"))
colnames(merged_Down_AME_mat) <- colnames(merged_Up_AME_mat) 

# convert matrices of -log10(padj) values to dataframe for writexl. 
merged_Up_AME_mat <- data.frame(motif_ID=row.names(merged_Up_AME_mat), merged_Up_AME_mat)
merged_Down_AME_mat <- data.frame(motif_ID=row.names(merged_Down_AME_mat), merged_Down_AME_mat)

Table_num <- "S7"
# Write to an Excel file with each table as a separate sheet
write_xlsx(list("MEME-AME Results"=Peaks_AME_Clusters, "-log10(padj)-nearby.Up.Peaks"=merged_Up_AME_mat, "-log10(padj)-nearby.Down.Peaks" =merged_Down_AME_mat), paste0(write_dir, "/Table", Table_num, "_ATAC_MEMEAME__enriched_motifs_in_nearby_RNA_Clusters_dynamic_peaks.xlsx"))
```

# GEO supplemental files
```{r}
# all genes counts matrix, normalized counts 
all_genes_counts_matrix <- counts(RNAddsall, normalize=T)
colnames(all_genes_counts_matrix) <-  paste0("Biorep", metaDataall$Biorep, "_", metaDataall$Treatment, "_day", metaDataall$Timepoint)
all_genes_counts_matrix <- all_genes_counts_matrix[ ,c(1,9,15,2,10,16,3,17,4,11,5,12,18,6,19,7,13,8,14)]
# export all genes counts matrix for GEO submission
write.csv(all_genes_counts_matrix, file="/lustre/fs4/risc_lab/store/jyeung/Senescence_Liposarcoma_GEO_submission/RNAseq/Normalized_by_DESeq2_RNA_counts_matrix.csv")
```
```{r}
# counts matrix for ATAC-seq peaks 
ATAC_normCounts_matrix <- counts(dds_filt, normalized=T)
ATAC_normCounts_matrix <- ATAC_normCounts_matrix[, c(1:5, 38, 40, 39, 41, 42, 28:32, 43:45, 6:7, 49:51, 11:15, 33:37, 46:48, 8:10, 52:54, 16:27)]
colnames(ATAC_normCounts_matrix) <- c("Cycling_Biorep1.1","Cycling_Biorep1.2","Cycling_Biorep2.1","Cycling_Biorep2.2","Cycling_Biorep2.3","Cycling_Biorep3.2","Cycling_Biorep3.3","Cycling_Biorep4.1","Cycling_Biorep4.2","Cycling_Biorep4.3","day3_Doxo_Biorep1.1","day3_Doxo_Biorep1.2","day3_Doxo_Biorep2.1","day3_Doxo_Biorep2.2","day3_Doxo_Biorep2.3","day10_Doxo_Biorep1.1","day10_Doxo_Biorep1.2","day10_Doxo_Biorep1.3","day14_Doxo_Biorep2.1","day14_Doxo_Biorep2.2","day14_Doxo_Biorep1.1","day14_Doxo_Biorep1.2","day14_Doxo_Biorep1.3","day21_Doxo_Biorep1.1","day21_Doxo_Biorep1.2","day21_Doxo_Biorep1.3","day21_Doxo_Biorep2.1","day21_Doxo_Biorep2.2","day3_Palbo_Biorep1.1","day3_Palbo_Biorep1.2","day3_Palbo_Biorep2.1","day3_Palbo_Biorep2.2","day3_Palbo_Biorep2.3","day10_Palbo_Biorep1.1","day10_Palbo_Biorep1.2","day10_Palbo_Biorep1.3","day14_Palbo_Biorep2.1","day14_Palbo_Biorep2.2","day14_Palbo_Biorep2.3","day14_Palbo_Biorep1.1","day14_Palbo_Biorep1.2","day14_Palbo_Biorep1.3","day21_Palbo_Biorep1.1","day21_Palbo_Biorep1.2","day21_Palbo_Biorep1.3","day21_Palbo_Biorep2.1","day21_Palbo_Biorep2.2","day21_Palbo_Biorep2.3","day28_Palbo_Biorep1.1","day28_Palbo_Biorep1.2","day28_Palbo_Biorep1.3","day28_Palbo_Biorep2.1","day28_Palbo_Biorep2.2","day28_Palbo_Biorep2.3")

# export all genes counts matrix for GEO submission
write.csv(ATAC_normCounts_matrix, file="/lustre/fs4/risc_lab/store/jyeung/Senescence_Liposarcoma_GEO_submission/ATACseq/Normalized_by_DESeq2_ATAC_counts_matrix.csv")
```

