title: "Figure S3: Signatures of enriched transcriptional regulators in therapy induced senescence, related to Figure 2"
author: "Joanna Yeung"
date: '2025-04-23'
output: html_document

```{r}
library(readr)
library(GenomicFeatures)
library(DESeq2)
library(apeglm)
library(gridExtra)
library(ggplot2)
library(pheatmap)
library(msigdbr)
library(stringr)
library(org.Hs.eg.db)
library(ggrepel)
library(rtracklayer)
library(cowplot)
library(readxl)
library(EnsDb.Hsapiens.v79)
library(viridis)
library(GenomicRanges)
library(rtracklayer)
library(RColorBrewer)
library(cowplot)
library(preprocessCore)
library(scales)
library(ChIPseeker)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)
library(BSgenome.Hsapiens.UCSC.hg38)
library(MotifDb)
library(motifmatchr)
library(Biostrings)
library(clusterProfiler)
library(pheatmap)
```

# annotation colors for pheatmap
```{r heatmap_colors}
hm_cols_RNA <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("Cycling"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD"), 
                Biorep=c("1"="orange", "2"="light blue", "3"="pink"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "DoxovsPalbo"="#CCEBC5"), 
                Clusters=c("2"="#66C2A5","5"="#A6D854","7"="#8DA0CB", "1"="#FC8D62", "4"="#E78AC3","6"="#FFD92F","3"="#E5C494"), 
                SASP = c("Yes"="dark green", "No"="white"),
                Cell_Cycle = c("Yes"="dark blue", "No"="white"), 
                p53_All=c("Yes"="#00B0F0", "No"="white"),
                NFKB_All=c("Yes"="red", "No"="white"))


hm_cols_ATAC <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("None"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "PvsD"="#CCEBC5"),
                Biorep=c("1"="orange", "2"="light blue", "3"="pink", "4"="light green"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "PvsD"="#CCEBC5"),
                Peak_anno=c("Distal"="#B15A28", "Gene Body"="#7F54A2", "Promoter"="#A7CEE2"), 
                Doxo=c("Yes"="#B3C543", "No"="white"),
                Palbo=c("Yes"="#E0ADCD", "No"="white")
                )
```

# load in workspace with DESeq results and rlog counts matrix of genes. 
```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure2/workspace.RData")
```

# function to get sequence under peaks for input into MEME-AME for motif enrichment analysis
```{r}
getSeqforMEME <- function(peakset, dir, filename){seq <- getSeq(BSgenome.Hsapiens.UCSC.hg38, peakset) #get sequence info under peaks
names(seq) <- peakset$name
writeXStringSet(seq, file = paste0(dir, filename, ".fa")) #write to FASTA format for input into MEME
}
```

# ATAC-seq: create input fasta files for MEME-AME analysis on all up or down peaks. 
```{r}
directory <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq"
for(i in 1:length(annoUp)){
getSeqforMEME(annoUp[[i]], dir=paste0(directory, "/annoUp/"), filename=names(annoUp)[i])
getSeqforMEME(annoDown[[i]], dir=paste0(directory, "/annoDown/"), filename=names(annoDown)[i])
}

# specify file paths for input to MEME-AME analysis for up peaks
folderlist <- gsub(".fa", "", dir(paste0(directory, "/annoUp"))) # folder names

writeLines(dir(paste0(directory, "/annoUp"), full.names=T, pattern=".fa$"), con=paste0(directory, "/annoUp/inputfiles_annoUp_DESeq_peaks.txt")) # directory paths to fasta files of Up peaks

writeLines(folderlist, con=paste0(directory, "/annoUp/folderlist_annoUp_DESeq_peaks.txt")) # write folder names

# specify file paths for input to MEME-AME analysis for down peaks
folderlist <- gsub(".fa", "", dir(paste0(directory, "/annoDown"))) # folder names

writeLines(dir(paste0(directory, "/annoDown"), full.names=T, pattern=".fa$"), con=paste0(directory, "/annoDown/inputfiles_annoDown_DESeq_peaks.txt")) # directory paths to fasta files of Up peaks

writeLines(folderlist, con=paste0(directory, "/annoDown/folderlist_annoDown_DESeq_peaks.txt")) # write folder names
```

### MEME-AME analysis on DESeq_Peaks pairwise comparison 
```{bash}
cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq

# all up peaks- background peak set set to corresponding down peaks of the same DESeq comparison. 
sbatch -p risc,hpc --array=1-16 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq/annoDown/inputfiles_annoDown_DESeq_peaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq/annoUp/inputfiles_annoUp_DESeq_peaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_output/DESeq/annoUp,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq/annoUp/folderlist_annoUp_DESeq_peaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm

# all down peaks- background peak set set to corresponding up peaks of the same DESeq comparison. 
sbatch -p risc,hpc --array=1-16 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq/annoUp/inputfiles_annoUp_DESeq_peaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq/annoDown/inputfiles_annoDown_DESeq_peaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_output/DESeq/annoDown,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq/annoDown/folderlist_annoDown_DESeq_peaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm
```


# ATAC-seq: import motif enrichment results under up peaks across all timepoint comparisons & drug treatment.
```{r annoUpPeaks_AME_df}
# extract ids of all expressed genes
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL
expressedGenes[grep("TP53$|TP63$|TP73$", expressedGenes)] <- c("P63", "P73", "P53")
expressedGenes[expressedGenes %in% "RELA"] <- "TF65"

# load in process_motif_data function. 
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS6/Functions/process_motif_data.RData")

annoUpPeaks_AME_df <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_output/DESeq/annoUp", expressed_genes = expressedGenes)

annoUpPeaks_AME_df$Comparison <- gsub("\\..*", "", row.names(annoUpPeaks_AME_df))
annoUpPeaks_AME_df$Comparison <- factor(annoUpPeaks_AME_df$Comparison, levels=unique(annoUpPeaks_AME_df$Comparison)[c(15,1,5,9,2,6,10,16,3,7,11,13,4,8,12,14)]) # column to identify DESeq comparison

annoUpPeaks_AME_df$Group <- ifelse(grepl("DoxovsCycling", annoUpPeaks_AME_df$Comparison), "DoxovsCycling", ifelse(grepl("PalbovsCycling", annoUpPeaks_AME_df$Comparison), "PalbovsCycling", ifelse(grepl("Doxovsday3Doxo", annoUpPeaks_AME_df$Comparison), "Doxovsday3Doxo", "Palbovsday3Palbo"))) # column to identify relative to cycling or day 3 comparison and drug treatment. 

annoUpPeaks_AME_df$Timepoint <-  gsub("(day[0-9]{1,2}).*", "\\1", annoUpPeaks_AME_df$Comparison) # column to identify which timepoint is compared to Cycling or day 3 controls, by removing everything after day[0-9] where [0-9] is 1 or 2 digits in the Comparison column. 
annoUpPeaks_AME_df$Timepoint <- factor(annoUpPeaks_AME_df$Timepoint, levels=c("day3", "day10", "day14", "day21", "day28"))
```

# ATAC-seq: import motif enrichment results under down peaks across all timepoint comparisons & drug treatment. 
```{r annoDownPeaks_AME_df}
annoDownPeaks_AME_df <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_output/DESeq/annoDown", expressed_genes = expressedGenes)

annoDownPeaks_AME_df$Comparison <- gsub("\\..*", "", row.names(annoDownPeaks_AME_df))
annoDownPeaks_AME_df$Comparison <- factor(annoDownPeaks_AME_df$Comparison, levels=unique(annoDownPeaks_AME_df$Comparison)[c(15,1,5,9,2,6,10,16,3,7,11,13,4,8,12,14)]) # column to identify DESeq comparison

annoDownPeaks_AME_df$Group <- ifelse(grepl("DoxovsCycling", annoDownPeaks_AME_df$Comparison), "DoxovsCycling", ifelse(grepl("PalbovsCycling", annoDownPeaks_AME_df$Comparison), "PalbovsCycling", ifelse(grepl("Doxovsday3Doxo", annoDownPeaks_AME_df$Comparison), "Doxovsday3Doxo", "Palbovsday3Palbo"))) # column to identify relative to cycling or day 3 comparison and drug treatment. 

annoDownPeaks_AME_df$Timepoint <-  gsub("(day[0-9]{1,2}).*", "\\1", annoDownPeaks_AME_df$Comparison) # column to identify which timepoint is compared to Cycling or day 3 controls, by removing everything after day[0-9] where [0-9] is 1 or 2 digits in the Comparison column. 
annoDownPeaks_AME_df$Timepoint <- factor(annoDownPeaks_AME_df$Timepoint, levels=c("day3", "day10", "day14", "day21", "day28"))
```

# plot differential upstream regs based on thresholding of the largest differences between overall sum of activation scores between Doxo & Palbo instead of thresholding. 

# run code necessary for matching top activated/inhibited transcriptional regulators with ATAC-seq motif enrichment results from up peaks across time. 
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/Figure4.Rmd", chunk_name="import_IPA_comparison_results")
```
```{r}
mat <- IPA_Compare_upstream_regs_mat$Palbo2
diff_regs <- names(rev(sort(rowSums(mat[ ,6:10]) - rowSums(mat[ ,1:5]))))[1:20]
diff_regs <- diff_regs[!diff_regs %in% c("arsenic trioxide", "cadmium chloride", "VH032", "IOX2")]

#pdf(paste0(pdf_dir, "/RNA_IPA_Palbo2_diff_regs_bydifference.pdf"), width=3, height=3)
pheatmap(mat[row.names(mat) %in% diff_regs[1:12], ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")

mat <- IPA_Compare_upstream_regs_mat$Palbo1
diff_regs <- names(rev(sort(rowSums(mat[ ,6:10]) - rowSums(mat[ ,1:5]))))[1:20]
diff_regs <- diff_regs[!diff_regs %in% c("arsenic trioxide", "cadmium chloride", "VH032", "IOX2")]
```

----------------------------------------------------------------------
# *FIGURES*
# create function that stores the -log10(padj) value of each motif ID for each DESeq comparison
```{r create_AME_peaks_matrix}
create_AME_peaks_matrix <- function(sig_peaks_df) {
  # Create an empty matrix for significant peaks
  unique_motifs <- unique(sig_peaks_df$motif_ID)
  comparisons <- levels(sig_peaks_df$Comparison)
  sig_peaks_mat <- matrix(nrow = length(unique_motifs), ncol = length(comparisons))
  
  rownames(sig_peaks_mat) <- unique_motifs # Row names are motif IDs
  colnames(sig_peaks_mat) <- comparisons   # Column names are DESeq comparisons
  
  # Populate the matrix with neglog10padj values
  for (i in 1:ncol(sig_peaks_mat)) {
    temp_comparison <- sig_peaks_df[sig_peaks_df$Comparison == comparisons[i], ]
    sig_peaks_mat[, i] <- temp_comparison[match(rownames(sig_peaks_mat), temp_comparison$motif_ID), ]$neglog10padj
    sig_peaks_mat[, i] <- ifelse(is.na(sig_peaks_mat[, i]), 0, sig_peaks_mat[, i])
  }
  
  return(sig_peaks_mat)
}
save(create_AME_peaks_matrix, file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/Functions/create_AME_peaks_matrix.RData")
```

# create function to plot motif enrichment results as dotplot where the fill color indicates -log10(padj) values & size of dot indicates %TP-%FP
```{r create_AME_dotplots}
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=50, size_limit=4.5){
  AME_df <- AME_df %>% mutate(Comparison = Timepoint)
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient(name="-log10(padj)", 
                        low="light yellow", high="dark orange",
                        limits=c(0, fill_limit), 
                        oob=scales::squish) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=c(0, size_limit), 
                          range=c(1, 10)) +  # Adjust dot sizes
    theme_minimal() +
    ggtitle(Title) +
    scale_x_discrete(labels = function(x) AME_df$Timepoint[match(x, AME_df$Comparison)]) +  # to replace Comparison with Timepoint labels on the x-axis. 
    facet_wrap(~Group, nrow=1, scales="free") +
    theme(
      axis.text.x = element_text(size=8, angle = 90, hjust=1),  # Rotate x-axis labels
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    )+
    theme_classic() +
    labs(x="Comparison", y="motif ID")
}
```
```{r}
save(create_AME_dotplots, file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/Functions/create_AME_dotplots.RData")
```

# ATAC-seq: plot heatmap of -log10(padj) values of top 50 enriched motifs in up Doxo peaks & up Palbo peaks across time.
```{r UpPeaks_AME_mat}
Palbo_UpPeaks_AME_mat <- create_AME_peaks_matrix(annoUpPeaks_AME_df[grepl("Palbo", annoUpPeaks_AME_df$Comparison), ])

# create matrix of -log10(padj) values of enriched motifs in all Palbo comparisons. Order by overall most to least enriched motifs across timepoints relative to Cycling or day 3. 
Palbo_UpPeaks_AME_mat <- list(vsCycling=Palbo_UpPeaks_AME_mat[order(rowSums(Palbo_UpPeaks_AME_mat[ ,1:5]), decreasing=T), ], # # order results by timepoints relative to Cycling
                              vsday3=Palbo_UpPeaks_AME_mat[order(rowSums(Palbo_UpPeaks_AME_mat[ ,6:9]), decreasing=T), ]) # # order results by timepoints relative to day3
                                                      
# create matrix of -log10(padj) values of enriched motifs in all Doxo comparisons. Order by overall most to least enriched motifs across timepoints relative to Cycling or day 3. 
Doxo_UpPeaks_AME_mat <- create_AME_peaks_matrix(annoUpPeaks_AME_df[grepl("Doxo", annoUpPeaks_AME_df$Comparison), ])

row.names(Doxo_UpPeaks_AME_mat)[grep("P53|P63|P73", row.names(Doxo_UpPeaks_AME_mat))] <- c("TP53", "TP73", "TP63")
                                                      
Doxo_UpPeaks_AME_mat <- list(vsCycling=Doxo_UpPeaks_AME_mat[order(rowSums(Doxo_UpPeaks_AME_mat[ ,1:4]), decreasing=T), ], # order results by timepoints relative to Cycling
                              vsday3=Doxo_UpPeaks_AME_mat[order(rowSums(Doxo_UpPeaks_AME_mat[ ,5:7]), decreasing=T), ]) # order results by timepoints relative to day3

# create matrix of -log10(padj) values of top 50 most enriched motifs in Palbo & top 50 most enriched motifs in Doxo (several of these motifs are enriched in both drug treatments)
top35_motifs <- unique(c(
row.names(Doxo_UpPeaks_AME_mat$vsCycling)[1:60], row.names(Palbo_UpPeaks_AME_mat$vsCycling)[1:35], row.names(Doxo_UpPeaks_AME_mat$vsday3)[1:10], row.names(Palbo_UpPeaks_AME_mat$vsday3)[1:35]))
top35_motifs_Palbo_Doxo_AME_Mat <- create_AME_peaks_matrix(annoUpPeaks_AME_df[annoUpPeaks_AME_df$motif_ID %in% top35_motifs, ])

top35_motifs_Palbo_Doxo_AME_Mat <- top35_motifs_Palbo_Doxo_AME_Mat[order(rowSums(top35_motifs_Palbo_Doxo_AME_Mat), decreasing=T), ]# order by overall most to least enriched motifs

# plot top 35 overall most enriched motifs of each drug category in all up peaks as heatmap of -log10(padj) values
outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3"
pdf(paste0(outdir, "/UpPeaks_AME_matf_test.pdf"), width=4.5, height=11)
pheatmap(top35_motifs_Palbo_Doxo_AME_Mat , cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(7), breaks = c(0,6,10,20,50,100,200), main="top 35 enriched motifs in up peaks")
dev.off()

# plot top 35 overall most enriched motifs of each drug category in all up peaks as dotplots
pdf(paste0(outdir, "/UpPeaks_AME_dotplots.pdf"), width=10, height=13)
create_AME_dotplots(annoUpPeaks_AME_df[annoUpPeaks_AME_df$motif_ID %in% top35_motifs, ], Title="top 35 enriched motifs in up peaks", fill_limit=200, size_limit=51)
dev.off()
```

# ATAC-seq: plot heatmap of -log10(padj) values of top 50 enriched motifs in Down Doxo peaks & Down Palbo peaks across time. 
```{r}
Palbo_DownPeaks_AME_mat <- create_AME_peaks_matrix(annoDownPeaks_AME_df[grepl("Palbo", annoDownPeaks_AME_df$Comparison), ])

# create matrix of -log10(padj) values of enriched motifs in all Palbo comparisons. Order by overall most to least enriched motifs across timepoints relative to Cycling or day 3. 
Palbo_DownPeaks_AME_mat <- list(vsCycling=Palbo_DownPeaks_AME_mat[order(rowSums(Palbo_DownPeaks_AME_mat[ ,1:5]), decreasing=T), ], # # order results by timepoints relative to Cycling
                              vsday3=Palbo_DownPeaks_AME_mat[order(rowSums(Palbo_DownPeaks_AME_mat[ ,6:9]), decreasing=T), ]) # # order results by timepoints relative to day3
                                                      
# create matrix of -log10(padj) values of enriched motifs in all Doxo comparisons. Order by overall most to least enriched motifs across timepoints relative to Cycling or day 3. 
Doxo_DownPeaks_AME_mat <- create_AME_peaks_matrix(annoDownPeaks_AME_df[grepl("Doxo", annoDownPeaks_AME_df$Comparison), ])
                                                      
Doxo_DownPeaks_AME_mat <- list(vsCycling=Doxo_DownPeaks_AME_mat[order(rowSums(Doxo_DownPeaks_AME_mat[ ,1:4]), decreasing=T), ], # order results by timepoints relative to Cycling
                              vsday3=Doxo_DownPeaks_AME_mat[order(rowSums(Doxo_DownPeaks_AME_mat[ ,5:7]), decreasing=T), ]) # order results by timepoints relative to day3

# create matrix of -log10(padj) values of top 50 most enriched motifs in Palbo & top 50 most enriched motifs in Doxo (several of these motifs are enriched in both drug treatments)
top35_motifs <- unique(c(row.names(Doxo_DownPeaks_AME_mat$vsCycling)[1:60], row.names(Palbo_DownPeaks_AME_mat$vsCycling)[1:35], row.names(Doxo_DownPeaks_AME_mat$vsday3)[1:10], row.names(Palbo_DownPeaks_AME_mat$vsday3)[1:35]))
top35_motifs_Palbo_Doxo_AME_Mat <- create_AME_peaks_matrix(annoDownPeaks_AME_df[annoDownPeaks_AME_df$motif_ID %in% top35_motifs, ])

top35_motifs_Palbo_Doxo_AME_Mat <- top35_motifs_Palbo_Doxo_AME_Mat[order(rowSums(top35_motifs_Palbo_Doxo_AME_Mat), decreasing=T), ]# order by overall most to least enriched motifs

# plot top 35 overall most enriched motifs of each drug category in all Down peaks as heatmap of -log10(padj) values
outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3"
pdf(paste0(outdir, "/DownPeaks_AME_mat.pdf"), width=4.5, height=11)
pheatmap(top35_motifs_Palbo_Doxo_AME_Mat , cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 300,length.out = 100), main="top 35 enriched motifs in Down peaks")
dev.off()

# plot top 35 overall most enriched motifs of each drug category in all Down peaks as dotplots
pdf(paste0(outdir, "/DownPeaks_AME_dotplots.pdf"), width=10, height=13)
create_AME_dotplots(annoDownPeaks_AME_df[annoDownPeaks_AME_df$motif_ID %in% top35_motifs, ], Title="top 35 enriched motifs in Down peaks", fill_limit=300, size_limit=51)
dev.off()
```

# RNA-seq + ATAC-seq: instead of differentially activated upstream regs between drug treatments, plot top activated & top inhibited upstream Transcriptional regulators 
```{r}
# read in database matching upstream regulators to molecule type and target molecules. 
IPA_TFs_db <- list() 
db_files <- dir("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/IPA_files_kmeans7_clusters/output/sigGenes_log2FC/databases", pattern="transcription_regulators_database.txt", full.names = T)
for(i in 1:length(db_files)){
IPA_TFs_db[[i]] <- read_delim(db_files[i], 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE, skip = 2)
}
names(IPA_TFs_db) <- gsub("_.*", "", dir("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/IPA_files_kmeans7_clusters/output/sigGenes_log2FC/databases", pattern="transcription_regulators_database.txt"))

# function to filter for top upstream regulators and plot activation scores as heatmap
plot_upstream_reg_heatmaps <- function(mat, cluster_name, tf_db, zscore_cols, atac_peaks, pdf_dir) {
  # Ensure the output directory exists
  if (!dir.exists(pdf_dir)) {
    dir.create(pdf_dir, recursive = TRUE)
  }

  # Order the matrix by highest to lowest overall activation z-scores
  ordered_mat <- mat[order(rowSums(zscore_cols), decreasing = TRUE), ]

  # Filter by transcriptional regulators
  filt_mat <- ordered_mat[row.names(ordered_mat) %in% tf_db, ]

  # Define filtering conditions
  filter_conditions <- list(
    list(
      name = paste0("RNA_IPA_",cluster_name, "_top_act_inh_upstream_regs"),
      data = filt_mat
    ),
    list(
      name = paste0("RNA_IPA_" ,cluster_name, "_top_act_inh_filt_by_ATAC_upstream_regs"),
      data = filt_mat[row.names(filt_mat) %in% row.names(atac_peaks), ]
    )
  )

  # Loop through each filtering condition and generate heatmaps
  for (condition in filter_conditions) {
    filt_data <- condition$data

    # Get top activated & inhibited upstream transcriptional regulators
    top_regs <- c(row.names(head(filt_data)), row.names(tail(filt_data)))

    # Define the output PDF filename
    pdf_filename <- paste0(pdf_dir, "/", condition$name, ".pdf")

    # Generate heatmap and save to PDF
    pdf(pdf_filename, width = 3, height = 3)
    pheatmap(
      filt_data[row.names(filt_data) %in% top_regs, ],
      cluster_cols = FALSE,
      cluster_rows = FALSE,
      colorRampPalette(rev(brewer.pal(n = 7, name = "PuOr")))(100),
      breaks = seq(-4, 4, length.out = 100),
      main = "Activation scores"
    )
    dev.off()
  }
}
```

# RNA-seq + ATAC-seq: generate plots of top activated and inhibited upstream transcriptional regulators for each cluster of sig changing upregulated genes. -- note Fig S3B
```{r}
plot_upstream_reg_heatmaps(
  mat = IPA_Compare_upstream_regs_mat$Doxo1, 
  cluster_name = "Doxo1", 
  tf_db = IPA_TFs_db$Doxo1$`Upstream Regulator`, 
  zscore_cols = IPA_Compare_upstream_regs_mat$Doxo1[, 1:5], 
  atac_peaks = Doxo_UpPeaks_AME_mat, 
  pdf_dir = "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3"
)

plot_upstream_reg_heatmaps(
  mat = IPA_Compare_upstream_regs_mat$Doxo2, 
  cluster_name = "Doxo2", 
  tf_db = IPA_TFs_db$Doxo2$`Upstream Regulator`, 
  zscore_cols = IPA_Compare_upstream_regs_mat$Doxo2[, 1:5], 
  atac_peaks = Doxo_UpPeaks_AME_mat, 
  pdf_dir = "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3"
)

plot_upstream_reg_heatmaps(
  mat = IPA_Compare_upstream_regs_mat$Palbo1, 
  cluster_name = "Palbo1", 
  tf_db = IPA_TFs_db$Palbo1$`Upstream Regulator`, 
  zscore_cols = IPA_Compare_upstream_regs_mat$Palbo1[, 6:10], 
  atac_peaks = Palbo_UpPeaks_AME_mat, 
  pdf_dir = "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3"
)

plot_upstream_reg_heatmaps(
  mat = IPA_Compare_upstream_regs_mat$Palbo2, 
  cluster_name = "Palbo2", 
  tf_db = IPA_TFs_db$Palbo2$`Upstream Regulator`, 
  zscore_cols = IPA_Compare_upstream_regs_mat$Palbo2[, 6:10], 
  atac_peaks = Palbo_UpPeaks_AME_mat, 
  pdf_dir = "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3"
)

plot_upstream_reg_heatmaps(
  mat = IPA_Compare_upstream_regs_mat$Shared, 
  cluster_name = "Shared", 
  tf_db = IPA_TFs_db$Shared$`Upstream Regulator`, 
  zscore_cols = IPA_Compare_upstream_regs_mat$Shared[, 1:10], 
  atac_peaks = Palbo_UpPeaks_AME_mat, 
  pdf_dir = "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3"
)
```

