---
title: "Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment"
author: "Joanna Yeung"
date: '2025-02-01'
output: html_document
---

# load in workspace with DESeq results and rlog counts matrix of genes. 
```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Global_transcriptome_and_chromatin_accessibility_is_dynamic_in_both_types_of_therapy_induced_senescence/workspace.RData")
```

# assign significant genes to relevant k-means clusters & drug + cell cycle arrest categories. 
```{r genes_in_clusters}
genes_in_clusters <- lapply(levels(clusterDF$Clusters), function(x){clusterDF[clusterDF$Clusters %in% x, ]$gene_id})
cluster_names <- c("CellCycle1", "CellCycle2", "Palbo1", "Palbo2", "Doxo1", "Doxo2", "Shared") 
names(genes_in_clusters) <- cluster_names
```

# ATAC-seq: annotation of all peaks near significantly changing genes for each cluster
```{r}
# import gtf file for genes
genes_anno <- import("/lustre/fs4/risc_lab/store/jyeung/references/gencode.v41.GRCh38.p13.annotation.gtf")
genes_anno <- genes_anno[genes_anno$type %in% "gene", ] # filter for regions that are only genes 
genes_anno$gene_id <- gsub('\\.[0-9]*$', '', genes_anno$gene_id) # remove numbers after . in gene id column 
genes_anno <- genes_anno[ ,c("source", "type", "gene_id", "gene_type", "gene_name", "level")] # keep only these columns

# get the genomic coordinates of significantly changing genes from each cluster (RNA-seq data)
genes_in_clusters_anno <- lapply(genes_in_clusters, function(genes_in_clusters){genes_anno[genes_anno$gene_name %in% genes_in_clusters]})

# make GRanges where the coordinates are 50kb upstream of gene TSS and 50kb downstream of end of gene. 
genes_in_clusters_anno_findpeaks <- genes_in_clusters_anno
for(i in 1:length(genes_in_clusters_anno)){
  start(genes_in_clusters_anno_findpeaks[[i]]) <- start(genes_in_clusters_anno[[i]])-50000
  end(genes_in_clusters_anno_findpeaks[[i]]) <- end(genes_in_clusters_anno[[i]])+50000
}

# find peaks that overlap within 50kb of significantly changing genes in clusters
peaks_in_geneclusters <- lapply(genes_in_clusters_anno_findpeaks, function(genes_in_clusters_anno_findpeaks){annomasterpeaks_pval0.2[annomasterpeaks_pval0.2 %over% genes_in_clusters_anno_findpeaks, ]})

# filter for just sig changing peaks within 50kb of sig changing genes in clusters. 
sigPeaks_in_geneclusters <- list()
for(i in 1:length(mergedDiffGR)){
sigPeaks_in_geneclusters[[i]] <-lapply(peaks_in_geneclusters, function(peaks_in_geneclusters){mergedDiffGR[[i]][mergedDiffGR[[i]] %over% peaks_in_geneclusters, ]}) 
}
names(sigPeaks_in_geneclusters) <- names(mergedDiffGR)

# create static peak set to use as background for MEME-AME analysis for peaks near gene clusters. 
# initialize empty nested list
staticPeaks_in_geneclusters <- vector(mode="list", length=2)
names(staticPeaks_in_geneclusters) <- c("Doxo", "Palbo")
for(i in 1:length(staticPeaks_in_geneclusters)){
  staticPeaks_in_geneclusters[[i]] <- vector(mode="list", length=length(peaks_in_geneclusters))
}
# find static peaks for each drug condition for each gene cluster. 
for(i in 1:length(peaks_in_geneclusters)){
staticPeaks_in_geneclusters$Doxo[[i]] <- peaks_in_geneclusters[[i]][!peaks_in_geneclusters[[i]] %over% c(sigPeaks_in_geneclusters$allUp_Doxo[[i]], sigPeaks_in_geneclusters$allDown_Doxo[[i]])]
staticPeaks_in_geneclusters$Palbo[[i]] <- peaks_in_geneclusters[[i]][!peaks_in_geneclusters[[i]] %over% c(sigPeaks_in_geneclusters$allUp_Palbo[[i]], sigPeaks_in_geneclusters$allDown_Palbo[[i]])]
}
```

# function to get sequence under peaks for input into MEME-AME for motif enrichment analysis
```{r}
getSeqforMEME <- function(peakset, dir, filename){seq <- getSeq(BSgenome.Hsapiens.UCSC.hg38, peakset) #get sequence info under peaks
names(seq) <- 1:length(seq)
writeXStringSet(seq, file = paste0(dir, filename, ".fa")) #write to FASTA format for input into MEME
}
```


# ATAC-seq: create input fasta files for MEME-AME analysis
```{r}
directory <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_input/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters"
# get sequences under significant peaks belonging to each hierarchical cluster
for(i in 1:length(sigPeaks_in_geneclusters)){
  for(j in 1:length(sigPeaks_in_geneclusters[[i]])){

filename <- paste0("/", names(sigPeaks_in_geneclusters)[i], "_sigPeaks_within_50kb_of_", names(sigPeaks_in_geneclusters[[i]])[j], "_genes") 
  
getSeqforMEME(sigPeaks_in_geneclusters[[i]][[j]],
dir=directory, filename=filename)
  }
}  
# get sequences under static peaks belonging to each hierarchical cluster
for(i in 1:length(staticPeaks_in_geneclusters)){
  for(j in 1:length(staticPeaks_in_geneclusters[[i]])){
    filename <- paste0("/",names(staticPeaks_in_geneclusters[i]), "_staticPeaks_within_50kb_of_", names(sigPeaks_in_geneclusters[[i]])[j], "_genes") 
    getSeqforMEME(staticPeaks_in_geneclusters[[i]][[j]], dir=directory, filename=filename)
  }
}
# specify file paths for input to MEME-AME analysis

# folder names 
folderlist <- gsub("_sigPeaks_within_50kb_of|.fa", "", dir(directory, pattern="sigPeaks"))

# sig peaks nearby sig genes 
writeLines(dir(directory, full.names=T, pattern="sigPeaks"), con=paste0(directory, "/inputfiles_sigPeaks_in_geneclusters.txt"))

# static peaks nearby sig genes 
writeLines(rep(dir(directory, full.names=T, pattern="staticPeaks"), 2), con=paste0(directory, "/inputfiles_staticPeaks_in_geneclusters.txt"))

# folder names
writeLines(folderlist, con=paste0(directory, "/folderlist_sigPeaks_in_geneclusters.txt"))
```

# ATAC-seq: MEME-AME analysis on sig changing peaks within 50kb near sig changing genes clusters (kmeans=7). 
```{bash}
cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_input

sbatch -p risc --array=1-28 --export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_input/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters/inputfiles_staticPeaks_in_geneclusters.txt,inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_input/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters/inputfiles_sigPeaks_in_geneclusters.txt,outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_output/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters,folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_input/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters/folderlist_sigPeaks_in_geneclusters.txt /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm
```
#ATAC-seq: import motif enrichment results of changing peaks within 50kb of each RNA cluster of changing genes. 
```{r import nearby peaks kmeans 7 RNA clusters}
#extract ids of all expressed genes
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL

# function to import motif data. 
process_motif_data <- function(base_dir, expressed_genes = NULL) {
  library(dplyr)
  library(readr)
  
  # Initialize an empty list to store the data frames
  Peaks_AME <- list()
  processed_dirs <- vector()
  setwd(base_dir)
  file_paths=dir()

  # Loop through each directory and process the files
  for (i in seq_along(file_paths)) {
    # Construct the path to the ame.tsv file
    ame_file_path <- paste0(file_paths[i], "/ame.tsv")
    
    # Check the file size
    file_info <- file.info(ame_file_path)
    if (file_info$size == 1.1 || is.na(file_info$size)) {
      # If the file size is 0 or file does not exist, skip to the next iteration
      next
    }
    
    # Read the ame.tsv file
    Peaks_AME[[length(Peaks_AME) + 1]] <- as.data.frame(read_tsv(file = ame_file_path, comment = "#"))
    
    # Add the current directory to the processed_dirs vector
    processed_dirs <- c(processed_dirs, file_paths[i])
  }
  
  # Set names for the list elements
  names(Peaks_AME) <- processed_dirs
  
  # Post-processing each dataframe
  for (i in seq_along(Peaks_AME)) {
    
    # Calculate TPvsFP and add as a new column
    Peaks_AME[[i]]$TPvsFP <- Peaks_AME[[i]]$`%TP` - Peaks_AME[[i]]$`%FP`
    
    # Remove everything after "_" in the motif_ID column
    Peaks_AME[[i]]$motif_ID <- gsub("\\_.*", "", Peaks_AME[[i]]$motif_ID)
  }
  
  # Convert the list of MEME-AME results into a single dataframe
  Peaks_AME <- Peaks_AME[!sapply(Peaks_AME, is.null)]
  Peaks_AME_df <- do.call(rbind, Peaks_AME)
  
  # Filter for motifs that are expressed, if provided
  if (!is.null(expressed_genes)) {
    Peaks_AME_df <- Peaks_AME_df %>%
      dplyr::filter(motif_ID %in% expressed_genes)
  }
colnames(Peaks_AME_df)[c(7,15,17)] <- c("adj_pvalue", "percentTP", "percentFP") # change column names without special symbols 
Peaks_AME_df$neglog10padj <- -log10(Peaks_AME_df$adj_pvalue) # add in column for -log10(adj pvalue)
  return(Peaks_AME_df)
}
save(process_motif_data, file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Functions/process_motif_data.RData")

Peaks_AME_Clusters <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_output/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters", expressed_genes = expressedGenes)

# Add column to specify which RNA cluster
Peaks_AME_Clusters$Cluster <- gsub('\\.[0-9]*$', '', row.names(Peaks_AME_Clusters)) 
Peaks_AME_Clusters$Cluster <- gsub("^[^_]+_[^_]+_", "", Peaks_AME_Clusters$Cluster) # remove everything before the 2nd "_"

Peaks_AME_Clusters$Comparison <- gsub(".*?(all(Up|Down)_(Palbo|Doxo)).*", "\\1", row.names(Peaks_AME_Clusters))
Peaks_AME_Clusters$Comparison <- as.factor(Peaks_AME_Clusters$Comparison)
Peaks_AME_Clusters$Direction <- ifelse(grepl("Up", Peaks_AME_Clusters$Comparison), "Up", "Down")
Peaks_AME_Clusters$Group <- gsub("^[^_]*_", "", Peaks_AME_Clusters$Comparison)
```
# ATAC-seq:convert dataframe of enriched motif results to matrix for input into plotting in pheatmap
```{r convert AME df to matrix}
# create function that stores the -log10(padj) value of each motif ID for each cluster
create_AME_peaks_matrix_byCluster <- function(sig_peaks_df) {
  # Create an empty matrix for significant peaks
  unique_motifs <- unique(sig_peaks_df$motif_ID)
  cluster <- unique(sig_peaks_df$Cluster)
  sig_peaks_mat <- matrix(nrow = length(unique_motifs), ncol = length(cluster))
  
  rownames(sig_peaks_mat) <- unique_motifs # Row names are motif IDs
  colnames(sig_peaks_mat) <- cluster   # Column names are DESeq comparisons
  
  # Populate the matrix with neglog10padj values
  for (i in 1:ncol(sig_peaks_mat)) {
    temp_comparison <- sig_peaks_df[sig_peaks_df$Cluster == cluster[i], ]
    sig_peaks_mat[, i] <- temp_comparison[match(rownames(sig_peaks_mat), temp_comparison$motif_ID), ]$neglog10padj
    sig_peaks_mat[, i] <- ifelse(is.na(sig_peaks_mat[, i]), 0, sig_peaks_mat[, i])
  }
  
  return(sig_peaks_mat)
}

AME_padj_mat <- list()
comparisons <- c("allDown_Doxo", "allDown_Palbo", "allUp_Doxo", "allUp_Palbo")
for(i in 1:length(comparisons)){
AME_padj_mat[[i]] <- create_AME_peaks_matrix_byCluster(Peaks_AME_Clusters[Peaks_AME_Clusters$Comparison %in% comparisons[i], ])
}
names(AME_padj_mat) <- c("allDown_Doxo", "allDown_Palbo", "allUp_Doxo", "allUp_Palbo")
```
--------------------------------------------------------------------------------------------------------------------------------------------

#*FIGURES*



# ATAC-seq, separate RNA clusters: plot motif enrichment results of all clusters as heatmap of p-adj values for each type (up, down, Doxo, Palbo) of sig changing peak 
```{r}
pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Up.Down.peaks.near.RNAclusters.enriched.motifs.pdf", width=3, height=9)

  pheatmap(AME_padj_mat[[1]], cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 300,length.out = 100), main="all Down Doxo: enriched motifs")

  pheatmap(AME_padj_mat[[2]], cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 30,length.out = 100), main="all Down Palbo: enriched motifs")

  pheatmap(AME_padj_mat[[3]], cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 40,length.out = 100), main="all Up Doxo: enriched motifs") 
  
  pheatmap(AME_padj_mat[[4]], cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 30,length.out = 100), main="all Up Palbo: enriched motifs")

dev.off()

# merge motif enrichment results from all up Doxo and all up Palbo peaks. 
merged_Up_AME_mat <- merge(
  data.frame(motif = rownames(AME_padj_mat$allUp_Doxo), AME_padj_mat$allUp_Doxo, row.names = NULL),
  data.frame(motif = rownames(AME_padj_mat$allUp_Palbo), AME_padj_mat$allUp_Palbo, row.names = NULL),
  by = "motif",
  all = TRUE
)

# Set row names back
row.names(merged_Up_AME_mat) <- merged_Up_AME_mat$motif
merged_Up_AME_mat$motif <- NULL
# Set NA values to 0. 
merged_Up_AME_mat[is.na(merged_Up_AME_mat)] <- 0

merged_Up_AME_mat <- as.matrix(merged_Up_AME_mat)

# merge motif enrichment results from all up Doxo and all up Palbo peaks. 
merged_Down_AME_mat <- merge(
  data.frame(motif = rownames(AME_padj_mat$allDown_Doxo), AME_padj_mat$allDown_Doxo, row.names = NULL),
  data.frame(motif = rownames(AME_padj_mat$allDown_Palbo), AME_padj_mat$allDown_Palbo, row.names = NULL),
  by = "motif",
  all = TRUE
)

# Set row names back
row.names(merged_Down_AME_mat) <- merged_Down_AME_mat$motif
merged_Down_AME_mat$motif <- NULL
# Set NA values to 0. 
merged_Down_AME_mat[is.na(merged_Down_AME_mat)] <- 0

merged_Down_AME_mat <- as.matrix(merged_Down_AME_mat)

pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Up.Down.mergedpeaks.near.RNAclusters.enriched.motifs.pdf", width=3, height=9)

pheatmap(merged_Down_AME_mat, cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 50,length.out = 100), main="all down peaks near RNA clusters: enriched motifs, left: Doxo, right: Palbo")

pheatmap(merged_Up_AME_mat, cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 30,length.out = 100), main="all up peaks near RNA clusters: enriched motifs, left: Doxo, right: Palbo")
dev.off()
```

# ATAC-seq, separate RNA clusters: plot motif enrichment results as dotplot indicating difference between %TP & %FP as dot size & fill color as padj values.  
```{r}
AME_padj_dots <- list()
comparisons <- c("allDown_Doxo", "allDown_Palbo", "allUp_Doxo", "allUp_Palbo")
for(i in 1:length(comparisons)){
AME_padj_dots[[i]] <- Peaks_AME_Clusters[Peaks_AME_Clusters$Comparison %in% comparisons[i], ]
AME_padj_dots[[i]] <- AME_padj_dots[[i]][AME_padj_dots[[i]]$motif_ID %in% row.names(AME_padj_mat[[i]][1:15, ]), ] # show only top 15 most significant padj values motif IDs
AME_padj_dots[[i]]$motif_ID <- factor(AME_padj_dots[[i]]$motif_ID, levels=rev(unique(AME_padj_dots[[i]]$motif_ID)))
}
names(AME_padj_dots) <- c("allDown_Doxo", "allDown_Palbo", "allUp_Doxo", "allUp_Palbo")
```

# ATAC-seq: create function to plot motif enrichment results as dotplot where the fill color indicates -log10(padj) values & size of dot indicates %TP-%FP
```{r create_AME_dotplots}
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=50, size_limit=4.5){
  AME_df <- AME_df %>% mutate(Comparison = Cluster)
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient(name="-log10(padj)", 
                        low="light yellow", high="dark orange",
                        limits=c(0, fill_limit), 
                        oob=scales::squish) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=c(0, size_limit), 
                          range=c(1, 10)) +  # Adjust dot sizes
    theme_minimal() +
    ggtitle(Title) +
    scale_x_discrete(labels = function(x) AME_df$Cluster[match(x, AME_df$Comparison)]) +  # to replace Comparison with Cluster labels on the x-axis. 
    facet_wrap(~Group, nrow=1, scales="free") +
    theme(
      axis.text.x = element_text(size=8, angle = 90, hjust=1),  # Rotate x-axis labels
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    )+
    theme_classic() +
    labs(x="Comparison", y="motif ID")}
``` 

```{r}
pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Up.Down.peaks.near.RNAclusters.enriched.motifs.dotplots.pdf", width=8, height=5)
for(i in 1:length(AME_padj_dots)){
p <- create_AME_dotplots(AME_padj_dots[[i]], Title=names(AME_padj_dots)[i]) 
print(p)
}
dev.off()
```

-------------------------------------------------------------------------------------------------------------------
# *FUNCTIONS*
# save functions for use in other RMarkdown files. 
```{r}
functions_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Functions"
save(process_motif_data, file=paste0(functions_dir, "/process_motif_data.RData"))

save(create_AME_peaks_matrix_byCluster, file=paste0(functions_dir, "/create_AME_peaks_matrix_byCluster.RData"))

# create_AME_dotplots with extra arguments to set fill limits and size limits for "%TP - %FP" & -log10(padj) values. 
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=50, size_limit=4.5 ){
  ggplot(AME_df, aes(x=Cluster, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient(name="-log10(padj)", 
                        low="light yellow", high="dark orange",
                        limits=c(0, fill_limit), 
                        oob=scales::squish) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=c(0, size_limit), 
                          range=c(1, 10)) +  # Adjust dot sizes
    theme_minimal() +
    ggtitle(Title) +
    theme(
      text = element_text(size=8), 
      axis.text.x = element_text(size=8), 
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    ) +
  theme_classic()+
    facet_wrap(~Cluster, scales="free", nrow=1)+
    labs(x="RNA cluster", y="motif ID")
}

save(create_AME_dotplots, file=paste0(functions_dir, "/create_AME_dotplots.RData"))
```
