---
title: "Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment"
author: "Joanna Yeung"
date: '2025-02-01'
output: html_document
---

# load in workspace with DESeq results and rlog counts matrix of genes. 
```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Global_transcriptome_and_chromatin_accessibility_is_dynamic_in_both_types_of_therapy_induced_senescence/workspace.RData")
```

# assign significant genes to relevant k-means clusters & drug + cell cycle arrest categories. 
```{r genes_in_clusters}
genes_in_clusters <- lapply(levels(clusterDF$Clusters), function(x){clusterDF[clusterDF$Clusters %in% x, ]$gene_id})
cluster_names <- c("CellCycle1", "CellCycle2", "Palbo1", "Palbo2", "Doxo1", "Doxo2", "Shared") 
names(genes_in_clusters) <- cluster_names
```

# ATAC-seq: annotation of all peaks near significantly changing genes for each cluster
```{r}
# import gtf file for genes
genes_anno <- import("/lustre/fs4/risc_lab/store/jyeung/references/gencode.v41.GRCh38.p13.annotation.gtf")
genes_anno <- genes_anno[genes_anno$type %in% "gene", ] # filter for regions that are only genes 
genes_anno$gene_id <- gsub('\\.[0-9]*$', '', genes_anno$gene_id) # remove numbers after . in gene id column 
genes_anno <- genes_anno[ ,c("source", "type", "gene_id", "gene_type", "gene_name", "level")] # keep only these columns

# get the genomic coordinates of significantly changing genes from each cluster (RNA-seq data)
genes_in_clusters_anno <- lapply(genes_in_clusters, function(genes_in_clusters){genes_anno[genes_anno$gene_name %in% genes_in_clusters]})

# make GRanges where the coordinates are 50kb upstream of gene TSS and 50kb downstream of end of gene. 
genes_in_clusters_anno_findpeaks <- genes_in_clusters_anno
for(i in 1:length(genes_in_clusters_anno)){
  start(genes_in_clusters_anno_findpeaks[[i]]) <- start(genes_in_clusters_anno[[i]])-50000
  end(genes_in_clusters_anno_findpeaks[[i]]) <- end(genes_in_clusters_anno[[i]])+50000
}

# find peaks that overlap within 50kb of significantly changing genes in clusters
peaks_in_geneclusters <- lapply(genes_in_clusters_anno_findpeaks, function(genes_in_clusters_anno_findpeaks){annomasterpeaks_pval0.2[annomasterpeaks_pval0.2 %over% genes_in_clusters_anno_findpeaks, ]})

# filter for just sig changing peaks within 50kb of sig changing genes in clusters. 
sigPeaks_in_geneclusters <- list()
for(i in 1:length(mergedDiffGR)){
sigPeaks_in_geneclusters[[i]] <-lapply(peaks_in_geneclusters, function(peaks_in_geneclusters){mergedDiffGR[[i]][mergedDiffGR[[i]] %over% peaks_in_geneclusters, ]}) 
}
names(sigPeaks_in_geneclusters) <- names(mergedDiffGR)

# create static peak set to use as background for MEME-AME analysis for peaks near gene clusters. 
# initialize empty nested list
staticPeaks_in_geneclusters <- vector(mode="list", length=2)
names(staticPeaks_in_geneclusters) <- c("Doxo", "Palbo")
for(i in 1:length(staticPeaks_in_geneclusters)){
  staticPeaks_in_geneclusters[[i]] <- vector(mode="list", length=length(peaks_in_geneclusters))
}
# find static peaks for each drug condition for each gene cluster. 
for(i in 1:length(peaks_in_geneclusters)){
staticPeaks_in_geneclusters$Doxo[[i]] <- peaks_in_geneclusters[[i]][!peaks_in_geneclusters[[i]] %over% c(sigPeaks_in_geneclusters$allUp_Doxo[[i]], sigPeaks_in_geneclusters$allDown_Doxo[[i]])]
staticPeaks_in_geneclusters$Palbo[[i]] <- peaks_in_geneclusters[[i]][!peaks_in_geneclusters[[i]] %over% c(sigPeaks_in_geneclusters$allUp_Palbo[[i]], sigPeaks_in_geneclusters$allDown_Palbo[[i]])]
}
```

# function to get sequence under peaks for input into MEME-AME for motif enrichment analysis
```{r}
getSeqforMEME <- function(peakset, dir, filename){seq <- getSeq(BSgenome.Hsapiens.UCSC.hg38, peakset) #get sequence info under peaks
names(seq) <- 1:length(seq)
writeXStringSet(seq, file = paste0(dir, filename, ".fa")) #write to FASTA format for input into MEME
}
```


# ATAC-seq: create input fasta files for MEME-AME analysis
```{r}
directory <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_input/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters"
# get sequences under significant peaks belonging to each hierarchical cluster
for(i in 1:length(sigPeaks_in_geneclusters)){
  for(j in 1:length(sigPeaks_in_geneclusters[[i]])){

filename <- paste0("/", names(sigPeaks_in_geneclusters)[i], "_sigPeaks_within_50kb_of_", names(sigPeaks_in_geneclusters[[i]])[j], "_genes") 
  
getSeqforMEME(sigPeaks_in_geneclusters[[i]][[j]],
dir=directory, filename=filename)
  }
}  
# get sequences under static peaks belonging to each hierarchical cluster
for(i in 1:length(staticPeaks_in_geneclusters)){
  for(j in 1:length(staticPeaks_in_geneclusters[[i]])){
    filename <- paste0("/",names(staticPeaks_in_geneclusters[i]), "_staticPeaks_within_50kb_of_", names(sigPeaks_in_geneclusters[[i]])[j], "_genes") 
    getSeqforMEME(staticPeaks_in_geneclusters[[i]][[j]], dir=directory, filename=filename)
  }
}
# specify file paths for input to MEME-AME analysis

# folder names 
folderlist <- gsub("_sigPeaks_within_50kb_of|.fa", "", dir(directory, pattern="sigPeaks"))

# sig peaks nearby sig genes 
writeLines(dir(directory, full.names=T, pattern="sigPeaks"), con=paste0(directory, "/inputfiles_sigPeaks_in_geneclusters.txt"))

# static peaks nearby sig genes 
writeLines(rep(dir(directory, full.names=T, pattern="staticPeaks"), 2), con=paste0(directory, "/inputfiles_staticPeaks_in_geneclusters.txt"))

# folder names
writeLines(folderlist, con=paste0(directory, "/folderlist_sigPeaks_in_geneclusters.txt"))
```

# ATAC-seq: MEME-AME analysis on sig changing peaks within 50kb near sig changing genes clusters (kmeans=7). 
```{bash}
cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_input

sbatch -p risc --array=1-28 --export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_input/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters/inputfiles_staticPeaks_in_geneclusters.txt,inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_input/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters/inputfiles_sigPeaks_in_geneclusters.txt,outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_output/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters,folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_input/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters/folderlist_sigPeaks_in_geneclusters.txt /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm
```

# ATAC-seq: make matrix of vst normalized counts under nearby upregulated peaks (within 50kb) of sig changing gene clusters. 
```{r}
# function to reorder counts matrix by drug category
reorder_by_drug_cat <- function(matrix,DoxoGR=DoxoGR, PalboGR=PalboGR) {
  
  # make dataframe to specify which drug treatment peak belongs to: Palbo, Doxo or both. 
  drug_cat <- data.frame(row.names=row.names(matrix), Palbo=ifelse(row.names(matrix) %in% PalboGR$name, "Yes", "No"), 
                         Doxo=ifelse(row.names(matrix) %in% DoxoGR$name, "Yes", "No"), 
                         Both=ifelse(row.names(matrix) %in% DoxoGR$name & row.names(matrix) %in% PalboGR$name, "Yes", "No"),
                         vsCyc=ifelse(row.names(matrix) %in% vsCyc$name, "Yes", "No"), 
                         vsday3=ifelse(row.names(matrix) %in% vsday3$name, "Yes", "No") )
  
  # order rows by Palbo, Doxo then both peaks
  drug_cat_list <- list(drug_cat[drug_cat$Palbo %in% "Yes" & drug_cat$Doxo %in% "No" & drug_cat$vsCyc %in% "Yes" & drug_cat$vsday3 %in% "No", ],
                        drug_cat[drug_cat$Palbo %in% "Yes" & drug_cat$Doxo %in% "No" & drug_cat$vsCyc %in% "Yes" & drug_cat$vsday3 %in% "Yes", ],
                        drug_cat[drug_cat$Palbo %in% "Yes" & drug_cat$Doxo %in% "No" & drug_cat$vsCyc %in% "No" & drug_cat$vsday3 %in% "Yes", ],
                        drug_cat[drug_cat$Both %in% "Yes" & drug_cat$vsCyc %in% "Yes" & drug_cat$vsday3 %in% "No", ],
                        drug_cat[drug_cat$Both %in% "Yes" & drug_cat$vsCyc %in% "Yes" & drug_cat$vsday3 %in% "Yes", ],
                        drug_cat[drug_cat$Both %in% "Yes" & drug_cat$vsCyc %in% "No" & drug_cat$vsday3 %in% "Yes", ],
                        drug_cat[drug_cat$Palbo %in% "No" & drug_cat$Doxo %in% "Yes" & drug_cat$vsCyc %in% "Yes" & drug_cat$vsday3 %in% "No", ], 
                        drug_cat[drug_cat$Palbo %in% "No" & drug_cat$Doxo %in% "Yes" & drug_cat$vsCyc %in% "Yes" & drug_cat$vsday3 %in% "Yes", ],
                        drug_cat[drug_cat$Palbo %in% "No" & drug_cat$Doxo %in% "Yes" & drug_cat$vsCyc %in% "No" & drug_cat$vsday3 %in% "Yes", ]
  )
  ordered_matrix <- list()
  for(i in 1:length(drug_cat_list)){
    ordered_matrix[[i]] <- matrix[row.names(matrix) %in% row.names(drug_cat_list[[i]]), ]
  }
  # merge back into master counts matrix
  ordered_matrix <- do.call(rbind, ordered_matrix)
}

# Up Peaks
# get changing peaks nearby each RNA-seq gene cluster
Uppeaks_in_geneclusters_heatmap <- list()
Up_geneclusters_anno <- list()
for(i in 1:length(peaks_in_geneclusters)){
  Uppeaks_in_geneclusters_heatmap[[i]] <- mergedDiff_hm_rmOutlier$UpPalbo.UpDoxo[row.names(mergedDiff_hm_rmOutlier$UpPalbo.UpDoxo) %in% peaks_in_geneclusters[[i]]$name, ]
  Up_geneclusters_anno[[i]] <- data.frame(row.names=row.names(Uppeaks_in_geneclusters_heatmap[[i]]))
}
# annotate which RNA_seq cluster peak is nearby to
for(i in 1:length(peaks_in_geneclusters)){
  for(j in 1:length(peaks_in_geneclusters)){
  Up_geneclusters_anno[[i]][ ,j] <- ifelse(row.names(Up_geneclusters_anno[[i]]) %in% peaks_in_geneclusters[[j]]$name, "Yes", "No")
  }
}
# reorder each RNA-seq cluster by drug category peak belongs to
Uppeaks_in_geneclusters_heatmap <- lapply(Uppeaks_in_geneclusters_heatmap, reorder_by_drug_cat, DoxoGR=mergedDiffGR$allUp_Doxo, PalboGR=mergedDiffGR$allUp_Palbo)

# merge clusters into 1 heatmap
Uppeaks_in_geneclusters_heatmap <- do.call(rbind, Uppeaks_in_geneclusters_heatmap)
# merge annotation of peaks into 1 heatmap
Up_geneclusters_anno <- do.call(rbind, Up_geneclusters_anno)
# remove duplicate peaks since a peak can belong to more than 1 cluster. 
Uppeaks_in_geneclusters_heatmap <- Uppeaks_in_geneclusters_heatmap[!duplicated(row.names(Uppeaks_in_geneclusters_heatmap)) & !row.names(Uppeaks_in_geneclusters_heatmap) %in% "", ]

# match order of annotation of peaks to the heatmap 
Up_geneclusters_anno<- Up_geneclusters_anno[match(row.names(Uppeaks_in_geneclusters_heatmap), row.names(Up_geneclusters_anno)), ]
colnames(Up_geneclusters_anno) <- names(peaks_in_geneclusters)

# annotate other peak features. 
UpannoRow <- data.frame(Doxo=ifelse(row.names(Uppeaks_in_geneclusters_heatmap) %in% mergedDiffGR$allUp_Doxo$name, "Yes", "No"), 
                      Palbo=ifelse(row.names(Uppeaks_in_geneclusters_heatmap) %in% mergedDiffGR$allUp_Palbo$name, "Yes", "No"), 
                      Peak_anno=simpleannomasterpeaks_pval0.2[match(row.names(Uppeaks_in_geneclusters_heatmap), simpleannomasterpeaks_pval0.2$name)]$annotation,
                      row.names = row.names(Uppeaks_in_geneclusters_heatmap))
# merge other annotation with annotation of RNA-seq clusters. 
UpannoRow <- cbind(UpannoRow, Up_geneclusters_anno)
```

#ATAC-seq: import motif enrichment results of changing peaks within 50kb of each RNA cluster of changing genes. 
```{r import nearby peaks kmeans 7 RNA clusters}
#extract ids of all expressed genes
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL
expressedGenes[grep("TP53$|TP63$|TP73$", expressedGenes)] <- c("P63", "P73", "P53")

# function to import motif data. 
process_motif_data <- function(base_dir, expressed_genes = NULL) {
  library(dplyr)
  library(readr)
  
  # Initialize an empty list to store the data frames
  Peaks_AME <- list()
  processed_dirs <- vector()
  setwd(base_dir)
  file_paths=dir()

  # Loop through each directory and process the files
  for (i in seq_along(file_paths)) {
    # Construct the path to the ame.tsv file
    ame_file_path <- paste0(file_paths[i], "/ame.tsv")
    
    # Check the file size
    file_info <- file.info(ame_file_path)
    if (file_info$size == 1.1 || is.na(file_info$size)) {
      # If the file size is 0 or file does not exist, skip to the next iteration
      next
    }
    
    # Read the ame.tsv file
    Peaks_AME[[length(Peaks_AME) + 1]] <- as.data.frame(read_tsv(file = ame_file_path, comment = "#"))
    
    # Add the current directory to the processed_dirs vector
    processed_dirs <- c(processed_dirs, file_paths[i])
  }
  
  # Set names for the list elements
  names(Peaks_AME) <- processed_dirs
  
  # Post-processing each dataframe
  for (i in seq_along(Peaks_AME)) {
    
    # Calculate TPvsFP and add as a new column
    Peaks_AME[[i]]$TPvsFP <- Peaks_AME[[i]]$`%TP` - Peaks_AME[[i]]$`%FP`
    
    # Remove everything after "_" in the motif_ID column
    Peaks_AME[[i]]$motif_ID <- gsub("\\_.*", "", Peaks_AME[[i]]$motif_ID)
  }
  
  # Convert the list of MEME-AME results into a single dataframe
  Peaks_AME <- Peaks_AME[!sapply(Peaks_AME, is.null)]
  Peaks_AME_df <- do.call(rbind, Peaks_AME)
  
  # Filter for motifs that are expressed, if provided
  if (!is.null(expressed_genes)) {
    Peaks_AME_df <- Peaks_AME_df %>%
      dplyr::filter(motif_ID %in% expressed_genes)
  }
colnames(Peaks_AME_df)[c(7,15,17)] <- c("adj_pvalue", "percentTP", "percentFP") # change column names without special symbols 
Peaks_AME_df$neglog10padj <- -log10(Peaks_AME_df$adj_pvalue) # add in column for -log10(adj pvalue)
  return(Peaks_AME_df)
}
save(process_motif_data, file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Functions/process_motif_data.RData")

Peaks_AME_Clusters <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/motif_analysis_output/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters", expressed_genes = expressedGenes)

# Add column to specify which RNA cluster
Peaks_AME_Clusters$Cluster <- gsub('\\.[0-9]*$', '', row.names(Peaks_AME_Clusters)) 
Peaks_AME_Clusters$Cluster <- gsub("^[^_]+_[^_]+_", "", Peaks_AME_Clusters$Cluster) # remove everything before the 2nd "_"

Peaks_AME_Clusters$Comparison <- gsub(".*?(all(Up|Down)_(Palbo|Doxo)).*", "\\1", row.names(Peaks_AME_Clusters))
Peaks_AME_Clusters$Comparison <- as.factor(Peaks_AME_Clusters$Comparison)
Peaks_AME_Clusters$Direction <- ifelse(grepl("Up", Peaks_AME_Clusters$Comparison), "Up", "Down")
Peaks_AME_Clusters$Group <- gsub("^[^_]*_", "", Peaks_AME_Clusters$Comparison)
```

# ATAC-seq:convert dataframe of enriched motif results to matrix for input into plotting in pheatmap
```{r convert AME df to matrix}
# create function that stores the -log10(padj) value of each motif ID for each cluster
create_AME_peaks_matrix_byCluster <- function(
    sig_peaks_df) {
  # Create an empty matrix for significant peaks
  unique_motifs <- unique(sig_peaks_df$motif_ID)
  cluster <- unique(sig_peaks_df$Cluster)
  sig_peaks_mat <- matrix(nrow = length(unique_motifs), ncol = length(cluster))
  
  rownames(sig_peaks_mat) <- unique_motifs # Row names are motif IDs
  colnames(sig_peaks_mat) <- cluster   # Column names are DESeq comparisons
  
  # Populate the matrix with neglog10padj values
  for (i in 1:ncol(sig_peaks_mat)) {
    temp_comparison <- sig_peaks_df[sig_peaks_df$Cluster == cluster[i], ]
    sig_peaks_mat[, i] <- temp_comparison[match(rownames(sig_peaks_mat), temp_comparison$motif_ID), ]$neglog10padj
    sig_peaks_mat[, i] <- ifelse(is.na(sig_peaks_mat[, i]), 0, sig_peaks_mat[, i])
  }
  
  return(sig_peaks_mat)
}

AME_padj_mat <- list()
comparisons <- c("allDown_Doxo", "allDown_Palbo", "allUp_Doxo", "allUp_Palbo")
for(i in 1:length(comparisons)){
AME_padj_mat[[i]] <- create_AME_peaks_matrix_byCluster(Peaks_AME_Clusters[Peaks_AME_Clusters$Comparison %in% comparisons[i], ])
AME_padj_mat[[i]] <- AME_padj_mat[[i]][ ,c(1,2,5,6,3,4,7)]
}
names(AME_padj_mat) <- c("allDown_Doxo", "allDown_Palbo", "allUp_Doxo", "allUp_Palbo")
```
--------------------------------------------------------------------------------------------------------------------------------------------

#*FIGURES*

# generate heatmap of vst counts under nearby upregulated peaks
```{r}
# merge technical replicates into 1 sample to make the heatmap look like overwhelming & stripy. 
dds_filt$group <- factor(paste0(dds_filt$Condition, dds_filt$Biorep))
#dds_filt_collapsed <- collapseReplicates(dds_filt, dds_filt$group)
dds_filt_collapsed <- collapseReplicates(dds_filt, dds_filt$Condition)
dds_filt_collapsed <- estimateSizeFactors(dds_filt_collapsed) 
dds_filt_collapsed <- dds_filt_collapsed[ ,c(1,9,2,4,6,10,3,5,7,8)] # reorder samples by time & drug
# generate vst transformed counts from collapsed DESeq counts. 
vstdds_collapsed <- vst(dds_filt_collapsed)
vstdds_matrix_collapsed <- assays(vstdds_collapsed)[[1]]

# create dataframe to annotate columns of counts matrix by color
annoCol_collapsed <- colData(dds_filt_collapsed)[ ,c( "Treatment", "Timepoint")]
annoCol_collapsed <- as.data.frame(annoCol_collapsed[c(1,9,2,4,6,10,3,5,7,8), ])

Upcollapsed_hm <- vstdds_matrix_collapsed[match(row.names( Uppeaks_in_geneclusters_heatmap), row.names(vstdds_matrix_collapsed)), ]


anno_colors <- list(Doxo=hm_cols_ATAC$Doxo, 
                    Palbo=hm_cols_ATAC$Palbo, 
                    Peak_anno=hm_cols_ATAC$Peak_anno, 
                    Timepoint=hm_cols_ATAC$Timepoint,
                    Treatment=hm_cols_ATAC$Treatment,
                CellCycle1=c("Yes"="#66C2A5", No="white"),
                CellCycle2=c("Yes"="#A6D854", No="white"),
                Palbo1=c("Yes"="#8DA0CB", No="white"),
                Palbo2=c("Yes"="#FC8D62", No="white"),
                Doxo1=c("Yes"="#E78AC3", No="white"),
                Doxo2=c("Yes"="#FFD92F", No="white"),
                Shared=c("Yes"="#E5C494", No="white")
)
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment"
pdf(paste0(pdf_dir, "/ATAC_Up.peaks.near.RNAclusters_mat.pdf"))
pheatmap(
  Upcollapsed_hm, 
  scale = "row",
  breaks=seq(-3, 3,length.out = 100),
  show_rownames = FALSE, 
  cluster_cols = FALSE, 
  cluster_rows= FALSE,
  annotation_col = annoCol_collapsed, 
  annotation_colors = anno_colors,
  annotation_row = UpannoRow[ , !colnames(UpannoRow) %in% "Peak_anno"], 
  main=paste0("Up peaks within 50kb of RNA-seq clusters:", nrow(Uppeaks_in_geneclusters_heatmap), " Peaks"
))
dev.off()
```

# ATAC-seq, separate RNA clusters: plot motif enrichment results of all clusters as heatmap of p-adj values for each type (up, down, Doxo, Palbo) of sig changing peak 
```{r}
pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Up.Down.peaks.near.RNAclusters.enriched.motifs.pdf", width=3, height=9)

  pheatmap(AME_padj_mat[[1]], cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 300,length.out = 100), main="all Down Doxo: enriched motifs")

  pheatmap(AME_padj_mat[[2]], cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 30,length.out = 100), main="all Down Palbo: enriched motifs")

  pheatmap(AME_padj_mat[[3]], cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 40,length.out = 100), main="all Up Doxo: enriched motifs") 
  
  pheatmap(AME_padj_mat[[4]], cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84", "#B30000"))(100), breaks = seq(0, 30,length.out = 100), main="all Up Palbo: enriched motifs")

dev.off()
```
```{r merged_nearby_enriched_motifs_mat}
# merge motif enrichment results from all up Doxo and all up Palbo peaks. 
merged_Up_AME_mat <- merge(
  data.frame(motif = rownames(AME_padj_mat$allUp_Doxo), AME_padj_mat$allUp_Doxo, row.names = NULL),
  data.frame(motif = rownames(AME_padj_mat$allUp_Palbo), AME_padj_mat$allUp_Palbo, row.names = NULL),
  by = "motif",
  all = TRUE
)

# Set row names back
row.names(merged_Up_AME_mat) <- merged_Up_AME_mat$motif
merged_Up_AME_mat$motif <- NULL
# Set NA values to 0. 
merged_Up_AME_mat[is.na(merged_Up_AME_mat)] <- 0

merged_Up_AME_mat <- as.matrix(merged_Up_AME_mat)

# merge motif enrichment results from all up Doxo and all up Palbo peaks. 
merged_Down_AME_mat <- merge(
  data.frame(motif = rownames(AME_padj_mat$allDown_Doxo), AME_padj_mat$allDown_Doxo, row.names = NULL),
  data.frame(motif = rownames(AME_padj_mat$allDown_Palbo), AME_padj_mat$allDown_Palbo, row.names = NULL),
  by = "motif",
  all = TRUE
)

# Set row names back
row.names(merged_Down_AME_mat) <- merged_Down_AME_mat$motif
merged_Down_AME_mat$motif <- NULL
# Set NA values to 0. 
merged_Down_AME_mat[is.na(merged_Down_AME_mat)] <- 0

merged_Down_AME_mat <- as.matrix(merged_Down_AME_mat)
```
```{r}
# filter out overall lowly enriched motifs. 
filtered_merged_Up_AME_mat <-merged_Up_AME_mat[apply(merged_Up_AME_mat, 1, max) >= 4.8, ]
filtered_merged_Down_AME_mat <-merged_Down_AME_mat[apply(merged_Down_AME_mat, 1, max) >= 9.3, ]
pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Up.Down.mergedpeaks.near.RNAclusters.enriched.motifs.pdf", width=6, height=12)
pheatmap(filtered_merged_Down_AME_mat, cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84",  "#EF6548", "#B30000", "#7F0000", "#7F0000"))(100), breaks = seq(0, 30,length.out = 100), main="all down peaks near RNA clusters: enriched motifs, left: Doxo, right: Palbo")

pheatmap(filtered_merged_Up_AME_mat, cluster_rows = T, cluster_cols=F, color = colorRampPalette(c("white",  "#FDBB84",  "#EF6548", "#B30000", "#7F0000", "#7F0000"))(100), breaks = seq(0, 30,length.out = 100), main="all up peaks near RNA clusters: enriched motifs, left: Doxo, right: Palbo")
dev.off()
```

# ATAC-seq, separate RNA clusters: plot motif enrichment results as dotplot indicating difference between %TP & %FP as dot size & fill color as padj values.  
```{r}
AME_padj_dots <- list()
comparisons <- c("allDown_Doxo", "allDown_Palbo", "allUp_Doxo", "allUp_Palbo")
AME_padj_dots <- list(allDown=Peaks_AME_Clusters[Peaks_AME_Clusters$Comparison %in% c("allDown_Doxo", "allDown_Palbo"), ],
                      allUp=Peaks_AME_Clusters[Peaks_AME_Clusters$Comparison %in% c("allUp_Doxo", "allUp_Palbo"), ])

AME_padj_dots$allDown$motif_ID <- factor(AME_padj_dots$allDown$motif_ID, levels=row.names(merged_Down_AME_mat)[(hclust(dist(merged_Down_AME_mat))$order)])

AME_padj_dots$allUp$motif_ID <- factor(AME_padj_dots$allUp$motif_ID, row.names(merged_Up_AME_mat)[(hclust(dist(merged_Up_AME_mat))$order)] )                
```

# ATAC-seq: create function to plot motif enrichment results as dotplot where the fill color indicates -log10(padj) values & size of dot indicates %TP-%FP
```{r create_AME_dotplots}
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=c(1,10), size_limit=4.5, midpoint=20){
  AME_df <- AME_df %>% mutate(Comparison = Cluster)
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient2(name="-log10(padj)", 
                        low="white", mid="light yellow", high="dark orange",
                        midpoint = midpoint,
                        limits=fill_limit) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=size_limit, 
                          range=c(1, 10)) +  # Adjust dot sizes
    theme_minimal() +
    ggtitle(Title) +
    scale_x_discrete(labels = function(x) AME_df$Cluster[match(x, AME_df$Comparison)]) +  # to replace Comparison with Cluster labels on the x-axis. 
    facet_wrap(~Group, nrow=1) +
    theme(
      axis.text.x = element_text(size=8, angle = 90, hjust=1),  # Rotate x-axis labels
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    )+
    theme_classic() +
    labs(x="Comparison", y="motif ID")}
``` 

```{r}
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment"
pdf(paste0(pdf_dir, "/Up.Down.mergedpeaks.near.RNAclusters.enriched.motifs.dotplots.pdf"), width=8, height=15)
create_AME_dotplots(AME_padj_dots$allUp, Title="nearby up peaks", fill_limit=c(0, 70), size_limit=c(0, 20.05), midpoint=2)
create_AME_dotplots(AME_padj_dots$allDown, Title="nearby down peaks", fill_limit=c(0, 305), size_limit=c(0, 39.4), midpoint=10)
dev.off()
```
# ATAC-seq: plot number of input peaks that went into foreground peak set for MEME-AME analysis: 
```{r}
input_sigpeaks_df <-  list()
for(i in 1:length(sigPeaks_in_geneclusters)){

input_sigpeaks_df[[i]] <- data.frame(Peaks=unlist(lapply(sigPeaks_in_geneclusters[[i]],length)), 
                                     RNA_Cluster=names(unlist(lapply(sigPeaks_in_geneclusters$allDown_Palbo,length))),
                                     Comparison=names(sigPeaks_in_geneclusters)[[i]], 
                                     MeanPeaks=mean(unlist(lapply(sigPeaks_in_geneclusters[[i]],length))))
}
input_sigpeaks_df <- do.call(rbind, input_sigpeaks_df)
input_sigpeaks_df$RNA_Cluster <- factor(input_sigpeaks_df$RNA_Cluster, levels=cluster_names)

colors <- hm_cols_RNA$Clusters
names(colors) <- cluster_names
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment"

pdf(paste0(pdf_dir, "/ATAC_Number_of_nearby.peaks_MEMEAME_input.pdf"), height=4, width=6)
ggplot(input_sigpeaks_df, aes(x=RNA_Cluster, y=Peaks, label=Peaks, fill=RNA_Cluster))+
  geom_bar(stat="identity", position="dodge")+
  geom_hline(data=input_sigpeaks_df, aes(yintercept = MeanPeaks), color="darkgrey", linetype="dashed")+
  facet_wrap(~Comparison)+geom_text()+
  scale_fill_manual(values=colors)+
    theme(
      axis.text.x = element_text(angle = 90, size = 8),
      axis.text.y = element_text(size = 8),
      axis.title.y = element_text(size = 10)
    ) +
  theme_classic()
dev.off()
```

-------------------------------------------------------------------------------------------------------------------
# *FUNCTIONS*
# save functions for use in other RMarkdown files. 
```{r}
functions_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Functions"
save(process_motif_data, file=paste0(functions_dir, "/process_motif_data.RData"))

save(create_AME_peaks_matrix_byCluster, file=paste0(functions_dir, "/create_AME_peaks_matrix_byCluster.RData"))

# create_AME_dotplots with extra arguments to set fill limits and size limits for "%TP - %FP" & -log10(padj) values. 
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=50, size_limit=4.5 ){
  ggplot(AME_df, aes(x=Cluster, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient(name="-log10(padj)", 
                        low="light yellow", high="dark orange",
                        limits=c(0, fill_limit), 
                        oob=scales::squish) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=c(0, size_limit), 
                          range=c(1, 10)) +  # Adjust dot sizes
    theme_minimal() +
    ggtitle(Title) +
    theme(
      text = element_text(size=8), 
      axis.text.x = element_text(size=8), 
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    ) +
  theme_classic()+
    facet_wrap(~Cluster, scales="free", nrow=1)+
    labs(x="RNA cluster", y="motif ID")
}

save(create_AME_dotplots, file=paste0(functions_dir, "/create_AME_dotplots.RData"))
```
