---
title: "Figure6"
author: "Joanna Yeung"
date: '2025-04-23'
output: html_document
---

```{r}
# data wrangling
library(preprocessCore)
library(readr)
library(readxl)
library(stringr)
library(dplyr)

# plotting
library(gridExtra)
library(ggplot2)
library(pheatmap)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(scales)
library(ggrepel)

# genomics data wrangling and analysis
library(rtracklayer)
library(Biostrings)
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(DESeq2)
library(apeglm)
library(ChIPseeker)
library(clusterProfiler)
library(msigdbr)
library(MotifDb)
library(motifmatchr)

# genome annotation databases
library(EnsDb.Hsapiens.v79)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(BSgenome.Hsapiens.UCSC.hg38)
library(org.Hs.eg.db)
```


# annotation colors for pheatmap
```{r}
hm_cols_RNA <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("Cycling"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "DoxovsPalbo"="#CCEBC5"), 
                Biorep=c("1"="orange", "2"="light blue", "3"="pink"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "DoxovsPalbo"="#CCEBC5"), 
                clusters_cutree8=brewer.pal(8, "Set2"), 
                Clusters=c("5"="#66C2A5","2"="#A6D854","1"="#FC8D62", "7"="#8DA0CB","4"="#E78AC3","6"="#E78AC3","3"="#FFD92F","8"="#E5C494"), 
                SASP = c("Yes"="dark green", "No"="white"),
                Cell_Cycle = c("Yes"="dark blue", "No"="white"), 
                p53_All=c("Yes"="#00B0F0", "No"="white"),
                NFKB_All=c("Yes"="red", "No"="white"))
names(hm_cols_RNA$clusters_cutree8) <- as.factor(1:8)


hm_cols_ATAC <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("None"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "PvsD"="#CCEBC5"),
                Biorep=c("1"="orange", "2"="light blue", "3"="pink", "4"="light green"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "PvsD"="#CCEBC5"),
                Peak_anno=c("Distal"="#B15A28", "Gene Body"="#7F54A2", "Promoter"="#A7CEE2"), 
                Doxo=c("Yes"="#B3C543", "No"="white"),
                Palbo=c("Yes"="#E0ADCD", "No"="white")
                )
```

# load in workspace with DESeq results and rlog counts matrix of genes. 
```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure2/workspace.RData")
```

#### universe (background gene set) for GO enrichment analysis is all genes expressed
```{r}
universe <- genes_anno_type[match(gsub('\\.[0-9]*$', '', row.names(acrossGroups)), genes_anno_type$ensembl_id), ]
```

# make function to specify code chunk to run from other Rmd file to avoid copying and pasting large lines of code. 
```{r}
# arguments: 
# Rmd_file=file path & name of Rmd file with code chunk
# chunk_name= name of code chunk to run from Rmd file. 
run_code_chunk <- function(Rmd_file, chunk_name){ 
  # Read the Rmd file
lines <- readLines(Rmd_file)

# Identify the start and end of the specific chunk
chunk_start <- grep(paste0("^```\\{r\\s*", chunk_name), lines)  # Find chunk start
chunk_end <- grep("^```$", lines)  # Find all chunk end lines

# Find the first `chunk_end` after `chunk_start`
chunk_end <- chunk_end[chunk_end > chunk_start][1]

  if (is.na(chunk_end)) {
    stop(paste("End of chunk", chunk_name, "not found. Check the Rmd file for missing delimiters."))
  }

# specify lines inside chunk, removing lines that make up the start and end of the chunk. 
chunk_to_run <- lines[(chunk_start[1]+1):(chunk_end[1]-1)]

# Evaluate the extracted code
  eval(parse(text = paste(chunk_to_run, collapse = "\n")), envir = .GlobalEnv)
}
```

# ATAC-seq: plot Tn5 insertion sites around RELA motifs to show shared increased accessibility at putative NF-kB binding sites (all up peaks)
```{r}
# filter for GRanges of up peaks in each drug and cell cycle arrest category 
Up_DrugandArrestCat <- list(
  PalboUpvsCyc=allUpvsCycling$Palbo[!allUpvsCycling$Palbo %over% allUpvsCycling$Doxo, ], 
  sharedUpvsCyc=allUpvsCycling$Palbo[allUpvsCycling$Palbo %over% allUpvsCycling$Doxo, ],
  DoxoUpvsCyc=allUpvsCycling$Doxo[!allUpvsCycling$Doxo %over% allUpvsCycling$Palbo, ], 
  PalboUpvsday3=allUpvsday3$Palbo[!allUpvsday3$Palbo %over% allUpvsday3$Doxo, ], 
  sharedUpvsday3=allUpvsday3$Palbo[allUpvsday3$Palbo %over% allUpvsday3$Doxo, ], 
  DoxoUpvsday3=allUpvsday3$Doxo[!allUpvsday3$Doxo %over% allUpvsday3$Palbo, ])

# filter for GRanges of up peaks in each drug and cell cycle arrest category but without specifying shared up peaks. 
Up_DrugandArrest_PvsD <- list(
  PalboUpvsCyc=allUpvsCycling$Palbo, 
  DoxoUpvsCyc=allUpvsCycling$Doxo, 
  PalboUpvsday3=allUpvsday3$Palbo,
  DoxoUpvsday3=allUpvsday3$Doxo)

# load in HOCOMOCO motif sequence database
load("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/motifs/HOCOMOCO_v11_matrix.RData")
# scan for RELA motifs
TF65_HMv11 <- matchPWM(HOCOMOCO_v11_matrix$TF65_HUMAN.H11MO.0.A, BSgenome.Hsapiens.UCSC.hg38, with.score=T)
# scan for NFKB1 motifs
NFKB1_HMv11 <- matchPWM(HOCOMOCO_v11_matrix$NFKB1_HUMAN.H11MO.1.B, BSgenome.Hsapiens.UCSC.hg38, with.score=T)

# filter for RELA motifs that overlap with up peaks
TF65_ATAC_Up_DrugandArrestCat <- lapply(Up_DrugandArrestCat, function(x){TF65_HMv11[TF65_HMv11 %over% x]})
TF65_ATAC_Up_DrugandArrest_PvsD <- lapply(Up_DrugandArrest_PvsD, function(x){TF65_HMv11[TF65_HMv11 %over% x]})

# Export each GRanges object as a BED file
for(i in 1:length(TF65_ATAC_Up_DrugandArrestCat)){
export.bed(TF65_ATAC_Up_DrugandArrestCat[[i]], con = paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds/", names(TF65_ATAC_Up_DrugandArrestCat)[i], "_TF65_HMv11.bed"))
}

for(i in 1:length(TF65_ATAC_Up_DrugandArrest_PvsD)){
export.bed(TF65_ATAC_Up_DrugandArrest_PvsD[[i]], con = paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds/", names(TF65_ATAC_Up_DrugandArrest_PvsD)[i], "_PvsD_TF65_HMv11.bed"))
}

# generate input files for deepTools scripts
# directory paths to bed files 
writeLines(dir("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds", full.names=T, pattern=".bed"), con="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds/regions.txt")

# filenames to be assigned to plots. 
filenames <- gsub(".bed", "", dir("//lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds", pattern=".bed"))
writeLines(filenames, con="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds/filename.txt")
```

# ATAC-seq: submit bash job to plot around TF65 motifs in all up peaks
```{bash}
mergedcuts="/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/Cycling_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day3_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day10_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day14_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day21_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day3_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day10_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day14_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day21_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day28_Palbo_scaleFactornorm_cutsCoverage.bw"
samplenames="Cycling day3_Doxo day10_Doxo day14_Doxo day21_Doxo day3_Palbo day10Palbo day14_Palbo day21_Palbo day28_Palbo"

cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/output_plots
sbatch -p risc,hpc --array=1-10 \
--export=bigwigs="$mergedcuts",\
regions=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds/regions.txt,\
samplenames="$samplenames",\
profile_outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/output_plots/plotProfile,\
heatmap_outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/output_plots/plotHeatmap,\
filename=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds/filename.txt \
/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/deeptools_Coverage_1kb.slurm
```

# ATAC-seq: plot Tn5 insertion sites around NFKB1 motifs to show shared increased accessibility at putative NF-kB binding sites (up peaks within 50kb of sig changing gene clusters )
```{r}
# merge RNA clusters of the same drug category.
merge_geneclusters <- list("CellCycle"=genes_in_clusters[[1]], 
                           "Palbo"=unlist(genes_in_clusters[2:3]), 
                           "Doxo"=unlist(genes_in_clusters[4:5]), 
                           "Shared"=genes_in_clusters[[6]]
)

# import gtf file for genes
genes_anno <- import("/lustre/fs4/risc_lab/store/jyeung/references/gencode.v41.GRCh38.p13.annotation.gtf")
genes_anno <- genes_anno[genes_anno$type %in% "gene", ] # filter for regions that are only genes 
genes_anno$gene_id <- gsub('\\.[0-9]*$', '', genes_anno$gene_id) # remove numbers after . in gene id column 
genes_anno <- genes_anno[ ,c("source", "type", "gene_id", "gene_type", "gene_name", "level")] # keep only these columns


# get the genomic coordinates of significantly changing genes from merged clusters (RNA-seq data)
merge_geneclusters_anno <- lapply(merge_geneclusters, function(merge_geneclusters){genes_anno[genes_anno$gene_name %in% merge_geneclusters]})

# make GRanges where the coordinates are 50kb upstream of gene TSS and 50kb downstream of end of gene. 
merge_geneclusters_anno_findpeaks <- merge_geneclusters_anno
for(i in 1:length(merge_geneclusters_anno)){
  start(merge_geneclusters_anno_findpeaks[[i]]) <- start(merge_geneclusters_anno[[i]])-50000
  end(merge_geneclusters_anno_findpeaks[[i]]) <- end(merge_geneclusters_anno[[i]])+50000
}

# find peaks that overlap within 50kb of significantly changing genes in clusters
peaks_in_geneclusters <- lapply(merge_geneclusters_anno_findpeaks, function(merge_geneclusters_anno_findpeaks){annomasterpeaks_pval0.2[annomasterpeaks_pval0.2 %over% merge_geneclusters_anno_findpeaks, ]})

# filter for sig changing peaks within 50kb of sig changing genes in clusters. 
Up_PvsD_gene_clusters <- list()
for(i in 1:length(Up_DrugandArrest_PvsD)){
Up_PvsD_gene_clusters[[i]] <-lapply(peaks_in_geneclusters, function(peaks_in_geneclusters){Up_DrugandArrest_PvsD[[i]][Up_DrugandArrest_PvsD[[i]] %over% peaks_in_geneclusters, ]}) 
}
names(Up_PvsD_gene_clusters) <- names(Up_DrugandArrest_PvsD)

# filter for NFKB1 motifs that overlap with up peaks nearby sig changing genes 
NFKB1_Up_PvsD_gene_clusters <- list()
for(i in 1:length(Up_PvsD_gene_clusters)){
NFKB1_Up_PvsD_gene_clusters[[i]] <- lapply(Up_PvsD_gene_clusters[[i]], function(x){NFKB1_HMv11[NFKB1_HMv11 %over% x]})
}
names(NFKB1_Up_PvsD_gene_clusters) <- names(Up_PvsD_gene_clusters)

directory <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds"
  
# Export each GRanges object as a BED file
for(i in 1:length(NFKB1_Up_PvsD_gene_clusters)){
  for(j in 1:length(NFKB1_Up_PvsD_gene_clusters[[i]])){
export.bed(NFKB1_Up_PvsD_gene_clusters[[i]][[j]], con = paste0(directory, "/", names(NFKB1_Up_PvsD_gene_clusters)[i], "_", names(NFKB1_Up_PvsD_gene_clusters[[i]])[j], "_genes", "_PvsD_NFKB1_HMv11.bed"))
  }
}


# generate input files for deepTools scripts
# directory paths to bed files 
writeLines(dir(directory, full.names=T, pattern="NFKB1.*\\.bed$"), con=paste0(directory, "/NFKB1_regions.txt"))

# filenames to be assigned to plots. 
filenames <- gsub(".bed", "", dir(directory, pattern="NFKB1.*\\.bed$"))
writeLines(filenames, con=paste0(directory, "/NFKB1_filename.txt"))
```

# ATAC-seq: submit bash job to plot around NFKB1 motifs in up peaks within 50kb of sig changing genes. 
```{bash}
mergedcuts="/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/Cycling_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day3_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day10_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day14_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day21_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day3_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day10_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day14_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day21_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day28_Palbo_scaleFactornorm_cutsCoverage.bw"
samplenames="Cycling day3_Doxo day10_Doxo day14_Doxo day21_Doxo day3_Palbo day10Palbo day14_Palbo day21_Palbo day28_Palbo"

cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/output_plots
sbatch -p risc,hpc --array=1-16 \
--export=bigwigs="$mergedcuts",\
regions=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds/NFKB1_regions.txt,\
samplenames="$samplenames",\
profile_outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/output_plots/plotProfile,\
heatmap_outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/output_plots/plotHeatmap,\
filename=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/input_beds/NFKB1_filename.txt \
/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/deepTools_files/deeptools_Coverage_1kb.slurm
```

# ATAC-seq: import motif enrichment results under ALL up peaks across all timepoint comparisons & drug treatment. 
```{r}
annoUpPeaks_AME_df <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_output/DESeq/annoUp", expressed_genes = expressedGenes)

annoUpPeaks_AME_df$Comparison <- gsub("\\..*", "", row.names(annoUpPeaks_AME_df))
annoUpPeaks_AME_df$Comparison <- factor(annoUpPeaks_AME_df$Comparison, levels=unique(annoUpPeaks_AME_df$Comparison)[c(15,1,5,9,2,6,10,16,3,7,11,13,4,8,12,14)]) # column to identify DESeq comparison

annoUpPeaks_AME_df$Group <- ifelse(grepl("DoxovsCycling", annoUpPeaks_AME_df$Comparison), "DoxovsCycling", ifelse(grepl("PalbovsCycling", annoUpPeaks_AME_df$Comparison), "PalbovsCycling", ifelse(grepl("Doxovsday3Doxo", annoUpPeaks_AME_df$Comparison), "Doxovsday3Doxo", "Palbovsday3Palbo"))) # column to identify relative to cycling or day 3 comparison and drug treatment. 

annoUpPeaks_AME_df$Timepoint <-  gsub("(day[0-9]{1,2}).*", "\\1", annoUpPeaks_AME_df$Comparison) # column to identify which timepoint is compared to Cycling or day 3 controls, by removing everything after day[0-9] where [0-9] is 1 or 2 digits in the Comparison column. 
annoUpPeaks_AME_df$Timepoint <- factor(annoUpPeaks_AME_df$Timepoint, levels=c("day3", "day10", "day14", "day21", "day28"))
```

# ATAC-seq: get peak ids for peaks that are true positive peaks for NFKB motif from motif enrichment analysis (MEME-AME)
```{r}
input_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_output/DESeq/annoUp"

out_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/NFKB_tp_peaks"

# generate input file containing directory paths to sequences.tsv file from MEME-AME output. 
writeLines(paste0(dir(input_dir, full.names = T), "/sequences.tsv"), con=paste0(out_dir, "/DESeq_annoUp_sequencefiles.txt"))
```

# ATAC-seq: extract peak ids of true positive peaks of NFKB from sequences.tsv file from MEME-AME output
```{bash}
cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/NFKB_tp_peaks

# for all NFKB motifs
sbatch -p risc --array=1-16 \
--export=outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/NFKB_tp_peaks,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/NFKB_tp_peaks/DESeq_annoUp_sequencefiles.txt,\
names=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq/annoUp/folderlist_annoUp_DESeq_peaks.txt,\
motif="NFKB" \
get_tp_peaks_motifs.slurm

rm $(ls|egrep '.err|.out')

# for all REL motifs
sbatch -p risc --array=1-16 \
--export=outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/NFKB_tp_peaks,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/NFKB_tp_peaks/DESeq_annoUp_sequencefiles.txt,\
names=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq/annoUp/folderlist_annoUp_DESeq_peaks.txt,\
motif="REL" \
get_tp_peaks_motifs.slurm

rm $(ls|egrep '.err|.out')
```

```{r}
# import peak ids of true positive down peaks containing any NFKB motif
setwd("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6/NFKB_tp_peaks")

file_ids <- dir()
all_NFKB_tp_peaks <- list()

for (i in seq_along(file_ids)) {
  file_path <- file_ids[i]  # Get full file name

  # Check if file is empty
  if (file.info(file_path)$size > 0) {
    all_NFKB_tp_peaks[[file_path]] <- read_tsv(file_path, col_names = FALSE) %>% pull(X1)
  }
}

names(all_NFKB_tp_peaks) <- gsub("_all_tp_Uppeaks_ids\\.tsv", "", names(all_NFKB_tp_peaks))
all_NFKB_tp_peaks <- all_NFKB_tp_peaks[!names(all_NFKB_tp_peaks) %in% c("DESeq_annoUp_sequencefiles.txt", "get_tp_peaks_motifs.slurm")]
```

#*FIGURES*

# ATAC-seq: plot motif enrichment results in the NFKB family for ALL up peaks across all timepoints and drug treatments (DESeq comparisons)
```{r}
# load in function to generate matrix of -log10(padj) values of motif enrichment results
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/Functions/create_AME_peaks_matrix.RData")

# function to create dotplots for motif enrichment results for each DESeq comparison. 
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=c(1,10), size_limit=4.5){
  AME_df <- AME_df %>% mutate(Comparison = Timepoint)
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient2(name="-log10(padj)", 
                        low="white", mid="light yellow", high="dark orange",
                        midpoint = 3,
                        limits=fill_limit) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=size_limit, 
                          range=c(1, 10)) +  # Adjust dot sizes
    theme_minimal() +
    ggtitle(Title) +
    scale_x_discrete(labels = function(x) AME_df$Timepoint[match(x, AME_df$Comparison)]) +  # to replace Comparison with Cluster labels on the x-axis. 
    facet_wrap(~Group, nrow=1, scales="free") +
    theme(
      axis.text.x = element_text(size=8, angle = 90, hjust=1),  # Rotate x-axis labels
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    )+
    theme_classic() +
    labs(x="Comparison", y="motif ID")}

NFkB_UpPeaks_AME_mat <- create_AME_peaks_matrix(annoUpPeaks_AME_df[grepl("NFKB|REL", annoUpPeaks_AME_df$motif_ID), ])
                                                      
NFkB_UpPeaks_AME_mat <- NFkB_UpPeaks_AME_mat[order(rowSums(NFkB_UpPeaks_AME_mat), decreasing=T), ] # order by overall most to least enriched NFkB motifs 

outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6"

# plot motif enrichment values of NFkB motifs in all Up peaks for all timepoints vs Cycling. 
pdf(paste0(outdir, "/ATAC_NFkB_UpPeaks_AME_dotplots.pdf"), width=8, height=4)
create_AME_dotplots(annoUpPeaks_AME_df[grepl("NFKB|REL", annoUpPeaks_AME_df$motif_ID), ], Title="NFkB Motifs enriched in Up peaks", fill_limit=c(2, 92), size_limit=c(0, 17.4))
dev.off()
```

# ATAC-seq: make sequence Logo from HOCOMOCO NFKB1 motif. 
```{r}
outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6"
library(seqLogo)
pdf(paste0(outdir, "/NFKB1_HUMAN.H11MO.1.B_HOCOMOCO_seqLogo.pdf"), width=6, height=4)
seqLogo(HOCOMOCO_v11_matrix$NFKB1_HUMAN.H11MO.1.B)
dev.off()
```

# plot venn diagram showing overlap of NFKB1 containing up peaks between Doxo & Palbo against number of all up peaks 
```{r}
# write function to create venn diagrams that are proportional to size of each category. 
create_proportional_venn <- function(sets, set_labels, colors = c("#E0ADCD", "#B3C543", "#A0ADCD")) {
  library(eulerr)
  # Check if sets and labels are provided correctly
  if (length(sets) != length(set_labels)) {
    stop("The number of sets and set_labels must be the same.")
  }
  
  # Prepare the data for eulerr
  venn_data <- sets
  names(venn_data) <- set_labels
  
  # Use eulerr to calculate the Venn diagram
  venn_fit <- euler(venn_data)
  
  # Plot the Venn diagram
  plot(venn_fit, 
       fills = colors, 
       labels = list(cex = 1, fontfamily = "Helvetica"),
       quantities = list(cex = 1),
       edges = list(lwd = 2))
}

# unique peak ids for Doxo vs Cycling up tp peaks with NFKB motifs
tp_uppeaks_Doxo <- unique(unlist(all_NFKB_tp_peaks[grepl("Doxo", names(all_NFKB_tp_peaks))]))

# unique peak ids for Palbo vs Cycling up tp peaks with NFKB motifs
tp_uppeaks_Palbo <- unique(unlist(all_NFKB_tp_peaks[grepl("Palbo", names(all_NFKB_tp_peaks))]))

# peak ids for all up peaks 
all_uppeaks_ids <- unique(c(mergedDiffGR$allUp_Doxo$name, mergedDiffGR$allUp_Palbo$name))

outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure6"

# plot venn diagram showing overlap of NFKB1 containing up peaks between Doxo & Palbo against number of all up peaks
pdf(paste0(outdir, "/ATAC_NFKB1_tp_uppeaks_PalbovsDoxo_overlap_venn.pdf"), width=6, height=4)
create_proportional_venn(sets=list(Palbo=tp_uppeaks_Palbo, Doxo=tp_uppeaks_Doxo, allup=all_uppeaks_ids), set_labels = c("Palbo", "Doxo", "all up"), colors = c("#E0ADCD", "#B3C543", "#F6F2EA"))
dev.off()
```