---
title: "FigureS4: Differentially enriched pathways between different types of therapy induced senescence, related to Figure 4"
author: "Joanna Yeung"
date: '2025-04-23'
output: html_document
---
```{r}
# data wrangling
library(preprocessCore)
library(readr)
library(readxl)
library(stringr)
library(dplyr)

# plotting
library(gridExtra)
library(ggplot2)
library(pheatmap)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(scales)
library(ggrepel)

# genomics data wrangling and analysis
library(rtracklayer)
library(Biostrings)
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(DESeq2)
library(apeglm)
library(ChIPseeker)
library(clusterProfiler)
library(msigdbr)
library(MotifDb)
library(motifmatchr)

# genome annotation databases
library(EnsDb.Hsapiens.v79)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(BSgenome.Hsapiens.UCSC.hg38)
library(org.Hs.eg.db)
```

# make function to specify code chunk to run from other Rmd file to avoid copying and pasting large lines of code. 
```{r}
# arguments: 
# Rmd_file=file path & name of Rmd file with code chunk
# chunk_name= name of code chunk to run from Rmd file. 
run_code_chunk <- function(Rmd_file, chunk_name){ 
  # Read the Rmd file
lines <- readLines(Rmd_file)

# Identify the start and end of the specific chunk
chunk_start <- grep(paste0("^```\\{r\\s*", chunk_name), lines)  # Find chunk start
chunk_end <- grep("^```$", lines)  # Find all chunk end lines

# Find the first `chunk_end` after `chunk_start`
chunk_end <- chunk_end[chunk_end > chunk_start][1]

  if (is.na(chunk_end)) {
    stop(paste("End of chunk", chunk_name, "not found. Check the Rmd file for missing delimiters."))
  }

# specify lines inside chunk, removing lines that make up the start and end of the chunk. 
chunk_to_run <- lines[(chunk_start[1]+1):(chunk_end[1]-1)]

# Evaluate the extracted code
  eval(parse(text = paste(chunk_to_run, collapse = "\n")), envir = .GlobalEnv)
}
```

# annotation colors for pheatmap
```{r}
hm_cols_RNA <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("Cycling"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD"), 
                Biorep=c("1"="orange", "2"="light blue", "3"="pink"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "DoxovsPalbo"="#CCEBC5"), 
                Clusters=c("2"="#66C2A5","5"="#A6D854","7"="#8DA0CB", "1"="#FC8D62", "4"="#E78AC3","6"="#FFD92F","3"="#E5C494"), 
                SASP = c("Yes"="dark green", "No"="white"),
                Cell_Cycle = c("Yes"="dark blue", "No"="white"), 
                p53_All=c("Yes"="#00B0F0", "No"="white"),
                NFKB_All=c("Yes"="red", "No"="white"))
names(hm_cols_RNA$clusters_cutree8) <- as.factor(1:8)


hm_cols_ATAC <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("None"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "PvsD"="#CCEBC5"),
                Biorep=c("1"="orange", "2"="light blue", "3"="pink", "4"="light green"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "PvsD"="#CCEBC5"),
                Peak_anno=c("Distal"="#B15A28", "Gene Body"="#7F54A2", "Promoter"="#A7CEE2"), 
                Doxo=c("Yes"="#B3C543", "No"="white"),
                Palbo=c("Yes"="#E0ADCD", "No"="white")
                )
```

# RNA-seq: run code chunks generating GO enrichment results
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/Figure4.Rmd", chunk_name="get_msigdbr")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/Figure4.Rmd", chunk_name="getGOenrichment function")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/Figure4.Rmd", chunk_name="genes_in_clusters")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/Figure4.Rmd", chunk_name="RNA-seq GO enrichment analysis")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/Figure4.Rmd", chunk_name="RNAseq GO enrichment results")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/Figure4.Rmd", chunk_name="RNAseq GSEA analysis")
```

# RNA-seq: plot padj values of GO enrichment results for top 5 drug distinct up genes
```{r}
# Still showing the pvalues of the other clusters
GO_results_mat_clusters <- list()
for(i in 1:length(GO_results_list)){

GO_results_mat_clusters[[i]] <- GO_results_list_mat[row.names(GO_results_list_mat) %in% GO_results_list[[i]]$ID[1:25], ]
GO_results_mat_clusters[[i]] <- GO_results_mat_clusters[[i]][order(GO_results_mat_clusters[[i]][ ,i], decreasing=T), ]
}

# plot padj values for top 5 GO terms for each cluster individually. Still showing the pvalues of the other clusters as columns but only the top GO terms from the cluster of interest. 
GO_results_mat_clusters <- list()
for(i in 1:length(GO_results_list)){

GO_results_mat_clusters[[i]] <- GO_results_list_mat[row.names(GO_results_list_mat) %in% GO_results_list[[i]]$ID[1:10], ]
GO_results_mat_clusters[[i]] <- GO_results_mat_clusters[[i]][order(GO_results_mat_clusters[[i]][ ,i], decreasing=T), ]
}
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS4"
pdf(paste0(pdf_dir, "/GO_results_Palbo1_clusters.pdf"), width=8, height=2)
pheatmap(GO_results_mat_clusters[[3]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 20,length.out = 100), cluster_rows = F, main=cluster_names[3], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()

pdf(paste0(pdf_dir, "/GO_results_Palbo2_clusters.pdf"), width=5.5, height=2)
pheatmap(GO_results_mat_clusters[[4]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 10,length.out = 100), cluster_rows = F, main=cluster_names[4], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()

pdf(paste0(pdf_dir, "/GO_results_Doxo1_clusters.pdf"), width=9, height=2)
pheatmap(GO_results_mat_clusters[[5]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 5,length.out = 100), cluster_rows = F, main=cluster_names[5], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()

pdf(paste0(pdf_dir, "/GO_results_Doxo2_clusters.pdf"), width=6.5, height=2)
pheatmap(GO_results_mat_clusters[[6]][1:5, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0, 8,length.out = 100), cluster_rows = F, main=cluster_names[6], annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()
```

# create function that gets GSEA enrichmentScores that trend opposite between drugs to figure out which enriched terms to plot runningScore plots with.
```{r get_opposing_GSEA_enrichment}
get_opposing_GSEA_enrichment <- function(Results_enriched=GSEA_Results_H$`Palbo14 vs Cycling0`, Results_opposite=GSEA_Results_H$`Doxo14 vs Cycling0`, names=c("Palbo", "Doxo"), database="Hallmark") {
  # Extract positively enriched terms in drug114 vs Cycling0
  GSEApos <- Results_enriched[
    Results_enriched$enrichmentScore > 0, 
    c("ID", "enrichmentScore", "p.adjust")
  ]
  
 # Find terms that are positively enriched in drug1 but negatively enriched in drug2
GSEAopp <- Results_opposite[
  Results_opposite$ID %in% GSEApos$ID & 
    Results_opposite$enrichmentScore < 0, 
  c("ID", "enrichmentScore", "p.adjust")
]

# Initialize an empty dataframe for opp_df
opp_df <- data.frame()

# Only proceed if GSEAopp has rows
if (nrow(GSEAopp) > 0) {
  opp_df <- data.frame(
    ID = GSEAopp$ID,
    Resultsopp_enrichmentScore = GSEAopp$enrichmentScore,
    Resultsenr_enrichmentScore = GSEApos[match(GSEAopp$ID, GSEApos$ID), ]$enrichmentScore,
    database = database
  )
  
  # Sort by drug1 enrichment score in descending order
  opp_df <- opp_df[order(opp_df$Resultsenr_enrichmentScore, decreasing = TRUE), ]
}
  # Find enriched terms in drug1 that are not enriched in drug2
  enriched_GSEA <- Results_enriched[
    !Results_enriched$ID %in% Results_opposite$ID, 
    c("ID", "enrichmentScore", "p.adjust")
  ]
  
  # Sort by drug1 enrichment score in descending order
  enriched_GSEA <- enriched_GSEA[order(enriched_GSEA$enrichmentScore, decreasing = TRUE), ]
  
  # Create dataframe for terms enriched in drug1 but not found in drug2
  notinopp_df <- data.frame(
    ID = enriched_GSEA$ID,
    Resultsopp_enrichmentScore = 0,
    Resultsenr_enrichmentScore = enriched_GSEA$enrichmentScore,
   database= database
  )
  
  # Merge results
  positiveenr_Results1 <- rbind(opp_df, notinopp_df)
  colnames(positiveenr_Results1) <- c("ID", paste0(names[2], "enrichmentScore"), paste0(names[1], "enrichmentScore"), "database")
  return(positiveenr_Results1 )
}
```

# RNA-seq: find GSEA enrichment Scores that are opposing between drug treatments
```{r}
# find positive enriched terms in day 14 Palbo vs Cycling
Palbo_enriched_GSEA <- rbind(get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_H$`Palbo14 vs Cycling0`, Results_opposite=GSEA_Results_H$`Doxo14 vs Cycling0`, names=c("Palbo", "Doxo"), database="Hallmark"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_KEGG$`Palbo14 vs Cycling0`, Results_opposite=GSEA_Results_KEGG$`Doxo14 vs Cycling0`, names=c("Palbo", "Doxo"), database="KEGG"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_REACT$`Palbo14 vs Cycling0`, Results_opposite=GSEA_Results_REACT$`Doxo14 vs Cycling0`, names=c("Palbo", "Doxo"), database="REACTOME"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_MF$`Palbo14 vs Cycling0`, Results_opposite=GSEA_Results_MF$`Doxo14 vs Cycling0`, names=c("Palbo", "Doxo"), database="GO:MF")
                            )
# find positive enriched terms in day 14 Doxo vs Cycling
Doxo_enriched_GSEA <- rbind(get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_H$`Doxo14 vs Cycling0`, Results_opposite=GSEA_Results_H$`Palbo14 vs Cycling0`, names=c("Doxo", "Palbo"), database="Hallmark"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_KEGG$`Doxo14 vs Cycling0`, Results_opposite=GSEA_Results_KEGG$`Palbo14 vs Cycling0`, names=c("Doxo", "Palbo"), database="KEGG"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_REACT$`Doxo14 vs Cycling0`, Results_opposite=GSEA_Results_REACT$`Palbo14 vs Cycling0`, names=c("Doxo", "Palbo"), database="REACTOME"), 
                            get_opposing_GSEA_enrichment(Results_enriched=GSEA_Results_MF$`Doxo14 vs Cycling0`, Results_opposite=GSEA_Results_MF$`Palbo14 vs Cycling0`, names=c("Doxo", "Palbo"), database="GO:MF")
                            )
# Palbo: order by highest to lowest enrichment scores and corresponding Doxo enrichment Score is negative
Palbo_enriched_GSEA <- Palbo_enriched_GSEA[Palbo_enriched_GSEA$DoxoenrichmentScore < 0, ]
Palbo_enriched_GSEA <- Palbo_enriched_GSEA[order(Palbo_enriched_GSEA$PalboenrichmentScore, decreasing=T), ]

# Doxo: order by highest to lowest enrichment scores and corresponding Palbo enrichment Score is negative
Doxo_enriched_GSEA <- Doxo_enriched_GSEA[order(Doxo_enriched_GSEA$DoxoenrichmentScore, decreasing=T) & Doxo_enriched_GSEA$PalboenrichmentScore < 0, ]
```

# RNA-seq: save GSEA running score enriched terms trending in opposite directions between drugs
```{r}
library(enrichplot)
library(patchwork)

# Define the gene sets and corresponding titles
genesetID <- c("ANGIOGENESIS", "EPITHELIAL MESENCHYMAL TRANSITION")
# Get padj value of GSEA results to be plotted
GSEA_padj <- data.frame(ID=GSEA_Results_H$`Palbo14 vs Cycling0`@result[GSEA_Results_H$`Palbo14 vs Cycling0`@result$ID %in% genesetID, ]$ID, 
                        palbo_p.adjust= GSEA_Results_H$`Palbo14 vs Cycling0`@result[GSEA_Results_H$`Palbo14 vs Cycling0`@result$ID %in% genesetID, ]$p.adjust, 
                        doxo_p.adjust= GSEA_Results_H$`Doxo14 vs Cycling0`@result[GSEA_Results_H$`Doxo14 vs Cycling0`@result$ID %in% genesetID, ]$p.adjust)

# Open PDF for saving plots
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS4"
pdf(paste0(pdf_dir, "/GSEA_day14Doxo_day14Palbo_vs_Cyc_Hallmark.pdf"), width=3, height=5)
# Loop through each gene set
for (i in 1:length(genesetID)) {
  
  # Generate GSEA plots for Doxo and Palbo
  p_doxo <- gseaplot(GSEA_Results_H$`Doxo14 vs Cycling0`, geneSetID = genesetID[i], by="runningScore", 
                      title = paste("day 14 Doxo vs Cycling:", genesetID[i]), 
                      color.line="#B3C543", color="#B3C543", color.vline ="#B3C543") +
            theme_classic() + ylim(-0.75, 0.85)
  
  p_palbo <- gseaplot(GSEA_Results_H$`Palbo14 vs Cycling0`, geneSetID = genesetID[i], by="runningScore", 
                      title = paste("day 14 Palbo vs Cycling:", genesetID[i]), 
                      color.line="#E0ADCD", color="#E0ADCD", color.vline ="#E0ADCD") +
            theme_classic() + ylim(-0.75, 0.85)
  
  # Plot the figures
  print(plot_grid(p_doxo, p_palbo, ncol=1))
}

# Close PDF
dev.off()

# Define the gene sets and corresponding titles
genesetID <- c("CROSSLINKING OF COLLAGEN FIBRILS", "ECM PROTEOGLYCANS", "INTEGRIN CELL SURFACE INTERACTIONS", "MEIOTIC RECOMBINATION", "DISEASES OF PROGRAMMED CELL DEATH", "DEPOSITION OF NEW CENPA CONTAINING NUCLEOSOMES AT THE CENTROMERE", "OXIDATIVE STRESS INDUCED SENESCENCE")
titles <- genesetID
# Get padj value of GSEA results to be plotted
GSEA_padj <- data.frame(ID=GSEA_Results_REACT$`Palbo14 vs Cycling0`@result[GSEA_Results_REACT$`Palbo14 vs Cycling0`@result$ID %in% genesetID, ]$ID, 
                        palbo_p.adjust= GSEA_Results_REACT$`Palbo14 vs Cycling0`@result[GSEA_Results_REACT$`Palbo14 vs Cycling0`@result$ID %in% genesetID, ]$p.adjust, 
                        doxo_p.adjust= GSEA_Results_REACT$`Doxo14 vs Cycling0`@result[GSEA_Results_REACT$`Doxo14 vs Cycling0`@result$ID %in% genesetID, ]$p.adjust)

# Open PDF for saving plots
pdf(paste0(pdf_dir, "/GSEA_day14Doxo_day14Palbo_vs_Cyc_REACTOME.pdf"), width=3, height=5)
# Loop through each gene set
for (i in 1:length(genesetID)) {
  
  # Generate GSEA plots for Doxo and Palbo
  p_doxo <- gseaplot(GSEA_Results_REACT$`Doxo14 vs Cycling0`, geneSetID = genesetID[i], by="runningScore", 
                      title = paste("day 14 Doxo vs Cycling:", genesetID[i]), 
                      color.line="#B3C543", color="#B3C543", color.vline ="#B3C543") +
            theme_classic() + ylim(-0.8, 0.85)
  
  p_palbo <- gseaplot(GSEA_Results_REACT$`Palbo14 vs Cycling0`, geneSetID = genesetID[i], by="runningScore", 
                      title = paste("day 14 Palbo vs Cycling:", genesetID[i]), 
                      color.line="#E0ADCD", color="#E0ADCD", color.vline ="#E0ADCD") +
            theme_classic() + ylim(-0.8, 0.85)
  
  # Plot the figures
  print(plot_grid(p_doxo, p_palbo, ncol=1))
}

# Close PDF
dev.off()
```

