---
title: "Untitled"
author: "Joanna Yeung"
date: "2026-04-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# load in workspace with DESeq results and rlog counts matrix of genes. 
```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure2/workspace.RData")
```

# ATAC-seq: import motif enrichment results under ALL up peaks across all timepoint comparisons & drug treatment. 
```{r}
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL
expressedGenes[grep("TP53$|TP63$|TP73$", expressedGenes)] <- c("P63", "P73", "P53")
expressedGenes[expressedGenes %in% "RELA"] <- "TF65"
annoUpPeaks_AME_df <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_output/DESeq/annoUp", expressed_genes = expressedGenes)

annoUpPeaks_AME_df$Comparison <- gsub("\\..*", "", row.names(annoUpPeaks_AME_df))
annoUpPeaks_AME_df$Comparison <- factor(annoUpPeaks_AME_df$Comparison, levels=unique(annoUpPeaks_AME_df$Comparison)[c(15,1,5,9,2,6,10,16,3,7,11,13,4,8,12,14)]) # column to identify DESeq comparison

annoUpPeaks_AME_df$Group <- ifelse(grepl("DoxovsCycling", annoUpPeaks_AME_df$Comparison), "DoxovsCycling", ifelse(grepl("PalbovsCycling", annoUpPeaks_AME_df$Comparison), "PalbovsCycling", ifelse(grepl("Doxovsday3Doxo", annoUpPeaks_AME_df$Comparison), "Doxovsday3Doxo", "Palbovsday3Palbo"))) # column to identify relative to cycling or day 3 comparison and drug treatment. 

annoUpPeaks_AME_df$Timepoint <-  gsub("(day[0-9]{1,2}).*", "\\1", annoUpPeaks_AME_df$Comparison) # column to identify which timepoint is compared to Cycling or day 3 controls, by removing everything after day[0-9] where [0-9] is 1 or 2 digits in the Comparison column. 
annoUpPeaks_AME_df$Timepoint <- factor(annoUpPeaks_AME_df$Timepoint, levels=c("day3", "day10", "day14", "day21", "day28"))
```

# ATAC-seq: plot motif enrichment results in the AP1 family for ALL up peaks across all timepoints and drug treatments (DESeq comparisons)
```{r}
# load in function to generate matrix of -log10(padj) values of motif enrichment results
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/Functions/create_AME_peaks_matrix.RData")

# function to create dotplots for motif enrichment results for each DESeq comparison. 
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=c(1,10), size_limit=4.5){
  AME_df <- AME_df %>% mutate(Comparison = Timepoint)
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient2(name="-log10(padj)", 
                        low="white", mid="light yellow", high="dark orange",
                        midpoint = 3,
                        limits=fill_limit) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=size_limit, 
                          range=c(1, 10)) +  # Adjust dot sizes
    theme_minimal() +
    ggtitle(Title) +
    scale_x_discrete(labels = function(x) AME_df$Timepoint[match(x, AME_df$Comparison)]) +  # to replace Comparison with Cluster labels on the x-axis. 
    facet_wrap(~Group, nrow=1, scales="free") +
    theme(
      axis.text.x = element_text(size=8, angle = 90, hjust=1),  # Rotate x-axis labels
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    )+
    theme_classic() +
    labs(x="Comparison", y="motif ID")}

AP1_UpPeaks_AME_mat <- create_AME_peaks_matrix(annoUpPeaks_AME_df[grepl("JUN|FOS|ATF", annoUpPeaks_AME_df$motif_ID), ])
                                                      
AP1_UpPeaks_AME_mat <- AP1_UpPeaks_AME_mat[order(rowSums(AP1_UpPeaks_AME_mat), decreasing=T), ] # order by overall most to least enriched AP1 motifs 

outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/additional_AP1_figures/Figures"

# plot motif enrichment values of AP1 motifs in all Up peaks for all timepoints
pdf(paste0(outdir, "/ATAC_AP1_UpPeaks_AME_dotplots.pdf"), width=10, height=4.5)
create_AME_dotplots(annoUpPeaks_AME_df[grepl("JUN|FOS", annoUpPeaks_AME_df$motif_ID), ], Title="AP1 Motifs enriched in Up peaks", fill_limit=c(2, 305), size_limit=c(0, 51))
dev.off()
```


#ATAC-seq: import motif enrichment results of changing peaks within 50kb of each RNA cluster of changing genes. 
```{r import nearby peaks kmeans 7 RNA clusters}
#extract ids of all expressed genes
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL
expressedGenes[grep("TP53$|TP63$|TP73$", expressedGenes)] <- c("P63", "P73", "P53")
expressedGenes[expressedGenes %in% "RELA"] <- "TF65"

# function to import motif data. 
process_motif_data <- function(base_dir, expressed_genes = NULL) {
  library(dplyr)
  library(readr)
  
  # Initialize an empty list to store the data frames
  Peaks_AME <- list()
  processed_dirs <- vector()
  setwd(base_dir)
  file_paths=dir()

  # Loop through each directory and process the files
  for (i in seq_along(file_paths)) {
    # Construct the path to the ame.tsv file
    ame_file_path <- paste0(file_paths[i], "/ame.tsv")
    
    # Check the file size
    file_info <- file.info(ame_file_path)
    if (file_info$size == 1.1 || is.na(file_info$size)) {
      # If the file size is 0 or file does not exist, skip to the next iteration
      next
    }
    
    # Read the ame.tsv file
    Peaks_AME[[length(Peaks_AME) + 1]] <- as.data.frame(read_tsv(file = ame_file_path, comment = "#"))
    
    # Add the current directory to the processed_dirs vector
    processed_dirs <- c(processed_dirs, file_paths[i])
  }
  
  # Set names for the list elements
  names(Peaks_AME) <- processed_dirs
  
  # Post-processing each dataframe
  for (i in seq_along(Peaks_AME)) {
    
    # Calculate TPvsFP and add as a new column
    Peaks_AME[[i]]$TPvsFP <- Peaks_AME[[i]]$`%TP` - Peaks_AME[[i]]$`%FP`
    
    # Remove everything after "_" in the motif_ID column
    Peaks_AME[[i]]$motif_ID <- gsub("\\_.*", "", Peaks_AME[[i]]$motif_ID)
  }
  
  # Convert the list of MEME-AME results into a single dataframe
  Peaks_AME <- Peaks_AME[!sapply(Peaks_AME, is.null)]
  Peaks_AME_df <- do.call(rbind, Peaks_AME)
  
  # Filter for motifs that are expressed, if provided
  if (!is.null(expressed_genes)) {
    Peaks_AME_df <- Peaks_AME_df %>%
      dplyr::filter(motif_ID %in% expressed_genes)
  }
colnames(Peaks_AME_df)[c(7,15,17)] <- c("adj_pvalue", "percentTP", "percentFP") # change column names without special symbols 
Peaks_AME_df$neglog10padj <- -log10(Peaks_AME_df$adj_pvalue) # add in column for -log10(adj pvalue)
  return(Peaks_AME_df)
}
save(process_motif_data, file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS6/Functions/process_motif_data.RData")

Peaks_AME_Clusters <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS6/motif_analysis_output/allDoxo_allPalbo_sigPeaks_kmeans7_RNA_clusters", expressed_genes = expressedGenes)

# Add column to specify which RNA cluster
Peaks_AME_Clusters$Cluster <- gsub('\\.[0-9]*$', '', row.names(Peaks_AME_Clusters)) 
Peaks_AME_Clusters$Cluster <- gsub("^[^_]+_[^_]+_", "", Peaks_AME_Clusters$Cluster) # remove everything before the 2nd "_"

Peaks_AME_Clusters$Comparison <- gsub(".*?(all(Up|Down)_(Palbo|Doxo)).*", "\\1", row.names(Peaks_AME_Clusters))
Peaks_AME_Clusters$Comparison <- as.factor(Peaks_AME_Clusters$Comparison)
Peaks_AME_Clusters$Direction <- ifelse(grepl("Up", Peaks_AME_Clusters$Comparison), "Up", "Down")
Peaks_AME_Clusters$Group <- gsub("^[^_]*_", "", Peaks_AME_Clusters$Comparison)
```

# ATAC-seq, separate RNA clusters: plot motif enrichment results as dotplot indicating difference between %TP & %FP as dot size & fill color as padj values.  
```{r}
AME_padj_dots <- list()
comparisons <- c("allDown_Doxo", "allDown_Palbo", "allUp_Doxo", "allUp_Palbo")
AME_padj_dots <- list(allDown=Peaks_AME_Clusters[Peaks_AME_Clusters$Comparison %in% c("allDown_Doxo", "allDown_Palbo"), ],
                      allUp=Peaks_AME_Clusters[Peaks_AME_Clusters$Comparison %in% c("allUp_Doxo", "allUp_Palbo"), ])
```

# ATAC-seq: create function to plot motif enrichment results as dotplot where the fill color indicates -log10(padj) values & size of dot indicates %TP-%FP
```{r create_AME_dotplots}
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=c(1,10), size_limit=4.5, midpoint=20){
  AME_df <- AME_df %>% mutate(Comparison = Cluster)
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient2(name="-log10(padj)", 
                        low="white", mid="light yellow", high="dark orange",
                        midpoint = midpoint,
                        limits=fill_limit) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=size_limit, 
                          range=c(1, 10)) +  # Adjust dot sizes
    theme_minimal() +
    ggtitle(Title) +
    scale_x_discrete(labels = function(x) AME_df$Cluster[match(x, AME_df$Comparison)]) +  # to replace Comparison with Cluster labels on the x-axis. 
    facet_wrap(~Group, nrow=1) +
    theme(
      axis.text.x = element_text(size=8, angle = 90, hjust=1),  # Rotate x-axis labels
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    )+
    theme_classic() +
    labs(x="Comparison", y="motif ID")}
``` 

```{r}
plot_nearbyUp.AP1_AME <- AME_padj_dots$allUp[grepl("JUN|FOS", AME_padj_dots$allUp$motif_ID) & AME_padj_dots$allUp$Cluster %in% "Shared_genes", ]

plot_nearbyDown.AP1_AME <- AME_padj_dots$allDown[grepl("JUN|FOS|ATF", AME_padj_dots$allDown$motif_ID) & AME_padj_dots$allDown$Cluster %in% "Shared_genes", ]
pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/additional_AP1_figures/Figures"
pdf(paste0(pdf_dir, "/Up.Down.mergedpeaks.near.RNAclusters.AP1.motifs.dotplots.pdf"), width=4, height=3)
create_AME_dotplots(plot_nearbyUp.AP1_AME, Title="nearby up peaks", fill_limit=c(0, 37), size_limit=c(0, 20.05), midpoint=2)
create_AME_dotplots(plot_nearbyDown.AP1_AME, Title="nearby down peaks", fill_limit=c(0, 305), size_limit=c(0, 39.4), midpoint=10)
dev.off()
```