---
title: "Figure A shared senescence signature driven by NF-KB signaling"
author: "Joanna Yeung"
date: '2025-01-27'
output: html_document
---
```{r}
# data wrangling
library(preprocessCore)
library(readr)
library(readxl)
library(stringr)
library(dplyr)

# plotting
library(gridExtra)
library(ggplot2)
library(pheatmap)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(scales)
library(ggrepel)

# genomics data wrangling and analysis
library(rtracklayer)
library(Biostrings)
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(DESeq2)
library(apeglm)
library(ChIPseeker)
library(clusterProfiler)
library(msigdbr)
library(MotifDb)
library(motifmatchr)

# genome annotation databases
library(EnsDb.Hsapiens.v79)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(BSgenome.Hsapiens.UCSC.hg38)
library(org.Hs.eg.db)
```


# annotation colors for pheatmap
```{r}
hm_cols_RNA <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("Cycling"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "DoxovsPalbo"="#CCEBC5"), 
                Biorep=c("1"="orange", "2"="light blue", "3"="pink"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "DoxovsPalbo"="#CCEBC5"), 
                clusters_cutree8=brewer.pal(8, "Set2"), 
                Clusters=c("5"="#66C2A5","2"="#A6D854","1"="#FC8D62", "7"="#8DA0CB","4"="#E78AC3","6"="#E78AC3","3"="#FFD92F","8"="#E5C494"), 
                SASP = c("Yes"="dark green", "No"="white"),
                Cell_Cycle = c("Yes"="dark blue", "No"="white"), 
                p53_All=c("Yes"="#00B0F0", "No"="white"),
                NFKB_All=c("Yes"="red", "No"="white"))
names(hm_cols_RNA$clusters_cutree8) <- as.factor(1:8)


hm_cols_ATAC <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("None"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "PvsD"="#CCEBC5"),
                Biorep=c("1"="orange", "2"="light blue", "3"="pink", "4"="light green"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "PvsD"="#CCEBC5"),
                Peak_anno=c("Distal"="#B15A28", "Gene Body"="#7F54A2", "Promoter"="#A7CEE2"), 
                Doxo=c("Yes"="#B3C543", "No"="white"),
                Palbo=c("Yes"="#E0ADCD", "No"="white")
                )
```

# load in workspace with DESeq results and rlog counts matrix of genes. 
```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Global_transcriptome_and_chromatin_accessibility_is_dynamic_in_both_types_of_therapy_induced_senescence/workspace.RData")
```

#### universe (background gene set) for GO enrichment analysis is all genes expressed
```{r}
universe <- genes_anno_type[match(gsub('\\.[0-9]*$', '', row.names(acrossGroups)), genes_anno_type$ensembl_id), ]
```

# make function to specify code chunk to run from other Rmd file to avoid copying and pasting large lines of code. 
```{r}
# arguments: 
# Rmd_file=file path & name of Rmd file with code chunk
# chunk_name= name of code chunk to run from Rmd file. 
run_code_chunk <- function(Rmd_file, chunk_name){ 
  # Read the Rmd file
lines <- readLines(Rmd_file)

# Identify the start and end of the specific chunk
chunk_start <- grep(paste0("^```\\{r\\s*", chunk_name), lines)  # Find chunk start
chunk_end <- grep("^```$", lines)  # Find all chunk end lines

# Find the first `chunk_end` after `chunk_start`
chunk_end <- chunk_end[chunk_end > chunk_start][1]

  if (is.na(chunk_end)) {
    stop(paste("End of chunk", chunk_name, "not found. Check the Rmd file for missing delimiters."))
  }

# specify lines inside chunk, removing lines that make up the start and end of the chunk. 
chunk_to_run <- lines[(chunk_start[1]+1):(chunk_end[1]-1)]

# Evaluate the extracted code
  eval(parse(text = paste(chunk_to_run, collapse = "\n")), envir = .GlobalEnv)
}
```

# RNA-seq: run code chunks generating GO enrichment results
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="getGOenrichment function")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="genes_in_clusters")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="RNA-seq GO enrichment analysis")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="RNAseq GO enrichment results")
```


# RNA-seq: run code chunks generating GSEA results
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="RNAseq GSEA analysis")
```
# RNA-seq: run code chunks to import IPA upstream regulators & pathways
```{r}
setwd("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/IPA_files_kmeans7_clusters/output")
IPA_upstream_regs <- list()
IPA_pathways <- list()
# read in information for enriched IPA pathways per hierarchical cluster (clustering based on scaled rlog normalized counts)
for(i in 1:length(cluster_names)){
IPA_pathways[[i]]<- read_tsv(paste0(cluster_names[i], "_IPA_pathways.txt"), skip=2)
# read in information for enriched upstream regulators 
IPA_upstream_regs[[i]] <- read.table(paste0(cluster_names[i], "_IPA_upstream_regs.txt"), sep="\t", skip=2, header=T)
IPA_upstream_regs[[i]] <- IPA_upstream_regs[[i]][!grepl("mir|miR|Mir", IPA_upstream_regs[[i]]$Upstream.Regulator), ] # remove upstream regulators that are miRNAs (uncertain if they will be informative & there is a lot of them)
}

names(IPA_upstream_regs) <- cluster_names
names(IPA_pathways) <- cluster_names

# filter out upstream regulators that are not expressed in this cell line. 

#extract ids of all expressed genes
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL
```

## plotting padj values of top IPA predicted upstream regulators of all clusters
```{r}
create_ipa_regs_matrix <- function(IPA_upstream_regs, cluster_names, filter_keyword = "complex", top_n = 25) {
  # Extract unique upstream regulators based on the filter keyword
  unique_upstream_regs <- unique(unlist(lapply(IPA_upstream_regs, function(x) {
    x$Upstream.Regulator[grepl(filter_keyword, x$Molecule.Type)][1:top_n]
  })))

  # Create a matrix to store -log10 p-values
  IPA_complex_regs_mat <- matrix(
    nrow = length(unique_upstream_regs),
    ncol = length(IPA_upstream_regs)
  )
  row.names(IPA_complex_regs_mat) <- unique_upstream_regs # Set row names to upstream regulators
  colnames(IPA_complex_regs_mat) <- cluster_names        # Set column names to cluster names

  # Fill in the matrix with -log10 p-values
  for (i in seq_len(ncol(IPA_complex_regs_mat))) {
    # Match and extract p-values for each cluster
    IPA_complex_regs_mat[, i] <- -log10(IPA_upstream_regs[[i]][
      match(row.names(IPA_complex_regs_mat), IPA_upstream_regs[[i]]$Upstream.Regulator),
      "p.value.of.overlap"
    ])

    # Replace NA values with 0 for heatmap compatibility
    IPA_complex_regs_mat[, i][is.na(IPA_complex_regs_mat[, i])] <- 0
  }

  return(IPA_complex_regs_mat)
}

IPA_complex_regs_mat <- create_ipa_regs_matrix(
  IPA_upstream_regs = IPA_upstream_regs,
  cluster_names = cluster_names,
  filter_keyword = "complex",
  top_n = 25
)

IPA_TF_regs_mat <- create_ipa_regs_matrix(
  IPA_upstream_regs = IPA_upstream_regs,
  cluster_names = cluster_names,
  filter_keyword = "transcription",
  top_n = 25
)

# plot heatmap of -log10 pvalue of overlap of enriched complex regulators
pheatmap(IPA_complex_regs_mat, cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 10,length.out = 100))

pheatmap(IPA_TF_regs_mat, cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 10,length.out = 100))
```

# RNA-seq: run code chunks to make matrix with log2FC values of significant genes
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="log2FC sig genes matrix")
```

# RNA-seq: P53 & NF-kB signatures
```{r}
# import p53 census target gene database from https://tp53.isb-cgc.org/target_genes
Fischer_Oncogene_2017_p53_census_target_genes <- read_excel("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/RNAseq/external_databases/Fischer_Oncogene_2017_p53_census_target_genes.xlsx")

# import Gilmore Lab's list of NF-kB target genes
Gilmore_NFKB_Targets <- as.data.frame(read_excel("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/RNAseq/external_databases/GilmoreLab_NFKB_Target_Genes_BostonUniversity.xlsx"))$`Human Gene Name`
Gilmore_NFKB_Targets <- unique(Gilmore_NFKB_Targets[!is.na(Gilmore_NFKB_Targets)])

# extract genes under Hallmark terms P53 PATHWAY & TNFA SIGNALING VIA NFKB
p53_Hallmark <- msigdbr_df_H[msigdbr_df_H$gs_name %in% "HALLMARK_P53_PATHWAY", ]$gene_symbol
NFKB_Hallmark <- msigdbr_df_H[msigdbr_df_H$gs_name %in% "HALLMARK_TNFA_SIGNALING_VIA_NFKB", ]$gene_symbol

# filter for p53 & NFKB target genes from IPA
IPA_targets <- list(p53=unlist(lapply(IPA_upstream_regs, function(x){str_split(x[grep("TP53$", x$Upstream.Regulator), ]$Target.Molecules.in.Dataset, ",")})),
                    NFKB=unlist(lapply(IPA_upstream_regs, function(x){unique(unlist(str_split(x[grep("NFKB", x$Upstream.Regulator), ]$Target.Molecules.in.Dataset, ",")))}))
                    )

# make log2FC matrix of p53 target genes going up with drug treatment
NFKB_sig_LFCMat <- sig_LFCMat[row.names(sig_LFCMat) %in% c(Gilmore_NFKB_Targets, NFKB_Hallmark, IPA_targets$NFKB) & ! row.names(sig_LFCMat) %in% row.names(clusterDF[clusterDF$Clusters %in% c("5", "2"), ]), ]

anno_sigLFC <- data.frame(
  row.names = colnames(sig_LFCMat),
  log2FC = c(rep("vs Cyc", 4), rep("vs day3", 3), rep("vs Cyc", 5), rep("vs day3", 4), rep("DoxovsPalbo",4)),
  Treatment = c(rep("Doxo", 7), rep("Palbo", 9), rep("DoxovsPalbo",4)),
  Timepoint = c("3", "10", "14", "21", "10", "14", "21", "3", "10", "14", "21", "28", "10", "14", "21", "28","3", "10", "14", "21")
)

# add annotation to significant genes based on if they are NF-kB or p53 targets. 
clusterDF$NFKB_All <- ifelse(clusterDF$gene_id %in% c(Gilmore_NFKB_Targets, NFKB_Hallmark, IPA_targets$NFKB), "Yes", "No") # annotate sig genes as NF-kB targets if they are part of Gilmore_NFKB_Targets or TNFA SIGNALING VIA NFKB Hallmark term 
clusterDF$p53_All <- ifelse(clusterDF$gene_id %in% c(Fischer_Oncogene_2017_p53_census_target_genes$`Gene Symbol`, p53_Hallmark), "Yes", "No")
```

# ATAC-seq: plot Tn5 insertion sites around RELA motifs to show shared increased accessibility at putative NF-kB binding sites (all up peaks)
```{r}
# filter for GRanges of up peaks in each drug and cell cycle arrest category 
Up_DrugandArrestCat <- list(
  PalboUpvsCyc=allUpvsCycling$Palbo[!allUpvsCycling$Palbo %over% allUpvsCycling$Doxo, ], 
  sharedUpvsCyc=allUpvsCycling$Palbo[allUpvsCycling$Palbo %over% allUpvsCycling$Doxo, ],
  DoxoUpvsCyc=allUpvsCycling$Doxo[!allUpvsCycling$Doxo %over% allUpvsCycling$Palbo, ], 
  PalboUpvsday3=allUpvsday3$Palbo[!allUpvsday3$Palbo %over% allUpvsday3$Doxo, ], 
  sharedUpvsday3=allUpvsday3$Palbo[allUpvsday3$Palbo %over% allUpvsday3$Doxo, ], 
  DoxoUpvsday3=allUpvsday3$Doxo[!allUpvsday3$Doxo %over% allUpvsday3$Palbo, ])

# filter for GRanges of up peaks in each drug and cell cycle arrest category but without specifying shared up peaks. 
Up_DrugandArrest_PvsD <- list(
  PalboUpvsCyc=allUpvsCycling$Palbo, 
  DoxoUpvsCyc=allUpvsCycling$Doxo, 
  PalboUpvsday3=allUpvsday3$Palbo,
  DoxoUpvsday3=allUpvsday3$Doxo)

# load in HOCOMOCO motif sequence database
load("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/motifs/HOCOMOCO_v11_matrix.RData")
# scan for RELA motifs
TF65_HMv11 <- matchPWM(HOCOMOCO_v11_matrix$TF65_HUMAN.H11MO.0.A, BSgenome.Hsapiens.UCSC.hg38, with.score=T)
# scan for NFKB1 motifs
NFKB1_HMv11 <- matchPWM(HOCOMOCO_v11_matrix$NFKB1_HUMAN.H11MO.1.B, BSgenome.Hsapiens.UCSC.hg38, with.score=T)

# filter for RELA motifs that overlap with up peaks
TF65_ATAC_Up_DrugandArrestCat <- lapply(Up_DrugandArrestCat, function(x){TF65_HMv11[TF65_HMv11 %over% x]})
TF65_ATAC_Up_DrugandArrest_PvsD <- lapply(Up_DrugandArrest_PvsD, function(x){TF65_HMv11[TF65_HMv11 %over% x]})

# Export each GRanges object as a BED file
for(i in 1:length(TF65_ATAC_Up_DrugandArrestCat)){
export.bed(TF65_ATAC_Up_DrugandArrestCat[[i]], con = paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds/", names(TF65_ATAC_Up_DrugandArrestCat)[i], "_TF65_HMv11.bed"))
}

for(i in 1:length(TF65_ATAC_Up_DrugandArrest_PvsD)){
export.bed(TF65_ATAC_Up_DrugandArrest_PvsD[[i]], con = paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds/", names(TF65_ATAC_Up_DrugandArrest_PvsD)[i], "_PvsD_TF65_HMv11.bed"))
}

# generate input files for deepTools scripts
# directory paths to bed files 
writeLines(dir("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds", full.names=T, pattern=".bed"), con="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds/regions.txt")

# filenames to be assigned to plots. 
filenames <- gsub(".bed", "", dir("//lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds", pattern=".bed"))
writeLines(filenames, con="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds/filename.txt")
```

# ATAC-seq: submit bash job to plot around TF65 motifs in all up peaks
```{bash}
mergedcuts="/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/Cycling_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day3_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day10_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day14_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day21_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day3_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day10_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day14_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day21_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day28_Palbo_scaleFactornorm_cutsCoverage.bw"
samplenames="Cycling day3_Doxo day10_Doxo day14_Doxo day21_Doxo day3_Palbo day10Palbo day14_Palbo day21_Palbo day28_Palbo"

cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/output_plots
sbatch -p risc,hpc --array=1-10 \
--export=bigwigs="$mergedcuts",\
regions=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds/regions.txt,\
samplenames="$samplenames",\
profile_outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/output_plots/plotProfile,\
heatmap_outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/output_plots/plotHeatmap,\
filename=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds/filename.txt \
/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/deeptools_Coverage_1kb.slurm
```

# ATAC-seq: plot Tn5 insertion sites around NFKB1 motifs to show shared increased accessibility at putative NF-kB binding sites (up peaks within 50kb of sig changing gene clusters )
```{r}
# merge RNA clusters of the same drug category.
merge_geneclusters <- list("CellCycle"=genes_in_clusters[[1]], 
                           "Palbo"=unlist(genes_in_clusters[2:3]), 
                           "Doxo"=unlist(genes_in_clusters[4:5]), 
                           "Shared"=genes_in_clusters[[6]]
)

# import gtf file for genes
genes_anno <- import("/lustre/fs4/risc_lab/store/jyeung/references/gencode.v41.GRCh38.p13.annotation.gtf")
genes_anno <- genes_anno[genes_anno$type %in% "gene", ] # filter for regions that are only genes 
genes_anno$gene_id <- gsub('\\.[0-9]*$', '', genes_anno$gene_id) # remove numbers after . in gene id column 
genes_anno <- genes_anno[ ,c("source", "type", "gene_id", "gene_type", "gene_name", "level")] # keep only these columns


# get the genomic coordinates of significantly changing genes from merged clusters (RNA-seq data)
merge_geneclusters_anno <- lapply(merge_geneclusters, function(merge_geneclusters){genes_anno[genes_anno$gene_name %in% merge_geneclusters]})

# make GRanges where the coordinates are 50kb upstream of gene TSS and 50kb downstream of end of gene. 
merge_geneclusters_anno_findpeaks <- merge_geneclusters_anno
for(i in 1:length(merge_geneclusters_anno)){
  start(merge_geneclusters_anno_findpeaks[[i]]) <- start(merge_geneclusters_anno[[i]])-50000
  end(merge_geneclusters_anno_findpeaks[[i]]) <- end(merge_geneclusters_anno[[i]])+50000
}

# find peaks that overlap within 50kb of significantly changing genes in clusters
peaks_in_geneclusters <- lapply(merge_geneclusters_anno_findpeaks, function(merge_geneclusters_anno_findpeaks){annomasterpeaks_pval0.2[annomasterpeaks_pval0.2 %over% merge_geneclusters_anno_findpeaks, ]})

# filter for sig changing peaks within 50kb of sig changing genes in clusters. 
Up_PvsD_gene_clusters <- list()
for(i in 1:length(Up_DrugandArrest_PvsD)){
Up_PvsD_gene_clusters[[i]] <-lapply(peaks_in_geneclusters, function(peaks_in_geneclusters){Up_DrugandArrest_PvsD[[i]][Up_DrugandArrest_PvsD[[i]] %over% peaks_in_geneclusters, ]}) 
}
names(Up_PvsD_gene_clusters) <- names(Up_DrugandArrest_PvsD)

# filter for NFKB1 motifs that overlap with up peaks nearby sig changing genes 
NFKB1_Up_PvsD_gene_clusters <- list()
for(i in 1:length(Up_PvsD_gene_clusters)){
NFKB1_Up_PvsD_gene_clusters[[i]] <- lapply(Up_PvsD_gene_clusters[[i]], function(x){NFKB1_HMv11[NFKB1_HMv11 %over% x]})
}
names(NFKB1_Up_PvsD_gene_clusters) <- names(Up_PvsD_gene_clusters)

directory <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds"
  
# Export each GRanges object as a BED file
for(i in 1:length(NFKB1_Up_PvsD_gene_clusters)){
  for(j in 1:length(NFKB1_Up_PvsD_gene_clusters[[i]])){
export.bed(NFKB1_Up_PvsD_gene_clusters[[i]][[j]], con = paste0(directory, "/", names(NFKB1_Up_PvsD_gene_clusters)[i], "_", names(NFKB1_Up_PvsD_gene_clusters[[i]])[j], "_genes", "_PvsD_NFKB1_HMv11.bed"))
  }
}


# generate input files for deepTools scripts
# directory paths to bed files 
writeLines(dir(directory, full.names=T, pattern="NFKB1.*\\.bed$"), con=paste0(directory, "/NFKB1_regions.txt"))

# filenames to be assigned to plots. 
filenames <- gsub(".bed", "", dir(directory, pattern="NFKB1.*\\.bed$"))
writeLines(filenames, con=paste0(directory, "/NFKB1_filename.txt"))
```

# ATAC-seq: submit bash job to plot around NFKB1 motifs in up peaks within 50kb of sig changing genes. 
```{bash}
mergedcuts="/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/Cycling_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day3_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day10_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day14_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day21_Doxo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day3_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day10_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day14_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day21_Palbo_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_RPGC_bigwigs/workspaces/day28_Palbo_scaleFactornorm_cutsCoverage.bw"
samplenames="Cycling day3_Doxo day10_Doxo day14_Doxo day21_Doxo day3_Palbo day10Palbo day14_Palbo day21_Palbo day28_Palbo"

cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/output_plots
sbatch -p risc,hpc --array=1-16 \
--export=bigwigs="$mergedcuts",\
regions=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds/NFKB1_regions.txt,\
samplenames="$samplenames",\
profile_outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/output_plots/plotProfile,\
heatmap_outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/output_plots/plotHeatmap,\
filename=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/input_beds/NFKB1_filename.txt \
/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/deepTools_files/deeptools_Coverage_1kb.slurm
```

# ATAC-seq: run code chunk to import MEME-AME results of changing peaks within 50kb of RNA clusters into dataframe
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment.Rmd", chunk_name="import nearby peaks kmeans 7 RNA clusters")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment/Supplemental_Figure_peaks_50kb_of_RNAclusters_heatmap_motif_enrichment.Rmd", chunk_name="convert AME df to matrix")
```

# ATAC-seq: import motif enrichment results under ALL up peaks across all timepoint comparisons & drug treatment. 
```{r}
annoUpPeaks_AME_df <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_output/DESeq/annoUp", expressed_genes = expressedGenes)

annoUpPeaks_AME_df$Comparison <- gsub("\\..*", "", row.names(annoUpPeaks_AME_df))
annoUpPeaks_AME_df$Comparison <- factor(annoUpPeaks_AME_df$Comparison, levels=unique(annoUpPeaks_AME_df$Comparison)[c(15,1,5,9,2,6,10,16,3,7,11,13,4,8,12,14)]) # column to identify DESeq comparison

annoUpPeaks_AME_df$Group <- ifelse(grepl("DoxovsCycling", annoUpPeaks_AME_df$Comparison), "DoxovsCycling", ifelse(grepl("PalbovsCycling", annoUpPeaks_AME_df$Comparison), "PalbovsCycling", ifelse(grepl("Doxovsday3Doxo", annoUpPeaks_AME_df$Comparison), "Doxovsday3Doxo", "Palbovsday3Palbo"))) # column to identify relative to cycling or day 3 comparison and drug treatment. 

annoUpPeaks_AME_df$Timepoint <-  gsub("(day[0-9]{1,2}).*", "\\1", annoUpPeaks_AME_df$Comparison) # column to identify which timepoint is compared to Cycling or day 3 controls, by removing everything after day[0-9] where [0-9] is 1 or 2 digits in the Comparison column. 
annoUpPeaks_AME_df$Timepoint <- factor(annoUpPeaks_AME_df$Timepoint, levels=c("day3", "day10", "day14", "day21", "day28"))
```

# ATAC-seq: get peak ids for peaks that are true positive peaks for E2F motif from motif enrichment analysis (MEME-AME)
```{r}
input_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_output/DESeq/annoUp"

out_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/NFKB_tp_peaks"

# generate input file containing directory paths to sequences.tsv file from MEME-AME output. 
writeLines(paste0(dir(input_dir, full.names = T), "/sequences.tsv"), con=paste0(out_dir, "/DESeq_annoUp_sequencefiles.txt"))
```

# ATAC-seq: extract peak ids of true positive peaks of NFKB from sequences.tsv file from MEME-AME output
```{bash}
cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/NFKB_tp_peaks

# for all NFKB motifs
sbatch -p risc --array=1-16 \
--export=outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/NFKB_tp_peaks,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/NFKB_tp_peaks/DESeq_annoUp_sequencefiles.txt,\
names=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoUp/folderlist_annoUp_DESeq_peaks.txt,\
motif="NFKB" \
get_tp_peaks_motifs.slurm

rm $(ls|egrep '.err|.out')

# for all REL motifs
sbatch -p risc --array=1-16 \
--export=outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/NFKB_tp_peaks,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/NFKB_tp_peaks/DESeq_annoUp_sequencefiles.txt,\
names=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/motif_analysis_input/DESeq/annoUp/folderlist_annoUp_DESeq_peaks.txt,\
motif="REL" \
get_tp_peaks_motifs.slurm

rm $(ls|egrep '.err|.out')
```

```{r}
# import peak ids of true positive down peaks containing any NFKB motif
setwd("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/NFKB_tp_peaks")

file_ids <- dir()
all_NFKB_tp_peaks <- list()

for (i in seq_along(file_ids)) {
  file_path <- file_ids[i]  # Get full file name

  # Check if file is empty
  if (file.info(file_path)$size > 0) {
    all_NFKB_tp_peaks[[file_path]] <- read_tsv(file_path, col_names = FALSE) %>% pull(X1)
  }
}

names(all_NFKB_tp_peaks) <- gsub("_all_tp_Uppeaks_ids\\.tsv", "", names(all_NFKB_tp_peaks))
all_NFKB_tp_peaks <- all_NFKB_tp_peaks[!names(all_NFKB_tp_peaks) %in% c("DESeq_annoUp_sequencefiles.txt", "get_tp_peaks_motifs.slurm")]
```

# *FIGURES*
# RNA-seq: save GSEA running score for day 14 Doxo vs Cycling's Hallmark TNFA SIGNALING VIA NFKB Pathway
```{r}
outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling"

# Get padj value of GSEA result for TNFA SIGNALING VIA NFKB pathway to add onto GSEA plots
GSEA_padj <- list(DoxovsCyc=GSEA_Results_H$`Doxo14 vs Cycling0`@result[GSEA_Results_H$`Doxo14 vs Cycling0`@result$ID %in% "TNFA SIGNALING VIA NFKB", ]$p.adjust, 
                  PalbovsCyc=GSEA_Results_H$`Palbo14 vs Cycling0`@result[GSEA_Results_H$`Palbo14 vs Cycling0`@result$ID %in% "TNFA SIGNALING VIA NFKB", ]$p.adjust, 
                  Doxovsday3=GSEA_Results_H$`Doxo14 vs Doxo3`@result[GSEA_Results_H$`Doxo14 vs Doxo3`@result$ID %in% "TNFA SIGNALING VIA NFKB", ]$p.adjust, 
                  Palbovsday3=GSEA_Results_H$`Palbo14 vs Palbo3`@result[GSEA_Results_H$`Palbo14 vs Palbo3`@result$ID %in% "TNFA SIGNALING VIA NFKB", ]$p.adjust)
                  
GSEA_padj <- lapply(GSEA_padj, function(x){format(round(x, 4), nsmall = 4)}) # round to 4 decimal places 

p1 <- gseaplot(GSEA_Results_H$`Doxo14 vs Cycling0`, geneSetID = "TNFA SIGNALING VIA NFKB", by="runningScore", title = "day 14 Doxo vs Cycling: TNFA SIGNALING VIA NFKB", color.line="#B3C543", color="#B3C543", color.vline="#B3C543")+theme_classic()+ylim(-0.1, 0.8)

p2 <- gseaplot(GSEA_Results_H$`Palbo14 vs Cycling0`, geneSetID = "TNFA SIGNALING VIA NFKB", by="runningScore", title = "day 14 Palbo vs Cycling: TNFA SIGNALING VIA NFKB", color="#E0ADCD", color.line="#E0ADCD", color.vline="#E0ADCD")+theme_classic()+ylim(-0.1, 0.8)

p3 <- gseaplot(GSEA_Results_H$`Doxo14 vs Doxo3`, geneSetID = "TNFA SIGNALING VIA NFKB", by="runningScore", title = "day 14 Doxo vs day 3 Doxo: TNFA SIGNALING VIA NFKB", color.line="#B3C543", color="#B3C543", color.vline="#B3C543")+theme_classic()+ylim(-0.1, 0.8)

p4 <- gseaplot(GSEA_Results_H$`Palbo14 vs Palbo3`, geneSetID = "TNFA SIGNALING VIA NFKB", by="runningScore", title = "day 14 Palbo vs day 3 Palbo: TNFA SIGNALING VIA NFKB", color="#E0ADCD", color.line="#E0ADCD", color.vline="#E0ADCD")+theme_classic()+ylim(-0.1, 0.8)

pdf(paste0(outdir, "/GSEA_day14Doxo_day14Palbo_Hallmark_TNFA_SIGNALING_VIA_NFKB_pathway.pdf"), width=3, height=5)
plot_grid(p1, p2, ncol=1)
plot_grid(p3, p4, ncol=1)
dev.off()

pdf(paste0(outdir, "/GSEA_day14Doxo_day14Palbo_Hallmark_TNFA_SIGNALING_VIA_NFKB_pathway_padj_labeled.pdf"), width=3, height=5)
plot_grid(p1+geom_text(label=paste0("padj=" ,GSEA_padj$DoxovsCyc), y=0.6, x=10000), p2+geom_text(label=paste0("padj=" ,GSEA_padj$PalbovsCyc), y=0.6, x=10000), ncol=1)
plot_grid(p3+geom_text(label=paste0("padj=" ,GSEA_padj$Doxovsday3), y=0.6, x=10000), p4+geom_text(label=paste0("padj=" ,GSEA_padj$Palbovsday3), y=0.6, x=10000), ncol=1)
dev.off()


gseaplot(GSEA_Results_H$`Doxo14 vs Doxo3`, geneSetID = "TNFA SIGNALING VIA NFKB", by="runningScore", title = "day 14 Doxo vs day 3 Doxo: TNFA SIGNALING VIA NFKB", color.line="#B3C543", color="#B3C543", color.vline="#B3C543")+theme_classic()+ylim(-0.1, 0.8)
```

# RNA-seq: plot padj values for top 10 GO terms for shared senescence up genes. 
```{r}
# Still showing the pvalues of the other clusters
GO_results_mat_clusters <- list()
for(i in 1:length(GO_results_list)){

GO_results_mat_clusters[[i]] <- GO_results_list_mat[row.names(GO_results_list_mat) %in% GO_results_list[[i]]$ID[1:25], ]
GO_results_mat_clusters[[i]] <- GO_results_mat_clusters[[i]][order(GO_results_mat_clusters[[i]][ ,i], decreasing=T), ]

}

pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling"

pdf(paste0(pdf_dir, "/RNA_GO_results_SharedUp_genes.pdf"), width=6, height=2.5)
pheatmap(GO_results_mat_clusters[[7]][1:10, ], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Purples"))(100), breaks = seq(0,10,length.out = 100), cluster_rows = F, main="GO enrichment: shared up genes", annotation_row=GO_anno, annotation_col = Cluster_anno, annotation_colors = GO_anno_colors, border_color = "black")
dev.off()
```

# RNA-seq: plot padj values for top 5 IPA upstream regs for each cluster individually. 
```{r}
# Still showing the pvalues of the other clusters as columns but only the top IPA upstream regs from the cluster of interest. 
IPA_TF_regs_clusters <- list()
IPA_complex_regs_clusters <- list()
for(i in 1:length(IPA_upstream_regs)){
  #transcription regulators
IPA_TF_regs_clusters[[i]] <-IPA_TF_regs_mat[row.names(IPA_TF_regs_mat) %in% IPA_upstream_regs[[i]]$Upstream.Regulator[grepl("transcription", IPA_upstream_regs[[i]]$Molecule.Type)][1:5], ]
IPA_TF_regs_clusters[[i]] <- IPA_TF_regs_clusters[[i]][order(IPA_TF_regs_clusters[[i]][ ,i], decreasing=T), ]
  # complex regulators
IPA_complex_regs_clusters[[i]] <-IPA_complex_regs_mat[row.names(IPA_complex_regs_mat) %in% IPA_upstream_regs[[i]]$Upstream.Regulator[grepl("complex", IPA_upstream_regs[[i]]$Molecule.Type)][1:5], ]
IPA_complex_regs_clusters[[i]] <- IPA_complex_regs_clusters[[i]][order(IPA_complex_regs_clusters[[i]][ ,i], decreasing=T), ] # sort by decreasing padj value
}

# make heatmap of top 5 IPA predicted upstream regulators for shared up genes. 
pdf(file=paste0(pdf_dir, "/RNA_IPA_TF_regs_SharedUp_genes.pdf"), width=3, height=2)
pheatmap(IPA_TF_regs_clusters[[7]], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 10,length.out = 100), cluster_rows = F, main="Shared up genes: predicted TF regulators", border_color = "black",  annotation_col=Cluster_anno, annotation_colors = GO_anno_colors)
dev.off()

pdf(file=paste0(pdf_dir, "/RNA_IPA_complex_regs_SharedUp_genes.pdf"), width=4, height=2)
pheatmap(IPA_complex_regs_clusters[[7]], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 10,length.out = 100), cluster_rows = F, main="Shared up genes: predicted complex regulators", border_color = "black", annotation_col=Cluster_anno, annotation_colors = GO_anno_colors)
dev.off()
```

# RNA-seq: plot venn diagram showing overlap of shared up genes that are p53 targets, NF-kB targets and SASP. 
```{r}
# write function to create venn diagrams that are proportional to size of each category. 
create_proportional_venn <- function(sets, set_labels, colors = c("#E0ADCD", "#B3C543", "#A0ADCD")) {
  library(eulerr)
  # Check if sets and labels are provided correctly
  if (length(sets) != length(set_labels)) {
    stop("The number of sets and set_labels must be the same.")
  }
  
  # Prepare the data for eulerr
  venn_data <- sets
  names(venn_data) <- set_labels
  
  # Use eulerr to calculate the Venn diagram
  venn_fit <- euler(venn_data)
  
  # Plot the Venn diagram
  plot(venn_fit, 
       fills = colors, 
       labels = list(cex = 1, fontfamily = "Helvetica"),
       quantities = list(cex = 1),
       edges = list(lwd = 2))
}
venn_data_shared <- list(row.names(clusterDF[clusterDF$Clusters %in% "3" & clusterDF$p53_All %in% "Yes", ]), 
                  row.names(clusterDF[clusterDF$Clusters %in% "3" & clusterDF$NFKB_All %in% "Yes", ]), 
                  row.names(clusterDF[clusterDF$Clusters %in% "3" & clusterDF$SASP%in% "Yes", ]), 
                  row.names(clusterDF[clusterDF$Clusters %in% "3", ]))

pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/RNA_p53vsNFKB_sharedgenes_vennDiagram.pdf", width=4, height=4)

create_proportional_venn(sets=venn_data_shared, 
                         set_labels = c("p53 targets", "NF-kB targets", "SASP", "Shared up genes"), 
                         colors = c("#00B0F0","red", "dark green", "#E5C494")
)
dev.off()
```

# plot heatmap of NF-kB target genes, clustered based on log2FC from day 3 treatment controls & Cycling controls
```{r}
pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/RNA_log2FC_NFKB_targets_mat.pdf", width=5, height=3)
pheatmap(NFKB_sig_LFCMat, show_rownames = T, show_colnames=F, cluster_cols = F, cluster_rows=F, colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100), breaks = seq(-6, 6,length.out = 100), annotation_row = clusterDF[ ,c("Clusters", "SASP", "Cell_Cycle", "p53_All")], annotation_col = anno_sigLFC, annotation_colors = hm_cols_RNA, border_color = NA, main="all up NF-kB target genes")

# plot showing only NFKB target genes belonging to shared up cluster of genes
pheatmap(NFKB_sig_LFCMat[row.names(NFKB_sig_LFCMat) %in% row.names(clusterDF[clusterDF$Clusters %in% "3", ]), ], show_rownames = T, show_colnames=F, cluster_cols = F, cluster_rows=T, colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100), breaks = seq(-6, 6,length.out = 100), annotation_row = clusterDF[ ,c("p53_All", "SASP")], annotation_col = anno_sigLFC, annotation_colors = hm_cols_RNA, border_color = NA, main="NF-kB Targets: Shared up genes")

# plot showing only NFKB target genes belonging to shared up cluster of genes & are SASP genes. 
pheatmap(NFKB_sig_LFCMat[row.names(NFKB_sig_LFCMat) %in% row.names(clusterDF[clusterDF$Clusters %in% "8" & clusterDF$SASP %in% "Yes", ]), ], show_rownames = T, show_colnames=F, cluster_cols = F, cluster_rows=T, colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100), breaks = seq(-6, 6,length.out = 100), annotation_row = clusterDF[ ,c("p53_All", "SASP")], annotation_col = anno_sigLFC, annotation_colors = hm_cols_RNA, border_color = NA, main="NF-kB SASP: Shared up genes")
dev.off()
```

# plot log2FC of NF-kB target genes at day 3 & 14 timepoints relative to Cycling
```{r}
p53_NFKB_SASP_genes <- clusterDF[clusterDF$p53_All %in% "Yes"|clusterDF$SASP %in% "Yes"|clusterDF$NFKB_All %in% "Yes", ]
p53_NFKB_SASP_genes <- row.names(p53_NFKB_SASP_genes[!p53_NFKB_SASP_genes$Clusters %in% c("5", "2"), ])

log2FC_p53_NFKB_SASP <- rbind(
          data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo3 vs Cycling0"], 
           Condition="day 3 Palbo", 
           Comparison="vs Cycling"),
  
          data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo14 vs Cycling0"], 
           Condition="day 14 Palbo", 
           Comparison="vs Cycling" ), 
           
           data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo21 vs Cycling0"], 
           Condition="day 21 Palbo", 
           Comparison="vs Cycling" ), 
           
           data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo28 vs Cycling0"], 
           Condition="day 28 Palbo", 
           Comparison="vs Cycling" ),
           
           data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Doxo3 vs Cycling0"], 
           Condition="day 3 Doxo", 
           Comparison="vs Cycling"),
      
      data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Doxo14 vs Cycling0"], 
           Condition="day 14 Doxo", 
           Comparison="vs Cycling"), 
      
          data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Doxo21 vs Cycling0"], 
           Condition="day 21 Doxo", 
           Comparison="vs Cycling"), 
      
      data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo14 vs Palbo3"], 
           Condition="day 14 Palbo",
           Comparison="vs day 3"), 
      
       data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo21 vs Palbo3"], 
           Condition="day 21 Palbo",
           Comparison="vs day 3" ), 
           
           data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo28 vs Palbo3"], 
           Condition="day 28 Palbo",
           Comparison="vs day 3" ),
      
      data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Doxo14 vs Doxo3"], 
           Condition="day 14 Doxo",
           Comparison="vs day 3"),
      
      data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Doxo21 vs Doxo3"], 
           Condition="day 21 Doxo",
           Comparison="vs day 3")
)

log2FC_p53_NFKB_SASP[ ,5:7] <- clusterDF[match(log2FC_p53_NFKB_SASP$genes, row.names(clusterDF)), c("SASP", "NFKB_All", "p53_All")]

# function to create violin plot of log2FC values of genes
create_violin_plot <- function(data=log2FC_p53_NFKB_SASP, plot_title="RNA: log2FC of p53 target genes") {
  library(ggplot2)
  library(ggbeeswarm)
  
  ggplot(
    data,
    aes(x = Condition, y = log2FC, fill = Condition)
  ) +
    geom_violin() +
    geom_quasirandom(
      color = "black", alpha = 0.2,
      method = "pseudorandom", # Aligns jitter points to violin shape
      groupOnX = TRUE,
      size = 0.5
    ) +
    stat_summary(
      fun = "median",
      geom = "crossbar",
      color = "white",
      width = 0.5
    ) +
    theme_classic() +
    theme(
      axis.text.x = element_text(angle = 90, size = 8),
      axis.text.y = element_text(size = 8),
      axis.title.y = element_text(size = 10)
    ) +
    facet_wrap(~Comparison, scales = "free_x") +
    scale_fill_manual(
      values = c(
        "day 3 Doxo" = "#B3C543",
        "day 14 Doxo" = "#B3C543",
        "day 21 Doxo" = "#B3C543",
        "day 3 Palbo" = "#E0ADCD",
        "day 14 Palbo" = "#E0ADCD",
        "day 21 Palbo" = "#E0ADCD",
        "day 28 Palbo" = "#E0ADCD"
      )
    ) +
    ylab("log2FC") +
    xlab("") +
    ggtitle(plot_title)+
    geom_hline(yintercept = 0, color="darkgrey", linetype="dashed")
}

pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling"

pdf(paste0(pdf_dir, "/RNA_day14_log2FC_NFKB_targets.pdf"), width=4, height=3)
library(ggbeeswarm)
create_violin_plot(
  data = log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$NFKB_All %in% "Yes" & log2FC_p53_NFKB_SASP$Condition %in% c("day 3 Doxo", "day 14 Doxo",  "day 3 Palbo", "day 14 Palbo") & log2FC_p53_NFKB_SASP$Comparison %in% "vs Cycling", ],
  plot_title = "RNA: log2FC of all upregulated NF-kB target genes"
)+ylim(-5,10) # after talking with Justin, we both agreed it looks better to include day 3 controls and plot relative to Cycling to show the increase over time in both drugs. 

shared_genes <- log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$genes %in% clusterDF[clusterDF$Clusters %in% "3", ]$gene_id, ]$genes
create_violin_plot(
  data = log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$NFKB_All %in% "Yes" & log2FC_p53_NFKB_SASP$Condition %in% c("day 3 Doxo", "day 14 Doxo",  "day 3 Palbo", "day 14 Palbo") & log2FC_p53_NFKB_SASP$Comparison %in% "vs Cycling" & log2FC_p53_NFKB_SASP$genes %in% shared_genes, ],
  plot_title = "RNA: log2FC of NF-kB target genes upregulated in both drugs"
)+ylim(-5,10) # after talking with Justin, we both agreed it looks better to include day 3 controls and plot relative to Cycling to show the increase over time in both drugs. 

Doxo_upgenes <- log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$genes %in% clusterDF[clusterDF$Clusters %in% c("4", "6"), ]$gene_id, ]$genes
create_violin_plot(
  data = log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$NFKB_All %in% "Yes" & log2FC_p53_NFKB_SASP$Condition %in% c("day 3 Doxo", "day 14 Doxo",  "day 3 Palbo", "day 14 Palbo") & log2FC_p53_NFKB_SASP$Comparison %in% "vs Cycling" & log2FC_p53_NFKB_SASP$genes %in% Doxo_upgenes, ],
  plot_title = "RNA: log2FC of NF-kB target genes upregulated in Doxo"
)+ylim(-5,10)

Palbo_upgenes <- log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$genes %in% clusterDF[clusterDF$Clusters %in% c("1", "7"), ]$gene_id, ]$genes
create_violin_plot(
  data = log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$NFKB_All %in% "Yes" & log2FC_p53_NFKB_SASP$Condition %in% c("day 3 Doxo", "day 14 Doxo",  "day 3 Palbo", "day 14 Palbo") & log2FC_p53_NFKB_SASP$Comparison %in% "vs Cycling" & log2FC_p53_NFKB_SASP$genes %in% Palbo_upgenes, ],
  plot_title = "RNA: log2FC of NF-kB target genes upregulated in Palbo"
)+ylim(-5,10)
dev.off()
```

# ATAC-seq, merged RNA clusters: plot motif enrichment results of all clusters as heatmap of p-adj values for each type (Doxo, Palbo, Shared) of sig changing peaks
```{r}
NFkB_AME_padj_dots <- list()
for(i in 1:length(comparisons)){
NFkB_AME_padj_dots[[i]] <- Peaks_AME_Clusters[Peaks_AME_Clusters$Comparison %in% comparisons[i], ]
NFkB_AME_padj_dots[[i]] <- NFkB_AME_padj_dots[[i]][NFkB_AME_padj_dots[[i]]$motif_ID %in% c("NFKB1", "NFKB2", "REL", "RELB"), ] # show only top 15 most significant padj values motif IDs
NFkB_AME_padj_dots[[i]]$motif_ID <- factor(NFkB_AME_padj_dots[[i]]$motif_ID, levels=rev(unique(NFkB_AME_padj_dots[[i]]$motif_ID)))
}
names(NFkB_AME_padj_dots) <- comparisons
NFkB_AME_padj_dots <- NFkB_AME_padj_dots[-which(sapply(NFkB_AME_padj_dots, nrow) ==0)]

create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=c(1,10), size_limit=4.5){
  AME_df <- AME_df %>% mutate(Comparison = Cluster)
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient2(name="-log10(padj)", 
                        low="white", mid="light yellow", high="dark orange",
                        midpoint = 3,
                        limits=fill_limit) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=size_limit, 
                          range=c(1, 10)) +  # Adjust dot sizes
    theme_minimal() +
    ggtitle(Title) +
    scale_x_discrete(labels = function(x) AME_df$Cluster[match(x, AME_df$Comparison)]) +  # to replace Comparison with Cluster labels on the x-axis. 
    facet_wrap(~Group, nrow=1) +
    theme(
      axis.text.x = element_text(size=8, angle = 90, hjust=1),  # Rotate x-axis labels
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    )+
    theme_classic() +
    labs(x="Comparison", y="motif ID")}

pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling/Up.Down.peaks.near.kmeans7RNAclusters.NFkB.motifs.dotplots.pdf", width=8, height=5)
plot_grid( create_AME_dotplots(NFkB_AME_padj_dots$allUp_Doxo, Title="Up Doxo", fill_limit = c(1,15.2), size_limit = c(0, 6.2)),
           create_AME_dotplots(NFkB_AME_padj_dots$allUp_Palbo, Title="Up Palbo", fill_limit = c(1,15.2), size_limit = c(0, 6.2)),
           nrow =1)
dev.off()
```

# ATAC-seq: plot motif enrichment results in the NFKB family for ALL up peaks across all timepoints and drug treatments (DESeq comparisons)
```{r}
# load in function to generate matrix of -log10(padj) values of motif enrichment results
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Supplemental_Figure_Summary_Motif_enrichment_results_all_sig_peaks/Functions/create_AME_peaks_matrix.RData")

# function to create dotplots for motif enrichment results for each DESeq comparison. 
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=c(1,10), size_limit=4.5){
  AME_df <- AME_df %>% mutate(Comparison = Timepoint)
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient2(name="-log10(padj)", 
                        low="white", mid="light yellow", high="dark orange",
                        midpoint = 3,
                        limits=fill_limit) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=size_limit, 
                          range=c(1, 10)) +  # Adjust dot sizes
    theme_minimal() +
    ggtitle(Title) +
    scale_x_discrete(labels = function(x) AME_df$Timepoint[match(x, AME_df$Comparison)]) +  # to replace Comparison with Cluster labels on the x-axis. 
    facet_wrap(~Group, nrow=1, scales="free") +
    theme(
      axis.text.x = element_text(size=8, angle = 90, hjust=1),  # Rotate x-axis labels
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    )+
    theme_classic() +
    labs(x="Comparison", y="motif ID")}

NFkB_UpPeaks_AME_mat <- create_AME_peaks_matrix(annoUpPeaks_AME_df[grepl("NFKB|REL", annoUpPeaks_AME_df$motif_ID), ])
                                                      
NFkB_UpPeaks_AME_mat <- NFkB_UpPeaks_AME_mat[order(rowSums(NFkB_UpPeaks_AME_mat), decreasing=T), ] # order by overall most to least enriched NFkB motifs 

outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling"

# plot motif enrichment values of NFkB motifs in all Up peaks for all timepoints vs Cycling. 
pdf(paste0(outdir, "/ATAC_NFkB_UpPeaks_AME_dotplots.pdf"), width=8, height=2.5)
create_AME_dotplots(annoUpPeaks_AME_df[grepl("NFKB|REL", annoUpPeaks_AME_df$motif_ID), ], Title="NFkB Motifs enriched in Up peaks", fill_limit=c(2, 92), size_limit=c(0, 17.4))
dev.off()
```

# ATAC-seq: make sequence Logo from HOCOMOCO NFKB1 motif. 
```{r}
outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling"
library(seqLogo)
pdf(paste0(outdir, "/NFKB1_HUMAN.H11MO.1.B_HOCOMOCO_seqLogo.pdf"), width=6, height=4)
seqLogo(HOCOMOCO_v11_matrix$NFKB1_HUMAN.H11MO.1.B)
dev.off()
```

# note: still need to adjust code to be NF-kB, not E2F. 
```{r}
# write function to create venn diagrams that are proportional to size of each category. 
create_proportional_venn <- function(sets, set_labels, colors = c("#E0ADCD", "#B3C543", "#A0ADCD")) {
  library(eulerr)
  # Check if sets and labels are provided correctly
  if (length(sets) != length(set_labels)) {
    stop("The number of sets and set_labels must be the same.")
  }
  
  # Prepare the data for eulerr
  venn_data <- sets
  names(venn_data) <- set_labels
  
  # Use eulerr to calculate the Venn diagram
  venn_fit <- euler(venn_data)
  
  # Plot the Venn diagram
  plot(venn_fit, 
       fills = colors, 
       labels = list(cex = 1, fontfamily = "Helvetica"),
       quantities = list(cex = 1),
       edges = list(lwd = 2))
}

# unique peak ids for Doxo vs Cycling up tp peaks with NFKB motifs
tp_uppeaks_Doxo <- unique(unlist(all_NFKB_tp_peaks[grepl("Doxo", names(all_NFKB_tp_peaks))]))

# unique peak ids for Palbo vs Cycling up tp peaks with NFKB motifs
tp_uppeaks_Palbo <- unique(unlist(all_NFKB_tp_peaks[grepl("Palbo", names(all_NFKB_tp_peaks))]))

# peak ids for all up peaks 
all_uppeaks_ids <- unique(c(mergedDiffGR$allUp_Doxo$name, mergedDiffGR$allUp_Palbo$name))

outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_A_shared_senescence_signature_driven_by_NFkB_signaling"

# plot venn diagram showing overlap of NFKB1 containing up peaks between Doxo & Palbo against number of all up peaks
pdf(paste0(outdir, "/ATAC_NFKB1_tp_uppeaks_PalbovsDoxo_overlap_venn.pdf"), width=6, height=4)
create_proportional_venn(sets=list(Palbo=tp_uppeaks_Palbo, Doxo=tp_uppeaks_Doxo, allup=all_uppeaks_ids), set_labels = c("Palbo", "Doxo", "all up"), colors = c("#E0ADCD", "#B3C543", "#F6F2EA"))
dev.off()
```