---
title: "Figure p53 activity is different depending on the presence of DNA damage"
author: "Joanna Yeung"
date: '2025-01-24'
output: html_document
---
```{r}
library(readr)
library(GenomicFeatures)
library(DESeq2)
library(apeglm)
library(gridExtra)
library(ggplot2)
library(pheatmap)
library(msigdbr)
library(stringr)
library(org.Hs.eg.db)
library(ggrepel)
library(rtracklayer)
library(cowplot)
library(readxl)
library(EnsDb.Hsapiens.v79)
library(viridis)
library(GenomicRanges)
library(rtracklayer)
library(RColorBrewer)
library(cowplot)
library(preprocessCore)
library(scales)
library(ChIPseeker)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)
library(BSgenome.Hsapiens.UCSC.hg38)
library(MotifDb)
library(motifmatchr)
library(Biostrings)
library(clusterProfiler)
library(pheatmap)
```

# load in workspace with DESeq results and rlog counts matrix of genes. 
```{r}
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Global_transcriptome_and_chromatin_accessibility_is_dynamic_in_both_types_of_therapy_induced_senescence/workspace.RData")
```

#### universe (background gene set) for GO enrichment analysis is all genes expressed
```{r}
universe <- genes_anno_type[match(gsub('\\.[0-9]*$', '', row.names(acrossGroups)), genes_anno_type$ensembl_id), ]
```

# make function to specify code chunk to run from other Rmd file to avoid copying and pasting large lines of code. 
```{r}
# arguments: 
# Rmd_file=file path & name of Rmd file with code chunk
# chunk_name= name of code chunk to run from Rmd file. 
run_code_chunk <- function(Rmd_file, chunk_name){ 
  # Read the Rmd file
lines <- readLines(Rmd_file)

# Identify the start and end of the specific chunk
chunk_start <- grep(paste0("^```\\{r\\s*", chunk_name), lines)  # Find chunk start
chunk_end <- grep("^```$", lines)  # Find all chunk end lines

# Find the first `chunk_end` after `chunk_start`
chunk_end <- chunk_end[chunk_end > chunk_start][1]

  if (is.na(chunk_end)) {
    stop(paste("End of chunk", chunk_name, "not found. Check the Rmd file for missing delimiters."))
  }

# specify lines inside chunk, removing lines that make up the start and end of the chunk. 
chunk_to_run <- lines[(chunk_start[1]+1):(chunk_end[1]-1)]

# Evaluate the extracted code
  eval(parse(text = paste(chunk_to_run, collapse = "\n")), envir = .GlobalEnv)
}
```

# RNA-seq: run code chunks generating GO enrichment results
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="getGOenrichment function")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="genes_in_clusters")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="RNA-seq GO enrichment analysis")

run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="RNAseq GO enrichment results")
```


# RNA-seq: GSEA Analysis
```{r}
ranked_res_forGSEA <- list()
for(i in 1:length(LFCResultsall_RNA)){
  
  rank <- LFCResultsall_RNA[[i]]
  
  # Assign a default p-value for genes with NA padj (maximum p-value)
  rank$pvalue[is.na(rank$padj)] <- 1 
  
  # Create a ranked list based on log2FoldChange and -log10(p-value)
  rank$rank_metric <- sign(rank$log2FoldChange) * -log10(rank$pvalue)
  
  # Sort by the ranking metric
  rank <- rank[order(rank$rank_metric, decreasing = TRUE), ]
  
  # Store the ranked values
  ranked_res_forGSEA[[i]] <- rank$rank_metric
  names(ranked_res_forGSEA[[i]]) <- row.names(rank)

  # Convert gene IDs to Entrez IDs if available
  entrez_ids <- genes_anno_type[match(names(ranked_res_forGSEA[[i]]), genes_anno_type$ensembl_id), ]$entrezid
  names(ranked_res_forGSEA[[i]]) <- ifelse(is.na(entrez_ids), names(ranked_res_forGSEA[[i]]), entrez_ids)
}

names(ranked_res_forGSEA) <- names(LFCResultsall_RNA)


# run GSEA analysis using: 
GSEA_Results_H <- list() # msigdb Hallmark database
GSEA_Results_REACT <- list() # msigdb KEGG database
GSEA_Results_REACT <- list() # msigdb Reactome database
GSEA_plots_H <- list()
GSEA_plots_REACT <- list()
GSEA_plots_REACT <- list()
for(i in 1:length(ranked_res_forGSEA)){
 GSEA_Results_H[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_H) 
 GSEA_Results_REACT[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_REACT) 
 GSEA_Results_REACT[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_REACT) 
}

names(GSEA_Results_H) <- names(LFCResultsall_RNA)
names(GSEA_Results_REACT) <- names(LFCResultsall_RNA)
names(GSEA_Results_REACT) <- names(LFCResultsall_RNA)

# generate running score plots for GSEA results
# GSEA Hallmark
for(i in 1:length(GSEA_Results_H)){
  maxplots <- length(GSEA_Results_H[[i]]$ID)
GSEA_plots_H[[i]] <- list()
GSEA_plots_KEGG[[i]] <- list()
GSEA_plots_REACT[[i]] <- list()
  for(j in 1:maxplots){
  GSEA_plots_H[[i]][[j]] <- gseaplot(GSEA_Results_H[[i]], geneSetID = j, by = "runningScore", title = paste0(names(GSEA_Results_H)[i], ": " , GSEA_Results_H[[i]]$ID[j]) )
  }
  names(GSEA_plots_H[[i]]) <- GSEA_Results_H[[i]]$ID[1:maxplots]
}

# GSEA KEGG
for (i in 1:length(GSEA_Results_KEGG)) {
  # Check if the current object is not empty
  if (!is.null(GSEA_Results_KEGG[[i]]) && length(GSEA_Results_KEGG[[i]]$ID) > 0) {
    maxplots <- length(GSEA_Results_KEGG[[i]]$ID)
    GSEA_plots_KEGG[[i]] <- list()  # Initialize a list to store plots for this object
    
    for (j in 1:maxplots) {
      GSEA_plots_KEGG[[i]][[j]] <- gseaplot(
        GSEA_Results_KEGG[[i]], 
        geneSetID = j, 
        by = "runningScore", 
        title = GSEA_Results_KEGG[[i]]$ID[j]
      )
    }
    # Assign names to the plots
    names(GSEA_plots_KEGG[[i]]) <- GSEA_Results_KEGG[[i]]$ID[1:maxplots]
  } else {
    # If empty, set the current slot in GSEA_plots_KEGG to NULL
    GSEA_plots_KEGG[[i]] <- NULL
  }
}

# GSEA Reactome
for (i in 1:length(GSEA_Results_REACT)) {
  # Check if the current object is not empty
  if (!is.null(GSEA_Results_REACT[[i]]) && length(GSEA_Results_REACT[[i]]$ID) > 0) {
    maxplots <- length(GSEA_Results_REACT[[i]]$ID)
    GSEA_plots_REACT[[i]] <- list()  # Initialize a list to store plots for this object
    
    for (j in 1:maxplots) {
      GSEA_plots_REACT[[i]][[j]] <- gseaplot(
        GSEA_Results_REACT[[i]], 
        geneSetID = j, 
        by = "runningScore", 
        title = GSEA_Results_REACT[[i]]$ID[j]
      )
    }
    # Assign names to the plots
    names(GSEA_plots_REACT[[i]]) <- GSEA_Results_REACT[[i]]$ID[1:maxplots]
  } else {
    # If empty, set the current slot in GSEA_plots_REACT to NULL
    GSEA_plots_REACT[[i]] <- NULL
  }
}

# assign names to plots based on which DESeq comparison GSEA was done on. 
names(GSEA_plots_H) <- names(LFCResultsall_RNA)
names(GSEA_plots_KEGG) <- names(which(unlist(lapply(GSEA_Results_KEGG, function(x){length(x$ID)})) >0))
names(GSEA_plots_REACT) <- names(which(unlist(lapply(GSEA_Results_REACT, function(x){length(x$ID)})) >0))
```

# RNA-seq: save GSEA running score plots to pdf
```{r}
pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/GSEA_plots_Hallmark.pdf", width=5, height=4)
GSEA_plots_H
dev.off()
```

# # RNA-seq: generate input files for IPA pathways. 
```{r}
# clusters of genes 
for(i in 1:length(genes_in_clusters)){
  
  IPA_input <- as.data.frame(genes_in_clusters[[i]])
  colnames(IPA_input) <- "ID"
  write.table(IPA_input, file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/IPA_files/", cluster_names[i], ".csv"), row.names=F, sep=",")
}
# DESeq Wald Test results for pairwise comparisons. 
for(i in 1:length(LFCResultsall_RNA)){
  write.table(as.data.frame(LFCResultsall_RNA[[i]]), file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/IPA_files/DESeq_results_input/", names(LFCResultsall_RNA)[i], ".csv"), sep=",")
}
```

# # RNA-seq: import IPA upstream regulators & pathways
```{r}
setwd("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/IPA_files/output/")
IPA_upstream_regs <- list()
IPA_pathways <- list()
# read in information for enriched IPA pathways per hierarchical cluster (clustering based on scaled rlog normalized counts)
for(i in 1:length(cluster_names)){
IPA_pathways[[i]]<- read_tsv(paste0(cluster_names[i], "_IPA_Pathways.txt"), skip=2)
# read in information for enriched upstream regulators 
IPA_upstream_regs[[i]] <- read.table(paste0(cluster_names[i], "_IPA_UpstreamRegs.txt"), sep="\t", skip=2, header=T)
IPA_upstream_regs[[i]] <- IPA_upstream_regs[[i]][!grepl("mir|miR|Mir", IPA_upstream_regs[[i]]$Upstream.Regulator), ] # remove upstream regulators that are miRNAs (uncertain if they will be informative & there is a lot of them)
}

# filter out upstream regulators that are not expressed in this cell line. 

#extract ids of all expressed genes
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL

# keep only upstream regulators that are part of expressed genes or have lower case letters (these are upstream regulators that could be miRNAs or protein complexes)
IPA_upstream_regs <- lapply(IPA_upstream_regs, function(IPA_upstream_regs){IPA_upstream_regs[c(grep('[:lower:]', IPA_upstream_regs$Upstream.Regulator), which(IPA_upstream_regs$Upstream.Regulator %in% expressedGenes)), ]})

names(IPA_upstream_regs) <- cluster_names

names(IPA_pathways) <- cluster_names
```

# RNA-seq: make matrix with log2FC values of significant genes
```{r}
run_code_chunk(Rmd_file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_Drug_distinct_responses_and_SASP/Figure_Drug_distinct_responses_and_SASP.Rmd", chunk_name="log2FC sig genes matrix")
```

# RNA-seq: P53 & NF-kB signatures
```{r}
# import p53 census target gene database from https://tp53.isb-cgc.org/target_genes
Fischer_Oncogene_2017_p53_census_target_genes <- read_excel("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/RNAseq/external_databases/Fischer_Oncogene_2017_p53_census_target_genes.xlsx")

# import Gilmore Lab's list of NF-kB target genes
Gilmore_NFKB_Targets <- as.data.frame(read_excel("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/RNAseq/external_databases/GilmoreLab_NFKB_Target_Genes_BostonUniversity.xlsx"))$`Human Gene Name`
Gilmore_NFKB_Targets <- unique(Gilmore_NFKB_Targets[!is.na(Gilmore_NFKB_Targets)])

# extract genes under Hallmark terms P53 PATHWAY & TNFA SIGNALING VIA NFKB
p53_Hallmark <- msigdbr_df_H[msigdbr_df_H$gs_name %in% "HALLMARK_P53_PATHWAY", ]$gene_symbol
NFKB_Hallmark <- msigdbr_df_H[msigdbr_df_H$gs_name %in% "HALLMARK_TNFA_SIGNALING_VIA_NFKB", ]$gene_symbol

# filter for p53 & NFKB target genes from IPA
IPA_targets <- list(p53=unlist(lapply(IPA_upstream_regs, function(x){str_split(x[grep("TP53$", x$Upstream.Regulator), ]$Target.Molecules.in.Dataset, ",")})),
                    NFKB=unlist(lapply(IPA_upstream_regs, function(x){unique(unlist(str_split(x[grep("NFKB", x$Upstream.Regulator), ]$Target.Molecules.in.Dataset, ",")))}))
                    )

# make log2FC matrix of p53 target genes going up with drug treatment
p53_sig_LFCMat <- sig_LFCMat[row.names(sig_LFCMat) %in% c(Fischer_Oncogene_2017_p53_census_target_genes$`Gene Symbol`, p53_Hallmark) & ! row.names(sig_LFCMat) %in% row.names(clusterDF[clusterDF$Clusters %in% c("5", "2"), ]), ]

anno_sigLFC <- data.frame(
  row.names = colnames(sig_LFCMat),
  log2FC = c(rep("vs Cyc", 4), rep("vs day3", 3), rep("vs Cyc", 5), rep("vs day3", 4), rep("DoxovsPalbo",4)),
  Treatment = c(rep("Doxo", 7), rep("Palbo", 9), rep("DoxovsPalbo",4)),
  Timepoint = c("3", "10", "14", "21", "10", "14", "21", "3", "10", "14", "21", "28", "10", "14", "21", "28","3", "10", "14", "21")
)

# add annotation to significant genes based on if they are NF-kB or p53 targets. 
clusterDF$NFKB_All <- ifelse(row.names(clusterDF) %in% c(Gilmore_NFKB_Targets, NFKB_Hallmark, IPA_targets$NFKB), "Yes", "No") # annotate sig genes as NF-kB targets if they are part of Gilmore_NFKB_Targets or TNFA SIGNALING VIA NFKB Hallmark term 
clusterDF$p53_All <- ifelse(row.names(clusterDF) %in% c(Fischer_Oncogene_2017_p53_census_target_genes$`Gene Symbol`, p53_Hallmark), "Yes", "No")
```

# make matrix of normalized counts across drug & time. 
```{r}
NormCountsMat <- counts(RNAddsall, normalized=T)
# convert row names of matrix of normalized counts of off to on genes to gene symbol format. If there is no gene symbol format, keep ensembl format. 
row.names(NormCountsMat) <- gsub('\\.[0-9]*$', '', row.names(NormCountsMat))
row.names(NormCountsMat) <- ifelse(genes_anno_type[match(row.names(NormCountsMat), genes_anno_type$ensembl_id), ]$gene_id %in% NA, row.names(NormCountsMat), genes_anno_type[match(row.names(NormCountsMat), genes_anno_type$ensembl_id), ]$gene_id)

# convert matrix of normalized counts of off to on genes into a dataframe with metaData information for ggplot to accept
NormCounts_df <- list()
for(i in 1:nrow(NormCountsMat)){
NormCounts_df[[i]] <- data.frame(Counts=NormCountsMat[i, ], sample=names((NormCountsMat[i, ])), Transcript= row.names(NormCountsMat)[i], metaDataall)
}
NormCounts_df <- do.call(rbind, NormCounts_df) # combine list into 1 giant dataframe
```


# ATAC-seq: plot Tn5 insertion sites around p53 motifs overlapping up peaks to show stronger p53 chromatin accessibility signature in Doxo compared to Palbo. 
```{r}
# filter for GRanges of up peaks in each drug and cell cycle arrest category 
Up_DrugandArrestCat <- list(
  PalboUpvsCyc=allUpvsCycling$Palbo[!allUpvsCycling$Palbo %over% allUpvsCycling$Doxo, ], 
  sharedUpvsCyc=allUpvsCycling$Palbo[allUpvsCycling$Palbo %over% allUpvsCycling$Doxo, ],
  DoxoUpvsCyc=allUpvsCycling$Doxo[!allUpvsCycling$Doxo %over% allUpvsCycling$Palbo, ], 
  PalboUpvsday3=allUpvsday3$Palbo[!allUpvsday3$Palbo %over% allUpvsday3$Doxo, ], 
  sharedUpvsday3=allUpvsday3$Palbo[allUpvsday3$Palbo %over% allUpvsday3$Doxo, ], 
  DoxoUpvsday3=allUpvsday3$Doxo[!allUpvsday3$Doxo %over% allUpvsday3$Palbo, ])

# filter for GRanges of up peaks in each drug and cell cycle arrest category but without specifying shared up peaks. 
Up_DrugandArrest_PvsD <- list(
  PalboUpvsCyc=allUpvsCycling$Palbo, 
  DoxoUpvsCyc=allUpvsCycling$Doxo, 
  PalboUpvsday3=allUpvsday3$Palbo,
  DoxoUpvsday3=allUpvsday3$Doxo)

# load in HOCOMOCO motif sequence database
load("/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/motifs/HOCOMOCO_v11_matrix.RData")
# scan for p53 motifs
p53_HMv11 <- matchPWM(HOCOMOCO_v11_matrix$P53_HUMAN.H11MO.0.A, BSgenome.Hsapiens.UCSC.hg38, with.score=T)

# filter for p53 motifs that overlap with up peaks
p53_ATAC_Up_DrugandArrestCat <- lapply(Up_DrugandArrestCat, function(x){p53_HMv11[p53_HMv11 %over% x]})
p53_ATAC_Up_DrugandArrest_PvsD <- lapply(Up_DrugandArrest_PvsD, function(x){p53_HMv11[p53_HMv11 %over% x]})

# print number of up peaks that overlap with p53 motifs. 
lapply(lapply(Up_DrugandArrestCat, function(x){x[x %over% p53_HMv11]}), length)

# Export each GRanges object as a BED file
for(i in 1:length(p53_ATAC_Up_DrugandArrestCat)){
export.bed(p53_ATAC_Up_DrugandArrestCat[[i]], con = paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/input_beds/", names(p53_ATAC_Up_DrugandArrestCat)[i], "_p53_HMv11.bed"))
}

for(i in 1:length(p53_ATAC_Up_DrugandArrest_PvsD)){
export.bed(p53_ATAC_Up_DrugandArrest_PvsD[[i]], con = paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/input_beds/", names(p53_ATAC_Up_DrugandArrest_PvsD)[i], "_PvsD_p53_HMv11.bed"))
}

# generate input files for deepTools scripts
# directory paths to bed files 
writeLines(dir("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/input_beds", full.names=T, pattern=".bed"), con="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/input_beds/regions.txt")

# filenames to be assigned to plots. 
filenames <- gsub(".bed", "", dir("//lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/input_beds", pattern=".bed"))
writeLines(filenames, con="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/input_beds/filename.txt")
```

# submit bash job to plot around Tn5 cut sites around p53 motifs in up peaks
```{bash}
mergedcuts="/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_bigwigs/scaleFactornorm_cutsCoverage_bigwigs/Cyclingmerged_scaleFactornorm_cutsCoverage.bw
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_bigwigs/scaleFactornorm_cutsCoverage_bigwigs/day3_Doxomerged_scaleFactornorm_cutsCoverage.bw
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_bigwigs/scaleFactornorm_cutsCoverage_bigwigs/day10_Doxomerged_scaleFactornorm_cutsCoverage.bw
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_bigwigs/scaleFactornorm_cutsCoverage_bigwigs/day14_Doxomerged_scaleFactornorm_cutsCoverage.bw
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_bigwigs/scaleFactornorm_cutsCoverage_bigwigs/day21_Doxomerged_scaleFactornorm_cutsCoverage.bw
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_bigwigs/scaleFactornorm_cutsCoverage_bigwigs/day3_Palbomerged_scaleFactornorm_cutsCoverage.bw
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_bigwigs/scaleFactornorm_cutsCoverage_bigwigs/day10_Palbomerged_scaleFactornorm_cutsCoverage.bw
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_bigwigs/scaleFactornorm_cutsCoverage_bigwigs/day14_Palbomerged_scaleFactornorm_cutsCoverage.bw
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_bigwigs/scaleFactornorm_cutsCoverage_bigwigs/day21_Palbomerged_scaleFactornorm_cutsCoverage.bw /lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/Tn5_offset_merged_bigwigs/scaleFactornorm_cutsCoverage_bigwigs/day28_Palbomerged_scaleFactornorm_cutsCoverage.bw"

samplenames="Cycling day3_Doxo day10_Doxo day14_Doxo day21_Doxo day3_Palbo day10Palbo day14_Palbo day21_Palbo day28_Palbo"

cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/output_plots
sbatch -p risc,hpc --array=1-10 \
--export=bigwigs="$mergedcuts",\
regions=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/input_beds/regions.txt,\
samplenames="$samplenames",\
profile_outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/output_plots/plotProfile,\
heatmap_outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/output_plots/plotHeatmap,\
filename=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/input_beds/filename.txt \
/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/deepTools_files/deeptools_Coverage_1kb.slurm
```

# *FIGURES*

# RNA-seq: save GSEA running score for day 14 Doxo vs Cycling's Hallmark p53 Pathway
```{r}
library(enrichplot)
library(patchwork)
pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/GSEA_day14Doxo_vs_Cyc_Hallmark_p53_pathway.pdf", width=3, height=2.5)
GSEA_plots_H$`Doxo14 vs Cycling0`$`P53 PATHWAY` +theme_classic()
dev.off()

# Get padj value of GSEA result for p53 pathway to add onto GSEA plots
GSEA_padj <- GSEA_Results_H$`Doxo14 vs Cycling0`@result[GSEA_Results_H$`Doxo14 vs Cycling0`@result$ID %in% "P53 PATHWAY", ]$p.adjust
GSEA_padj <- format(round(GSEA_padj, 4), nsmall = 4) # round to 4 decimal places 

# plot GSEA RES
p1 <- gseaplot(GSEA_Results_H$`Doxo14 vs Cycling0`, geneSetID = "P53 PATHWAY", by="runningScore", title = "day 14 Doxo vs Cycling: P53 PATHWAY", color.line="#B3C543", color="#B3C543", color.vline ="#B3C543")+theme_classic()+ylim(-0.4, 0.7)+geom_text(label=paste0("padj=" ,GSEA_padj), y=0.6, x=10000)

p2 <- gseaplot(GSEA_Results_H$`Palbo14 vs Cycling0`, geneSetID = "P53 PATHWAY", by="runningScore", title = "day 14 Palbo vs Cycling: P53 PATHWAY", color="#E0ADCD", color.line="#E0ADCD")+theme_classic()+ylim(-0.4, 0.7)

pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/GSEA_day14Doxo_day14Palbo_vs_Cyc_Hallmark_p53_pathway.pdf", width=3, height=5)
plot_grid(p1, p2, ncol=1)
dev.off()
```

# plot heatmap, clustered based on log2FC from day 3 treatment controls & Cycling controls
```{r}
# when plotting all p53 targets that are sig changing, annotate based on which database they were found in: Fischer or Hallmark, P53 PATHWAY term. 
p53_genes_anno <- clusterDF[clusterDF$p53_All %in% "Yes", ]
p53_genes_anno$Fischer_p53 <- ifelse(row.names(p53_genes_anno) %in% Fischer_Oncogene_2017_p53_census_target_genes$`Gene Symbol`, "Yes", "No")
p53_genes_anno$Hallmark_p53 <- ifelse(row.names(p53_genes_anno) %in% p53_Hallmark, "Yes", "No")
hm_cols_RNA$Fischer_p53 <- c("Yes"="#1E7DA5", No="white")
hm_cols_RNA$Hallmark_p53 <- c("Yes"="#1E7DA5", No="white")

pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/RNA_log2FC_p53_Hallmark_mat.pdf", width=5, height=3.5)
pheatmap(p53_sig_LFCMat, show_rownames = T, show_colnames=F, cluster_cols = F, cluster_rows=F, colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100), breaks = seq(-6, 6,length.out = 100), annotation_row = p53_genes_anno[ ,c("Clusters", "Fischer_p53", "Hallmark_p53", "SASP", "NFKB_All")], annotation_col = anno_sigLFC, annotation_colors = hm_cols_RNA, border_color = NA, main="all up p53 target genes")

# plot showing only p53 target genes under p53 Hallmark term from GO enrichment. 

p53_genes <- sort(row.names(p53_sig_LFCMat[row.names(p53_sig_LFCMat) %in% unlist(str_split(GO_enrichment_clusters_HallMark_df$Doxo1[GO_enrichment_clusters_HallMark_df$Doxo1$ID %in% "P53 PATHWAY", ]$geneID, "/")), ])) # filter for p53 pathway Hallmark genes that are significantly changing

pheatmap(p53_sig_LFCMat[match(p53_genes, row.names(p53_sig_LFCMat)) , ], show_rownames = T, show_colnames=F, cluster_cols = F, cluster_rows=T, colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100), breaks = seq(-6, 6,length.out = 100), annotation_row = clusterDF[ ,c("Clusters", "NFKB_All", "SASP")], annotation_col = anno_sigLFC, annotation_colors = hm_cols_RNA, border_color = NA, main="Hallmark p53 pathway: Doxo enriched up genes") # with log2FC legend limits set to -6, 6. 

pheatmap(p53_sig_LFCMat[match(p53_genes, row.names(p53_sig_LFCMat)), ], show_rownames = T, show_colnames=F, cluster_cols = F, cluster_rows=F, colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100), breaks = seq(-4, 4,length.out = 100), annotation_row = clusterDF[ ,c("NFKB_All", "SASP")], annotation_col = anno_sigLFC, annotation_colors = hm_cols_RNA, border_color = NA, main="Hallmark p53 pathway: Doxo enriched up genes") # with log2FC legend limits set to -4, 4 & gene ids ordered in alphabetical order
dev.off()
```
# plot: venn diagram showing overlap of Doxo or Palbo enriched up genes that are p53 targets, NF-kB targets and SASP. 
```{r}
# write function to create venn diagrams that are proportional to size of each category. 
create_proportional_venn <- function(sets, set_labels, colors = c("#E0ADCD", "#B3C543", "#A0ADCD")) {
  library(eulerr)
  # Check if sets and labels are provided correctly
  if (length(sets) != length(set_labels)) {
    stop("The number of sets and set_labels must be the same.")
  }
  
  # Prepare the data for eulerr
  venn_data <- sets
  names(venn_data) <- set_labels
  
  # Use eulerr to calculate the Venn diagram
  venn_fit <- euler(venn_data)
  
  # Plot the Venn diagram
  plot(venn_fit, 
       fills = colors, 
       labels = list(cex = 1, fontfamily = "Helvetica"),
       quantities = list(cex = 1),
       edges = list(lwd = 2))
}
venn_data_Doxo <- list(row.names(clusterDF[clusterDF$Clusters %in% c("4", "6") & clusterDF$p53_All %in% "Yes", ]), 
                  row.names(clusterDF[clusterDF$Clusters %in% c("4", "6") & clusterDF$NFKB_All %in% "Yes", ]), 
                  row.names(clusterDF[clusterDF$Clusters %in% c("4", "6") & clusterDF$SASP%in% "Yes", ]), 
                  row.names(clusterDF[clusterDF$Clusters %in% c("4", "6"), ]))

venn_data_Palbo <- list(row.names(clusterDF[clusterDF$Clusters %in% c("1", "7") & clusterDF$p53_All %in% "Yes", ]), 
                  row.names(clusterDF[clusterDF$Clusters %in% c("1", "7") & clusterDF$NFKB_All %in% "Yes", ]), 
                  row.names(clusterDF[clusterDF$Clusters %in% c("1", "7") & clusterDF$SASP%in% "Yes", ]), 
                  row.names(clusterDF[clusterDF$Clusters %in% c("1", "7"), ]))

pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/RNA_p53vsNFKB_genes_vennDiagram.pdf", width=4, height=4)

create_proportional_venn(sets=venn_data_Doxo, 
                         set_labels = c("p53 targets", "NF-kB targets", "SASP", "Doxo enriched up genes"), 
                         colors = c("#00B0F0","red", "dark green", "#B3C543")
)

create_proportional_venn(sets=venn_data_Palbo, 
                         set_labels = c("p53 targets", "NF-kB targets", "SASP", "Palbo enriched up genes"), 
                         colors = c("#00B0F0","red", "dark green", "#E0ADCD")
)
dev.off()
```

# plot log2FC of p53 target genes at across all timepoints relative to Cycling. 
```{r}
p53_NFKB_SASP_genes <- clusterDF[clusterDF$p53_All %in% "Yes"|clusterDF$SASP %in% "Yes"|clusterDF$NFKB_All %in% "Yes", ]
p53_NFKB_SASP_genes <- row.names(p53_NFKB_SASP_genes[!p53_NFKB_SASP_genes$Clusters %in% c("5", "2"), ])

log2FC_p53_NFKB_SASP <- rbind(
          data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo3 vs Cycling0"], 
           Condition="day 3 Palbo", 
           Comparison="vs Cycling"),
  
          data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo14 vs Cycling0"], 
           Condition="day 14 Palbo", 
           Comparison="vs Cycling" ), 
           
           data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo21 vs Cycling0"], 
           Condition="day 21 Palbo", 
           Comparison="vs Cycling" ), 
           
           data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo28 vs Cycling0"], 
           Condition="day 28 Palbo", 
           Comparison="vs Cycling" ),
           
           data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Doxo3 vs Cycling0"], 
           Condition="day 3 Doxo", 
           Comparison="vs Cycling"),
      
      data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Doxo14 vs Cycling0"], 
           Condition="day 14 Doxo", 
           Comparison="vs Cycling"), 
      
          data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Doxo21 vs Cycling0"], 
           Condition="day 21 Doxo", 
           Comparison="vs Cycling"), 
      
      data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo14 vs Palbo3"], 
           Condition="day 14 Palbo",
           Comparison="vs day 3"), 
      
       data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo21 vs Palbo3"], 
           Condition="day 21 Palbo",
           Comparison="vs day 3" ), 
           
           data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Palbo28 vs Palbo3"], 
           Condition="day 28 Palbo",
           Comparison="vs day 3" ),
      
      data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Doxo14 vs Doxo3"], 
           Condition="day 14 Doxo",
           Comparison="vs day 3"),
      
      data.frame(genes= p53_NFKB_SASP_genes , 
           log2FC=sig_LFCMat[row.names(sig_LFCMat) %in% p53_NFKB_SASP_genes, "Doxo21 vs Doxo3"], 
           Condition="day 21 Doxo",
           Comparison="vs day 3")
)

log2FC_p53_NFKB_SASP[ ,5:7] <- clusterDF[match(log2FC_p53_NFKB_SASP$genes, row.names(clusterDF)), c("SASP", "NFKB_All", "p53_All")]

# function to create violin plot of log2FC values of genes
create_violin_plot <- function(data=log2FC_p53_NFKB_SASP, plot_title="RNA: log2FC of p53 target genes") {
  library(ggplot2)
  library(ggbeeswarm)
  
  ggplot(
    data,
    aes(x = Condition, y = log2FC, fill = Condition)
  ) +
    geom_violin() +
    geom_quasirandom(
      color = "black", alpha = 0.2,
      method = "pseudorandom", # Aligns jitter points to violin shape
      groupOnX = TRUE,
      size = 0.5
    ) +
    stat_summary(
      fun = "median",
      geom = "crossbar",
      color = "white",
      width = 0.5
    ) +
    theme_classic() +
    theme(
      axis.text.x = element_text(angle = 90, size = 8),
      axis.text.y = element_text(size = 8),
      axis.title.y = element_text(size = 10)
    ) +
    facet_wrap(~Comparison, scales = "free_x") +
    scale_fill_manual(
      values = c(
        "day 3 Doxo" = "#B3C543",
        "day 14 Doxo" = "#B3C543",
        "day 21 Doxo" = "#B3C543",
        "day 3 Palbo" = "#E0ADCD",
        "day 14 Palbo" = "#E0ADCD",
        "day 21 Palbo" = "#E0ADCD",
        "day 28 Palbo" = "#E0ADCD"
      )
    ) +
    ylab("log2FC") +
    xlab("") +
    ggtitle(plot_title)
}

pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/RNA_log2FC_p53_targets.pdf", width=7, height=3)
library(ggbeeswarm)
create_violin_plot(
  data = log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$p53_All %in% "Yes", ],
  plot_title = "RNA: log2FC of p53 target genes"
)
dev.off()

pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/RNA_day3_log2FC_p53_targets.pdf", width=4, height=3)
library(ggbeeswarm)
create_violin_plot(
  data = log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$p53_All %in% "Yes" & log2FC_p53_NFKB_SASP$Condition %in% c("day 3 Doxo", "day 3 Palbo"), ],
  plot_title = "RNA: log2FC of p53 target genes"
)
dev.off()

# plot only for day 14
pdf("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/RNA_day14_log2FC_p53_targets.pdf", width=4, height=3)
library(ggbeeswarm)
create_violin_plot(
  data = log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$p53_All %in% "Yes" & log2FC_p53_NFKB_SASP$Condition %in% c("day 14 Doxo", "day 14 Palbo"), ],
  plot_title = "RNA: log2FC of p53 target genes"
)
dev.off()

# get significance of log2FC values between day 14 Palbo & day 14 Doxo vs Cycling based on paired student t-test

# Check normality of differences 
d14_Palbo_log2FC <- log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$p53_All %in% "Yes" & log2FC_p53_NFKB_SASP$Condition %in% "day 14 Palbo" & log2FC_p53_NFKB_SASP$Comparison %in% "vs Cycling", ]$log2FC # log2FC values of p53 targets in day 14 Palbo vs Cycling

d14_Doxo_log2FC <- log2FC_p53_NFKB_SASP[log2FC_p53_NFKB_SASP$p53_All %in% "Yes" & log2FC_p53_NFKB_SASP$Condition %in% "day 14 Doxo" & log2FC_p53_NFKB_SASP$Comparison %in% "vs Cycling", ]$log2FC # log2FC values of p53 targets in day 14 Doxo vs Cycling

differences <- d14_Palbo_log2FC - d14_Doxo_log2FC

# Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(differences)
print(shapiro_test)

# Visualize distribution of differences
ggplot(data.frame(differences), aes(x = differences)) +
  geom_histogram(binwidth = 0.2, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Log2FC Differences", x = "Difference (Drug1 - Drug2)", y = "Count")

# Perform Paired t-test if normality assumption holds
if (shapiro_test$p.value > 0.05) {
  t_test_result <- t.test(d14_Palbo_log2FC, d14_Doxo_log2FC, paired = TRUE)
  print(t_test_result)
} else {
  # Use Wilcoxon signed-rank test if normality is violated
  wilcox_test_result <- wilcox.test(d14_Palbo_log2FC, d14_Doxo_log2FC, paired = TRUE)
  print(wilcox_test_result)
}

```

# RNA-seq Plot normalized counts of well-known p53 targets, MDM2 & CDKN1A over time. 
```{r}
NormCounts_df$Timepoint <- as.numeric(as.character(NormCounts_df$Timepoint))
# save plot to pdf
pdf(file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage/RNA_NormCountsMat_MDM2_CDKN1A.pdf", width=8, height=2)

plot_grid(ggplot(NormCounts_df[NormCounts_df$Transcript %in% "CDKN1A", ], aes(x=Timepoint, y=Counts, color=Treatment, shape=Biorep))+geom_point(size=3)+scale_color_manual(values=hm_cols_RNA$Treatment)+theme_minimal()+ylab("Normalized Counts")+xlab("Length of Treatment (days)")+ggtitle("CDKN1A")+scale_y_continuous(n.breaks=5)+scale_x_continuous(n.breaks=5)+theme_classic()+theme(
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8),
      axis.title.y = element_text(size = 10), 
      axis.title.x = element_text(size = 10)
    ),

ggplot(NormCounts_df[NormCounts_df$Transcript %in% "MDM2", ], aes(x=Timepoint, y=Counts, color=Treatment, shape=Biorep))+geom_point(size=3)+scale_color_manual(values=hm_cols_RNA$Treatment)+theme_minimal()+ylab("Normalized Counts")+xlab("Length of Treatment (days)")+ggtitle("MDM2")+scale_y_continuous(n.breaks=5)+scale_x_continuous(n.breaks=5)+theme_classic()+theme(
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8),
      axis.title.y = element_text(size = 10), 
      axis.title.x = element_text(size = 10)
    )+ylim(0,150000)
)

dev.off()
```

# ATAC-seq: make sequence Logo from HOCOMOCO p53 motif. 
```{r}
outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure_p53_activity_is_different_depending_on_the_presence_of_DNA_damage"
library(seqLogo)
pdf(paste0(outdir, "/P53_HUMAN.H11MO.0.A_HOCOMOCO_seqLogo.pdf"), width=6, height=4)
seqLogo(HOCOMOCO_v11_matrix$P53_HUMAN.H11MO.0.A)
dev.off()
```



