---
title: "Figure 4: Differential regulatory pathways distinguish Doxo and Palbo driven therapy-induced senescence"
author: "Joanna Yeung"
date: '2025-01-27'
output: html_document
---

```{r}
# data wrangling
library(preprocessCore)
library(readr)
library(readxl)
library(stringr)
library(dplyr)

# plotting
library(gridExtra)
library(ggplot2)
library(pheatmap)
library(cowplot)
library(viridis)
library(RColorBrewer)
library(scales)
library(ggrepel)

# genomics data wrangling and analysis
library(rtracklayer)
library(Biostrings)
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(DESeq2)
library(apeglm)
library(ChIPseeker)
library(clusterProfiler)
library(msigdbr)
library(MotifDb)
library(motifmatchr)

# genome annotation databases
library(EnsDb.Hsapiens.v79)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(BSgenome.Hsapiens.UCSC.hg38)
library(org.Hs.eg.db)
```

# make function to specify code chunk to run from other Rmd file to avoid copying and pasting large lines of code. 
```{r}
# arguments: 
# Rmd_file=file path & name of Rmd file with code chunk
# chunk_name= name of code chunk to run from Rmd file. 
run_code_chunk <- function(Rmd_file, chunk_name){ 
  # Read the Rmd file
lines <- readLines(Rmd_file)

# Identify the start and end of the specific chunk
chunk_start <- grep(paste0("^```\\{r\\s*", chunk_name), lines)  # Find chunk start
chunk_end <- grep("^```$", lines)  # Find all chunk end lines

# Find the first `chunk_end` after `chunk_start`
chunk_end <- chunk_end[chunk_end > chunk_start][1]

  if (is.na(chunk_end)) {
    stop(paste("End of chunk", chunk_name, "not found. Check the Rmd file for missing delimiters."))
  }

# specify lines inside chunk, removing lines that make up the start and end of the chunk. 
chunk_to_run <- lines[(chunk_start[1]+1):(chunk_end[1]-1)]

# Evaluate the extracted code
  eval(parse(text = paste(chunk_to_run, collapse = "\n")), envir = .GlobalEnv)
}
```

# annotation colors for pheatmap
```{r}
hm_cols_RNA <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("Cycling"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD"), 
                Biorep=c("1"="orange", "2"="light blue", "3"="pink"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "DoxovsPalbo"="#CCEBC5"), 
                Clusters=c("2"="#66C2A5","5"="#A6D854","7"="#8DA0CB", "1"="#FC8D62", "4"="#E78AC3","6"="#FFD92F","3"="#E5C494"), 
                SASP = c("Yes"="dark green", "No"="white"),
                Cell_Cycle = c("Yes"="dark blue", "No"="white"), 
                p53_All=c("Yes"="#00B0F0", "No"="white"),
                NFKB_All=c("Yes"="red", "No"="white"))
names(hm_cols_RNA$clusters_cutree8) <- as.factor(1:8)


hm_cols_ATAC <- list(Timepoint=c("0"="#440154FF", "3"="#3B528BFF", "10"= "#4DA6A3", "14"="#21908CFF", "21"="#5DC863FF", "28"="#FDE725FF"), 
                Treatment=c("None"="#F8766D", "Doxo"="#B3C543", "Palbo"="#E0ADCD", "PvsD"="#CCEBC5"),
                Biorep=c("1"="orange", "2"="light blue", "3"="pink", "4"="light green"), 
                log2FC=c("vs Cyc"="#FBB4AE","vs day3"="#B3CDE3", "PvsD"="#CCEBC5"),
                Peak_anno=c("Distal"="#B15A28", "Gene Body"="#7F54A2", "Promoter"="#A7CEE2"), 
                Doxo=c("Yes"="#B3C543", "No"="white"),
                Palbo=c("Yes"="#E0ADCD", "No"="white")
                )
```

# download msigdb databases, to be used for GO enrichment analysis. 
```{r get_msigdbr}
# msigdb GO Biological Processes database
msigdbr_df_GO_BP <- as.data.frame(msigdbr(species = "Homo sapiens", category="C5", subcategory ="GO:BP"))
msigdbr_t2g_GO_BP = msigdbr_df_GO_BP %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_GO_BP$gs_name <- sub("GOBP_", "", msigdbr_t2g_GO_BP$gs_name)
msigdbr_t2g_GO_BP$gs_name <- gsub("_", " ", msigdbr_t2g_GO_BP$gs_name)

# msigdb GO Molecular Function Processes database
msigdbr_df_GO_MF <- as.data.frame(msigdbr(species = "Homo sapiens", category="C5", subcategory ="GO:MF"))
msigdbr_t2g_GO_MF = msigdbr_df_GO_MF %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_GO_MF$gs_name <- sub("GOMF_", "", msigdbr_t2g_GO_MF$gs_name)
msigdbr_t2g_GO_MF$gs_name <- gsub("_", " ", msigdbr_t2g_GO_MF$gs_name)

# msigdb Hallmark database
msigdbr_df_H <- as.data.frame(msigdbr(species = "Homo sapiens", category="H", subcategory =""))
msigdbr_t2g_H = msigdbr_df_H %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_H$gs_name <- sub("HALLMARK_", "", msigdbr_t2g_H$gs_name)
msigdbr_t2g_H$gs_name <- gsub("_", " ", msigdbr_t2g_H$gs_name)

# msigdb KEGG pathway database
msigdbr_df_KEGG <- as.data.frame(msigdbr(species = "Homo sapiens", category="C2", subcategory ="CP:KEGG"))
msigdbr_t2g_KEGG = msigdbr_df_KEGG %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_KEGG$gs_name <- sub("KEGG_", "", msigdbr_t2g_KEGG$gs_name)
msigdbr_t2g_KEGG$gs_name <- gsub("_", " ", msigdbr_t2g_KEGG$gs_name)

# msigdb Reactome database
msigdbr_df_REACT <- as.data.frame(msigdbr(species = "Homo sapiens", category="C2", subcategory="CP:REACTOME"))
msigdbr_t2g_REACT = msigdbr_df_REACT %>% dplyr::distinct(gs_name, entrez_gene) %>% as.data.frame()
msigdbr_t2g_REACT$gs_name <- sub("REACTOME_", "", msigdbr_t2g_REACT$gs_name)
msigdbr_t2g_REACT$gs_name <- gsub("_", " ", msigdbr_t2g_REACT$gs_name)
```


#### universe (background gene set) for GO enrichment analysis is all genes expressed
```{r}
universe <- genes_anno_type[match(gsub('\\.[0-9]*$', '', row.names(acrossGroups)), genes_anno_type$ensembl_id), ]
```

# GO enrichment analysis on k-means clusters from sigMat (significantly changing genes across all conditions, based on Likelihood ratio test)

## *function to get GO enriched terms*:
## arguments:
### gene= vector of genes to test for GO enrichment (gene ensembl format)
### universe=background gene set 
### Database=database containing groups of genes to GO annotation term e.g. msigdb GO:BP
```{r getGOenrichment function}
getGOenrichment <-function(gene,universe,Database){
  #gene_entrezID <-  AnnotationDbi::select(EnsDb.Hsapiens.v79, gene[!is.na(gene)], "ENTREZID", "SYMBOL") #convert gene IDs from SYMBOL format to ENTREZ ID format (was using v79, new code is using v107)
  gene_entrezID <- genes_anno_type[match(gene, genes_anno_type$gene_id), ]
  gene_go <- enricher(gene_entrezID$entrezid, TERM2GENE=Database, universe=universe) #enrichment analysis with hypergeometric test
  gene_go <- setReadable(gene_go, OrgDb = org.Hs.eg.db, keyType="ENTREZID") #convert enrichment results from ENTREZ ID format to SYMBOL for easier interpretation
return(gene_go)
} 
```

# assign significant genes to relevant k-means clusters & drug + cell cycle arrest categories. 
```{r genes_in_clusters}
# load in workspace with DESeq results and rlog counts matrix of genes. 
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure2/workspace.RData")
genes_in_clusters <- lapply(levels(clusterDF$Clusters), function(x){clusterDF[clusterDF$Clusters %in% x, ]$gene_id})
cluster_names <- c("CellCycle1", "CellCycle2", "Palbo1", "Palbo2", "Doxo1", "Doxo2", "Shared") 
names(genes_in_clusters) <- cluster_names
```

# RNA-seq: GO enrichment on significant genes
```{r RNA-seq GO enrichment analysis}
# GO enrichment analysis with msigdb GO:BP database:
GOBP_enrichment_clusters <- list()
# GO enrichment analysis with msigdb REACT database:
GO_enrichment_clusters_KEGG <- list()
# GO enrichment analysis with msigdb Hallmark database:
GO_enrichment_clusters_HallMark <- list()
# GO enrichment analysis with msigdb REACT database:
GO_enrichment_clusters_REACT <- list()
# GO enrichment analysis with msigdb GO:MF database:
GO_enrichment_clusters_MF <- list()

for(i in 1:length(genes_in_clusters)){
  
GOBP_enrichment_clusters[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_GO_BP) # do GO enrichment analysis with background gene set set as all gene expressed, database used is msigdb GO:BP

GO_enrichment_clusters_KEGG[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_KEGG)

GO_enrichment_clusters_HallMark[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_H)

GO_enrichment_clusters_REACT[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_REACT)

GO_enrichment_clusters_MF[[i]] <- getGOenrichment(genes_in_clusters[[i]], as.character(universe$entrezid),msigdbr_t2g_GO_MF)
}
```

### RNA-seq: format GO enrichment terms from each cluster into matrix of -log10(padj values)
```{r RNAseq GO enrichment results}
# convert each clusters' GO enrichment terms into dataframes
GO_enrichment_clusters_KEGG_df <- list()
 GO_enrichment_clusters_REACT_df <- list()
 GO_enrichment_clusters_HallMark_df <- list()
 GOBP_enrichment_clusters_df <- list()
GO_enrichment_clusters_MF_df <- list()
for(i in 1:length(GO_enrichment_clusters_REACT)){
  GO_enrichment_clusters_KEGG_df[[i]] <- as.data.frame(GO_enrichment_clusters_KEGG[[i]])
   GO_enrichment_clusters_REACT_df[[i]] <- as.data.frame(GO_enrichment_clusters_REACT[[i]])
 GO_enrichment_clusters_HallMark_df[[i]] <- as.data.frame(GO_enrichment_clusters_HallMark[[i]])
 GOBP_enrichment_clusters_df[[i]] <- as.data.frame(GOBP_enrichment_clusters[[i]])
  GO_enrichment_clusters_MF_df[[i]] <- as.data.frame(GO_enrichment_clusters_MF[[i]])
}

 # indicate names of each list as the cluster number
 names(GO_enrichment_clusters_KEGG_df) <- cluster_names 
 names(GO_enrichment_clusters_REACT_df) <- cluster_names 
 names(GO_enrichment_clusters_HallMark_df) <- cluster_names 
 names(GOBP_enrichment_clusters_df) <- cluster_names 
names(GO_enrichment_clusters_MF_df) <- cluster_names 
 
 # remove any cluster of genes that have 0 GO enrichment terms
GO_enrichment_clusters_KEGG_df <-  GO_enrichment_clusters_KEGG_df[sapply( GO_enrichment_clusters_KEGG_df, nrow)>0]
GO_enrichment_clusters_REACT_df <-  GO_enrichment_clusters_REACT_df[sapply( GO_enrichment_clusters_REACT_df, nrow)>0]
GO_enrichment_clusters_HallMark_df <-  GO_enrichment_clusters_HallMark_df[sapply( GO_enrichment_clusters_HallMark_df, nrow)>0]
GOBP_enrichment_clusters_df <-  GOBP_enrichment_clusters_df[sapply( GOBP_enrichment_clusters_df, nrow)>0]
GO_enrichment_clusters_MF_df <-  GO_enrichment_clusters_MF_df[sapply( GO_enrichment_clusters_MF_df, nrow)>0]
 
 # indicate which cluster and database each dataframe comes from 
 for(i in 1:length(GO_enrichment_clusters_KEGG_df)){
  GO_enrichment_clusters_KEGG_df[[i]]$Cluster <- names(GO_enrichment_clusters_KEGG_df)[i]
  GO_enrichment_clusters_KEGG_df[[i]]$Database <- "KEGG"
 }
  for(i in 1:length(GO_enrichment_clusters_REACT_df)){
   GO_enrichment_clusters_REACT_df[[i]]$Cluster <-  names(GO_enrichment_clusters_REACT_df)[i]
   GO_enrichment_clusters_REACT_df[[i]]$Database <- "REACTOME"
  }
  for(i in 1:length(GO_enrichment_clusters_HallMark_df)){
 GO_enrichment_clusters_HallMark_df[[i]]$Cluster <- names(GO_enrichment_clusters_HallMark_df)[i]
 GO_enrichment_clusters_HallMark_df[[i]]$Database <- "Hallmark"
  }
 
  for(i in 1:length(GOBP_enrichment_clusters_df)){
 GOBP_enrichment_clusters_df[[i]]$Cluster <- names(GOBP_enrichment_clusters_df)[i]
  GOBP_enrichment_clusters_df[[i]]$Database <- "GO:BP"
  }

  for(i in 1:length(GO_enrichment_clusters_MF_df)){
 GO_enrichment_clusters_MF_df[[i]]$Cluster <- names(GO_enrichment_clusters_MF_df)[i]
 GO_enrichment_clusters_MF_df[[i]]$Database <- "GO:MF"
  }

# combine GO enrichment results from different databases into 1 dataframe for each cluster of genes. 
GO_results_list <- vector()
GO_results_list$CellCycle1 <- rbind(GO_enrichment_clusters_KEGG_df$CellCycle1,
                                    GO_enrichment_clusters_HallMark_df$CellCycle1,
                                    GO_enrichment_clusters_REACT_df$CellCycle1,
                                    GOBP_enrichment_clusters_df$CellCycle1,
                                    GO_enrichment_clusters_MF_df$CellCycle1)

GO_results_list$CellCycle2 <- rbind(GO_enrichment_clusters_KEGG_df$CellCycle2,
                                    GO_enrichment_clusters_HallMark_df$CellCycle2,
                                    GO_enrichment_clusters_REACT_df$CellCycle2,
                                    GOBP_enrichment_clusters_df$CellCycle2,
                                    GO_enrichment_clusters_MF_df$CellCycle2)

GO_results_list$Palbo1 <- rbind(GO_enrichment_clusters_KEGG_df$Palbo1,     
                                GO_enrichment_clusters_HallMark_df$Palbo1 , 
                                GO_enrichment_clusters_REACT_df$Palbo1, 
                                GOBP_enrichment_clusters_df$Palbo1, 
                                GO_enrichment_clusters_MF_df$Palbo1)

GO_results_list$Palbo2 <- rbind(GO_enrichment_clusters_KEGG_df$Palbo2,
                                GO_enrichment_clusters_HallMark_df$Palbo2,
                                GO_enrichment_clusters_REACT_df$Palbo2,
                                GOBP_enrichment_clusters_df$Palbo2, 
                                GO_enrichment_clusters_MF_df$Palbo2)

GO_results_list$Doxo1 <- rbind(GO_enrichment_clusters_KEGG_df$Doxo1,
                               GO_enrichment_clusters_HallMark_df$Doxo1,
                               GO_enrichment_clusters_REACT_df$Doxo1,
                               GOBP_enrichment_clusters_df$Doxo1, 
                               GO_enrichment_clusters_MF_df$Doxo1)

GO_results_list$Doxo2 <- rbind(GO_enrichment_clusters_KEGG_df$Doxo2,
                               GO_enrichment_clusters_HallMark_df$Doxo2,
                               GO_enrichment_clusters_REACT_df$Doxo2,
                               GOBP_enrichment_clusters_df$Doxo2, 
                               GO_enrichment_clusters_MF_df$Doxo2)

GO_results_list$Shared <- rbind(GO_enrichment_clusters_KEGG_df$Shared,
                                GO_enrichment_clusters_HallMark_df$Shared,
                                GO_enrichment_clusters_REACT_df$Shared,
                                GOBP_enrichment_clusters_df$Shared, 
                                GO_enrichment_clusters_MF_df$Shared)

# order GO enrichment results from different databases by padj values
for(i in 1:length(GO_results_list)){
  GO_results_list[[i]] <- GO_results_list[[i]][order(GO_results_list[[i]]$p.adjust), ]
}

# combine dataframes of each cluster of gene's GO enrichment results into 1 master dataframe
GO_results_df <- do.call(rbind, GO_results_list)

# create matrix of -log10pvalues of the top 50 enriched pathways for each cluster

GO_results_list_mat <- matrix(nrow=length(unique(unlist(lapply(GO_results_list, function(GO_results_list){GO_results_list$ID[1:50]})))), ncol=length(GO_results_list)) # make empty matrix, number of columns are the clusters of genes and number of total top GO terms of all clusters

row.names(GO_results_list_mat) <- unique(unlist(lapply(GO_results_list, function(GO_results_list){GO_results_list$ID[1:50]}))) # row names are top 12 GO enrichment terms for each cluster combined
colnames(GO_results_list_mat) <- cluster_names  # column names are clusters

# pathways that are not enriched in the cluster are listed as NA so just change the value to 0 so that it can be plotted in heatmap
for(i in 1:ncol(GO_results_list_mat)){
GO_results_list_mat[ ,i] <- -log10(GO_results_list[[i]][match(row.names(GO_results_list_mat), GO_results_list[[i]]$ID), ]$p.adjust)
GO_results_list_mat[ ,i] <- ifelse(GO_results_list_mat[ ,i] %in% NA, 0, GO_results_list_mat[ ,i])
}
GO_results_list_mat <- GO_results_list_mat[!is.na(row.names(GO_results_list_mat)), ] # remove padj values that are NA 

# specify GO results matrix annotation colors: based on ggsci:pal_aaas palette. 
GO_anno_colors <-  list(database=c("GO:BP"="#3B4992FF", "Hallmark"="#EE0000FF", "KEGG"="#008B45FF", "REACTOME" ="#631879FF", "GO:MF"="#808180FF"), 
                        Cluster=c("CellCycle1"="#66C2A5", "CellCycle2"="#A6D854", "Palbo1"="#8DA0CB", "Palbo2"= "#FC8D62","Doxo1"="#E78AC3","Doxo2"="#FFD92F", "Shared"="#E5C494"))
names(GO_anno_colors$Cluster) <- cluster_names 

# annotation dataframe for rows of heatmap to indicate which database GO term came from 
GO_anno <- data.frame(row.names=row.names(GO_results_list_mat), database=GO_results_df[match(row.names(GO_results_list_mat), GO_results_df$ID), ]$Database)
# annotation dataframe for columns of heatmap to color by Cluster colors. 
Cluster_anno <- data.frame(Cluster=cluster_names , row.names = colnames(GO_results_list_mat))
```

# RNA-seq: GSEA Analysis -- note: for Fig S4
```{r RNAseq GSEA analysis}
ranked_res_forGSEA <- list()
for(i in 1:length(LFCResultsall_RNA)){
  
  rank <- LFCResultsall_RNA[[i]]
  
  # Assign a default p-value for genes with NA padj (maximum p-value)
  rank$pvalue[is.na(rank$padj)] <- 1 
  
  # Create a ranked list based on log2FoldChange and -log10(p-value)
  rank$rank_metric <- sign(rank$log2FoldChange) * -log10(rank$pvalue)
  
  # Sort by the ranking metric
  rank <- rank[order(rank$rank_metric, decreasing = TRUE), ]
  
  # Store the ranked values
  ranked_res_forGSEA[[i]] <- rank$rank_metric
  names(ranked_res_forGSEA[[i]]) <- row.names(rank)

  # Convert gene IDs to Entrez IDs if available
  entrez_ids <- genes_anno_type[match(names(ranked_res_forGSEA[[i]]), genes_anno_type$ensembl_id), ]$entrezid
  names(ranked_res_forGSEA[[i]]) <- ifelse(is.na(entrez_ids), names(ranked_res_forGSEA[[i]]), entrez_ids)
}

names(ranked_res_forGSEA) <- names(LFCResultsall_RNA)

# run GSEA analysis using: 
GSEA_Results_H <- list() # msigdb Hallmark database
GSEA_Results_KEGG <- list() # msigdb KEGG database
GSEA_Results_REACT <- list() # msigdb Reactome database
GSEA_Results_MF <- list() # msigdb GO:MF database
for(i in 1:length(ranked_res_forGSEA)){
 GSEA_Results_H[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_H) 
 GSEA_Results_KEGG[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_KEGG) 
 GSEA_Results_REACT[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_REACT) 
 GSEA_Results_MF[[i]] <- GSEA(ranked_res_forGSEA[[i]], TERM2GENE = msigdbr_t2g_GO_MF) 
}

names(GSEA_Results_H) <- names(LFCResultsall_RNA)
names(GSEA_Results_KEGG) <- names(LFCResultsall_RNA)
names(GSEA_Results_REACT) <- names(LFCResultsall_RNA)
names(GSEA_Results_MF) <- names(LFCResultsall_RNA)
```

# RNA-seq: generate input files for IPA pathways. 
```{r}
# clusters of genes 
for(i in 1:length(genes_in_clusters)){
  
  IPA_input <- as.data.frame(genes_in_clusters[[i]])
  colnames(IPA_input) <- "ID"
  write.table(IPA_input, file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/IPA_files_kmeans7_clusters/input/", cluster_names[i], ".csv"), row.names=F, sep=",")
}
# DESeq Wald Test results for pairwise comparisons. 
for(i in 1:length(LFCResultsall_RNA)){
  write.table(as.data.frame(LFCResultsall_RNA[[i]]), file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/IPA_files/DESeq_results_input/", names(LFCResultsall_RNA)[i], ".csv"), sep=",")
}

# log2FC values of each sig changing gene clusters. 
for(i in 1:length(drug_cat_list)){
table <- sig_LFCMat[row.names(sig_LFCMat) %in% row.names(drug_cat_list[[i]]), ]
    write.table(table, file=paste0("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/IPA_files_kmeans7_clusters/input/sigGenes_log2FC/", cluster_names[i], "_log2FC.csv"), sep=",", row.names = T)
}
```

# RNA-seq: import IPA upstream regulators & pathways
```{r import IPA results}
IPA_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/IPA_files_kmeans7_clusters/output"
IPA_upstream_regs <- list()
IPA_pathways <- list()
# read in information for enriched IPA pathways per hierarchical cluster (clustering based on scaled rlog normalized counts)
for(i in 1:length(cluster_names)){
IPA_pathways[[i]]<- read_tsv(paste0(IPA_dir, "/", cluster_names[i], "_IPA_pathways.txt"), skip=2)
# read in information for enriched upstream regulators 
IPA_upstream_regs[[i]] <- read.table(paste0(IPA_dir, "/", cluster_names[i], "_IPA_upstream_regs.txt"), sep="\t", skip=2, header=T)
IPA_upstream_regs[[i]] <- IPA_upstream_regs[[i]][!grepl("mir|miR|Mir", IPA_upstream_regs[[i]]$Upstream.Regulator), ] # remove upstream regulators that are miRNAs (uncertain if they will be informative & there is a lot of them)
}

names(IPA_upstream_regs) <- cluster_names
names(IPA_pathways) <- cluster_names

# filter out upstream regulators that are not expressed in this cell line. 

#extract ids of all expressed genes
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL

# keep only upstream regulators that are part of expressed genes or have lower case letters (these are upstream regulators that could be miRNAs or protein complexes)
#IPA_upstream_regs <- lapply(IPA_upstream_regs, function(IPA_upstream_regs){IPA_upstream_regs[c(grep('[:lower:]', IPA_upstream_regs$Upstream.Regulator), which(IPA_upstream_regs$Upstream.Regulator %in% expressedGenes)), ]})
```

# RNA-seq: plotting padj values of top IPA predicted upstream regulators of all clusters -- note: For Fig 5B & Fig S2B
```{r IPA results matrix}
create_ipa_regs_matrix <- function(IPA_upstream_regs, cluster_names, filter_keyword = "complex", top_n = 25) {
  # Extract unique upstream regulators based on the filter keyword
  unique_upstream_regs <- unique(unlist(lapply(IPA_upstream_regs, function(x) {
    x$Upstream.Regulator[grepl(filter_keyword, x$Molecule.Type)][1:top_n]
  })))

  # Create a matrix to store -log10 p-values
  IPA_complex_regs_mat <- matrix(
    nrow = length(unique_upstream_regs),
    ncol = length(IPA_upstream_regs)
  )
  row.names(IPA_complex_regs_mat) <- unique_upstream_regs # Set row names to upstream regulators
  colnames(IPA_complex_regs_mat) <- cluster_names        # Set column names to cluster names

  # Fill in the matrix with -log10 p-values
  for (i in seq_len(ncol(IPA_complex_regs_mat))) {
    # Match and extract p-values for each cluster
    IPA_complex_regs_mat[, i] <- -log10(IPA_upstream_regs[[i]][
      match(row.names(IPA_complex_regs_mat), IPA_upstream_regs[[i]]$Upstream.Regulator),
      "p.value.of.overlap"
    ])

    # Replace NA values with 0 for heatmap compatibility
    IPA_complex_regs_mat[, i][is.na(IPA_complex_regs_mat[, i])] <- 0
  }

  return(IPA_complex_regs_mat)
}

IPA_group_regs_mat <- create_ipa_regs_matrix(
  IPA_upstream_regs = IPA_upstream_regs,
  cluster_names = cluster_names,
  filter_keyword = "group",
  top_n = 25
)

IPA_TF_regs_mat <- create_ipa_regs_matrix(
  IPA_upstream_regs = IPA_upstream_regs,
  cluster_names = cluster_names,
  filter_keyword = "transcription",
  top_n = 25
)

# plot heatmap of -log10 pvalue of overlap of enriched complex regulators
pheatmap(IPA_group_regs_mat, cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 10,length.out = 100))

pheatmap(IPA_TF_regs_mat, cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 10,length.out = 100))
```

# RNA-seq: import comparison analysis of IPA upstream regulators and pathways for sig changing genes
```{r import_IPA_comparison_results}
setwd("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/IPA_files_kmeans7_clusters/output/sigGenes_log2FC")
IPA_Compare_upstream_regs <- list()
IPA_Compare_pathways <- list()
# read in information for enriched IPA pathways per hierarchical cluster (clustering based on scaled rlog normalized counts)

upstream_regs_files <- dir(pattern="_IPA_upstream_regs.txt")
pathways_files <- dir(pattern="_IPA_pathways.txt")
for(i in 1:length(pathways_files)){
# read in information for enriched pathways
IPA_Compare_pathways[[i]]<- read_tsv(pathways_files[i], skip=2)
# read in information for enriched upstream regulators 
IPA_Compare_upstream_regs[[i]] <- read_tsv(upstream_regs_files[i], skip=2)

  IPA_Compare_pathways[[i]] <- IPA_Compare_pathways[[i]] %>%
  mutate(across(where(is.character), ~ replace(.x, .x == "N/A", "0"))) %>%
  mutate(across(-`Canonical Pathways`, as.numeric)) # Convert all except the first column to numeric
  
  IPA_Compare_upstream_regs[[i]] <- IPA_Compare_upstream_regs[[i]] %>%
  mutate(across(where(is.character), ~ replace(.x, .x == "N/A", "0"))) %>%
    mutate(across(-`Upstream Regulators`, as.numeric)) # Convert all except the first column to numeric
}

names(IPA_Compare_upstream_regs) <- gsub("_.*", "", pathways_files)
names(IPA_Compare_pathways) <- gsub("_.*", "", pathways_files)

# convert IPA results into matrix for plotting. 
IPA_Compare_pathways_mat <- lapply(IPA_Compare_pathways, function(x){
  mat <- as.matrix(x[ ,-1])
  row.names(mat) <- x$`Canonical Pathways`
 remove <- which(rowSums(mat != 0) == 0) # Find rows where the entire row (all columns) is exactly 0 to remove upstream regulators that have no activation score across all columns (DESeq comparisons)
    return(mat[-remove, ])
})

IPA_Compare_upstream_regs_mat <- lapply(IPA_Compare_upstream_regs, function(x){
  mat <- as.matrix(x[ ,-1])
  row.names(mat) <- x$`Upstream Regulators`
  remove <- which(rowSums(mat != 0) == 0) # Find rows where the entire row (all columns) is exactly 0 to remove upstream regulators that have no activation score across all columns (DESeq comparisons)
  return(mat[-remove, ])
})
```

# RNA-seq: make matrix with log2FC values of significant genes
```{r log2FC sig genes matrix}
# get log2FC values of significantly changing genes over time
sig_LFCMat <- matrix(ncol=length(LFCResultsall_RNA), nrow=nrow(clusterDF))
colnames(sig_LFCMat) <- names(LFCResultsall_RNA)
row.names(sig_LFCMat) <-row.names(LFCResultsall_RNA[[10]][row.names(LFCResultsall_RNA[[10]]) %in% sigGenes, ])
for(i in 1:ncol(sig_LFCMat)){
sig_LFCMat[ ,i] <- LFCResultsall_RNA[[i]][row.names(LFCResultsall_RNA[[i]]) %in% sigGenes, ]$log2FoldChange
}

row.names(sig_LFCMat) <-  make.unique(ifelse(genes_anno_type[match(row.names(sig_LFCMat), genes_anno_type$ensembl_id), ]$gene_id %in% NA, row.names(sig_LFCMat), genes_anno_type[match(row.names(sig_LFCMat), genes_anno_type$ensembl_id), ]$gene_id))

sig_LFCMat <- sig_LFCMat[match(row.names(ordered_sigMat), row.names(sig_LFCMat)), ] # order based on k-means & hierarchical ordering of rlog counts. 

# get log2FC values of all expressed genes over time 
LFCMat <- matrix(ncol=length(LFCResultsall_RNA), nrow=nrow(LFCResultsall_RNA[[1]]))
colnames(LFCMat) <- names(LFCResultsall_RNA)
row.names(LFCMat) <-row.names(LFCResultsall_RNA[[1]])
for(i in 1:ncol(LFCMat)){
LFCMat[ ,i] <- LFCResultsall_RNA[[i]]$log2FoldChange
}

row.names(LFCMat) <-  ifelse(genes_anno_type[match(row.names(LFCMat), genes_anno_type$ensembl_id), ]$gene_id %in% NA, row.names(LFCMat), genes_anno_type[match(row.names(LFCMat), genes_anno_type$ensembl_id), ]$gene_id)


anno_sigLFC <- data.frame(
  row.names = colnames(sig_LFCMat),
  log2FC = c(rep("vs Cyc", 4), rep("vs day3", 3), rep("vs Cyc", 5), rep("vs day3", 4), rep("DoxovsPalbo",4)),
  Treatment = c(rep("Doxo", 7), rep("Palbo", 9), rep("DoxovsPalbo",4)),
  Timepoint = c("3", "10", "14", "21", "10", "14", "21", "3", "10", "14", "21", "28", "10", "14", "21", "28","3", "10", "14", "21")
)
```

# ATAC-seq: create input fasta files for MEME-AME analysis on all up or down peaks. 
```{r}
directory <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/motif_analysis_input/DESeq"
outdir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_input"


# specify file paths for input to MEME-AME analysis for up peaks
folderlist <- c("day10PalbovsDoxoUpPeaks_vsCycling", 
                "day10PalbovsDoxoUpPeaks_vsday3", 
                "day14PalbovsDoxoUpPeaks_vsCycling", 
                "day14PalbovsDoxoUpPeaks_vsday3",
                "day21PalbovsDoxoUpPeaks_vsCycling", 
                "day21PalbovsDoxoUpPeaks_vsday3",
                "day28Palbovsday21DoxoUpPeaks_vsCycling", 
                "day28Palbovsday21DoxoUpPeaks_vsday3",
                "day3PalbovsDoxoUpPeaks_vsCycling")

writeLines(dir(paste0(directory, "/annoUp"), full.names=T, pattern="Palbovs"), con=paste0(outdir, "/inputfiles_annoUp_DESeq_Palbopeaks.txt")) # directory paths to fasta files of Up peaks

writeLines(c(dir(paste0(directory, "/annoUp"), full.names=T, pattern="Doxovs")[1:6], dir(paste0(directory, "/annoUp"), full.names=T, pattern="Doxovs")[5:7]), con=paste0(outdir, "/inputfiles_annoUp_DESeq_Doxopeaks.txt")) # directory paths to fasta files of Up peaks 
# since Doxo doesn't have day 28 timepoint, repeat day 21 DESeq comparison files for comparing to day28 Palbo. 

writeLines(folderlist, con=paste0(outdir, "/folderlist_annoUp_DESeq_PalbovsDoxoPeaks.txt")) # write folder names
```

### ATAC-seq: MEME-AME analysis on DESeq_Peaks pairwise comparison between up Palbo and up Doxo peaks to find differences in enriched motifs between upregulated drug distinct peaks. 
```{bash}
cd /lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_input

# what is enriched in Palbo up peaks that is different from Doxo up peaks? set Doxo up peaks as background. 
sbatch -p risc,hpc --array=1-9 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_input/inputfiles_annoUp_DESeq_Doxopeaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_input/inputfiles_annoUp_DESeq_Palbopeaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_output/annoUp_PalbovsDoxo,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_input/folderlist_annoUp_DESeq_PalbovsDoxoPeaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm

# what is enriched in Doxo up peaks that is different from Palbo up peaks? set Palbo up peaks as background.  
sbatch -p risc,hpc --array=1-9 \
--export=backgroundpeakslist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_input/inputfiles_annoUp_DESeq_Palbopeaks.txt,\
inputfiles=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_input/inputfiles_annoUp_DESeq_Doxopeaks.txt,\
outdir=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_output/annoUp_DoxovsPalbo,\
folderlist=/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_input/folderlist_annoUp_DESeq_PalbovsDoxoPeaks.txt \
/lustre/fs4/risc_lab/scratch/jyeung/LS_PDvsDoxo_2/ATACseq/post_peakcalling_analysis/06222024/motif_analysis/motif_analysis_input/MEME_AME.slurm
```

# ATAC-seq: import MEME-AME results into dataframe: MEME-AME was done to find differential motifs under up peaks between Doxo & Palbo at different timepoints relative to cycling or day 3. 
```{r import_differential_MEMEAME_results}
expressedGenes <- LFCResultsall_RNA[[1]]$SYMBOL
expressedGenes[grep("TP53$|TP63$|TP73$", expressedGenes)] <- c("P63", "P73", "P53")
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS6/Functions/process_motif_data.RData")

# differential motifs: Doxo vs Palbo
annoUp_DvsP_AME <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_output/annoUp_DoxovsPalbo", expressed_genes = expressedGenes)

# Add column to specify which RNA cluster
annoUp_DvsP_AME$Comparison <- gsub('\\.[0-9]*$', "",  row.names(annoUp_DvsP_AME)) 
annoUp_DvsP_AME$Comparison <- factor(annoUp_DvsP_AME$Comparison, levels=unique(annoUp_DvsP_AME$Comparison)[c(9,1,3,5,7,2,4,6,8)])
# Add column to specify whether up peaks are relative to cycling or day 3 controls. 
annoUp_DvsP_AME$Group <- ifelse(grepl("vsCycling", annoUp_DvsP_AME$Comparison), "vsCycling", "vsday3")

# differential motifs: Palbo vs Doxo
annoUp_PvsD_AME <- process_motif_data("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/motif_analysis_output/annoUp_PalbovsDoxo", expressed_genes = expressedGenes)

# Add column to specify which RNA cluster
annoUp_PvsD_AME$Comparison <- gsub('\\.[0-9]*$', "",  row.names(annoUp_PvsD_AME)) 
annoUp_PvsD_AME$Comparison <- factor(annoUp_PvsD_AME$Comparison, unique(annoUp_DvsP_AME$Comparison)[c(9,1,3,5,7,2,4,6,8)])

# Add column to specify whether up peaks are relative to cycling or day 3 controls. 
annoUp_PvsD_AME$Group <- ifelse(grepl("vsCycling", annoUp_PvsD_AME$Comparison), "vsCycling", "vsday3")
```

-------------------------------------------------------------------------------------------------------------------------
# *FIGURES*

# RNA-seq: plot heatmap of drug distinct genes, clustered based on log2FC from day 3 treatment controls & Cycling controls
```{r}
sig_LFCMat_drugdistinct <- sig_LFCMat[row.names(sig_LFCMat) %in% clusterDF[clusterDF$Clusters %in% c("7", "1", "4", "6"), ]$gene_id & !row.names(sig_LFCMat) %in% c("Y_RNA", "Metazoa_SRP"), ]

pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4"

pdf(paste0(pdf_dir, "/RNA_log2FC_drug_distinct_sig_LFCmat.pdf"), width=4.35, height=5)
pheatmap(sig_LFCMat_drugdistinct, show_rownames = F, show_colnames=F, cluster_cols = F, cluster_rows=F, colorRampPalette(rev(brewer.pal(n = 7, name ="RdBu")))(100), breaks = seq(-4.5, 4.5,length.out = 100), annotation_row = clusterDF[ ,c("SASP", "Clusters")], annotation_col = anno_sigLFC, annotation_colors = hm_cols_RNA, border_color = NA, main="drug distinct upreg genes")
dev.off()
```

# RNA-seq: plotting padj values of top IPA predicted upstream regulators of drug distinct gene clusters. 
```{r}
filter_upstream_regulators <- function(mat, threshold_neg1 = -0.5, threshold_pos1 = 0.5,  threshold_neg2 = -0.5, threshold_pos2 = 0.5) {
  # Identify upstream regulators that are differentially regulated between Doxo & Palbo
  filt_negDoxo_posPalbo <- which(rowSums(mat[, 1:5]) <= threshold_neg1 & rowSums(mat[, 6:10]) >= threshold_pos1)
  filt_posDoxo_negPalbo <- which(rowSums(mat[, 1:5]) >= threshold_pos2 & rowSums(mat[, 6:10]) <= threshold_neg2)
  
  # Combine filtered indices
  filt_mat <- mat[c(filt_negDoxo_posPalbo, filt_posDoxo_negPalbo), , drop = FALSE]
  
  # Remove random drugs/chemicals (upstream regulators that contain at least 2 uppercase letters)
  filt_mat <- filt_mat[grep("[A-Z].*[A-Z]", rownames(filt_mat)), , drop = FALSE]
  
  return(filt_mat)
}

# plot most differentially activated upstream regulators of Palbo1 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Palbo1, threshold_neg1 = -8.5, threshold_pos1 = 10,  threshold_neg2 = -6, threshold_pos2 = 8) # set differential thresholds for Palbo vs Doxo. 
filt_mat <- filt_mat[!row.names(filt_mat) %in% c("SB203580", "LY294002", "SB-431542"), ]  # remove upstream regulators that are drugs/chemicals

length(row.names(filt_mat)[row.names(filt_mat) %in% expressedGenes])/length(row.names(filt_mat)) # double check that the upstream regulators are expressed genes
# only"TGF beta" dropped out which is fine because its a family, not a specific gene. 

pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4"
# plot heatmap of -log10 pvalue of overlap of enriched complex regulators
pdf(paste0(pdf_dir, "/RNA_IPA_Palbo1_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,6:10]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-3, 3,length.out = 100), main="Activation scores")
dev.off() # order by top to bottom sum of scores. 

# plot most differentially activated upstream regulators of Palbo2 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Palbo2, threshold_neg1 = 1.2, threshold_pos1 = 6.2, threshold_neg2 = -0.5, threshold_pos2 =0.5)

length(row.names(filt_mat)[row.names(filt_mat) %in% expressedGenes])/length(row.names(filt_mat)) # double check that the upstream regulators are expressed genes

# finding genes that did not appear in expressed genes using their ENSG identifier to double check it really isn't being expressed. 
# normCounts[grepl("ENSG00000112116", row.names(normCounts)), ]

filt_mat <- filt_mat[grep("IL5|IL17F|VH032|IOX2", rownames(filt_mat), invert = T), , drop = FALSE]
nrow(filt_mat) # filter out chemicals and non-expressed genes 

# only"TGF beta" dropped out which is fine because its a family, not a specific gene. 

pdf(paste0(pdf_dir, "/RNA_IPA_Palbo2_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,6:10]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")
dev.off()

# plot most differentially activated upstream regulators of Doxo1 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Doxo1, threshold_neg1 = -4.2, threshold_pos1 = 4.2, threshold_neg2 = -4.2, threshold_pos2 =4.2)

length(row.names(filt_mat)[row.names(filt_mat) %in% expressedGenes])/length(row.names(filt_mat)) # double check that the upstream regulators are expressed genes
row.names(filt_mat)[!row.names(filt_mat) %in% expressedGenes]
filt_mat <- filt_mat[grep("GW9662|PG3-Oc", rownames(filt_mat), invert = T), , drop = FALSE]  # filter out chemicals and non-expressed genes. AURK and p38 MAPK are encoded by AURKA, MAPK14 respectively 

pdf(paste0(pdf_dir, "/RNA_IPA_Doxo1_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,1:5]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")
dev.off()

# plot most differentially activated upstream regulators of Doxo2 upregulated genes
filt_mat <- filter_upstream_regulators(IPA_Compare_upstream_regs_mat$Doxo2, threshold_neg1 = -6.5, threshold_pos1 = 0,  threshold_neg2 = -2, threshold_pos2 = 0.5)

length(row.names(filt_mat)[row.names(filt_mat) %in% expressedGenes])/length(row.names(filt_mat)) # double check that the upstream regulators are expressed genes
row.names(filt_mat)[!row.names(filt_mat) %in% expressedGenes]
normCounts[grep("ENSG00000182968|ENSG00000134595|ENSG00000222028", row.names(normCounts)), ]
filt_mat <- filt_mat[grep("GW9662|PG3-Oc|PSMB11|SOX1|SOX3", rownames(filt_mat), invert = T), , drop = FALSE]  # filter out chemicals and non-expressed genes. AURK and p38 MAPK are encoded by AURKA, MAPK14 respectively 

pdf(paste0(pdf_dir, "/RNA_IPA_Doxo2_upstream_regs_Comparison.pdf"), width=3, height=3)
pheatmap(filt_mat[order(rowSums(filt_mat[ ,1:5]), decreasing=T), ], cluster_cols = F, cluster_rows=F,  colorRampPalette(rev(brewer.pal(n = 7, name ="PuOr")))(100), breaks = seq(-2.5, 2.5,length.out = 100), main="Activation scores")
dev.off()

```

# ATAC-seq: plot differentially enriched motifs between Palbo-up vs Doxo-up peaks and Doxo-up vs Palbo-up peaks.
```{r}
# load in required functions 
load("/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/FigureS3/Functions/create_AME_peaks_matrix.RData")

# get top enriched motifs for Doxo vs Palbo MEME-AME analysis results
AME_padj_mat_DvsP <- list()
comparisons <- levels(annoUp_DvsP_AME$Comparison)

for(i in 1:length(comparisons)){
AME_padj_mat_DvsP[[i]] <- create_AME_peaks_matrix(annoUp_DvsP_AME[annoUp_DvsP_AME$Comparison %in% comparisons[i], ])

AME_padj_mat_DvsP[[i]] <- AME_padj_mat_DvsP[[i]][order(rowSums(AME_padj_mat_DvsP[[i]]), decreasing=T), ] # sort motif IDs by overall significance across clusters. 
}

names(AME_padj_mat_DvsP) <- comparisons

AME_padj_dots_DvsP <- list()
for(i in 1:length(comparisons)){
AME_padj_dots_DvsP[[i]] <- annoUp_DvsP_AME[annoUp_DvsP_AME$Comparison %in% comparisons[i], ]

max_number <- ifelse(nrow(AME_padj_mat_DvsP[[i]]) >=7, 7, nrow(AME_padj_mat_DvsP[[i]]))
AME_padj_dots_DvsP[[i]] <- AME_padj_dots_DvsP[[i]][AME_padj_dots_DvsP[[i]]$motif_ID %in% row.names(AME_padj_mat_DvsP[[i]][1:max_number, ]), ] # show only top 15 most significant padj values motif IDs
AME_padj_dots_DvsP[[i]]$motif_ID <- factor(AME_padj_dots_DvsP[[i]]$motif_ID, levels=rev(unique(AME_padj_dots_DvsP[[i]]$motif_ID)))
}

AME_padj_dots_DvsP <- do.call(rbind, AME_padj_dots_DvsP)


# get top enriched motifs for Doxo vs Palbo MEME-AME analysis results
AME_padj_mat_PvsD <- list()
comparisons <- levels(annoUp_PvsD_AME$Comparison)

for(i in 1:length(comparisons)){
AME_padj_mat_PvsD[[i]] <- create_AME_peaks_matrix(annoUp_PvsD_AME[annoUp_PvsD_AME$Comparison %in% comparisons[i], ])

AME_padj_mat_PvsD[[i]] <- AME_padj_mat_PvsD[[i]][order(rowSums(AME_padj_mat_PvsD[[i]]), decreasing=T), ] # sort motif IDs by overall significance across clusters. 
}

names(AME_padj_mat_PvsD) <- comparisons

AME_padj_dots_PvsD <- list()
for(i in 1:length(comparisons)){
AME_padj_dots_PvsD[[i]] <- annoUp_PvsD_AME[annoUp_PvsD_AME$Comparison %in% comparisons[i], ]

max_number <- ifelse(nrow(AME_padj_mat_PvsD[[i]]) >=7, 7, nrow(AME_padj_mat_PvsD[[i]]))
AME_padj_dots_PvsD[[i]] <- AME_padj_dots_PvsD[[i]][AME_padj_dots_PvsD[[i]]$motif_ID %in% row.names(AME_padj_mat_PvsD[[i]][1:max_number, ]), ] # show only top 15 most significant padj values motif IDs
AME_padj_dots_PvsD[[i]]$motif_ID <- factor(AME_padj_dots_PvsD[[i]]$motif_ID, levels=rev(unique(AME_padj_dots_PvsD[[i]]$motif_ID)))
}

AME_padj_dots_PvsD <- do.call(rbind, AME_padj_dots_PvsD)

# create function for plotting motif enrichment results as dotplot
create_AME_dotplots <- function(AME_df=AME_padj_dots, Title=Title, fill_limit=50, size_limit=4.5 ){
  ggplot(AME_df, aes(x=Comparison, y=motif_ID, size=TPvsFP)) +
    geom_point(aes(fill=neglog10padj), colour="black", pch=21, alpha=0.8) +
    scale_fill_gradient(name="-log10(padj)", 
                        low="light yellow", high="dark orange",
                        limits=c(0, fill_limit), 
                        oob=scales::squish) + # Adjust fill scale
    scale_size_continuous(name="%TP - %FP", 
                          limits=c(0, size_limit), 
                          range=c(1, 10)) +  # Adjust dot sizes
    ggtitle(Title) +
    theme_classic()+
    theme(
      text = element_text(size=8), 
      axis.text.x = element_text(size=8, angle=90), 
      strip.text.x = element_text(size=8, face="bold"),
      strip.text.y = element_text(size=8, face="bold")
    ) +
    facet_wrap(~Group, nrow=1, scales="free") +
    labs(x="Comparison", y="motif ID")
}


pdf_dir <- "/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4"
pdf(paste0(pdf_dir, "/ATAC_PalbovsDoxo_top_differential_motifs_UpPeaks.dotplots.pdf"), width=5, height=7)
create_AME_dotplots(AME_padj_dots_DvsP, Title="Doxo vs Palbo Up peaks", fill_limit = 200, size_limit = 35) 
create_AME_dotplots(AME_padj_dots_PvsD, Title="Palbo vs Doxo Up peaks", fill_limit = 100, size_limit = 25) 
dev.off()
```

# RNA-seq: plot padj values for top 5 IPA upstream regs for each cluster individually. -- Fig 5B & Fig S2B 
```{r}
# Still showing the pvalues of the other clusters as columns but only the top IPA upstream regs from the cluster of interest. 
IPA_TF_regs_clusters <- list()
IPA_group_regs_clusters <- list()
for(i in 1:length(IPA_upstream_regs)){
  #transcription regulators
IPA_TF_regs_clusters[[i]] <-IPA_TF_regs_mat[row.names(IPA_TF_regs_mat) %in% IPA_upstream_regs[[i]]$Upstream.Regulator[grepl("transcription", IPA_upstream_regs[[i]]$Molecule.Type)][1:5], ]
IPA_TF_regs_clusters[[i]] <- IPA_TF_regs_clusters[[i]][order(IPA_TF_regs_clusters[[i]][ ,i], decreasing=T), ]
  # complex regulators
IPA_group_regs_clusters[[i]] <-IPA_group_regs_mat[row.names(IPA_group_regs_mat) %in% IPA_upstream_regs[[i]]$Upstream.Regulator[grepl("group", IPA_upstream_regs[[i]]$Molecule.Type)][1:5], ]
IPA_group_regs_clusters[[i]] <- IPA_group_regs_clusters[[i]][order(IPA_group_regs_clusters[[i]][ ,i], decreasing=T), ] # sort by decreasing padj value
}

# make heatmap of top 5 IPA predicted upstream regulators for drug distinct upregulated genes
pdf(file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/RNA_IPA_TF_regs_drug_distinct_up_genes.pdf", width=3, height=2)
pheatmap(IPA_TF_regs_clusters[[6]], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 100,length.out = 100), cluster_rows = F, main="Shared down genes: predicted TF regulators", border_color = "black",  annotation_col=Cluster_anno, annotation_colors = GO_anno_colors)
dev.off()

pdf(file="/lustre/fs4/risc_lab/scratch/jyeung/Manuscript_Figures/Figure4/RNA_IPA_group_regs_SharedDown_genes.pdf", width=3.8, height=2)
pheatmap(IPA_group_regs_clusters[[1]], cluster_cols = F, colorRampPalette(brewer.pal(n = 7, name ="Greens"))(100), breaks = seq(0, 100,length.out = 100), cluster_rows = F, main="Shared down genes: predicted group regulators", border_color = "black", annotation_col=Cluster_anno, annotation_colors = GO_anno_colors)
dev.off()
```


